### 2017-10-01 manuscript.Rmd
1-10-2017 first draft

---
title             : "Equivalence Testing for Psychological Science"
shorttitle        : "Equivalence Testing"

author: 
  - name          : "Daniel Lakens"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands"
    email         : "D.Lakens@tue.nl"
  - name          : "Anne M. Scheel"
    affiliation   : "1"
  - name          : "Peder Isager"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Eindhoven University of Technology"

author_note: >
  We would like to thank Courtney Soderbergh for creating the first version of the TOST function to test two independent proportions.

abstract: >
  Psychologists  need to be able to test for the absence of an effect. Using the Two-One-Sided Tests (TOST) procedure, researchers can easily test whether the observed effects are too small to be meaningful. By specifying a smallest effect size of interest (SESOI) researchers test whether observed effects are surprisingly closer to zero, assuming there was an effect the consider meaningful. We explain a range of approaches to determine the SESOI in psychological science, and provide detailed examples of how equivalence tests should be performed and reported.

keywords          : "Equivalence Testing, NHST, power, TOST"
wordcount         : "X"

bibliography      : ["r-references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

```{r include = FALSE}
library("papaja")
```

# Justifying the Smallest Effect Size of Interest

Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI). The SESOI is sometimes based on just noticeable differences, which can be objectively determined. Most often, however, it is a subjective decision that varies across individuals and time. Both these approaches are valid, and even when a just noticeable difference can be objectively determined, researchers might choose to set the SESOI to a larger value. In many research areas, the SESOI is based on a cost-benefit analysis. Since both costs and benefits are necessarily relative, the SESOI will depend on the researcher who designs the study. The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base. The goal is not to determine a SESOI that is universally valid: The goal is to set a SESOI such that inferences based on it answer a meaninful question.

##Objective Justifications of a SESOI

An objectively determined SESOI should be based on quantifiable theoretical predictions, such as computational models. Sometimes, the only theoretical prediction is that an effect should be noticeable. In such circumstances, the SESOI can be set based on just noticeable differences. For example, Burriss and colleagues (2015) examined whether women displayed an increase in redness in the face during the fertile phase of their ovulatory cycle. The hypothesis was that a slighly redder skin signals greater attractiveness and physical health, and sending this signal to men yields an evolutionary advantage. This hypothesis requires that the increase in redness can be detected with the naked eye by men. Penke et al (2017) collected data from X women, and showed that there was indeed an increase in redness of the facial skin of woman during their fertile period. However, this increase was not large enough to be noticeable with the naken eye by men, thus falsifying their hypothesis. Because the just noticeable difference in redness of the skin can be measured, it is possible to objectively establish the SESOI.

Another example of an objectively determined SESOI can be found in Button and colleagues (2015) where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking X patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.

##Subjective justifications of a SESOI

We distinguish between three categories of subjective justifications for SESOI. First, researchers can use benchmarks. For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988). Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small (Maxwell, Lau, & Howard, 2015). Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided.

```{r, results="hide"}
library(pwr)
alpha <- 0.05
N <- 100
pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d
```

Second, researchers can determine the SESOI based on the literature. Ideally, researchers would specify the SESOI in their research, but this is not yet common practice. It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study. Simonsohn (2015) recently proposed to set the SESOI to 33% of the effect size an earlier study could detect based on the sample size. For example, consider a study where 100 participants answered a question, which was analyzed with an one-sample t-test. For a two-sided test with an alpha of .05 this test had 33% power to detect a d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`.

Another justifiable choice we would like to propose here is to use the smallest observed effect size that could have been statistically significant in the original study as the SESOI in a replication study. Based only on the alpha level and the sample size, we can calculate the criticial test value (e.g., *t*, *F*, *Z*). This critical test value can also be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*. All observed effect sizes smaller than the critical effect size would not be statistically significant in an original study, given the alpha and sample size. By setting the SESOI to the critical effect size, an equivalence test can reject all observed effect sizes that could have been detected in an earlier study.

```{r, results="hide"}
library(TOSTER)
powerTOSTone(alpha=0.05, statistical_power=0.9, low_eqbound_d=-0.33, high_eqbound_d=0.33)
```

Third, researchers can set the SESOI based on the resources they have available. The amount of data you can collect limits the inferences you can make. Given a an alpha level and a sample size, researchers can calculate the smallest effect size that can be rejected with the desired power in an equivalence test. For example, a researcher who plans to perform a two-sided one-sample t-test using an alpha of 5%, based on data from 100 observations, has 90% power declare equivalence when testing a SESOI of d = 0.33. Whether or not a test against a SESOI based on the available resources contributes to the scientific literature depends. One should not expect a test based on 11 observations, which provides 90% power to reject effects larger than d = 1, is particularly informative, given that most effects in psychology are substantially smaller than d = 1. But by transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes.

<!-- It would be nice to adapt the current equivalence testing power functions, such that you can not just calculate N, but also the SESOI -->

```{r}
p <- 0.05
N <- 100
crit_d <- abs(qt(p/2, (N*2)-2))/sqrt(N/2)
```

<!-- This is just an example of a comment, and a formula -->
$$t _ { L } = \frac { M _ { 1} - M _ { 2} - \Delta _ { L } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }, t _ { U } = \frac { M _ { 1} - M _ { 2} - \Delta _ { U } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }
$$

## Examples

#Equivalence test for Meta-analysis

Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where a trivial difference is specified as an effect size smaller then d = 0.1. The present a table with Cohen's d and se is reproduced below:

|Grades|d + se|
|---|---|
|Grade 2	|0.06 +/- 0.003|
|Grade 3	|0.04 +/- 0.002|
|Grade 4	|-0.01 +/- 0.002|
|Grade 5	|-0.01 +/- 0.002|
|Grade 6	|-0.01 +/- 0.002|
|Grade 7	|-0.02 +/- 0.002|
|Grade 8	|-0.02 +/- 0.002|
|Grade 9	|-0.01 +/- 0.003|
|Grade 10	|0.04 +/- 0.003|
|Grade 11	|0.06 +/- 0.003|

```{r, fig.width=6}
library("TOSTER")
TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.05)
```

We see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1. However, note that all of the effects are also statistically significant - so, the the effects are statistically different from zero, and practically equivalent.

*Hyde, J. S., Lindberg, S. M., Linn, M. C., Ellis, A. B., & Williams, C. C. (2008). Gender similarities characterize math performance. Science, 321(5888), 494-495.*

&nbsp;

&nbsp;

#Independent Groups Equivalence Test

&nbsp;

##Independent Groups Student's Equivalence Test

Eskine (2013) showed that participants who had been exposed to organic food were substantially harsher in their moral judgments relative to those exposed to control (d = 0.81, 95% CI: [0.19, 1.45]). A replication by Moery & Calin-Jageman (2016, Study 2) did not observe a significant effect (Control: n = 95, M = 5.25, SD = 0.95, Organic Food: n = 89, M = 5.22, SD = 0.83). Following Simonsohn's (2015) recommendation the equivalence bound was set to the effect size the original study had 33% power to detect (with n = 21 in each condition, this means the equivalence bound is d = 0.48, which equals a difference of 0.384 on a 7-point scale given the sample sizes and a pooled standard deviation of 0.894). Using a TOST equivalence test with alpha = 0.05, assuming equal variances, and equivalence bounds of d = -0.43 and d = 0.43 is significant, t(182) = -2.69, p = 0.004. We can reject effects larger than d = 0.43.

```{r, fig.width=6}
TOSTtwo.raw(m1=5.25,m2=5.22,sd1=0.95,sd2=0.83,n1=95,n2=89,low_eqbound=-0.384, high_eqbound=0.384, alpha = 0.05, var.equal=TRUE)
```

*Moery, E., & Calin-Jageman, R. J. (2016). Direct and Conceptual Replications of Eskine (2013): Organic Food Exposure Has Little to No Effect on Moral Judgments and Prosocial Behavior. Social Psychological and Personality Science, 7(4), 312-319. https://doi.org/10.1177/1948550616639649 *

&nbsp;

##Independent Groups Welch's Equivalence Test

Deary, Thorpe, Wilson, Starr, and Whally (2003) report the IQ scores of 79,376 children in Scotland (39,343 girls and 40,033 boys). The IQ score for girls (M = 100.64, SD = 14.1) girls and boys (100.48, SD = 14.9) was non-significant. With such a huge sample, we can examine whether the data is equivalent using very small equivalence bounds, such as -0.05 and 0.05. Because sample sizes are unequal, and variances are unequal, Welch's t-test is used, and we can conclude IQ scores are indeed equivalent, based on equivalence bounds of -0.05 and 0.05.

```{r, fig.width=6}
TOSTtwo(m1=100.64,m2=100.48,sd1=14.1,sd2=14.9,n1=39343,n2=40033,low_eqbound_d=-0.05, high_eqbound_d=0.05, alpha = 0.05, var.equal=FALSE)
```

*Deary, I. J., Thorpe, G., Wilson, V., Starr, J. M., & Whalley, L. J. (2003). Population sex differences in IQ at age 11: The Scottish mental survey 1932. Intelligence, 31(6), 533-542.*

&nbsp;

&nbsp;

#One-Sample Equivalence Test

&nbsp;

Lakens, Semin, & Foroni (2012) examined whether the color of Chinese ideographs (white vs. black) would bias whether participants judged that the ideograph correctly translated positive, neutral, or negative words. In Study 4 the prediction was that participants would judge negative words to be correctly translated by black ideographs above guessing average, but no effects were predicted for translations in the other 5 between subject conditions in the 2 (ideograph color: white vs. black) x 3 (word meaning: positive, neutral, negative) design. The table below is a summary of the data in all 6 conditions:

|Color	|Valence	|M	|%	|SD	|t	|df	|p	|d|
|---|---|---|---|---|---|---|---|---|
|White	|Positive	|6.05	|50.42	|1.50	|0.15	|20	|.89	|.03|
|White	|Negative	|6.68	|55.67	|1.70	|1.75	|18	|.10	|.40|
|White	|Neutral	|5.95	|49.58	|1.40	|0.87	|19	|.87	|.04|
|Black	|Positive	|6.45	|53.75	|1.91	|1.06	|19	|.30	|.24|
|Black	|Negative	|6.95	|57.92	|1.15	|3.71	|19	|.00	|.80|
|Black	|Neutral	|5.71	|47.58	|1.79	|0.73	|20	|.47	|.16|

With 19 to 21 participants in each between subject condition, the study had 80% power to detect equivalence with equivalence bounds of -0.68 to d = 0.68.

```{r}
powerTOSTone(alpha=0.05, statistical_power=0.8, low_eqbound_d=-0.68, high_eqbound_d=0.68)
```

When we perform 5 equivalence tests using equivalence bounds of -0.68 and 0.68, using a Holm-Bonferroni sequential procedure to control error rates, we can conclude statistical equivalence for the data in row 1, 3, and 6, but the conclusion is undetermined for tests 2 and 4, where the data is neither significantly different from guessing average, nor statistically equivalent. For example, performing the test for the data in row 6:

```{r, fig.width=6}
TOSTone(m=5.71,mu=6,sd=1.79,n=20,low_eqbound_d=-0.68, high_eqbound_d=0.68, alpha=0.05)
```

It is clear I should have been more tentative in my conclusions in this study. Not only can I not conclude equivalence in some of the conditions, the equivalence bound I had 80% power to detect is very large, meaning the possibility that there are theoretically interesting but smaller effects remains.

*Lakens, D., Semin, G. R., & Foroni, F. (2012). But for the bad, there would not be good: Grounding valence in brightness through shared relational structures. Journal of Experimental Psychology: General, 141(3), 584-594. https://doi.org/10.1037/a0026468 *

&nbsp;

&nbsp;

#Equivalence test for correlations

&nbsp;

Olson, Fazio, and Hermann (2007) reported correlations between implicit and explicit measures of self-esteem, such as the IAT, Rosenberg's self-esteem scale, a feeling thermometer, and trait ratings. In Study 1 71 participants completed the self-esteem measures. Because no equivalence bounds are mentioned, we can see which equivalence bounds the researchers would have 80% power to detect.

```{r, fig.width=6}
powerTOSTr(alpha=0.05, statistical_power=0.8, low_eqbound_r=-0.24, high_eqbound_r=0.24)
```
With 71 pairs of observations between measures, the researchers have 80% power for equivalence bounds of r = -0.24 and r = 0.24.

The correlations observed by Olson et al (2007), Study 1, are presented in the table below (significant correlations are flagged by an asteriks).

|Measure |IAT |Rosenberg|Feeling thermometer| Trait ratings|
|---|---|---|---|---|
|IAT |- |-.12 |-.09 |-.06|
|Rosenberg| | - |.62*| .09|
|Feeling thermometer| | | - |.29*|
|Trait ratings| | | |-|

We can test each correlation, for example the correlation between the IAT and the Rosenberg self-esteem scale of -0.12, given 71 participants, and using equivalence bounds of r_U = 0.24 and r_L = -0.24.

```{r, fig.width=6}
TOSTr(n=71, r=-0.06, low_eqbound_r=-0.24, high_eqbound_r=0.24, alpha=0.05)
```
We see that none of the correlations can be declared to be statistically equivalent. Instead, the tests yield undetermined outcomes: the correlations are not significantly different from 0, nor statistically equivalent.

*Olson, M. A., Fazio, R. H., & Hermann, A. D. (2007). Reporting tendencies underlie discrepancies between implicit and explicit measures of self-esteem. Psychological Science, 18(4), 287-291.*

# Discussion

\newpage

# References

Button, K. S., Kounali, D., Thomas, L., Wiles, N. J., Peters, T. J., Welton, N. J., . Lewis, G. (2015). Minimal clinically important difference on the Beck Depression Inventory - II according to the patient's perspective. Psychological Medicine, 45(15), 3269-3279. https://doi.org/10.1017/S0033291715001270

Burriss, R. P., Troscianko, J., Lovell, P. G., Fulford, A. J. C., Stevens, M., Quigley, R., . Rowland, H. M. (2015). Changes in Women's Facial Skin Color over the Ovulatory Cycle are Not Detectable by the Human Visual System. PLOS ONE, 10(7), e0130093. https://doi.org/10.1371/journal.pone.0130093

```{r create_r-references}
r_refs(file = "r-references.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

### 2017-10-02 manuscript.Rmd
2-10-2017

@@
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["equivalence.bib"]
 
@@
 ```
+
+Psychologists should be able to falsify predictions.
+A common prediction in psychological research is that an effect exists in the population that differs from zero.
+For example, we might predict that priming American Asian women with their Asian identify will increase their performence on a math test compared to women in the control condition who are primed with their female identity.
+To design a study that allows for strong inferences (@platt_strong_1964) it is therefore important to specify which test result would *disprove* the hypothesis.
+
+An equivalence test can be used to test whether the observed effect is surprisingly close to zero, assuming a meaningful effect exists in the population.
+The test is a simple variation on the widely used null hypothesis significance tests, with the small difference that instead of testing the observed effect against zero, it is tested against a smallest effect size of interest (SESOI).
+For example, if an improvement on a math test of 5% or more is considered meaningful, one could test against a SESOI of 5%.
+If the observed effect is suprisingly closer to 0, assuming a true effect of 5% existed, one can declare statistical equivalence, or reject effects that are large enough to be considered meaningful.
+
+<!-- If we use Dienes & McLatchie's 5 examples, it would make sense to add a Figure like a forest plot, where we present the effect size estimates (raw or standardized) indicated by a square, lines that represent the 90% CI, and another symbol or | to indicate the SESOI, sor all examples.
+We can then also use the same Figure in the intro to explain the four possible outcomes of equivalence testing).
+Although we should do all tests in R, we also should try to create Jamovi files that contain the data and analyses - another reason to follow Zoltan in using the Social Psychology special issue studies, because the data should be shared, and we can import the raw data into Jamovi files, save these (with the analysis) and share them.
+-->
 
@@
 This hypothesis requires that the increase in redness can be detected with the naked eye by men.
-Penke et al (2017) collected data from X women, and showed that there was indeed an increase in redness of the facial skin of woman during their fertile period.
+@burriss_changes_2015 collected data from 22 women and showed that there was indeed an increase in redness of the facial skin of woman during their fertile period.
 However, this increase was not large enough to be noticeable with the naken eye by men, thus falsifying their hypothesis.
@@
 
-Another example of an objectively determined SESOI can be found in Button and colleagues (2015) where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking X patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
+Another example of an objectively determined SESOI can be found in @button_minimal_2015 where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
 

### 2017-10-03 twostproptwo.R
added resources

@@ -1,0 +1,74 @@
+#' TOST function for two propotions (raw scores)
+#' @param proprop1 proportion of group 1
+#' @param proprop2 proportion of group 2
+#' @param n1 sample size in group 1
+#' @param n2 sample size in group 2
+#' @param margin equivalence bounds (e.g., 0.1) expressed in raw scale units (e.g., scalepoints)
+#' @param alpha alpha level (default = 0.05)
+#' @return Returns TOST z-value 1, TOST p-value 1, TOST z-value 2, TOST p-value 2, low equivalence bound, high equivalence bound, Lower limit confidence interval TOST, Upper limit confidence interval TOST
+#' @importFrom stats pnorm pt qnorm qt
+#' @importFrom graphics abline plot points segments title
+#' @examples
+#' ## Equivalence test for two independent proportions equal to .65 and .70, with 100 samples per group, an equivalence margin of .1 
+#' ## and alpha = 0.05.
+#' TOSTtwo.prop(prop1 = .65, prop2 = .70, n1 = 100, n2 = 100, margin = .1, alpha = .05)
+#' @section References:
+#' Tunes da Silva, G., Logan, B. R., & Klein, J. P. (2008). Methods for Equivalence and Noninferiority Testing. Biology of Blood Marrow Transplant, 15(1 Suppl), 120-127.
+#' Yin, G. (2012). Clinical Trial Design: Bayesian and Frequentist Adaptive Methods. Hoboken, New Jersey: John Wiley & Sons, Inc.
+#' @export
+
+TOSTtwo.prop <- function(prop1, prop2, n1, n2, margin, alpha) {
+  if(missing(alpha)) {
+    alpha <- 0.05
+  }
+  if(margin < 0) {
+    margin <- abs(margin)
+  }
+  
+  prop_dif <- prop1 - prop2
+  prop_se <- sqrt((prop1*(1-prop1))/n1 + (prop2*(1-prop2))/n2)
+  
+  ##defining z-statistic
+  z_lower_numerator <- prop_dif + margin
+  z_upper_numerator <- prop_dif - margin
+  z_lower <- z_lower_numerator/prop_se
+  z_upper <- z_upper_numerator/prop_se
+  
[…]

### 2017-10-03 manuscript.Rmd
test
auto zotero library update using fetch_zotero_refs function
3-10-2017

@@
 library("papaja")
+
+# papaja:::fetch_zotero_refs(
+#   x = "1786197"
+#   , bib_name = "equivalence.bib"
+#   , API_key = "F7JGNtCxZGT6q87Eeg2zNWiP"
+#   , lib_type = "group"
+# )
 ```
@@
 For example, we might predict that priming American Asian women with their Asian identify will increase their performence on a math test compared to women in the control condition who are primed with their female identity.
-To design a study that allows for strong inferences (@platt_strong_1964) it is therefore important to specify which test result would *disprove* the hypothesis.
+To be able to falsify this hypothesies, and design a study that allows for strong inferences @platt_strong_1964, it is important to specify which test result would *disprove* the hypothesis.
 
@@
 The test is a simple variation on the widely used null hypothesis significance tests, with the small difference that instead of testing the observed effect against zero, it is tested against a smallest effect size of interest (SESOI).
-For example, if an improvement on a math test of 5% or more is considered meaningful, one could test against a SESOI of 5%.
-If the observed effect is suprisingly closer to 0, assuming a true effect of 5% existed, one can declare statistical equivalence, or reject effects that are large enough to be considered meaningful.
+For example, if a difference on a math test of 5% or more is considered meaningful, one could test against a SESOI of 5%.
+If the observed effect is suprisingly closer to 0, assuming a true effect of 5% or larger existed, one can declare statistical equivalence, or reject the presence of effects that are large enough to be considered meaningful.
+An equivalence test can be used to falsify hypotheses.
+
+One can conclude statistical equivalence whenever a 90% confidence interval around the effect size estimate does not contain the SESOI, or when two one-sided tests (i.e., examining whether the difference on the math test is more than -5%, but less than 5%) are statistically significant.
+By combining null-hypothesis tests and equivalence tests, four combinations of results can be observed (see Figure 1).
+Two outcomes are straightforward to interpret, namely when the effect is equivalent to zero, and not different from zero (an effect that is not meaningful), and when the effect is not equivalent, and different from zero (a meaningful effect).
+In addition, the effect can be different from zero and too small to be considered meaningful (an effect which is not meaningful) or the effect can neither differ from zero, not be equivalent to zero (an undetermined result, where more data is needed).
+
+Even though equivalence tests are just a small variation of traditional Frequentist tests (testing against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations (@lakens_equivalence_2017).
+In this article, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.
 
@@
 Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI).
-The SESOI is sometimes based on just noticeable differences, which can be objectively determined.
+The SESOI can sometimes based on just noticeable differences, which can be objectively determined.
 Most often, however, it is a subjective decision that varies across individuals and time.
 Both these approaches are valid, and even when a just noticeable difference can be objectively determined, researchers might choose to set the SESOI to a larger value.
-In many research areas, the SESOI is based on a cost-benefit analysis.
+In many research areas, the SESOI is best based on a cost-benefit analysis.
 Since both costs and benefits are necessarily relative, the SESOI will depend on the researcher who designs the study.
 The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base.
-The goal is not to determine a SESOI that is universally valid: The goal is to set a SESOI such that inferences based on it answer a meaninful question.
+Researchers should not aim to determine a SESOI that is universally valid.
+The goal is to set a SESOI such that inferences based on it answer a meaningful question.
 

### 2017-10-03 manuscript.Rmd
correct references

@@
 For example, we might predict that priming American Asian women with their Asian identify will increase their performence on a math test compared to women in the control condition who are primed with their female identity.
-To be able to falsify this hypothesies, and design a study that allows for strong inferences @platt_strong_1964, it is important to specify which test result would *disprove* the hypothesis.
+To be able to falsify this hypothesies, and design a study that allows for strong inferences [@platt_strong_1964], it is important to specify which test result would *disprove* the hypothesis.
 
@@
 
-Even though equivalence tests are just a small variation of traditional Frequentist tests (testing against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations (@lakens_equivalence_2017).
+Even though equivalence tests are just a small variation of traditional Frequentist tests (testing against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017].
 In this article, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.
@@
 In such circumstances, the SESOI can be set based on just noticeable differences.
-For example, Burriss and colleagues (2015) examined whether women displayed an increase in redness in the face during the fertile phase of their ovulatory cycle.
+For example, Burriss and colleagues [-@burriss_changes_2015] examined whether women displayed an increase in redness in the face during the fertile phase of their ovulatory cycle.
 The hypothesis was that a slighly redder skin signals greater attractiveness and physical health, and sending this signal to men yields an evolutionary advantage.
 This hypothesis requires that the increase in redness can be detected with the naked eye by men.
-@burriss_changes_2015 collected data from 22 women and showed that there was indeed an increase in redness of the facial skin of woman during their fertile period.
+They collected data from 22 women and showed that there was indeed an increase in redness of the facial skin of woman during their fertile period.
 However, this increase was not large enough to be noticeable with the naken eye by men, thus falsifying their hypothesis.

### 2017-10-04 Dienes_example1.R
Added 1st draft of equivalence tests for Dienes example 1

@@ -1,0 +1,33 @@
+# TOST on the values from Gibson, Losee, and Vitello example
+
+Gibson.m1 <- 0.59
+Gibson.m2 <- 0.53
+Gibson.sd1 <- 0.23
+Gibson.sd2 <- 0.23
+Gibson.n1 <- 52
+Gibson.n2 <- 54
+
+Shih.m1 <- 0.54
+Shih.n2 <- 0.43
+Shih.sd1 <- 0.17
+Shih.sd2 <- 0.16
+Shih.n1 <- 16
+Shih.n2 <- 14
+
+### Calculate subjective bound critera from original study data ###
+
+# Original had 33% Power to detect effect of d=0.57 (GPower result)
+
+# Original study smallest detectable effect size
+# Critical t for original study was:
+crit_t = qt(1-.05/2, 29)
+# So d becomes:
+crit_d = crit_t*sqrt(1/16 + 1/14)
+
+### TOST for the different bounds ###
+
+# TOST for 33% Power rule:
+TOSTtwo(m1 = Gibson.m1, m2 = Gibson.m2, sd1 = Gibson.sd1, sd2 = Gibson.sd2, n1 = Gibson.n1, n2 = Gibson.n2, low_eqbound_d = -0.57, high_eqbound_d = 0.57, alpha = .05)
+
+# TOST for smallest detectable effect in original study
+TOSTtwo(m1 = 0.59, m2 = 0.53, sd1 = 0.23, sd2 = 0.23, n1 = 52, n2 = 54, low_eqbound_d = -crit_d, high_eqbound_d = crit_d, alpha = .05)

### 2017-10-04 Dienes_example1.R
Finished first draft of Dienes 2017 example 1

@@ -1,11 +1,13 @@
 # TOST on the values from Gibson, Losee, and Vitello example
 
-Gibson.m1 <- 0.59
-Gibson.m2 <- 0.53
-Gibson.sd1 <- 0.23
+#### Data ####
+
+Gibson.m1 <- 0.63
+Gibson.m2 <- 0.51
+Gibson.sd1 <- 0.22
 Gibson.sd2 <- 0.23
-Gibson.n1 <- 52
-Gibson.n2 <- 54
+Gibson.n1 <- 40
+Gibson.n2 <- 43
 
 Shih.m1 <- 0.54
 Shih.n2 <- 0.43
@@ -13,21 +15,43 @@ Shih.sd1 <- 0.17
 Shih.sd2 <- 0.16
 Shih.n1 <- 16
 Shih.n2 <- 14
+Shih.t <- 2.02
 
-### Calculate subjective bound critera from original study data ###
+#### Calculate subjective bound critera from original study data ####
 
-# Original had 33% Power to detect effect of d=0.57 (GPower result)
+# Original had 33% Power to detect effect of d=0.57 (G*Power result)
+t33 <- 0.57
 
 # Original study smallest detectable effect size
-# Critical t for original study was:
-crit_t = qt(1-.05/2, 29)
-# So d becomes:
-crit_d = crit_t*sqrt(1/16 + 1/14)
 
-### TOST for the different bounds ###
+crit_t = qt(1-.05/2, 29) # Critical t for original study
+crit_d = crit_t*sqrt(1/16 + 1/14) # Convert critical t to d
+
+# Original study d
+Shih.d <- Shih.t*sqrt(1/16 + 1/14)
+
+#### TOST for the different bounds ####
 
 # TOST for 33% Power rule:
-TOSTtwo(m1 = Gibson.m1, m2 = Gibson.m2, sd1 = Gibson.sd1, sd2 = Gibson.sd2, n1 = Gibson.n1, n2 = Gibson.n2, low_eqbound_d = -0.57, high_eqbound_d = 0.57, alpha = .05)
+TOSTtwo(m1 = Gibson.m1, 
+        m2 = Gibson.m2, 
+        sd1 = Gibson.sd1, 
+        sd2 = Gibson.sd2, 
+        n1 = Gibson.n1, 
+        n2 = Gibson.n2, 
+        low_eqbound_d = -t33, 
+        high_eqbound_d = t33, alpha = .05)
 
[…]

### 2017-10-04 manuscript.tex

\documentclass[english,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

  \usepackage{ifthen} % Only add declarations when endfloat package is loaded
  \ifthenelse{\equal{\string man}{\string man}}{%
   \DeclareDelayedFloatFlavor{ThreePartTable}{table} % Make endfloat play with longtable
   % \DeclareDelayedFloatFlavor{ltable}{table} % Make endfloat play with lscape
   \DeclareDelayedFloatFlavor{lltable}{table} % Make endfloat play with lscape & longtable
  }{}%

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Equivalence Testing for Psychological Science},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Equivalence Testing for Psychological Science}

  \shorttitle{Equivalence Testing}

  \author{Daniel Lakens\textsuperscript{1}, Anne M. Scheel\textsuperscript{1}, \& Peder Isager\textsuperscript{1}}

  \def\affdep{{"", "", ""}}%
  \def\affcity{{"", "", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Eindhoven University of Technology  }

  \authornote{
    \newcounter{author}
    We would like to thank Courtney Soderbergh for creating the first
    version of the TOST function to test two independent proportions.

                      Correspondence concerning this article should be addressed to Daniel Lakens, Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands. E-mail: \href{mailto:D.Lakens@tue.nl}{\nolinkurl{D.Lakens@tue.nl}}
                                    }

  \abstract{Psychologists need to be able to test for the absence of an effect.
Using the Two-One-Sided Tests (TOST) procedure, researchers can easily
test whether the observed effects are too small to be meaningful. By
specifying a smallest effect size of interest (SESOI) researchers test
whether observed effects are surprisingly closer to zero, assuming there
was an effect the consider meaningful. We explain a range of approaches
to determine the SESOI in psychological science, and provide detailed
examples of how equivalence tests should be performed and reported.}
  \keywords{Equivalence Testing, NHST, power, TOST \\

    \indent Word count: X
  }

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

Psychologists should be able to falsify predictions. A common prediction
in psychological research is that an effect exists in the population
that differs from zero. For example, we might predict that priming
American Asian women with their Asian identify will increase their
performence on a math test compared to women in the control condition
who are primed with their female identity. To be able to falsify this
hypothesies, and design a study that allows for strong inferences
({\textbf{???}}), it is important to specify which test result would
\emph{disprove} the hypothesis.

An equivalence test can be used to test whether the observed effect is
surprisingly close to zero, assuming a meaningful effect exists in the
population. The test is a simple variation on the widely used null
hypothesis significance tests, with the small difference that instead of
testing the observed effect against zero, it is tested against a
smallest effect size of interest (SESOI). For example, if a difference
on a math test of 5\% or more is considered meaningful, one could test
against a SESOI of 5\%. If the observed effect is suprisingly closer to
0, assuming a true effect of 5\% or larger existed, one can declare
statistical equivalence, or reject the presence of effects that are
large enough to be considered meaningful. An equivalence test can be
used to falsify hypotheses.

One can conclude statistical equivalence whenever a 90\% confidence
interval around the effect size estimate does not contain the SESOI, or
when two one-sided tests (i.e., examining whether the difference on the
math test is more than -5\%, but less than 5\%) are statistically
significant. By combining null-hypothesis tests and equivalence tests,
four combinations of results can be observed (see Figure 1). Two
outcomes are straightforward to interpret, namely when the effect is
equivalent to zero, and not different from zero (an effect that is not
meaningful), and when the effect is not equivalent, and different from
zero (a meaningful effect). In addition, the effect can be different
from zero and too small to be considered meaningful (an effect which is
not meaningful) or the effect can neither differ from zero, not be
equivalent to zero (an undetermined result, where more data is needed).

Even though equivalence tests are just a small variation of traditional
Frequentist tests (testing against the SESOI, instead of 0), their use
was limited until the availability of user-friendly software to perform
the calculations (Lakens, 2017). In this article, we discuss different
approaches to determining the SESOI for psychological research, and
provide detailed reproducible examples of how to perform power analyses
when designing equivalence tests, and statistical re-analyses of
published psychology experiments.

\section{Justifying the Smallest Effect Size of
Interest}\label{justifying-the-smallest-effect-size-of-interest}

Equivalence tests are performed against a value that is considered the
smallest effect size of interest (SESOI). The SESOI can sometimes based
on just noticeable differences, which can be objectively determined.
Most often, however, it is a subjective decision that varies across
individuals and time. Both these approaches are valid, and even when a
just noticeable difference can be objectively determined, researchers
might choose to set the SESOI to a larger value. In many research areas,
the SESOI is best based on a cost-benefit analysis. Since both costs and
benefits are necessarily relative, the SESOI will depend on the
researcher who designs the study. The goal of setting a SESOI is to
clearly justify why designing a study that has a high probability of
rejecting effects larger than a specified value contributes to our
knowledge base. Researchers should not aim to determine a SESOI that is
universally valid. The goal is to set a SESOI such that inferences based
on it answer a meaningful question.

\subsection{Objective Justifications of a
SESOI}\label{objective-justifications-of-a-sesoi}

An objectively determined SESOI should be based on quantifiable
theoretical predictions, such as computational models. Sometimes, the
only theoretical prediction is that an effect should be noticeable. In
such circumstances, the SESOI can be set based on just noticeable
differences. For example, Burriss and colleagues ({\textbf{???}})
examined whether women displayed an increase in redness in the face
during the fertile phase of their ovulatory cycle. The hypothesis was
that a slighly redder skin signals greater attractiveness and physical
health, and sending this signal to men yields an evolutionary advantage.
This hypothesis requires that the increase in redness can be detected
with the naked eye by men. They collected data from 22 women and showed
that there was indeed an increase in redness of the facial skin of woman
during their fertile period. However, this increase was not large enough
to be noticeable with the naken eye by men, thus falsifying their
hypothesis. Because the just noticeable difference in redness of the
skin can be measured, it is possible to objectively establish the SESOI.

Another example of an objectively determined SESOI can be found in
({\textbf{???}}) where the minimal clinically important difference on
the Beck Depression Inventory - II was determined by asking 1039
patients when they subjectively felt less depressed (i.e., when they
personally noticed an improvement) and relating this to the
corresponding difference score on the depression inventory.

\subsection{Subjective justifications of a
SESOI}\label{subjective-justifications-of-a-sesoi}

We distinguish between three categories of subjective justifications for
SESOI. First, researchers can use benchmarks. For example, one might set
the SESOI to a standardized effect size of d = 0.5, which would allow
one to reject effect as large or larger than a \enquote{medium} effect
size (Cohen, 1988). Similarly, effect sizes smaller than a Cohen's d of
0.1 are sometimes considered trivially small (Maxwell, Lau, \& Howard,
2015). Relying on a benchmark is the weakest possible justification of a
SESOI, and should be avoided.

Second, researchers can determine the SESOI based on the literature.
Ideally, researchers would specify the SESOI in their research, but this
is not yet common practice. It is thus up to researchers who build on
earlier work to decide which effect size is too small to be meaningful,
given an earlier study. Simonsohn (2015) recently proposed to set the
SESOI to 33\% of the effect size an earlier study could detect based on
the sample size. For example, consider a study where 100 participants
answered a question, which was analyzed with an one-sample t-test. For a
two-sided test with an alpha of .05 this test had 33\% power to detect a
d = 0.15.

Another justifiable choice we would like to propose here is to use the
smallest observed effect size that could have been statistically
significant in the original study as the SESOI in a replication study.
Based only on the alpha level and the sample size, we can calculate the
criticial test value (e.g., \emph{t}, \emph{F}, \emph{Z}). This critical
test value can also be transformed to a standardized effect size (e.g.,
\(d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }\)),
which can thus be interpreted as a \emph{critical effect size}. All
observed effect sizes smaller than the critical effect size would not be
statistically significant in an original study, given the alpha and
sample size. By setting the SESOI to the critical effect size, an
equivalence test can reject all observed effect sizes that could have
been detected in an earlier study.

Third, researchers can set the SESOI based on the resources they have
available. The amount of data you can collect limits the inferences you
can make. Given a an alpha level and a sample size, researchers can
calculate the smallest effect size that can be rejected with the desired
power in an equivalence test. For example, a researcher who plans to
perform a two-sided one-sample t-test using an alpha of 5\%, based on
data from 100 observations, has 90\% power declare equivalence when
testing a SESOI of d = 0.33. Whether or not a test against a SESOI based
on the available resources contributes to the scientific literature
depends. One should not expect a test based on 11 observations, which
provides 90\% power to reject effects larger than d = 1, is particularly
informative, given that most effects in psychology are substantially
smaller than d = 1. But by transparently reporting the effects one can
detect and reject, based on the study design, researchers can
communicate the information their study contributes.

\[t _ { L } = \frac { M _ { 1} - M _ { 2} - \Delta _ { L } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }, t _ { U } = \frac { M _ { 1} - M _ { 2} - \Delta _ { U } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }
\]

\subsection{Examples}\label{examples}

\section{Equivalence test for
Meta-analysis}\label{equivalence-test-for-meta-analysis}

Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect
sizes for gender differences in mathematics tests across the 7 million
students in the US represent trivial differences, where a trivial
difference is specified as an effect size smaller then d = 0.1. The
present a table with Cohen's d and se is reproduced below:

\begin{longtable}[]{@{}ll@{}}
\toprule
Grades & d + se\tabularnewline
\midrule
\endhead
Grade 2 & 0.06 +/- 0.003\tabularnewline
Grade 3 & 0.04 +/- 0.002\tabularnewline
Grade 4 & -0.01 +/- 0.002\tabularnewline
Grade 5 & -0.01 +/- 0.002\tabularnewline
Grade 6 & -0.01 +/- 0.002\tabularnewline
Grade 7 & -0.02 +/- 0.002\tabularnewline
Grade 8 & -0.02 +/- 0.002\tabularnewline
Grade 9 & -0.01 +/- 0.003\tabularnewline
Grade 10 & 0.04 +/- 0.003\tabularnewline
Grade 11 & 0.06 +/- 0.003\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{}
\end{figure}

\begin{verbatim}
## Using alpha = 0.05 the meta-analysis was significant, Z = 20, p = 2.753624e-89
## Using alpha = 0.05 the equivalence test was significant, Z = -13.33333, p = 7.406413e-41TOST results:
##   Z-value 1 p-value 1 Z-value 2    p-value 2
## 1  53.33333         0 -13.33333 7.406413e-41
## 
## Equivalence bounds (Cohen's d):
##   low bound d high bound d
## 1        -0.1          0.1
## 
## TOST confidence interval:
##   Lower Limit 90% CI Upper Limit 90% CI
## 1         0.05506544         0.06493456
\end{verbatim}

We see that indeed, all effect size estimates are measured with such
high precision, we can conclude that they fall within the equivalence
bound of d = -0.1 and d = 0.1. However, note that all of the effects are
also statistically significant - so, the the effects are statistically
different from zero, and practically equivalent.

\emph{Hyde, J. S., Lindberg, S. M., Linn, M. C., Ellis, A. B., \&
Williams, C. C. (2008). Gender similarities characterize math
performance. Science, 321(5888), 494-495.}

~

~

\section{Independent Groups Equivalence
Test}\label{independent-groups-equivalence-test}

~

\subsection{Independent Groups Student's Equivalence
Test}\label{independent-groups-students-equivalence-test}

Eskine (2013) showed that participants who had been exposed to organic
food were substantially harsher in their moral judgments relative to
those exposed to control (d = 0.81, 95\% CI: {[}0.19, 1.45{]}). A
replication by Moery \& Calin-Jageman (2016, Study 2) did not observe a
significant effect (Control: n = 95, M = 5.25, SD = 0.95, Organic Food:
n = 89, M = 5.22, SD = 0.83). Following Simonsohn's (2015)
recommendation the equivalence bound was set to the effect size the
original study had 33\% power to detect (with n = 21 in each condition,
this means the equivalence bound is d = 0.48, which equals a difference
of 0.384 on a 7-point scale given the sample sizes and a pooled standard
deviation of 0.894). Using a TOST equivalence test with alpha = 0.05,
assuming equal variances, and equivalence bounds of d = -0.43 and d =
0.43 is significant, t(182) = -2.69, p = 0.004. We can reject effects
larger than d = 0.43.

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{}
\end{figure}

\begin{verbatim}
## Using alpha = 0.05 Student's t-test was non-significant, t(182) = 0.2274761, p = 0.8203089
## Using alpha = 0.05 the equivalence test based on Student's t-test was significant, t(182) = -2.684218, p = 0.003970479TOST results:
##   t-value 1    p-value 1 t-value 2   p-value 2  df
## 1   3.13917 0.0009885653 -2.684218 0.003970479 182
## 
## Equivalence bounds (raw scores):
##   low bound raw high bound raw
## 1        -0.384          0.384
## 
## TOST confidence interval:
##   Lower Limit 90% CI raw Upper Limit 90% CI raw
## 1             -0.1880364              0.2480364
\end{verbatim}

\emph{Moery, E., \& Calin-Jageman, R. J. (2016). Direct and Conceptual
Replications of Eskine (2013): Organic Food Exposure Has Little to No
Effect on Moral Judgments and Prosocial Behavior. Social Psychological
and Personality Science, 7(4), 312-319.
\url{https://doi.org/10.1177/1948550616639649} }

~

\subsection{Independent Groups Welch's Equivalence
Test}\label{independent-groups-welchs-equivalence-test}

Deary, Thorpe, Wilson, Starr, and Whally (2003) report the IQ scores of
79,376 children in Scotland (39,343 girls and 40,033 boys). The IQ score
for girls (M = 100.64, SD = 14.1) girls and boys (100.48, SD = 14.9) was
non-significant. With such a huge sample, we can examine whether the
data is equivalent using very small equivalence bounds, such as -0.05
and 0.05. Because sample sizes are unequal, and variances are unequal,
Welch's t-test is used, and we can conclude IQ scores are indeed
equivalent, based on equivalence bounds of -0.05 and 0.05.

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{}
\end{figure}

\begin{verbatim}
## Using alpha = 0.05 Welch's t-test was non-significant, t(79260.94) = 1.554136, p = 0.1201559
## Using alpha = 0.05 the equivalence test based on Welch's t-test  was significant, t(79260.94) = -5.490723, p = 2.007585e-08TOST results:
##   t-value 1    p-value 1 t-value 2    p-value 2       df
## 1  8.598995 4.092652e-18 -5.490723 2.007585e-08 79260.94
## 
## Equivalence bounds (Cohen's d):
##   low bound d high bound d
## 1       -0.05         0.05
## 
## Equivalence bounds (raw scores):
##   low bound raw high bound raw
## 1    -0.7252758      0.7252758
## 
## TOST confidence interval:
##   Lower Limit 90% CI raw Upper Limit 90% CI raw
## 1           -0.009341433              0.3293414
\end{verbatim}

\emph{Deary, I. J., Thorpe, G., Wilson, V., Starr, J. M., \& Whalley, L.
J. (2003). Population sex differences in IQ at age 11: The Scottish
mental survey 1932. Intelligence, 31(6), 533-542.}

~

~

\section{One-Sample Equivalence Test}\label{one-sample-equivalence-test}

~

Lakens, Semin, \& Foroni (2012) examined whether the color of Chinese
ideographs (white vs.~black) would bias whether participants judged that
the ideograph correctly translated positive, neutral, or negative words.
In Study 4 the prediction was that participants would judge negative
words to be correctly translated by black ideographs above guessing
average, but no effects were predicted for translations in the other 5
between subject conditions in the 2 (ideograph color: white vs.~black) x
3 (word meaning: positive, neutral, negative) design. The table below is
a summary of the data in all 6 conditions:

\begin{longtable}[]{@{}lllllllll@{}}
\toprule
Color & Valence & M & \% & SD & t & df & p & d\tabularnewline
\midrule
\endhead
White & Positive & 6.05 & 50.42 & 1.50 & 0.15 & 20 & .89 &
.03\tabularnewline
White & Negative & 6.68 & 55.67 & 1.70 & 1.75 & 18 & .10 &
.40\tabularnewline
White & Neutral & 5.95 & 49.58 & 1.40 & 0.87 & 19 & .87 &
.04\tabularnewline
Black & Positive & 6.45 & 53.75 & 1.91 & 1.06 & 19 & .30 &
.24\tabularnewline
Black & Negative & 6.95 & 57.92 & 1.15 & 3.71 & 19 & .00 &
.80\tabularnewline
Black & Neutral & 5.71 & 47.58 & 1.79 & 0.73 & 20 & .47 &
.16\tabularnewline
\bottomrule
\end{longtable}

With 19 to 21 participants in each between subject condition, the study
had 80\% power to detect equivalence with equivalence bounds of -0.68 to
d = 0.68.

\begin{verbatim}
## The required sample size to achieve 80 % power with equivalence bounds of -0.68 and 0.68 is 19
\end{verbatim}

\begin{verbatim}
## [1] 18.52043
\end{verbatim}

When we perform 5 equivalence tests using equivalence bounds of -0.68
and 0.68, using a Holm-Bonferroni sequential procedure to control error
rates, we can conclude statistical equivalence for the data in row 1, 3,
and 6, but the conclusion is undetermined for tests 2 and 4, where the
data is neither significantly different from guessing average, nor
statistically equivalent. For example, performing the test for the data
in row 6:

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{}
\end{figure}

\begin{verbatim}
## Using alpha = 0.05 the NHST one-sample t-test was non-significant, t(19) = -0.724536, p = 0.4775649
## Using alpha = 0.05 the equivalence test was significant, t(19) = 2.316516, p = 0.01592727TOST results:
##   t-value 1  p-value 1 t-value 2    p-value 2 df
## 1  2.316516 0.01592727 -3.765588 0.0006543016 19
## 
## Equivalence bounds (Cohen's d):
##   low bound d high bound d
## 1       -0.68         0.68
## 
## Equivalence bounds (raw scores):
##   low bound raw high bound raw
## 1       -1.2172         1.2172
## 
## TOST confidence interval:
##   Lower Limit 90% CI raw Upper Limit 90% CI raw
## 1             -0.9820961              0.4020961
\end{verbatim}

It is clear I should have been more tentative in my conclusions in this
study. Not only can I not conclude equivalence in some of the
conditions, the equivalence bound I had 80\% power to detect is very
large, meaning the possibility that there are theoretically interesting
but smaller effects remains.

\emph{Lakens, D., Semin, G. R., \& Foroni, F. (2012). But for the bad,
there would not be good: Grounding valence in brightness through shared
relational structures. Journal of Experimental Psychology: General,
141(3), 584-594. \url{https://doi.org/10.1037/a0026468} }

~

~

\section{Equivalence test for
correlations}\label{equivalence-test-for-correlations}

~

Olson, Fazio, and Hermann (2007) reported correlations between implicit
and explicit measures of self-esteem, such as the IAT, Rosenberg's
self-esteem scale, a feeling thermometer, and trait ratings. In Study 1
71 participants completed the self-esteem measures. Because no
equivalence bounds are mentioned, we can see which equivalence bounds
the researchers would have 80\% power to detect.

\begin{verbatim}
## The required sample size to achieve 80 % power with equivalence bounds of -0.24 and 0.24 is 71 pairs
\end{verbatim}

\begin{verbatim}
## [1] 70.05703
\end{verbatim}

With 71 pairs of observations between measures, the researchers have
80\% power for equivalence bounds of r = -0.24 and r = 0.24.

The correlations observed by Olson et al (2007), Study 1, are presented
in the table below (significant correlations are flagged by an
asteriks).

\begin{longtable}[]{@{}lllll@{}}
\toprule
Measure & IAT & Rosenberg & Feeling thermometer & Trait
ratings\tabularnewline
\midrule
\endhead
IAT & - & -.12 & -.09 & -.06\tabularnewline
Rosenberg & & - & .62* & .09\tabularnewline
Feeling thermometer & & & - & .29*\tabularnewline
Trait ratings & & & & -\tabularnewline
\bottomrule
\end{longtable}

We can test each correlation, for example the correlation between the
IAT and the Rosenberg self-esteem scale of -0.12, given 71 participants,
and using equivalence bounds of r\_U = 0.24 and r\_L = -0.24.

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-11-1.pdf}
\caption{}
\end{figure}

\begin{verbatim}
## Using alpha = 0.05 the NHST t-test was non-significant, p = 0.6191585
## Using alpha = 0.05 the equivalence test was non-significant, p = 0.06386793TOST results:
##     p-value 1  p-value 2
## 1 0.005971455 0.06386793
## 
## Equivalence bounds (r):
##   low bound r high bound r
## 1       -0.24         0.24
## 
## TOST confidence interval:
##   Lower Limit 90% CI raw Upper Limit 90% CI raw
## 1             -0.2538652              0.1384997
\end{verbatim}

We see that none of the correlations can be declared to be statistically
equivalent. Instead, the tests yield undetermined outcomes: the
correlations are not significantly different from 0, nor statistically
equivalent.

\emph{Olson, M. A., Fazio, R. H., \& Hermann, A. D. (2007). Reporting
tendencies underlie discrepancies between implicit and explicit measures
of self-esteem. Psychological Science, 18(4), 287-291.}

\section{Discussion}\label{discussion}

\newpage

\section{References}\label{references}

Button, K. S., Kounali, D., Thomas, L., Wiles, N. J., Peters, T. J.,
Welton, N. J., . Lewis, G. (2015). Minimal clinically important
difference on the Beck Depression Inventory - II according to the
patient's perspective. Psychological Medicine, 45(15), 3269-3279.
\url{https://doi.org/10.1017/S0033291715001270}

Burriss, R. P., Troscianko, J., Lovell, P. G., Fulford, A. J. C.,
Stevens, M., Quigley, R., . Rowland, H. M. (2015). Changes in Women's
Facial Skin Color over the Ovulatory Cycle are Not Detectable by the
Human Visual System. PLOS ONE, 10(7), e0130093.
\url{https://doi.org/10.1371/journal.pone.0130093}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-lakens_equivalence_2017}{}
Lakens, D. (2017). Equivalence Tests: A Practical Primer for t Tests,
Correlations, and Meta-Analyses. \emph{Social Psychological and
Personality Science}, \emph{8}(4), 355--362.
doi:\href{https://doi.org/10.1177/1948550617697177}{10.1177/1948550617697177}

\end{document}

### 2017-10-04 Dienes_example1.R
added power analysis

@@ -16,7 +16,9 @@ Shih.n2 <- 14
 
 ### Calculate subjective bound critera from original study data ###
 
-# Original had 33% Power to detect effect of d=0.57 (GPower result)
+# Original had 33% Power to detect effect of d=0.57:
+library(pwr)
+pwr.t.test(power=0.33,sig.level=0.05,n=15,type="two.sample",alternative="two.sided")
 
 # Original study smallest detectable effect size
 # Critical t for original study was:
@@ -27,7 +29,16 @@ crit_d = crit_t*sqrt(1/16 + 1/14)
 ### TOST for the different bounds ###
 
 # TOST for 33% Power rule:
+library(devtools)
+install_github("Lakens/TOSTER", force = TRUE)
 TOSTtwo(m1 = Gibson.m1, m2 = Gibson.m2, sd1 = Gibson.sd1, sd2 = Gibson.sd2, n1 = Gibson.n1, n2 = Gibson.n2, low_eqbound_d = -0.57, high_eqbound_d = 0.57, alpha = .05)
 
 # TOST for smallest detectable effect in original study
 TOSTtwo(m1 = 0.59, m2 = 0.53, sd1 = 0.23, sd2 = 0.23, n1 = 52, n2 = 54, low_eqbound_d = -crit_d, high_eqbound_d = crit_d, alpha = .05)
+
+#How much power did the replication have? 
+#For 33% power value
+powerTOSTtwo(alpha=0.05, N = Gibson.n1, low_eqbound_d=-0.57, high_eqbound_d=0.57)
+
+#For crit_d (note this will almost always be less strict than 33% threshold)
+powerTOSTtwo(alpha=0.05, N = Gibson.n1, low_eqbound_d=-crit_d, high_eqbound_d=crit_d)

### 2017-10-13 TOSTtwo.prop.R
Added 95% CIs and NHST results to the figure output

@@ -1,0 +1,73 @@
+#' TOST function for two proportions (raw scores)
+#' @param prop1 proportion of group 1
+#' @param prop2 proportion of group 2
+#' @param n1 sample size in group 1
+#' @param n2 sample size in group 2
+#' @param low_eqbound lower equivalence bounds (e.g., -0.1) expressed in proportions
+#' @param high_eqbound upper equivalence bounds (e.g., 0.1) expressed in proportions
+#' @param alpha alpha level (default = 0.05)
+#' @return Returns TOST z-value 1, TOST p-value 1, TOST z-value 2, TOST p-value 2, low equivalence bound, high equivalence bound, Lower limit confidence interval TOST, Upper limit confidence interval TOST
+#' @importFrom stats pnorm pt qnorm qt
+#' @importFrom graphics abline plot points segments title
+#' @examples
+#' ## Equivalence test for two independent proportions equal to .65 and .70, with 100 samples per group, a lower equivalence bound of -0.1,
+#' ## a higher equivalence bound of 0.1, and an alpha of 0.05.
+#' TOSTtwo.prop(prop1 = .65, prop2 = .70, n1 = 100, n2 = 100, low_eqbound = -0.1, high_eqbound = 0.1, alpha = .05)
+#' @section References:
+#' Tunes da Silva, G., Logan, B. R., & Klein, J. P. (2008). Methods for Equivalence and Noninferiority Testing. Biology of Blood Marrow Transplant, 15(1 Suppl), 120-127.
+#' Yin, G. (2012). Clinical Trial Design: Bayesian and Frequentist Adaptive Methods. Hoboken, New Jersey: John Wiley & Sons, Inc.
+#' @export
+
+TOSTtwo.prop <- function(prop1, prop2, n1, n2, low_eqbound, high_eqbound, alpha) {
+  if(missing(alpha)) {
+    alpha <- 0.05
+  }
+  prop_dif <- prop1 - prop2
+  prop_se <- sqrt((prop1*(1-prop1))/n1 + (prop2*(1-prop2))/n2)
+  
+  #calculating z-statistic
+  z1 <- (prop_dif - low_eqbound)/prop_se
[…]

### 2017-10-14 NHST_and_TOST.Rmd
Added RMarkdown file simulating NHST and TOST and plotting p-values.

---
title: "Simulating NHST and TOST"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```
#Simulating one-sample t-tests

```{r}
require(TOSTER)
n <- 100 #set sample size
low_eqbound=-0.3
high_eqbound=0.3
sd_true<-1
m_true<-0.3
mu<-0.3
alpha<-0.05
nSims <- 1000000 #number of simulated experiments
pttest<-numeric(nSims) #set up empty container for all simulated p-values
ptost<-numeric(nSims) #set up empty container for all simulated p-values
tlist<-numeric(nSims) #set up empty container for all simulated p-values
p1list<-numeric(nSims) #set up empty container for all simulated p-values
p2list<-numeric(nSims) #set up empty container for all simulated p-values
LLlist<-numeric(nSims)
ULlist<-numeric(nSims)

for(i in 1:nSims){ #for each simulated experiment
  x<-rnorm(n = n, mean = m_true, sd = sd_true) #Simulate data with specified mean, standard deviation, and sample size
  m<-mean(x)
  sd<-sd(x)
  degree_f<-n-1
  t1<-(m-mu-low_eqbound)/(sd/sqrt(n))# t-test
  p1<-pt(t1, degree_f, lower=FALSE) 
  p1list[i]<-p1
  t2<-(m-mu-high_eqbound)/(sd/sqrt(n)) #t-test
  p2<-pt(t2, degree_f, lower=TRUE) 
  p2list[i]<-p2
  t<-(m-mu)/(sd/sqrt(n))
  tlist[i]<-t
  pttest[i]<-2*pt(-abs(t), df=degree_f)
  LL90<-(m-mu)-qt(1-alpha, degree_f)*(sd/sqrt(n))
  LLlist[i]<-LL90
  UL90<-(m-mu)+qt(1-alpha, degree_f)*(sd/sqrt(n))
  ULlist[i]<-UL90
  ptost[i]<-max(p1,p2)
  ttost<-ifelse(abs(t1) < abs(t2), t1, t2) #Get lowest t-value for summary TOST result
}

#P-value distribution TOST
hist(ptost, breaks=20, xlim=c(0,1))
sum(ptost<0.05)/nSims #Power for TOST

sum(LLlist>=low_eqbound&ULlist<=high_eqbound)/nSims #Same power or Type 1 error, now based on CI

#Combine p-values from TOST and NHST

p<-data.frame(ptost,pttest)
plot(ptost,pttest)

```

### 2017-10-14 NHST_and_TOST.Rmd
update simulation (N = 80% power given equivalence bounds, p-dist plot).

@@
 ```{r}
+options(scipen=999)
 library(knitr)
@@
 require(TOSTER)
-n <- 100 #set sample size
+n <- 50 #set sample size
 low_eqbound=-0.3
@@
 sd_true<-1
-m_true<-0.3
-mu<-0.3
+m_true<-0.0
+mu<-0.0
 alpha<-0.05
+
+require("TOSTER")
+n<-powerTOSTone(alpha=alpha, statistical_power=0.8, low_eqbound_d=-low_eqbound, high_eqbound=0.3)
+n<-ceiling(n)
+
 nSims <- 1000000 #number of simulated experiments
@@
 p<-data.frame(ptost,pttest)
+#look only at significant TOST results
+p_sig_TOST<-p[which(ptost<alpha),]
+#Plot the p-values for the NHST t-test, but only those for significant TOT results. 
+hist(p_sig_TOST$pttest, breaks=100)
+
+#How often is a test significant AND equivalent?
+nrow(p[which(ptost<alpha & pttest<alpha),])/nSims
+
+#plot p-values against each other for TOST and NHST
 plot(ptost,pttest)

### 2017-10-14 NHST_and_TOST.Rmd
add html

@@
 output:
+  html_document: default
+  html_notebook: default
   pdf_document: default
-  html_notebook: default
-  html_document: default
 editor_options: 

### 2017-10-16 NSNEQ_Moon_Example.Rmd
Created first draft of Moon-Shih example as RMarkdown and added Moon data from OSF for analysis in Jamovi. Currently, Jamovi analysis does not reproduce analysis in R.

---
title: "NSNEQ_Moon_Example"
author: "Peder M Isager"
date: "2017 M10 16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Example of effect that is neither significant nor equivalent

Shih, Pittinsky, and Ambady (1999) conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity. They found a 5% increase in correct answers compared to an un-primed control group of Asian-American women, d = 0.26, Student t(30) = 0.76, p = 0.45, two-sided). A replication by Moon & Roeder (2014) observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, Welch t(96.62) = 1.15, p = 0.26, two-sided). Having failed to obtain a significant result, we are left with two alternatives:

1. The data indicate the absence of any effect, and we can accept the null hypothesis.
2. The data are not sensitive enough to tell us whether an effect is present or abscent.

In order to distinguish between these two hypotheses, we calculate the TOST. We first need to define our SESOI (and thus, our bounds). Following Simonsohn's (2015) recommendation for replications, we can set this to the effect size that the original study had 33% power to detect. using the "pwr" package in R, we can calculate this in the following way:

``` {r}
# Calculate effect size that Shih, Pittinsky, and Ambady (1999) had 33% power to detect
d33 <- pwr::pwr.t2n.test(n1          = 16, 
                         n2          = 16, 
                         power       = 0.33, 
                         sig.level   = 0.05, 
                         alternative = "two.sided")$d 
d33
```

Using the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using Cohen's d = 0.5549 as equivalence bounds. We find that the TOST is non-significant, t(96.62) = 1.64, p = 0.053. Thus, we cannot reject the possibility of the true effect being larger than our SESOI.

```{r}
TOSTER::TOSTtwo(m1  = 0.46, 
                m2  = 0.50, 
                sd1 = 0.17, 
                sd2 = 0.18, 
                n1  = 53, 
                n2  = 48, 
                low_eqbound  = -d33, 
                high_eqbound = d33, 
                alpha        = .05)

#Jamovi results

Moon_data <- read.csv("Moon_data_asianvscontrol.csv")
Moon_data_edit <- Moon_data[!Moon_data$identity_salience==3,]

TOSTER::dataTOSTtwo(data  = Moon_data_edit, 
                    deps  = "accuracy", 
                    group = "identity_salience", 
                    low_eqbound_d  = -0.5549, 
                    high_eqbound_d = 0.5549, 
                    plots     = TRUE, 
                    var_equal = FALSE)
```

### 2017-10-16 Dienes_example3.R
Added study 1 from Banerjee/Brandt, restructured previous code

@@ -1,175 +1,204 @@
 ###==========================================================================###
-################ Daniel: TOSTtwo + BFs from standardised effects ###############
+library(TOSTER)
+library(pwr)
 ###==========================================================================###
 
-#' TOST function for an independent t-test (Cohen's d)
-#' @param m1 mean of group 1
-#' @param m2 mean of group 2
-#' @param sd1 standard deviation of group 1
-#' @param sd2 standard deviation of group 2
-#' @param n1 sample size in group 1
-#' @param n2 sample size in group 2
-#' @param low_eqbound_d lower equivalence bounds (e.g., -0.5) expressed in standardized mean difference (Cohen's d)
-#' @param high_eqbound_d upper equivalence bounds (e.g., 0.5) expressed in standardized mean difference (Cohen's d)
-#' @param alpha alpha level (default = 0.05)
-#' @param var.equal logical variable indicating whether equal variances assumption is assumed to be TRUE or FALSE.  Defaults to FALSE.
-#' @return Returns TOST t-value 1, TOST p-value 1, TOST t-value 2, TOST p-value 2, degrees of freedom, low equivalence bound, high equivalence bound, low equivalence bound in Cohen's d, high equivalence bound in Cohen's d, Lower limit confidence interval TOST, Upper limit confidence interval TOST
-#' @importFrom stats pnorm pt qnorm qt
-#' @importFrom graphics abline plot points segments title
-#' @examples
-#' ## Eskine (2013) showed that participants who had been exposed to organic
-#' ## food were substantially harsher in their moral judgments relative to
-#' ## those exposed to control (d = 0.81, 95% CI: [0.19, 1.45]). A
-#' ## replication by Moery & Calin-Jageman (2016, Study 2) did not observe
[…]

### 2017-10-17 app.R
added app - can be basis for a TOST shiny app

@@ -1,0 +1,198 @@
+#To do: Scale y-axis p-curve
+#       Include Likelihood H1/H0
+
+library(shiny)
+library(shinythemes)
+ui <- fluidPage(theme=shinytheme("flatly"),
+                titlePanel("Distribution of Cohen's d, p-values, and power curves for an independent two-tailed t-test"),
+                sidebarLayout(
+                  sidebarPanel(numericInput("N", "Participants per group:", 50, min = 1, max = 1000),
+                               sliderInput("d", label = HTML("Cohen's d (&delta;)"), min = 0, max = 2, value = 0.5, step= 0.01),
+                               sliderInput("p_upper", "alpha, or p-value (upper limit):", min = 0, max = 1, value = 0.05, step= 0.005),
+                               uiOutput("p_low"),
+                               h4(textOutput("pow0")),br(),
+                               h4(textOutput("pow1")),br(),
+                               h4(textOutput("pow2")),br(),
+                               h4("The three other plots indicate power for a range of alpha levels (top right), sample sizes per group (bottom left), and effect sizes (bottom right). The bottom right figure illustrates the point that when the true effect size of a study is unknown, the power of a study is best thought of as a curve, not as a single value."),br()
+                  ),
+                  mainPanel(
+                    plotOutput("plot_d", width = "1004px", height = "500px"),
+                    splitLayout(style = "border: 1px solid silver:", cellWidths = c(500,500),cellHeights = c(800,800), 
+                                plotOutput("pdf"),
+                                plotOutput("cdf")
+                    ),
+                    splitLayout(style = "border: 1px solid silver:", cellWidths = c(500,500), 
[…]

### 2017-10-17 NSNEQ_Moon_Example.Rmd
Finished RMarkdown draft of Moon example, now with Jamovi code fixed, and the bound criterion based on hypothetical grade limits.

@@
 date: "2017 M10 16"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 We first need to define our SESOI (and thus, our bounds).
-Following Simonsohn's (2015) recommendation for replications, we can set this to the effect size that the original study had 33% power to detect.
-using the "pwr" package in R, we can calculate this in the following way:
+Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
+A+ = 93.75% to 100%).
+We can decide that we are only interested in test differences so far as they correspond to at least a 1 gradepoint increase or decrease.
+We can convert this percentage difference into a corresponding Cohen's d for Moon & Roeder's (2014) data by multiplying 6.25/100 on the pooled standard deviation of the groups.
+
+$$d = \frac{Diff}{SD_{pooled}}$$  
+$$SD_{pooled} = \sqrt \frac{sd_{1}^2 + sd_{2}^2}{2}$$
 
 ``` {r}
-# Calculate effect size that Shih, Pittinsky, and Ambady (1999) had 33% power to detect
-d33 <- pwr::pwr.t2n.test(n1          = 16, 
-                         n2          = 16, 
-                         power       = 0.33, 
-                         sig.level   = 0.05, 
-                         alternative = "two.sided")$d 
-d33
+
+sesoi     <- 0.065                        # <- Set the smallest effect size of interest
+sd.pooled <- sqrt((0.17^2 + 0.18^2) / 2)  # <- Calculate the pooled standard deviation
+sesoi.d   <- 0.065 / sd.pooled            # <- Convert the raw SESOI to Cohen's d
+
+print(sesoi.d)
+
 ```
 
-Using the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using Cohen's d = 0.5549 as equivalence bounds.
-We find that the TOST is non-significant, t(96.62) = 1.64, p = 0.053.
-Thus, we cannot reject the possibility of the true effect being larger than our SESOI.
+Using the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using +/- the value of `sesoi.d` (`r sesoi.d`) as our equivalence bounds.
+We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24.
+Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.
 
 ```{r}
+
+# Below are the code with rounded and ~accurate values.
+We need to decide which to use.
+The accurate one will correspond better to Jamovi.
+The rounded one will correspont better to the values given above.
+I (Peder) prefer the rounded.
 TOSTER::TOSTtwo(m1  = 0.46,
@@
                 n2  = 48,
-                low_eqbound  = -d33, 
-                high_eqbound = d33, 
+                low_eqbound  = -sesoi.d,
+                high_eqbound = sesoi.d,
                 alpha        = .05)
 
-#Jamovi results
+TOSTER::TOSTtwo(m1  = 0.4585130,
+                m2  = 0.4962693,
+                sd1 = 0.1749011,
+                sd2 = 0.1770452,
+                n1  = 53,
+                n2  = 48,
+                low_eqbound  = -sesoi.d,
+                high_eqbound = sesoi.d,
+                alpha        = .05)
+```
+
+```{r}
+#Jamovi results - This is just included to illustrate the correspondence between R results and what readers will see in Jamovi
 
@@
                     group = "identity_salience",
-                    low_eqbound_d  = -0.5549, 
-                    high_eqbound_d = 0.5549, 
+                    low_eqbound_d  = -sesoi.d,
+                    high_eqbound_d = sesoi.d,
                     plots     = TRUE,

### 2017-10-17 TOSTtwo.prop.R
Small edit to proportion test - added a report message stating significanse, p and z, like in TOSTtwo.

@@ -55,6 +55,10 @@ TOSTtwo.prop <- function(prop1, prop2, n1, n2, low_eqbound, high_eqbound, alpha)
   segments(CI_lb95,0.5,CI_ub95,0.5, lwd=1)
   title(main=paste("Equivalence bounds ",round(low_eqbound,digits=3)," and ",round(high_eqbound,digits=3),"\nProportion Difference = ",round(prop_dif,3)," \n TOST: ", 100*(1-alpha*2),"% CI [",round(CI_lb,digits=3),";",round(CI_ub,digits=3),"] ", TOSToutcome, " \n NHST: ", 100*(1-alpha),"% CI [",round(CI_lb95,digits=3),";",round(CI_ub95,digits=3),"] ", ZTESToutcome, sep=""), cex.main=1)
   
+  message(cat("Using alpha = ",alpha," Fishers exact z-test was ",ZTESToutcome,", z = ",z,", p = ",ztest,sep=""))
+  cat("\n")
+  message(cat("Using alpha = ",alpha," the equivalence test based on Fishers exact z-test was ",TOSToutcome,", z = ",ztost,", p = ",ptost,sep=""))
+  
   TOSTresults<-data.frame(z1,p1,z2,p2)
   colnames(TOSTresults) <- c("z-value 1","p-value 1","z-value 2","p-value 2")
   bound_results<-data.frame(low_eqbound,high_eqbound)

### 2017-10-17 SNEQ_Lynott_example.Rmd
Completed first draft of the significant non-equivalent proportions test from Lynott et al. Skipped the reference to values from Williams & Barge, since they are not used in any calcuations, and rephrased the bound criterion to smallest effect size that we can detect given the current sample. An alternative would be to calculate the smallest detectable effect size given the original sample.

---
title: "SNEQ_Lynott_example"
author: "Peder M Isager"
date: "October 17, 2017"
output: html_document
---

# Example of effect that is significant and not equivalent

Lynott et al. (2014), replicating Williams & Bargh (2008) conducted a study to investigate the effect of being exposed to physical warmth on interpersonal warmth and prosocial behavior. They observed  that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same. Calculating a z-test for the difference between the proportions, this effect is statistically significant (Diff = -0.067, z = -1.98, p = 0.024). Having obtained a significant result, we are left with two alternatives:

1. The difference is small enough that even though it is *statisitcally* significant, we can consider it *practically* equivalent to zero.
2. The difference is substantial enough that we cannot rule out that it is also practically significant.

In order to distinguish between these two hypotheses, we calculate the TOST. We first need to define our SESOI (and thus, our bounds). In this case, it is not really clear what an objective cirterion for the smallest effect size of interest could be. However, we could decide that we do not care about an effect if the difference is smaller than the smallest effect we could potentially detect with the current sample. This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds. To convert the critical z value to the critical difference in proportions for this study, we multiply critical z with the standard error:

```{r}
se     <- sqrt((0.512 * (1 - 0.512)) / 427 + (0.579 * (1 - 0.579)) / 434)  # Calculate the (unpooled) standard error
crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05

crit.diff <- crit.z * se  # Convert z to the critical difference

print(crit.diff)
```

Using the TOSTtwo.prop functiion in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Lynott et al. (2014), assuming an aplha of 0.05, and using +/- the value of `crit.diff` (`r crit.diff`) as our equivalence bounds. We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24. Thus, in addition to begin statistically different from zero, we cannot reject the possibility that the true difference really is larger than our SESOI.

```{r}

download.file(url = "https://raw.githubusercontent.com/Lakens/EquivalenceTestingForPsychologicalScience/master/TOSTtwo.prop.R", destfile = "TOSTtwo.prop.R")
source(file = "TOSTtwo.prop.R")

TOSTtwo.prop(prop1 = 0.512, 
             prop2 = 0.579, 
             n1    = 427, 
             n2    = 434, 
             low_eqbound  = -crit.diff, 
             high_eqbound = crit.diff, 
             alpha = 0.05)

```

### 2017-10-18 NSNEQ_Moon_Example.Rmd
minor comments

@@
 ```
-
-# Example of effect that is neither significant nor equivalent
+<!-- I don't think we need to say 'Welch's t-test. We might throw in a quick remakr that by default, the TOST package reports Welch's t-test (following Delacre et al., 2017) -->  
+# Example of an effect that is neither statistically significant nor equivalent
 
@@
 
-1. The data indicate the absence of any effect, and we can accept the null hypothesis.
-2. The data are not sensitive enough to tell us whether an effect is present or abscent.
+1. The data indicate the absence of a meaningful effect, and we can accept the null hypothesis.
+2. The data are not sensitive enough to tell us whether a meaningful effect is present or absent.
 
-In order to distinguish between these two hypotheses, we calculate the TOST.
-We first need to define our SESOI (and thus, our bounds).
+In order to distinguish between these two hypotheses, we first need to define a SESOI (and thus, our equivalence bounds).
 Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
@@
 ```  
+<!-- So the TOST results below should not be typed in, but calculated in R, like you do with sesoi.d -->
 

### 2017-10-18 NHST_and_TOST.Rmd
NHST_TOST p-value plots, comments Lynott example

@@
 #P-value distribution TOST
-hist(ptost, breaks=20, xlim=c(0,1))
+hist(ptost, breaks=500, xlim=c(0,1))
 sum(ptost<0.05)/nSims #Power for TOST
@@
 
+#plot p-value distribution for TOST - looks peculiar
+#Dat can only be so far removed from one bound, before it gets closer to the other bound. So p = 0.0000001 often not possible 
+# (need to widen bounds for that, or increase N)
+hist(p_sig_TOST$ptost, breaks=100)
+
 #How often is a test significant AND equivalent?

### 2017-10-18 NSNEQ_Moon_Example.Rmd
My last edits to Moon example before Daniel's comments. Remoced details of original as it was unnecessary for the example.

--- 2017-10-17 NSNEQ_Moon_Example.Rmd
@@
 
-Shih, Pittinsky, and Ambady (1999) conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
-They found a 5% increase in correct answers compared to an un-primed control group of Asian-American women, d = 0.26, Student t(30) = 0.76, p = 0.45, two-sided).
-A replication by Moon & Roeder (2014) observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, Welch t(96.62) = 1.15, p = 0.26, two-sided).
+Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
+They observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, Welch t(96.62) = 1.15, p = 0.26, two-sided).
 Having failed to obtain a significant result, we are left with two alternatives:
@@
 
-Using the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using +/- the value of `sesoi.d` (`r sesoi.d`) as our equivalence bounds.
+Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using +/- the value of `sesoi.d` (`r sesoi.d`) as our equivalence bounds.
 We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24.

### 2017-10-18 SNEQ_Lynott_example.Rmd

@@
 Lynott et al.
-(2014), replicating Williams & Bargh (2008) conducted a study to investigate the effect of being exposed to physical warmth on interpersonal warmth and prosocial behavior.
+(2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on ubsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
 They observed that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same.
@@
 
-1. The difference is small enough that even though it is *statisitcally* significant, we can consider it *practically* equivalent to zero.
+1. The difference is small enough that even though it is *statistically* significant, we can consider it *practically* equivalent to zero.
 2. The difference is substantial enough that we cannot rule out that it is also practically significant.
@@
 
-Using the TOSTtwo.prop functiion in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Lynott et al.
+Using the TOSTtwo.prop function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Lynott et al.
 (2014), assuming an aplha of 0.05, and using +/- the value of `crit.diff` (`r crit.diff`) as our equivalence bounds.

### 2017-10-18 manuscript.Rmd

@@
 
-An equivalence test can be used to test whether the observed effect is surprisingly close to zero, assuming a meaningful effect exists in the population.
-The test is a simple variation on the widely used null hypothesis significance tests, with the small difference that instead of testing the observed effect against zero, it is tested against a smallest effect size of interest (SESOI).
-For example, if a difference on a math test of 5% or more is considered meaningful, one could test against a SESOI of 5%.
-If the observed effect is suprisingly closer to 0, assuming a true effect of 5% or larger existed, one can declare statistical equivalence, or reject the presence of effects that are large enough to be considered meaningful.
-An equivalence test can be used to falsify hypotheses.
+An equivalence test can be used to test whether the observed effect is surprisingly small, assuming a meaningful effect exists in the population.
+The test is a simple variation on the widely used null hypothesis significance tests.
+To understand the idea behind equivalence tests, it is useful to realize the null hypothesis we test against can be any numerical value.
+When we compare two groups, we often test whether the difference between groups is zero.
 
-One can conclude statistical equivalence whenever a 90% confidence interval around the effect size estimate does not contain the SESOI, or when two one-sided tests (i.e., examining whether the difference on the math test is more than -5%, but less than 5%) are statistically significant.
+Imagine a researcher who who is interested in voluntary participation in a national program to train young infants' motor skills.
+The researcher wants to test whether more boys than girls are brought in to the training program by their parents.
+One could test whether the difference betwen boys and girls is zero.
+However, this would ignore that the human sex ratio of boys to girls is not exactly 1:1, or an odds ratio of 1.
+On average, slightly more boys than girls are born.
+Given that a null hypothesis of exactly zero is unlikely, the researcher tests whether the difference between boys and girls is surprising, assuming an odds ratio in the range of 0.93 to 1.07.
+If the null hypothesis can be rejected (indicating the effect is surprisingly smaller than OR 0.93, or sirprisingly larger than 1.07), the researcher decides to act as if there is a real gender difference in participation in the motor skill training program.
+This test against is known as a *minimal effect test*.
+
+However, even if the researcher would consider any odds ratio statistically smaller than 0.93 or larger than 1.07 a rejection of the null hypothesis, not every difference is large enough to matter.
+After an extensive discussion with experts, the researcher decides that any effect smaller than an odds ratio of 0.8 or larger than an odds ratio of 1.2 is substantial enough to move money away from the core training program, and address the gender imbalance in the voluntary participation rate.
+To test whether the gender difference is not large enough to matter, the researcher can perform an *equivalence test*, which tests whether the observed gender difference is statistically larger than an odds ratio of 0.8, *and* statistically smaller than an odds ratio of 1.2.
+If this is the case, the researcher can act as if the gender difference is too small to intervene.
+
+One can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI; the values 0.8 and 1.2 in the previous example).
+When two one-sided tests (i.e., one testing whether the odds ratio is larger than 0.8 the other testing whether the odds ration is smaller than 1.2) are statistically significant, we can conclude statistical equivalence.
 By combining null-hypothesis tests and equivalence tests, four combinations of results can be observed (see Figure 1).
@@
 
-Even though equivalence tests are just a small variation of traditional Frequentist tests (testing against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017].
-In this article, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.
-
-<!-- If we use Dienes & McLatchie's 5 examples, it would make sense to add a Figure like a forest plot, where we present the effect size estimates (raw or standardized) indicated by a square, lines that represent the 90% CI, and another symbol or | to indicate the SESOI, sor all examples.
-We can then also use the same Figure in the intro to explain the four possible outcomes of equivalence testing).
-Although we should do all tests in R, we also should try to create Jamovi files that contain the data and analyses - another reason to follow Zoltan in using the Social Psychology special issue studies, because the data should be shared, and we can import the raw data into Jamovi files, save these (with the analysis) and share them.
--->
+Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017].
+In this article, we provide several examples of equivalence tests which illustrate the procedure.
+Furthermore, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.
 
@@
 Most often, however, it is a subjective decision that varies across individuals and time.
-Both these approaches are valid, and even when a just noticeable difference can be objectively determined, researchers might choose to set the SESOI to a larger value.
-In many research areas, the SESOI is best based on a cost-benefit analysis.
+The SESOI is ideally based on a cost-benefit analysis.
 Since both costs and benefits are necessarily relative, the SESOI will depend on the researcher who designs the study.
@@
 ## Examples
+
+#Adressing Trello cards
+
+For the example where the results are inconclusive, explain how adding more data, will, assuming the true effect is 0, lead to a smaller confidence interval, and thus will lead to an informative test.
+However, with increasing data, the effect could also slightly increase, to above or at the SESOI level, and then the effect will be significant and not equivalent.
+With massive amounts of data, many results will be significant and equivalent.
+
+Boundary hacking should go into the discussion.
 

### 2017-10-18 NSEQ_Brandt_Example.Rmd
First draft for non-significant+equivalent example

---
title: "Brandt et al. example: equivalent & not significant"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Equivalence bounds based on critical effect sizes

```{r include = FALSE}
library(TOSTER)

# original study: Banerjee et al., study 1
orig.m1 <- 4.71
orig.m2 <- 5.3
orig.sd1 <- 0.85
orig.sd2 <- 0.97
orig.d <- 0.65
orig.t <- 2.03
orig.p <- 0.049
orig.N <- 40 #group size unknown, therefore equal n assumed
orig.df <- orig.N-2

# replication study: Brandt et al., study 1
rep.m1 <- 4.79
rep.m2 <- 4.66
rep.sd1 <- 1.09
rep.sd2 <- 1.19
rep.n1 <- 49
rep.n2 <- 51
rep.t <- 0.56
rep.p <- .574
rep.df <- rep.n1 + rep.n2 - 2
rep.d <- 0.11

# Calculate critical effect (d): smallest effect orig. study had the power to detect
t.crit = qt(1-.05/2, (orig.N-2))
d.crit = t.crit*sqrt((1/(orig.N/2)) + 1/(orig.N/2))

Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
```

Another justifiable choice we would like to propose is to use the smallest observed effect size that could have been statistically significant in the original study. Based only on the alpha level and the sample size, we can calculate the criticial test value (e.g., $t$, $F$, $Z$). This critical test value can be transformed to a standardized effect size (e.g., $d_{crit} = t_{crit} \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*[^1]. Observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study. Based on this we can guess that the authors were not interested in effects smaller than this critical effect size, and thus use it as the SESOI. An equivalence test with these bounds can reject all observed effect sizes that the original study had the power to detect.

As an example, Banerjee, Chatterjee, & Sinha (2012) reported that participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$). A close replication by Brandt, IJzerman, & Blanken (2014) found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).  The smallest effect the original study could have detected is $d_{crit}= `r round(t.crit, 2)` \sqrt { \frac { 1} {`r orig.N/2`} + \frac { 1} {`r orig.N/2`} }=`r round(d.crit, 2)`$. Using this as our SESOI for a TOST with Welch's t-test for independent samples --- resulting in equivalence bounds of ${\Delta}_L=`r round(-d.crit, 2)`$ and ${\Delta}_U=`r round(d.crit, 2)`$ --- we indeed find that the effect reported by the replication study is equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.

```{r, fig.width=6}
TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
```

[^1]: This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect. This procedure will thus result in effect sizes that are substantially larger than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect.

### 2017-10-18 TOSTtwo.prop.R
Corrected minor errors and added output to list in TOSTtwo.prop.R

@@ -55,10 +55,10 @@ TOSTtwo.prop <- function(prop1, prop2, n1, n2, low_eqbound, high_eqbound, alpha)
   segments(CI_lb95,0.5,CI_ub95,0.5, lwd=1)
   title(main=paste("Equivalence bounds ",round(low_eqbound,digits=3)," and ",round(high_eqbound,digits=3),"\nProportion Difference = ",round(prop_dif,3)," \n TOST: ", 100*(1-alpha*2),"% CI [",round(CI_lb,digits=3),";",round(CI_ub,digits=3),"] ", TOSToutcome, " \n NHST: ", 100*(1-alpha),"% CI [",round(CI_lb95,digits=3),";",round(CI_ub95,digits=3),"] ", ZTESToutcome, sep=""), cex.main=1)
   
-  message(cat("Using alpha = ",alpha," Fishers exact z-test was ",ZTESToutcome,", z = ",z,", p = ",ztest,sep=""))
+  message(cat("Using alpha = ", alpha," Fishers exact z-test was ", ZTESToutcome,", z = ", z,", p = ", ztest * 2, sep=""))
   cat("\n")
   message(cat("Using alpha = ",alpha," the equivalence test based on Fishers exact z-test was ",TOSToutcome,", z = ",ztost,", p = ",ptost,sep=""))
-  
+  cat("\n")
   TOSTresults<-data.frame(z1,p1,z2,p2)
   colnames(TOSTresults) <- c("z-value 1","p-value 1","z-value 2","p-value 2")
   bound_results<-data.frame(low_eqbound,high_eqbound)
@@ -73,5 +73,5 @@ TOSTtwo.prop <- function(prop1, prop2, n1, n2, low_eqbound, high_eqbound, alpha)
   cat("\n")
   cat("TOST confidence interval:\n")
   print(CIresults)
-  invisible(list(TOST_z1=z1,TOST_p1=p1,TOST_z2=z2,TOST_p2=p2, alpha=alpha,low_eqbound=low_eqbound,high_eqbound=high_eqbound, LL_CI_TOST=CI_lb,UL_CI_TOST=CI_ub))
[…]

### 2017-10-18 NSNEQ_Moon_Example.Rmd

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 ```
-<!-- I don't think we need to say 'Welch's t-test. We might throw in a quick remakr that by default, the TOST package reports Welch's t-test (following Delacre et al., 2017) -->  
 # Example of an effect that is neither statistically significant nor equivalent
@@
 Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
-They observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, Welch t(96.62) = 1.15, p = 0.26, two-sided).
+They observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, t(96.62) = 1.15, p = 0.26, two-sided).
 Having failed to obtain a significant result, we are left with two alternatives:
@@
 We can decide that we are only interested in test differences so far as they correspond to at least a 1 gradepoint increase or decrease.
-We can convert this percentage difference into a corresponding Cohen's d for Moon & Roeder's (2014) data by multiplying 6.25/100 on the pooled standard deviation of the groups.
+Thus, our SESOI becomes 6.25%, or 0.0625.
 
-$$d = \frac{Diff}{SD_{pooled}}$$  
-$$SD_{pooled} = \sqrt \frac{sd_{1}^2 + sd_{2}^2}{2}$$
-
-``` {r}
-
-sesoi     <- 0.065                        # <- Set the smallest effect size of interest
-sd.pooled <- sqrt((0.17^2 + 0.18^2) / 2)  # <- Calculate the pooled standard deviation
-sesoi.d   <- 0.065 / sd.pooled            # <- Convert the raw SESOI to Cohen's d
-
-print(sesoi.d)
-
-```  
-<!-- So the TOST results below should not be typed in, but calculated in R, like you do with sesoi.d -->
-
-Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using +/- the value of `sesoi.d` (`r sesoi.d`) as our equivalence bounds.
-We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24.
-Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.
+Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder's (2014) data, assuming an aplha of 0.05, and using +/- the SESOI of 0.0625 as our equivalence bounds.
 
@@
 
-# Below are the code with rounded and ~accurate values.
-We need to decide which to use.
-The accurate one will correspond better to Jamovi.
-The rounded one will correspont better to the values given above.
-I (Peder) prefer the rounded.
-TOSTER::TOSTtwo(m1  = 0.46,
-                m2  = 0.50,
-                sd1 = 0.17,
-                sd2 = 0.18,
-                n1  = 53,
-                n2  = 48,
-                low_eqbound  = -sesoi.d,
-                high_eqbound = sesoi.d,
-                alpha        = .05)
+#  We first load the relevant datafile into R
+moon.data    <- read.csv("NSNEQ_Moon_Example_data.csv")
 
-TOSTER::TOSTtwo(m1  = 0.4585130,
-                m2  = 0.4962693,
-                sd1 = 0.1749011,
-                sd2 = 0.1770452,
-                n1  = 53,
-                n2  = 48,
-                low_eqbound  = -sesoi.d,
-                high_eqbound = sesoi.d,
+#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
+mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
+sd.asian     <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
+sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
+n.asian      <- sum(moon.data$identity_salience == 1)
+n.control    <- sum(moon.data$identity_salience == 2)
+
+#  We can now run the TOST!
+#  Performing the TOST using the 'TOSTtwo.raw' function allows us to specify equivalence bounds in raw values
+tost <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
+                            m2  = mean.control,
+                            sd1 = sd.asian,
+                            sd2 = sd.control,
+                            n1  = n.asian,
+                            n2  = n.control,
+                            low_eqbound  = -0.0625,
+                            high_eqbound =  0.0625,
                             alpha = .05)
@@
 
-```{r}
-#Jamovi results - This is just included to illustrate the correspondence between R results and what readers will see in Jamovi
-
-Moon_data <- read.csv("Moon_data_asianvscontrol.csv")
-Moon_data_edit <- Moon_data[!Moon_data$identity_salience==3,]
-
-TOSTER::dataTOSTtwo(data  = Moon_data_edit,
-                    deps  = "accuracy",
-                    group = "identity_salience",
-                    low_eqbound_d  = -sesoi.d,
-                    high_eqbound_d = sesoi.d,
-                    plots     = TRUE,
-                    var_equal = FALSE)
-```
+We find that the TOST is non-significant, t(`r round(tost$TOST_df, 2)`) = `r round(min(abs(c(tost$TOST_t1, tost$TOST_t2))), 2)`, p = `r round(max(tost$TOST_p1, tost$TOST_p2), 2)`.
+Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.

### 2017-10-19 NSEQ_Lynott_example.Rmd
Lynott example - Finished draft, including adding the data file, a description file, and a jamovi file. Current example replicates Lynott SPSS .sav+.sps but not values in paper. Notification letter to the authors has been composed.

--- 2017-10-18 SNEQ_Lynott_example.Rmd
@@
 In this case, it is not really clear what an objective cirterion for the smallest effect size of interest could be.
-However, we could decide that we do not care about an effect if the difference is smaller than the smallest effect we could potentially detect with the current sample.
+However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
 This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds.
-To convert the critical z value to the critical difference in proportions for this study, we multiply critical z with the standard error:
+To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
 
 ```{r}
-se     <- sqrt((0.512 * (1 - 0.512)) / 427 + (0.579 * (1 - 0.579)) / 434)  # Calculate the (unpooled) standard error
+se     <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
 crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05
@@
 (2014), assuming an aplha of 0.05, and using +/- the value of `crit.diff` (`r crit.diff`) as our equivalence bounds.
-We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24.
-Thus, in addition to begin statistically different from zero, we cannot reject the possibility that the true difference really is larger than our SESOI.
 
@@
 
-TOSTtwo.prop(prop1 = 0.512, 
-             prop2 = 0.579, 
-             n1    = 427, 
-             n2    = 434, 
+#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria. 
+lynott.data <- read.csv("SNEQ_Lynott_example_data.csv")
+lynott.data <- lynott.data[lynott.data$include1 == 1,]
+
+#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+n.hot        <- sum(lynott.data$packtype == 1)
+n.cold       <- sum(lynott.data$packtype == 0)
+selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
+selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
+
+#  We can now run the TOST!
+tost <- TOSTtwo.prop(prop1 = selfish.hot, 
+                     prop2 = selfish.cold, 
+                     n1    = n.hot, 
+                     n2    = n.cold, 
                      low_eqbound  = -crit.diff, 
@@
 ```
+
+We find that the TOST is `r tost$TOST_outcome`, z = `r round(min(abs(c(tost$TOST_z1, tost$TOST_z2))), 2)`, p `r ifelse(round(max(abs(c(tost$TOST_p1, tost$TOST_p2))), 2) >= 0.01, paste("=", round(max(abs(c(tost$TOST_p1), tost$TOST_p2))), 2), "< 0.01")`.
+Thus, in addition to beinn statistically different from zero, we cannot reject the possibility that the true difference really is larger than our SESOI.

### 2017-10-19 NSEQ_Brandt_Example.Rmd

@@
 # replication study: Brandt et al., study 1
-rep.m1 <- 4.79
-rep.m2 <- 4.66
-rep.sd1 <- 1.09
-rep.sd2 <- 1.19
+rep.m1 <- 4.7857
+rep.m2 <- 4.6569
+rep.sd1 <- 1.0897
+rep.sd2 <- 1.1895
 rep.n1 <- 49

### 2017-10-21 NSEQ_Lynott_example.Rmd
Fixed direction of the odds ratio test from hot/cold to cold/hot, to match the test in Lynott et al.

@@
 #  We can now run the TOST!
-tost <- TOSTtwo.prop(prop1 = selfish.hot, 
-                     prop2 = selfish.cold, 
-                     n1    = n.hot, 
-                     n2    = n.cold, 
+tost <- TOSTtwo.prop(prop1 = selfish.cold, 
+                     prop2 = selfish.hot, 
+                     n1    = n.cold, 
+                     n2    = n.hot, 
                      low_eqbound  = -crit.diff, 

### 2017-10-23 NSEQ_Lynott_example.Rmd
Peder - Fixed minor edits in Lynott
* Small fixes in Lynott Rmd
* Renamed files after fixing calculations
Peder - Changed example to non-inferiority
Math is the same, but explanation has been changed, and only upper bound (z2 and p2) is used in conclusion.

@@
 
-# Example of effect that is significant and not equivalent
+# Example of effect that is not significant and is inferior to SESOI
 
 Lynott et al.
-(2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on ubsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
+(2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
 They observed that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same.
-Calculating a z-test for the difference between the proportions, this effect is statistically significant (Diff = -0.067, z = -1.98, p = 0.024).
-Having obtained a significant result, we are left with two alternatives:
+Calculating a z-test for the difference between the proportions, this effect is not statistically significant (Diff = -0.067, z = -1.98, p = 0.024).
+Having obtained a non-significant result, we are left with the same two alternatives as in example 3.
 
-1. The difference is small enough that even though it is *statistically* significant, we can consider it *practically* equivalent to zero.
-2. The difference is substantial enough that we cannot rule out that it is also practically significant.
-
-In order to distinguish between these two hypotheses, we calculate the TOST.
-We first need to define our SESOI (and thus, our bounds).
-In this case, it is not really clear what an objective cirterion for the smallest effect size of interest could be.
+In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be.
 However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
 This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds.
-To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
+To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error.
 
@@
 
-Using the TOSTtwo.prop function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Lynott et al.
-(2014), assuming an aplha of 0.05, and using +/- the value of `crit.diff` (`r crit.diff`) as our equivalence bounds.
+Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study.
+In this case, we can still use a TOST to analyse the data, but we only need to consider the test against the upper bound because an effect with CIs crossing the lower bound can still be *lower* than the original effect.
 
@@
 
-#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria. 
-lynott.data <- read.csv("SNEQ_Lynott_example_data.csv")
+#  We first load the data...
+lynott.data <- read.csv("NSEQ_Lynott_example_data.csv")
+#  ... and only include cases that did not meet the authors' exclusion criteria. 
 lynott.data <- lynott.data[lynott.data$include1 == 1,]
 
-#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+#  We then calculate the summary statistics from the data that we need in the TOST calculation below.
 n.hot        <- sum(lynott.data$packtype == 1)
@@
 
-We find that the TOST is `r tost$TOST_outcome`, z = `r round(min(abs(c(tost$TOST_z1, tost$TOST_z2))), 2)`, p `r ifelse(round(max(abs(c(tost$TOST_p1, tost$TOST_p2))), 2) >= 0.01, paste("=", round(max(abs(c(tost$TOST_p1), tost$TOST_p2))), 2), "< 0.01")`.
-Thus, in addition to beinn statistically different from zero, we cannot reject the possibility that the true difference really is larger than our SESOI.
+We find that the test against the upper bound is `r ifelse(tost$TOST_p2 < tost$alpha, "significant", "non-significant")`, $z =$ `r round(tost$TOST_z2, 2)`, $p =$ `r ifelse(round(tost$TOST_p2, 3) >= 0.001, paste("=", round(tost$TOST_p2, 2)), "< 0.001")`.
+Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.

### 2017-10-23 H0H1_plot.R
plot to visuale "directions"/regions/bounds of testing for equivalence tests, NHST, minimal effects tests, and inferiority tests

@@ -1,0 +1,58 @@
+library(ggplot2)
+library(gridExtra)
+
+plotheight <- 0.9
+lowerbound <- -0.2
+upperbound <- 0.2
+df <- data.frame()
+
+baseplot <-   ggplot(df) +
+  scale_y_continuous(limits = c(0,plotheight+0.02), breaks=NULL) + # no y-axis will be displayed
+  scale_x_continuous(limits = c(-0.85,0.85), breaks=c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8), name = "effect size (d)") +
+  theme_classic() +
+  theme(plot.title = element_text(size = rel(1.3), face = "bold"), #font size & appearance for plot titles
+        axis.title.y = element_blank(), #remove title of y-axis
+        axis.title.x = element_text(size=rel(1.1), lineheight = 0.5), #font size for x-axis label
+        plot.margin=unit(c(0.5,0.8,0.5,0.8),"cm")) #add padding around each plot to make them look nicer when combined; margin order: top, right, bottom, left
+
+#equivalence test plot
+eqplot <- baseplot +  
+  ggtitle("A: Equivalence test") +
+  annotate("segment", x = lowerbound, xend = lowerbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for lower bound
+  annotate("segment", x = upperbound, xend = upperbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for upper bound
+  annotate("rect", xmin = -Inf, xmax = lowerbound, ymin = -Inf, ymax = plotheight, fill = "red", alpha = .2, color = NA) + #shading for lower area
+  annotate("rect", xmin = upperbound, xmax = Inf, ymin = -Inf, ymax = plotheight, fill = "red", alpha = .2, color = NA) + #shading for upper area
+  annotate("text", size = rel(5), x=-0.6, y=plotheight/2, parse=TRUE, label="H[0]") + #label for lower area (H0)
+  annotate("text", size = rel(5), x=0.6, y=plotheight/2, parse=TRUE, label="H[0]") + #label for upper area (H0)
+  annotate("text", size = rel(5), x=-0, y=plotheight/2, parse=TRUE, label="H[1]", hjust = 0.3) #label for equivalence area
+
+#NHST plot
+NHSTplot <- baseplot +  
+  ggtitle("B: Classic NHST (two-sided)") +
[…]

### 2017-10-24 Empirical studies citing Lakens (2017).Rmd
Peder - summary of papers citing Lakens
* Reviewed the literature citing Lakens (2017) equivalence primer
* Wrote a short summary for each paper

---
title: "Empirical studies citing Lakens' (2017) equivalence primer"
author: "Peder M Isager"
date: "October 24, 2017"
output: html_document
---

The following studies are all empirical studies that cites Lakens' (2017) equivalence primer. Studies that appear to be reviews or theoretical pieces were not included. A short description of the study, and a short description of the bound criterion used, is provided for each study.

## Kunz et al. (2017)

The authors study how hematologic properties of blood changes during long-duration spaceflight. Equivalence tests were applied in a within subject design to test whether various hematologic properties would remain stable over prolonged periods (days - weeks) of spaceflight.

Equivalence bounds were set based on the within-person coefficient of variation for each hematologic parameter in a group of 852 non-institutionalized subjects, reported by Lacher et al. (2012).

## Mirabella et al. (2017)

The authors study inhibitory control's reliance upon a right-lateralized brain pathway by studying inhibition function in left and right lateralized Parkinson patients.

Equivalence bounds were set to the bounds that the study had 80% power to reject a true effect of 0 with.

## Schumann et al. (2017)

The authors assessed moderators of successful computer-mediated intergroup contact (CMIC).

Equivalence bounds were based on Cohen's d benchmarks of −0.50 and d = 0.50.

## Protzko and Schooler (2017)

Studied the relationship between researcher "caliber" (defined as h-index), and ability to successfully replicate (defined as effect size in replication).

Equivalence bounds were not formally defined. Instead, 90% CIs were calculated and found to be very narrow (beta of +/- .005 around a mean of .00003). The authors then state (something to the effect of) that effects within this interval is practically equivalent to zero.

## Petsko and Bondenhausen (2017)

The authors attempted to replicate the race-crime congruency effect, and study the influence of sexual orientation on this effect.

Equivalence bounds were set to the bounds that the study had 80% power to reject a true effect of 0 with (using the power calculator for equivalence tests).

## Müller and Gregorick (2017)

The authors study the effect of a feedback program on student grammar in academic writing.

Equivalence bounds were set to d = +/- 0.25 standardized gain (reduction in errors). Their explanation is that this represents an educationally meaningful effect, but the justification appears to be based on expert advice (Wolf, 1986) rather than theoretical or empirical evidence.

## Mc Gee et al. (2017)

The authors studied the difference between age groups on the Sense of Coherence scale.

Equivalence bounds were set to d = +/- 0.4. Sample size and power of 80% is mentioned, but does not make it clear exactly how the authors set the bounds: *"Given the current sample size and an achieved power of 80%, modest lower and upper equivalence bounds of Cohen’s d = ¡0.4 and d = 0.4 were set as the smallest effect size of interest"*.

## Kordsmeyer and Penke (2017)

The authors studied the relationship between bodily assumetries and indicators of quantitative mating success.

Equivalence bounds were based on mean effects from previous meta-analyses of similar effects.

## Matheson et al (2017)

The authors compare different methods for assessing PET receptor binding (of a specific ligand to a specific protein) compared to gold standard.

Equivalence tests are referenced for future research, but not utilized in the study.

## Wimmer & Poldrack (2017)

The authors studied the effect of time distribution on learning effects in feedback training.

Equivalence bounds appears to be set based on the effect that the studies have 80% power to detect, but it is not perfectly clear based on the language in the paper whether this is the case: *"We used bounds of Cohen’s d = 0.51 (Study1) or d = 0.53 and d = 0.54 (Study 2), where power to detect an effect in the included group of participants is estimated to be 80%."*

Used TOSTER in R

## Geerse et al. (2017)

the authors studied the reliability of the eKinect v2 Sensor in quantifying gait, based on foot placement locations.

Equivalence bounds were based on the smallest "limits of agreement" (?) in a previous study on similar effects by the same authors.

Used TOSTER in Jamovi

## Brown et al. (2017)

The authors studied the effect of behavioral inhibition system (BIS) activation on likelihood to convict in a lab-constructed jury situation.

Equivalence bounds were based on Cohen's d benchmarks of −0.50 and d = 0.50. This is assumed, but the authors never explicitly state the justification behind the chosen equivalence bound.

### 2017-10-26 examples_plotgrid.R
Peder - 2x2 grid of example TOST plots
* Added code for plotting the meta-, Brandt-, Moon-, and Lynott examples, using ggplot2, purrr, and gridExtra.

@@ -1,0 +1,192 @@
+#  Plot example effects for AMPPS equivalence paper
+
+####  Setup ----------------------------------------
+
+library(purrr)
+library(gridExtra)
+library(ggplot2)
+
+library("TOSTER")
+download.file(url = "https://raw.githubusercontent.com/Lakens/TOSTER/master/R/TOSTtwo.prop.R", destfile = "TOSTtwo.prop.R")
+source(file = "TOSTtwo.prop.R")
+
+plot.scaler = 2
+
+####  Base plot ----------------------------------------
+
+baseplot <- ggplot(data.frame()) +
+  scale_x_continuous(breaks=NULL) +
+  coord_flip() + # flip coordinates (puts labels on y axis)
+  theme_classic() + # use a white background
+  geom_hline(yintercept=0, lty=1) +
+  theme(axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+####  Meta analysis ----------------------------------------
+
+gr2 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr3 <- TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr4 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr5 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr6 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr7 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr8 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr9 <- TOSTmeta(ES = -0.01, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
[…]

### 2017-10-27 TOST_simulator.R
Peder - Created TOST simulator script
* Runs a set amount of TOST tests
* Different criterions for SESOI/bounds can be specified.
* Plot of TOST p-value vs NHST p-value is created by default.

@@ -1,0 +1,181 @@
+
+####  Setup ----------------------------------------
+
+library(TOSTER)
+library(purrr)
+source("../../TOSTER/R/TOSTtwo.R") #  Source modified TOSTtwo from pederisager Fork in order to get necessary values from TOST
+library(ggplot2)
+
+sims <- list()
+
+
+# Group 1 parameters
+mean1 <- 0    # The mean effect
+std1  <- 10   # The standard deviation
+n1    <- 10   # The number of subjects in each group
+
+# Group 2 parameters
+mean2 <- 0    # The mean effect
+std2  <- 10   # The standard deviation
+n2    <- 10   # The number of subjects in each group
+
+# TOST
+alpha <- 0.05
+tails <- 2 # 1 = one-tailed, 2 = two-tailed
+
+
+# Set simulation parameters
+
+nsims <- 10000  # Number of simulations
+bound.criterion <- "tost.current.power"  # Choose between "crit.val", "benchmark", "d33", "current.power", "tost.current.power"
+
+#  When using citerion "crit.val", the critical t value based on the parameters of the current sample are used as bounds
+crit.t <- qt(p = 1-(alpha/tails), df = n1+n2-2) # Calculate critical t
+
+#  When using criterion "benchmark", bounds = benhmark
+benchmark <- 0.5
+
+#  When using criterion "d33", small telescops is applied, group n of previous study is random 10-100, and bounds are set to effect that previous study had tlscop.pwr power to detect. 
+tlscop.pwr <- 0.33
+
+#  When using criterion "current.power", effect that current study had  pwr.thrsh power to detect are used as bounds
+pwr.thrsh <- 0.80
+crnt.pwr  <- pwr::pwr.t2n.test(n1 = n1,
+                               n2 = n2, 
+                               power = pwr.thrsh, 
+                               sig.level = alpha, 
+                               alternative = "two.sided")$d # Calculate effect size that original study have 33% power to detect
+
+# When using criterion "tost.current.power", effect is set to bounds that current study has 80% power for
+tost.pwr <- 0.80
+crnt.tost.pwr <- powerTOSTtwo(alpha = alpha, statistical_power = tost.pwr, N = n1)[2]
+
+
+
+####  Run simulation  ----------------------------------------
+
+for (sim in 1:nsims) {
+  control <- rnorm(n = n1, mean = mean1, sd = std1)
[…]

### 2017-10-27 TOST_simulator.R
Peder - Fixed path to TOSTtwo
* Modified TOSTtwo is now downloaded from github fork so script can be used independent of local filesystem.

@@ -3,7 +3,8 @@
 
 library(TOSTER)
 library(purrr)
-source("../../TOSTER/R/TOSTtwo.R") #  Source modified TOSTtwo from pederisager Fork in order to get necessary values from TOST
+download.file(url = "https://raw.githubusercontent.com/pederisager/TOSTER/TOST-modification/R/TOSTtwo.R", destfile = "TOSTtwo.prop.R")
+source(file = "TOSTtwo.prop.R") #  Source modified TOSTtwo from pederisager Fork in order to get necessary values from TOST
 library(ggplot2)
 
 sims <- list()
@@ -151,7 +152,7 @@ abline(v=0.05, lty=2)
 abline(h = 0.05, lty=2)
 
 
-#  Plot subset of sims in forestplot 
+##  Plot subset of sims in forestplot 
 # es <- diff
 # l_eqb <- map_dbl(sims, "low_eqbound")
 # u_eqb <- map_dbl(sims, "high_eqbound")

### 2017-10-30 NSNEQ_Moon_Example.Rmd
Added footnote
* Added footnote conserning the difference between using a raw effect, the Cohen's d, and the Cohen's d based on the SD of another study.

@@
 
-```{r}
+```{r echo=FALSE}
 
@@
 We find that the TOST is non-significant, t(`r round(tost$TOST_df, 2)`) = `r round(min(abs(c(tost$TOST_t1, tost$TOST_t2))), 2)`, p = `r round(max(tost$TOST_p1, tost$TOST_p2), 2)`.
-Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.
+Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote]
+
+[^footnote]: In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
+We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d (equivalence bounds = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`).
+Since this is a replication of a previous study, we could also have chosen calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
+Now we would no longer be answering a question about a raw effect independent of the samples.
+Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
+Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d, and thus the bounds, would also be only sligthly smaller (bounds = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`).
+As standard deviations of the original study compared to the replication becomes larger, the bounds become more narrow.

### 2017-11-01 H0H1_plot.R
removed tick labels on x-axis

@@ -8,7 +8,8 @@ df <- data.frame()
 
 baseplot <-   ggplot(df) +
   scale_y_continuous(limits = c(0,plotheight+0.02), breaks=NULL) + # no y-axis will be displayed
-  scale_x_continuous(limits = c(-0.85,0.85), breaks=c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8), name = "effect size (d)") +
+  scale_x_continuous(limits = c(-0.85,0.85), breaks=c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8), 
+                     labels=c("", "", "", "", "0", "", "", "", ""), name = "effect size") +
   theme_classic() +
   theme(plot.title = element_text(size = rel(1.3), face = "bold"), #font size & appearance for plot titles
         axis.title.y = element_blank(), #remove title of y-axis

### 2017-11-02 manuscript.Rmd
manuscript update

@@
 A common prediction in psychological research is that an effect exists in the population that differs from zero.
-For example, we might predict that priming American Asian women with their Asian identify will increase their performence on a math test compared to women in the control condition who are primed with their female identity.
-To be able to falsify this hypothesies, and design a study that allows for strong inferences [@platt_strong_1964], it is important to specify which test result would *disprove* the hypothesis.
+For example, we might predict that priming American Asian women with their Asian identity will increase their performance on a math test compared to women who are primed with their female identity.
+To be able to falsify this hypothesis, and design a study that allows for strong inferences [@platt_strong_1964], it is important to specify which test result would *disprove* the hypothesis.
 
 An equivalence test can be used to test whether the observed effect is surprisingly small, assuming a meaningful effect exists in the population.
-The test is a simple variation on the widely used null hypothesis significance tests.
+The test is a simple variation of the widely used null hypothesis significance tests (NHST).
 To understand the idea behind equivalence tests, it is useful to realize the null hypothesis we test against can be any numerical value.
-When we compare two groups, we often test whether the difference between groups is zero.
+When we compare two groups, we often test whether the difference between these groups is zero, but we may sometimes want to test against other values than zero.
+Imagine a researcher who is interested in voluntary participation in a national program to train young infants' motor skills.
+The researcher wants to test whether more boys than girls are brought into the program by their parents.
+One could test whether the difference between boys and girls is zero.
+However, this would ignore the fact that the human sex ratio of newborn boys to girls is not exactly 1:1, and we should not expect 50% of participants to be boys.
+On average, 103 boys are born for every 100 girls (CIA, 2017), so approximately 50.74% of applicants should be boys, and 49.26% should be girls (with 0.5074/0.4926≈1.03).
+If boys and girls were exactly equally likely to be brought into the program by their parents, we would not expect a difference of zero, but 1.5% more boys.
+This can be used as a null hypothesis: Rather than testing against a difference in proportions of 0, the researcher tests against 0.015.
 
-Imagine a researcher who who is interested in voluntary participation in a national program to train young infants' motor skills.
-The researcher wants to test whether more boys than girls are brought in to the training program by their parents.
-One could test whether the difference betwen boys and girls is zero.
-However, this would ignore that the human sex ratio of boys to girls is not exactly 1:1, or an odds ratio of 1.
-On average, slightly more boys than girls are born.
-Given that a null hypothesis of exactly zero is unlikely, the researcher tests whether the difference between boys and girls is surprising, assuming an odds ratio in the range of 0.93 to 1.07.
-If the null hypothesis can be rejected (indicating the effect is surprisingly smaller than OR 0.93, or sirprisingly larger than 1.07), the researcher decides to act as if there is a real gender difference in participation in the motor skill training program.
-This test against is known as a *minimal effect test*.
+Even though 0.015 might be a more sensible value for a null hypothesis, our researcher knows that the true sex ratio in their population will differ slightly from that in the entire world.
+Rather than testing a point null hypothesis (setting $H_0$ to one specific value), the researcher decides to define a range of values around the difference in proportions of 0.015 that can be considered trivially small, even when the true ratio in the population is not exactly 0, or even 0.015.
+The researcher can for example test if the difference is smaller than -0.005, or larger than 0.035.
+This test against two bounds, with $H_0$ as a range rather than one value, is known as a *minimal effects test* (Murphy, Myors, & Wolach, 2014).(Murphy et al.
+2014)
 
-However, even if the researcher would consider any odds ratio statistically smaller than 0.93 or larger than 1.07 a rejection of the null hypothesis, not every difference is large enough to matter.
-After an extensive discussion with experts, the researcher decides that any effect smaller than an odds ratio of 0.8 or larger than an odds ratio of 1.2 is substantial enough to move money away from the core training program, and address the gender imbalance in the voluntary participation rate.
-To test whether the gender difference is not large enough to matter, the researcher can perform an *equivalence test*, which tests whether the observed gender difference is statistically larger than an odds ratio of 0.8, *and* statistically smaller than an odds ratio of 1.2.
-If this is the case, the researcher can act as if the gender difference is too small to intervene.
+However, even if the researcher would consider any difference in percentages smaller than $-0.5 \%$ or larger than $3.5 \%$ a rejection of the null hypothesis, not every difference is large enough to matter in practice.
+As long as gender differences in the participation rate are not too extreme, there is no need to spend money to address the gender imbalance in the voluntary participation rate, and all money can be spent on the core training program.
+To test whether the gender difference in participants is *not large enough to matter, the researcher can perform an *equivalence test*.
+Equivalence tests examine whether the presence of gender differences that are large enough to matter can be rejected.
+After an extensive discussion with experts, the researcher decides that as long as the difference in proportions is not larger than 6% the researcher can act as if the gender difference is too small to intervene.
+Given an expected true difference in the population of 0.015, the researcher will tests if the observed difference in the data is smaller than 0.075, and larger than -0.055.
+If differences more extreme than both these boundary values can be rejected in two one-sided tests, the researcher will conclude statistical equivalence, and no money will be spent on addressing a gender difference in participation.
 
-One can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI; the values 0.8 and 1.2 in the previous example).
-When two one-sided tests (i.e., one testing whether the odds ratio is larger than 0.8 the other testing whether the odds ration is smaller than 1.2) are statistically significant, we can conclude statistical equivalence.
-By combining null-hypothesis tests and equivalence tests, four combinations of results can be observed (see Figure 1).
-Two outcomes are straightforward to interpret, namely when the effect is equivalent to zero, and not different from zero (an effect that is not meaningful), and when the effect is not equivalent, and different from zero (a meaningful effect).
-In addition, the effect can be different from zero and too small to be considered meaningful (an effect which is not meaningful) or the effect can neither differ from zero, not be equivalent to zero (an undetermined result, where more data is needed).
+One can conclude statistical equivalence whenever a 90% confidence interval[a][b][c][d][e][f] around the observed estimate does not contain the smallest effect size of interest (SESOI[g]).
+Figure 1 illustrates the possible conclusions we can draw from null hypothesis, minimal effect, and equivalence tests.
 
-Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017].
+Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, instead of a value of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017].
 In this article, we provide several examples of equivalence tests which illustrate the procedure.
@@
 Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI).
-The SESOI can sometimes based on just noticeable differences, which can be objectively determined.
+The SESOI can sometimes be based on just noticeable differences, which can be objectively determined.
 Most often, however, it is a subjective decision that varies across individuals and time.
 The SESOI is ideally based on a cost-benefit analysis.
-Since both costs and benefits are necessarily relative, the SESOI will depend on the researcher who designs the study.
+Since both costs and benefits are necessarily subjective, the SESOI will depend on the researcher who designs the study.
 The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base.
-Researchers should not aim to determine a SESOI that is universally valid.
-The goal is to set a SESOI such that inferences based on it answer a meaningful question.
+A SESOI should be chosen such that inferences based on it answer a meaningful question.
 
@@
 For example, Burriss and colleagues [-@burriss_changes_2015] examined whether women displayed an increase in redness in the face during the fertile phase of their ovulatory cycle.
-The hypothesis was that a slighly redder skin signals greater attractiveness and physical health, and sending this signal to men yields an evolutionary advantage.
+The hypothesis was that a slightly redder skin signals greater attractiveness and physical health, and sending this signal to men yields an evolutionary advantage.
 This hypothesis requires that the increase in redness can be detected with the naked eye by men.
 They collected data from 22 women and showed that there was indeed an increase in redness of the facial skin of woman during their fertile period.
-However, this increase was not large enough to be noticeable with the naken eye by men, thus falsifying their hypothesis.
+However, this increase was not large enough to be noticeable with the naked eye by men, thus falsifying their hypothesis.
 Because the just noticeable difference in redness of the skin can be measured, it is possible to objectively establish the SESOI.
@@
 First, researchers can use benchmarks.
-For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988).
+For example, one might set the SESOI to a standardized effect size of d = 0.5[h][i][j][k], which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988).
 Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small (Maxwell, Lau, & Howard, 2015).
@@
 
-Second, researchers can determine the SESOI based on the literature.
-Ideally, researchers would specify the SESOI in their research, but this is not yet common practice.
+Second, researchers can determine the SESOI based on previous work in the literature.
+Ideally, researchers would specify their SESOI in studies they report, but this is not yet common practice.
 It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study.
-Simonsohn (2015) recently proposed to set the SESOI to 33% of the effect size an earlier study could detect based on the sample size.
-For example, consider a study where 100 participants answered a question, which was analyzed with an one-sample t-test.
-For a two-sided test with an alpha of .05 this test had 33% power to detect a d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`.
+Simonsohn (2015) recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have 33% power to detect based on the sample size.
+This is called the 'small telescopes' approach.
+For example, consider a study where `r N` participants answered a question, which was analyzed with an one-sample t-test.
+For a two-sided test with an alpha of `r alpha` this test had 33% power to detect a d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`.
 
-Another justifiable choice we would like to propose here is to use the smallest observed effect size that could have been statistically significant in the original study as the SESOI in a replication study.
-Based only on the alpha level and the sample size, we can calculate the criticial test value (e.g., *t*, *F*, *Z*).
+Another justifiable choice for the SESOI we would like to propose here is to use the smallest observed effect size that could have been statistically significant in the original study.
+In other words, we decide that effects that could not have yielded a *p* < 0.05 in the original study (if this is your preferred alpha threshold) will not be considered meaningful in the replication either, even if effects of this size are statistically significant in the replication study.
+Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
 This critical test value can also be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*.
-All observed effect sizes smaller than the critical effect size would not be statistically significant in an original study, given the alpha and sample size.
+All observed effect sizes smaller than the critical effect size would not be statistically significant in an original study, given the alpha and sample size of that study.
 By setting the SESOI to the critical effect size, an equivalence test can reject all observed effect sizes that could have been detected in an earlier study.
@@
 
-Third, researchers can set the SESOI based on the resources they have available.
+Third[l][m], researchers can set the SESOI based on the resources they have available.
 The amount of data you can collect limits the inferences you can make.
-Given a an alpha level and a sample size, researchers can calculate the smallest effect size that can be rejected with the desired power in an equivalence test.
-For example, a researcher who plans to perform a two-sided one-sample t-test using an alpha of 5%, based on data from 100 observations, has 90% power declare equivalence when testing a SESOI of d = 0.33.
-Whether or not a test against a SESOI based on the available resources contributes to the scientific literature depends.
-One should not expect a test based on 11 observations, which provides 90% power to reject effects larger than d = 1, is particularly informative, given that most effects in psychology are substantially smaller than d = 1.
+Given a an alpha level and a sample size, researchers can calculate the smallest effect size that one has sufficient power to detect.[n] This approach is suggested here to provide researchers with a way to think about setting the SESOI for a study in a line of research where no quantifiable predictions are made, and where no previous studies have been performed.
+Observing statistical equivalence for a SESOI based on this approach allows one to reject effects that can be reliably detected with the given sample size.
+For example, a researcher who plans to perform a two-sided one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364.
+Using equivalence bounds of d = -0.364 and d = 0.364 in an equivalence test allows one to reject effects the study had high statistical power to detect.
+In this case, concluding equivalence would suggest one can reject effects that can reasonably be detected with N = 100.
+Whether or not this is interesting depends in part on how typical the chosen sample size is for a research area.
+One should not expect an equivalence test based on 15 observations testing against a SESOI of d = -1 and d = 1 to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds.
 But by transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes.
-
-<!-- It would be nice to adapt the current equivalence testing power functions, such that you can not just calculate N, but also the SESOI -->
+When there are no previous studies to build on, and no theoretical predictions, specifying a SESOI based on the alpha level and a sample size can provide a starting point for a discussion on what a reasonable SESOI is.
+Nevertheless, the value of setting a SESOI based on this procedure is limited, and the equivalence test does not reject values that are theoretically interesting.
 
@@
 
-<!-- This is just an example of a comment, and a formula -->
-$$t _ { L } = \frac { M _ { 1} - M _ { 2} - \Delta _ { L } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }, t _ { U } = \frac { M _ { 1} - M _ { 2} - \Delta _ { U } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }
-$$
+## Example 1: Statistically Equivalent and Not Statistically Different
 
-## Examples
+```{r include = FALSE}
+library(TOSTER)
+
+# original study: Banerjee et al., study 1
+orig.m1 <- 4.71
+orig.m2 <- 5.3
+orig.sd1 <- 0.85
+orig.sd2 <- 0.97
+orig.d <- 0.65
+orig.t <- 2.03
+orig.p <- 0.049
+orig.N <- 40 #group size unknown, therefore equal n assumed
+orig.df <- orig.N-2
+
+# replication study: Brandt et al., study 1
+rep.m1 <- 4.7857
+rep.m2 <- 4.6569
+rep.sd1 <- 1.0897
+rep.sd2 <- 1.1895
+rep.n1 <- 49
+rep.n2 <- 51
+rep.t <- 0.56
+rep.p <- .574
+rep.df <- rep.n1 + rep.n2 - 2
+rep.d <- 0.11
+
+# Calculate critical effect (d): smallest effect orig. study had the power to detect
+t.crit = qt(1-.05/2, (orig.N-2))
+d.crit = t.crit*sqrt((1/(orig.N/2)) + 1/(orig.N/2))
+
+Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
+```
+
+Banerjee, Chatterjee, & Sinha (2012) reported that participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
+A close replication by Brandt, IJzerman, & Blanken (2014) found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
+The smallest effect size that could have been statistically significant in the original study is $d_{crit}= `r round(t.crit, 2)` \sqrt { \frac { 1} {`r orig.N/2`} + \frac { 1} {`r orig.N/2`} }=`r round(d.crit, 2)`$.
+Using this as our SESOI for a TOST with Welch's *t*-test for independent samples --- resulting in equivalence bounds of ${\Delta}_L=`r round(-d.crit, 2)`$ and ${\Delta}_U=`r round(d.crit, 2)`$ --- we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.
+
+```{r, fig.width=6}
+TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
+```
+
+[^1]: This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect[o].
+This procedure will thus result in equivalence bounds that are substantially larger than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect[p][q][r][s][t][u][v][w].
+
+## Example 2: Statistically Equivalent and Statistically Different
+
+```{r, fig.width=6}
+library("TOSTER")
+TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = -0.01, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = 0.04, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+```
+Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1.
+For example, in grade 3 the difference is d = 0.04, with a standard error of 0.002.
+When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of d = -0.1 and d = 0.1, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1, and can be considered trivially small, according to the bounds set by the authors.
+However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge.
+This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance[x].
+
+## Example 3: Not Statistically Equivalent and Not Statistically Different
+
+Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
+They observed a slightly reversed difference between the Asian primed group (*n* = 53, *M* = 0.46, *SD* = 0.17) and the control (*n* = 48, *M* = 0.50, *SD* - 0.18) which was not significant, d = -0.23, *t*(96.62) = 1.15, *p* = 0.26, two-sided).
+A non-significant null-hypothesis test leaves two possible conclusions:
+
+1. The data suggest the absence of a meaningful effect, and we can reject effects large enough to be meaningful.
+2. The data are not sensitive enough to tell us whether a meaningful effect is present or absent.
+
+Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
+A+ = 93.75% to 100%).
+We can decide that we are only interested in test differences so far as they correspond to at least a 1 grade point increase or decrease.
+Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.
+
+Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by inserting the summary statistics from Moon & Roeder's (2014) data, assuming an alpha of 0.05, and using +/- the SESOI of 0.0625 as our equivalence bounds.[y][z][aa][ab][ac]
+
+```{r echo=FALSE}
+
+#  We first load the relevant datafile into R
+moon.data        <- read.csv("NSNEQ_Moon_Example_data.csv")
+
+#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
+mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
+sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
+sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
+n.asian          <- sum(moon.data$identity_salience == 1)
+n.control        <- sum(moon.data$identity_salience == 2)
+
+#  We can now run the TOST!
+#  Performing the TOST using the 'TOSTtwo.raw' function allows us to specify equivalence bounds in raw values
+tost <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
+                                m2  = mean.control,
+                                sd1 = sd.asian,
+                                sd2 = sd.control,
+                                n1  = n.asian,
+                                n2  = n.control,
+                                low_eqbound  = -0.0625,
+                                high_eqbound =  0.0625,
+                                alpha = .05)
+```
+
+We find that the TOST is non-significant, t(`r round(tost$TOST_df, 2)`) = `r round(min(abs(c(tost$TOST_t1, tost$TOST_t2))), 2)`, p = `r round(max(tost$TOST_p1, tost$TOST_p2), 2)`.
+Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote]
+
+[^footnote]:[ad] In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
+We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d (equivalence bounds = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`).
+Since this is a replication of a previous study, we could also have chosen calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
+Now we would no longer be answering a question about a raw effect independent of the samples.
+Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
+Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d, and thus the bounds, would also be only sligthly smaller (bounds = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`).
+As standard deviations of the original study compared to the replication becomes larger, the bounds become more narrow.
+
+## Example 4: Statistically Inferior [ae][af]and Not Statistically Different
+
+Lynott et al.
+(2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
+They observed that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same.
+Calculating a z-test for the difference between the proportions, this effect is statistically significant (Diff = -0.067, z = -1.98, p = 0.024[ag]).
+Having obtained a significant result, we are left with two alternatives:
+
+1. The difference is small enough that even though it is *statistically* significant, we can consider it *practically* equivalent to zero.
+2. The difference is substantial enough that we cannot rule out that it is also practically significant.
+
+In this case, it is not really clear what an objective cirterion for the smallest effect size of interest could be.
+However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
+This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds.
+To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
+
+```{r}
+se         <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
+crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05
+
+crit.diff <- crit.z * se  # Convert z to the critical difference
+
+print(crit.diff)
+```
+
+Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study.
+In this case, we can still use a TOST to analyse the data, but we only need to consider the test against the upper bound because an effect with CIs crossing the lower bound can still be *lower* than the original effect.
+
+```{r}
+
+download.file(url = "https://raw.githubusercontent.com/Lakens/EquivalenceTestingForPsychologicalScience/master/TOSTtwo.prop.R", destfile = "TOSTtwo.prop.R")
+source(file = "TOSTtwo.prop.R")
+
+#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
+lynott.data <- read.csv("NSEQ_Lynott_example_data.csv")
+lynott.data <- lynott.data[lynott.data$include1 == 1,]
+
+#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+n.hot            <- sum(lynott.data$packtype == 1)
+n.cold           <- sum(lynott.data$packtype == 0)
+selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
+selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
+
+#  We can now run the TOST!
+tost <- TOSTtwo.prop(prop1 = selfish.cold,
+                         prop2 = selfish.hot,
+                         n1        = n.cold,
+                         n2        = n.hot,
+                         low_eqbound  = -crit.diff,
+                         high_eqbound = crit.diff,
+                         alpha = 0.05)
+
+```
+
+We find that the test against the upper bound is `r ifelse(tost$TOST_p2 < tost$alpha, "significant", "non-significant")`, $z =$ `r round(tost$TOST_z2, 2)`, $p =$ `r ifelse(round(tost$TOST_p2, 3) >= 0.001, paste("=", round(tost$TOST_p2, 2)), "< 0.001")`.
+Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.
+## Example 5: Not Statistically Equivalent and Statistically Different
+Example 5: Correlations[ah]?
+
+Would be nice to have 1 bounds based on effect sizes you have power to detect.
+And one on just noticeable differences.
+This: http://wiki.ihe.net/images/e/e5/Measurement_of_pain.pdf suggests that  this: Berry, H., Huskisson, E.
+C.
+Clin.
+Trials J.
+1972, 9, 13.
+12. Hardy, J. D., Wolff, H. G., Goodell, H. Pain Sensations and
+Reactions. Baltimore, 1952. shows that a VAS for pain can be divided into 20 just noticeable differences' Or we use
 
@@
 
-Boundary hacking should go into the discussion.
+# Discussion
 
-#Equivalence test for Meta-analysis
+The given examples and figures XXX to YYY illustrate that equivalence testing is a relatively simple and flexible technique: As long as we can calculate a confidence interval around a mean effect, we can compare it to one or several bounds.
+In that sense, the result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman[ai] & Serlin, 1998; Tryon, 2001).
+This ease of use might tempt researchers to set equivalence bounds after looking at their data.
+For example, we showed above that Moon & Roeder's replication of the identity-priming study by Shih et al.
+(1999; example 3) does not exclude effects that are large enough to affect a student's final grade on a test.
+However, if we look at the $90 \%$ CI (figure XXX), we can see that the data would have been statistically equivalent if the bounds had been based on a slightly larger SESOI, e.g.
+a mean difference of 0.1.
+It is important to point out that this strategy --- setting equivalence bounds *based on the obtained results* --- may be considered as "boundary hacking": Regardless of the underlying true effect, any finite CI will exclude some values, which means that this procedure will lead to capitalization on chance.
 
-Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where a trivial difference is specified as an effect size smaller then d = 0.1.
-The present a table with Cohen's d and se is reproduced below:
+In order to avoid fooling ourselves more often than $\alpha \%$ of the time, equivalence bounds should be specified before results are known (ideally as part of a preregistration; cf.
+CONSORT guidelines, Piaggio et al., 2006).
+A perhaps helpful guiding question may be: "Will testing against these bounds yield interesting information in the long run?" In most cases[aj],
 
-|Grades|d + se|
-|---|---|
-|Grade 2	|0.06 +/- 0.003|
-|Grade 3	|0.04 +/- 0.002|
-|Grade 4	|-0.01 +/- 0.002|
-|Grade 5	|-0.01 +/- 0.002|
-|Grade 6	|-0.01 +/- 0.002|
-|Grade 7	|-0.02 +/- 0.002|
-|Grade 8	|-0.02 +/- 0.002|
-|Grade 9	|-0.01 +/- 0.003|
-|Grade 10	|0.04 +/- 0.003|
-|Grade 11	|0.06 +/- 0.003|
+#Raw[ak] versus standardized equivalence bounds
 
-```{r, fig.width=6}
-library("TOSTER")
-TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.05)
-```
+Equivalence bounds can be set in terms of standardized effect sizes, or raw differences.
+For example, the SESOI might be a raw mean difference of 0.5 points on a 7-point scale, or a Cohen’s d of 0.5.
+The main difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as [FORMULA (X1-X2)/SD].
+The observed standard deviation randomly varies across samples.
+In practice, this means that when you use standardized differences as bounds (e.g., d = 0.35) the equivalence test depends on the standard deviation in the sample.
+The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of d = 0.5 when the standard deviation is 1, but a standardized equivalence bound of d = 1 then the standard deviation is 0.5.
 
-We see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1.
-However, note that all of the effects are also statistically significant - so, the the effects are statistically different from zero, and practically equivalent.
-
-*Hyde, J. S., Lindberg, S. M., Linn, M. C., Ellis, A. B., & Williams, C. C. (2008). Gender similarities characterize math performance. Science, 321(5888), 494-495.*
-
-&nbsp;
-
-&nbsp;
-
-#Independent Groups Equivalence Test
-
-&nbsp;
-
-##Independent Groups Student's Equivalence Test
-
-Eskine (2013) showed that participants who had been exposed to organic food were substantially harsher in their moral judgments relative to those exposed to control (d = 0.81, 95% CI: [0.19, 1.45]).
-A replication by Moery & Calin-Jageman (2016, Study 2) did not observe a significant effect (Control: n = 95, M = 5.25, SD = 0.95, Organic Food: n = 89, M = 5.22, SD = 0.83).
-Following Simonsohn's (2015) recommendation the equivalence bound was set to the effect size the original study had 33% power to detect (with n = 21 in each condition, this means the equivalence bound is d = 0.48, which equals a difference of 0.384 on a 7-point scale given the sample sizes and a pooled standard deviation of 0.894).
-Using a TOST equivalence test with alpha = 0.05, assuming equal variances, and equivalence bounds of d = -0.43 and d = 0.43 is significant, t(182) = -2.69, p = 0.004.
-We can reject effects larger than d = 0.43.
-
-```{r, fig.width=6}
-TOSTtwo.raw(m1=5.25,m2=5.22,sd1=0.95,sd2=0.83,n1=95,n2=89,low_eqbound=-0.384, high_eqbound=0.384, alpha = 0.05, var.equal=TRUE)
-```
-
-*Moery, E., & Calin-Jageman, R.
-J.
-(2016).
-Direct and Conceptual Replications of Eskine (2013): Organic Food Exposure Has Little to No Effect on Moral Judgments and Prosocial Behavior.
-Social Psychological and Personality Science, 7(4), 312-319.
-https://doi.org/10.1177/1948550616639649 *
-
-&nbsp;
-
-##Independent Groups Welch's Equivalence Test
-
-Deary, Thorpe, Wilson, Starr, and Whally (2003) report the IQ scores of 79,376 children in Scotland (39,343 girls and 40,033 boys).
-The IQ score for girls (M = 100.64, SD = 14.1) girls and boys (100.48, SD = 14.9) was non-significant.
-With such a huge sample, we can examine whether the data is equivalent using very small equivalence bounds, such as -0.05 and 0.05.
-Because sample sizes are unequal, and variances are unequal, Welch's t-test is used, and we can conclude IQ scores are indeed equivalent, based on equivalence bounds of -0.05 and 0.05.
-
-```{r, fig.width=6}
-TOSTtwo(m1=100.64,m2=100.48,sd1=14.1,sd2=14.9,n1=39343,n2=40033,low_eqbound_d=-0.05, high_eqbound_d=0.05, alpha = 0.05, var.equal=FALSE)
-```
-
-*Deary, I. J., Thorpe, G., Wilson, V., Starr, J. M., & Whalley, L. J. (2003). Population sex differences in IQ at age 11: The Scottish mental survey 1932. Intelligence, 31(6), 533-542.*
-
-&nbsp;
-
-&nbsp;
-
-#One-Sample Equivalence Test
-
-&nbsp;
-
-Lakens, Semin, & Foroni (2012) examined whether the color of Chinese ideographs (white vs.
-black) would bias whether participants judged that the ideograph correctly translated positive, neutral, or negative words.
-In Study 4 the prediction was that participants would judge negative words to be correctly translated by black ideographs above guessing average, but no effects were predicted for translations in the other 5 between subject conditions in the 2 (ideograph color: white vs.
-black) x 3 (word meaning: positive, neutral, negative) design.
-The table below is a summary of the data in all 6 conditions:
-
-|Color	|Valence	|M	|%	|SD	|t	|df	|p	|d|
-|---|---|---|---|---|---|---|---|---|
-|White	|Positive	|6.05	|50.42	|1.50	|0.15	|20	|.89	|.03|
-|White	|Negative	|6.68	|55.67	|1.70	|1.75	|18	|.10	|.40|
-|White	|Neutral	|5.95	|49.58	|1.40	|0.87	|19	|.87	|.04|
-|Black	|Positive	|6.45	|53.75	|1.91	|1.06	|19	|.30	|.24|
-|Black	|Negative	|6.95	|57.92	|1.15	|3.71	|19	|.00	|.80|
-|Black	|Neutral	|5.71	|47.58	|1.79	|0.73	|20	|.47	|.16|
-
-With 19 to 21 participants in each between subject condition, the study had 80% power to detect equivalence with equivalence bounds of -0.68 to d = 0.68.
-
-```{r}
-powerTOSTone(alpha=0.05, statistical_power=0.8, low_eqbound_d=-0.68, high_eqbound_d=0.68)
-```
-
-When we perform 5 equivalence tests using equivalence bounds of -0.68 and 0.68, using a Holm-Bonferroni sequential procedure to control error rates, we can conclude statistical equivalence for the data in row 1, 3, and 6, but the conclusion is undetermined for tests 2 and 4, where the data is neither significantly different from guessing average, nor statistically equivalent.
-For example, performing the test for the data in row 6:
-
-```{r, fig.width=6}
-TOSTone(m=5.71,mu=6,sd=1.79,n=20,low_eqbound_d=-0.68, high_eqbound_d=0.68, alpha=0.05)
-```
-
-It is clear I should have been more tentative in my conclusions in this study.
-Not only can I not conclude equivalence in some of the conditions, the equivalence bound I had 80% power to detect is very large, meaning the possibility that there are theoretically interesting but smaller effects remains.
-
-*Lakens, D., Semin, G.
-R., & Foroni, F.
-(2012).
-But for the bad, there would not be good: Grounding valence in brightness through shared relational structures.
-Journal of Experimental Psychology: General, 141(3), 584-594.
-https://doi.org/10.1037/a0026468 *
-
-&nbsp;
-
-&nbsp;
-
-#Equivalence test for correlations
-
-&nbsp;
-
-Olson, Fazio, and Hermann (2007) reported correlations between implicit and explicit measures of self-esteem, such as the IAT, Rosenberg's self-esteem scale, a feeling thermometer, and trait ratings.
-In Study 1 71 participants completed the self-esteem measures.
-Because no equivalence bounds are mentioned, we can see which equivalence bounds the researchers would have 80% power to detect.
-
-```{r, fig.width=6}
-powerTOSTr(alpha=0.05, statistical_power=0.8, low_eqbound_r=-0.24, high_eqbound_r=0.24)
-```
-With 71 pairs of observations between measures, the researchers have 80% power for equivalence bounds of r = -0.24 and r = 0.24.
-
-The correlations observed by Olson et al (2007), Study 1, are presented in the table below (significant correlations are flagged by an asteriks).
-
-|Measure |IAT |Rosenberg|Feeling thermometer| Trait ratings|
-|---|---|---|---|---|
-|IAT |- |-.12 |-.09 |-.06|
-|Rosenberg| | - |.62*| .09|
-|Feeling thermometer| | | - |.29*|
-|Trait ratings| | | |-|
-
-We can test each correlation, for example the correlation between the IAT and the Rosenberg self-esteem scale of -0.12, given 71 participants, and using equivalence bounds of r_U = 0.24 and r_L = -0.24.
-
-```{r, fig.width=6}
-TOSTr(n=71, r=-0.06, low_eqbound_r=-0.24, high_eqbound_r=0.24, alpha=0.05)
-```
-We see that none of the correlations can be declared to be statistically equivalent.
-Instead, the tests yield undetermined outcomes: the correlations are not significantly different from 0, nor statistically equivalent.
-
-*Olson, M. A., Fazio, R. H., & Hermann, A. D. (2007). Reporting tendencies underlie discrepancies between implicit and explicit measures of self-esteem. Psychological Science, 18(4), 287-291.*
-
-# Discussion
+Both raw equivalence bounds and standardized equivalence bounds have specific benefits and limitations (for a discussion, see Baguley, 2009).
+When raw mean differences are meaningful and of theoretical interest, it makes sense to set equivalence bounds based on raw effect sizes.
+When the raw mean difference is of less theoretical importance, or different measures are used across research lines, it is often easier to set equivalence bounds based on standardized differences.
+Researchers should realize that equivalence bounds based on raw differences or standardized differences ask slightly different questions, and justify their choice for an equivalence bound.
+When setting equivalence bounds based on earlier research, such as in replication studies, equivalence bounds based on raw or standardized differences would ideally give the same result, and large differences in standard deviations between studies are as important to interpret as large differences in means.
 
@@
 
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
+Lakens, D. (2017). Equivalence Tests: A Practical Primer for t Tests, Correlations, and Meta-Analyses. Social Psychological and Personality Science, 8(4), 355–362.
 

### 2017-11-03 H0H1_plot.R
Changed font size to better match RMarkdown manuscript and added \Delta to x-axes

@@ -1,3 +1,4 @@
+####  Setup ----------------------------------------
 library(ggplot2)
 library(gridExtra)
 
@@ -6,54 +7,61 @@ lowerbound <- -0.2
 upperbound <- 0.2
 df <- data.frame()
 
+####  Base plot ------------------------------------
 baseplot <-   ggplot(df) +
   scale_y_continuous(limits = c(0,plotheight+0.02), breaks=NULL) + # no y-axis will be displayed
-  scale_x_continuous(limits = c(-0.85,0.85), breaks=c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8), 
-                     labels=c("", "", "", "", "0", "", "", "", ""), name = "effect size") +
   theme_classic() +
-  theme(plot.title = element_text(size = rel(1.3), face = "bold"), #font size & appearance for plot titles
+  theme(plot.title = element_text(size = rel(1), face = "bold"), #font size & appearance for plot titles
         axis.title.y = element_blank(), #remove title of y-axis
-        axis.title.x = element_text(size=rel(1.1), lineheight = 0.5), #font size for x-axis label
+        axis.title.x = element_text(size=rel(0.7), lineheight = 0.5), #font size for x-axis label
         plot.margin=unit(c(0.5,0.8,0.5,0.8),"cm")) #add padding around each plot to make them look nicer when combined; margin order: top, right, bottom, left
 
 #equivalence test plot
 eqplot <- baseplot +  
   ggtitle("A: Equivalence test") +
+  scale_x_continuous(limits = c(-0.85,0.85), breaks=c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8),
+                     labels=c("", "", "", bquote(Delta[L]), "0", bquote(Delta[U]), "", "", ""), name = "effect size") +
   annotate("segment", x = lowerbound, xend = lowerbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for lower bound
   annotate("segment", x = upperbound, xend = upperbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for upper bound
[…]

### 2017-11-08 corr_Kahane_Example.Rmd
Correlation example with power-based bounds: Study 4 from Kahane et al. (2015); raw data file & Rmd writeup

---
title: "Kahane et al. example: correlation, power-based bounds"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Equivalence bounds based on power

```{r include = FALSE}
library(TOSTER)
library(pwr)

## Kahane, Everett, Earp, Farias, & Savulescu (2015)
##
## Study 4, 5.2.1. ii., p. 203: "There  was  no  relationship  between  perceived wrongness in 
## personal dilemmas and in the ‘greater good’ dilemmas (r = -.04, p = .53): that is, people
## who were more ‘utilitarian’ in the personal dilemmas were not more likely to be more truly
## utilitarian in the other dilemmas, and vice versa. This lack of relationship held even when
## controlling for primary psychopathy in a partial second order correlation (r = -.07, p = .32)."

Kahane.data <- read.csv("Kahane_etal_study4_cleaned.csv", header = TRUE, sep = ",")
Kahane.data <- as.data.frame(Kahane.data)

# sample size
Kahane.N <- min(sum(!is.na(Kahane.data$ClassicSacrificialDilemmas)), sum(!is.na(Kahane.data$GreaterGoodDilemmas)))

# correlation b/w perceived wrongness in personal dilemmas and 'greater good' dilemmas
Kahane.r.GG <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$estimate
Kahane.p <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$p.value

# correlation the study had 80% power to detect
Kahane.r.80 <- (pwr.r.test(n = Kahane.N, r = NULL, sig.level = 0.05, power = 0.8, alternative = "two.sided"))$r

# TOST with bounds set to the correlation the study had 80% power to detect
Kahane <- TOSTr(n = Kahane.N, r = Kahane.r.GG, low_eqbound_r = -Kahane.r.80, high_eqbound_r = Kahane.r.80, alpha = 0.05)
```

In a paper published in 2015, Kahane and colleagues investigate moral dilemma vignettes in which participants have to decide whether or not they would sacrifice one person's life to save several other lives, or judge how morally admissible such actions are. Traditionally, greater endorsement for sacrificing a life to save others has been interpreted as a more "utilitarian" moral orientation, i.e. a stronger concern for the greater good (in total, fewer people lose their lives). In a number of studies, Kahane et al. contest this interpretation, for example by showing that greater endorsement for such "utilitarian" choices correlates with sub-clinical psychopathy, and, crucially, by comparing the traditionally used vignettes with a set of new ones that pit partial motivations against an impartial concern for the greater good (e.g. buying a new mobile phone vs. donating the money to save lives in a distant country, study 4).

Kahane et al. find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$[^footnote]). They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p. 206). 
This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had $80\%$ power to detect effects of $r = `r round(Kahane.r.80, 2)`$. Given bounds of $\Delta_{L} = `r -round(Kahane.r.80, 2)`$ and $\Delta_{U} = `r round(Kahane.r.80, 2)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(max(Kahane$TOST_p1, Kahane$TOST_p2), 3)`$. This means that values smaller than $r = `r -round(Kahane.r.80, 2)`$ and larger than $r = `r round(Kahane.r.80, 2)`$ can be rejected at an alpha level of $5\%$ and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.

[^footnote]: Study 4 in Kahane et al. (2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of only $N = `r Kahane.N`$.

### 2017-11-08 NSNEQ_Moon_Example.Rmd
Wrote short text box on TOST in R for manuscript

@@
 
-Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder's (2014) data, assuming an aplha of 0.05, and using +/- the SESOI of 0.0625 as our equivalence bounds.
+We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using +/- the SESOI of 0.0625 as our equivalence bounds.
 

### 2017-11-08 manuscript.tex
Wrote short text box on TOST in R for manuscript

@@
 that differs from zero. For example, we might predict that priming
-American Asian women with their Asian identify will increase their
-performence on a math test compared to women in the control condition
-who are primed with their female identity. To be able to falsify this
-hypothesies, and design a study that allows for strong inferences
-({\textbf{???}}), it is important to specify which test result would
-\emph{disprove} the hypothesis.
+American Asian women with their Asian identity will increase their
+performance on a math test compared to women who are primed with their
+female identity. To be able to falsify this hypothesis, and design a
+study that allows for strong inferences ({\textbf{???}}), it is
+important to specify which test result would \emph{disprove} the
+hypothesis.
 
 An equivalence test can be used to test whether the observed effect is
-surprisingly close to zero, assuming a meaningful effect exists in the
-population. The test is a simple variation on the widely used null
-hypothesis significance tests, with the small difference that instead of
-testing the observed effect against zero, it is tested against a
-smallest effect size of interest (SESOI). For example, if a difference
-on a math test of 5\% or more is considered meaningful, one could test
-against a SESOI of 5\%. If the observed effect is suprisingly closer to
-0, assuming a true effect of 5\% or larger existed, one can declare
-statistical equivalence, or reject the presence of effects that are
-large enough to be considered meaningful. An equivalence test can be
-used to falsify hypotheses.
+surprisingly small, assuming a meaningful effect exists in the
+population. The test is a simple variation of the widely used null
+hypothesis significance tests (NHST). To understand the idea behind
+equivalence tests, it is useful to realize the null hypothesis we test
+against can be any numerical value. When we compare two groups, we often
+test whether the difference between these groups is zero, but we may
+sometimes want to test against other values than zero. Imagine a
+researcher who is interested in voluntary participation in a national
+program to train young infants' motor skills. The researcher wants to
+test whether more boys than girls are brought into the program by their
+parents. One could test whether the difference between boys and girls is
+zero. However, this would ignore the fact that the human sex ratio of
+newborn boys to girls is not exactly 1:1, and we should not expect 50\%
+of participants to be boys. On average, 103 boys are born for every 100
+girls (CIA, 2017), so approximately 50.74\% of applicants should be
+boys, and 49.26\% should be girls (with 0.5074/0.4926≈1.03). If boys and
+girls were exactly equally likely to be brought into the program by
+their parents, we would not expect a difference of zero, but 1.5\% more
+boys. This can be used as a null hypothesis: Rather than testing against
+a difference in proportions of 0, the researcher tests against 0.015.
+
+Even though 0.015 might be a more sensible value for a null hypothesis,
+our researcher knows that the true sex ratio in their population will
+differ slightly from that in the entire world. Rather than testing a
+point null hypothesis (setting \(H_0\) to one specific value), the
+researcher decides to define a range of values around the difference in
+proportions of 0.015 that can be considered trivially small, even when
+the true ratio in the population is not exactly 0, or even 0.015. The
+researcher can for example test if the difference is smaller than
+-0.005, or larger than 0.035. This test against two bounds, with \(H_0\)
+as a range rather than one value, is known as a \emph{minimal effects
+test} (Murphy, Myors, \& Wolach, 2014).(Murphy et al. 2014)
+
+However, even if the researcher would consider any difference in
+percentages smaller than \(-0.5 \%\) or larger than \(3.5 \%\) a
+rejection of the null hypothesis, not every difference is large enough
+to matter in practice. As long as gender differences in the
+participation rate are not too extreme, there is no need to spend money
+to address the gender imbalance in the voluntary participation rate, and
+all money can be spent on the core training program. To test whether the
+gender difference in participants is not large enough to matter, the
+researcher can perform an \emph{equivalence test}. Equivalence tests
+examine whether the presence of gender differences that are large enough
+to matter can be rejected. After an extensive discussion with experts,
+the researcher decides that as long as the difference in proportions is
+not larger than 6\% the researcher can act as if the gender difference
+is too small to intervene. Given an expected true difference in the
+population of 0.015, the researcher will tests if the observed
+difference in the data is smaller than 0.075, and larger than -0.055. If
+differences more extreme than both these boundary values can be rejected
+in two one-sided tests, the researcher will conclude statistical
+equivalence, and no money will be spent on addressing a gender
+difference in participation.
 
 One can conclude statistical equivalence whenever a 90\% confidence
-interval around the effect size estimate does not contain the SESOI, or
-when two one-sided tests (i.e., examining whether the difference on the
-math test is more than -5\%, but less than 5\%) are statistically
-significant. By combining null-hypothesis tests and equivalence tests,
-four combinations of results can be observed (see Figure 1). Two
-outcomes are straightforward to interpret, namely when the effect is
-equivalent to zero, and not different from zero (an effect that is not
-meaningful), and when the effect is not equivalent, and different from
-zero (a meaningful effect). In addition, the effect can be different
-from zero and too small to be considered meaningful (an effect which is
-not meaningful) or the effect can neither differ from zero, not be
-equivalent to zero (an undetermined result, where more data is needed).
+interval around the observed estimate does not contain the smallest
+effect size of interest (SESOI). Figure 1 illustrates the possible
+conclusions we can draw from null hypothesis, minimal effect, and
+equivalence tests.
 
 Even though equivalence tests are just a small variation of traditional
-Frequentist tests (testing against the SESOI, instead of 0), their use
-was limited until the availability of user-friendly software to perform
-the calculations (Lakens, 2017). In this article, we discuss different
-approaches to determining the SESOI for psychological research, and
-provide detailed reproducible examples of how to perform power analyses
-when designing equivalence tests, and statistical re-analyses of
-published psychology experiments.
+frequentist null-hypothesis significance tests (and test against the
+SESOI, instead of a value of 0), their use was limited until the
+availability of user-friendly software to perform the calculations
+(Lakens, 2017). In this article, we provide several examples of
+equivalence tests which illustrate the procedure. Furthermore, we
+discuss different approaches to determining the SESOI for psychological
+research, and provide detailed reproducible examples of how to perform
+power analyses when designing equivalence tests, and statistical
+re-analyses of published psychology experiments.
 
@@
 Equivalence tests are performed against a value that is considered the
-smallest effect size of interest (SESOI). The SESOI can sometimes based
-on just noticeable differences, which can be objectively determined.
-Most often, however, it is a subjective decision that varies across
-individuals and time. Both these approaches are valid, and even when a
-just noticeable difference can be objectively determined, researchers
-might choose to set the SESOI to a larger value. In many research areas,
-the SESOI is best based on a cost-benefit analysis. Since both costs and
-benefits are necessarily relative, the SESOI will depend on the
-researcher who designs the study. The goal of setting a SESOI is to
-clearly justify why designing a study that has a high probability of
-rejecting effects larger than a specified value contributes to our
-knowledge base. Researchers should not aim to determine a SESOI that is
-universally valid. The goal is to set a SESOI such that inferences based
-on it answer a meaningful question.
+smallest effect size of interest (SESOI). The SESOI can sometimes be
+based on just noticeable differences, which can be objectively
+determined. Most often, however, it is a subjective decision that varies
+across individuals and time. The SESOI is ideally based on a
+cost-benefit analysis. Since both costs and benefits are necessarily
+subjective, the SESOI will depend on the researcher who designs the
+study. The goal of setting a SESOI is to clearly justify why designing a
+study that has a high probability of rejecting effects larger than a
+specified value contributes to our knowledge base. A SESOI should be
+chosen such that inferences based on it answer a meaningful question.
 
@@
 during the fertile phase of their ovulatory cycle. The hypothesis was
-that a slighly redder skin signals greater attractiveness and physical
+that a slightly redder skin signals greater attractiveness and physical
 health, and sending this signal to men yields an evolutionary advantage.
@@
 during their fertile period. However, this increase was not large enough
-to be noticeable with the naken eye by men, thus falsifying their
+to be noticeable with the naked eye by men, thus falsifying their
 hypothesis. Because the just noticeable difference in redness of the
@@
 
-Second, researchers can determine the SESOI based on the literature.
-Ideally, researchers would specify the SESOI in their research, but this
-is not yet common practice. It is thus up to researchers who build on
-earlier work to decide which effect size is too small to be meaningful,
-given an earlier study. Simonsohn (2015) recently proposed to set the
-SESOI to 33\% of the effect size an earlier study could detect based on
-the sample size. For example, consider a study where 100 participants
-answered a question, which was analyzed with an one-sample t-test. For a
-two-sided test with an alpha of .05 this test had 33\% power to detect a
-d = 0.15.
+Second, researchers can determine the SESOI based on previous work in
+the literature. Ideally, researchers would specify their SESOI in
+studies they report, but this is not yet common practice. It is thus up
+to researchers who build on earlier work to decide which effect size is
+too small to be meaningful, given an earlier study. Simonsohn (2015)
+recently proposed to set the SESOI to the effect size that an earlier
+study on the same effect would have 33\% power to detect based on the
+sample size. This is called the \enquote{small telescopes} approach. For
+example, consider a study where 100 participants answered a question,
+which was analyzed with an one-sample t-test. For a two-sided test with
+an alpha of 0.05 this test had 33\% power to detect a d = 0.15.
 
-Another justifiable choice we would like to propose here is to use the
-smallest observed effect size that could have been statistically
-significant in the original study as the SESOI in a replication study.
-Based only on the alpha level and the sample size, we can calculate the
-criticial test value (e.g., \emph{t}, \emph{F}, \emph{Z}). This critical
-test value can also be transformed to a standardized effect size (e.g.,
+Another justifiable choice for the SESOI we would like to propose here
+is to use the smallest observed effect size that could have been
+statistically significant in the original study. In other words, we
+decide that effects that could not have yielded a \emph{p} \textless{}
+0.05 in the original study (if this is your preferred alpha threshold)
+will not be considered meaningful in the replication either, even if
+effects of this size are statistically significant in the replication
+study. Based only on the alpha level and the sample size, we can
+calculate the critical test value (e.g., \emph{t}, \emph{F}, \emph{Z}).
+This critical test value can also be transformed to a standardized
+effect size (e.g.,
 \(d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }\)),
@@
 statistically significant in an original study, given the alpha and
-sample size. By setting the SESOI to the critical effect size, an
-equivalence test can reject all observed effect sizes that could have
-been detected in an earlier study.
+sample size of that study. By setting the SESOI to the critical effect
+size, an equivalence test can reject all observed effect sizes that
+could have been detected in an earlier study.
 
@@
 can make. Given a an alpha level and a sample size, researchers can
-calculate the smallest effect size that can be rejected with the desired
-power in an equivalence test. For example, a researcher who plans to
-perform a two-sided one-sample t-test using an alpha of 5\%, based on
-data from 100 observations, has 90\% power declare equivalence when
-testing a SESOI of d = 0.33. Whether or not a test against a SESOI based
-on the available resources contributes to the scientific literature
-depends. One should not expect a test based on 11 observations, which
-provides 90\% power to reject effects larger than d = 1, is particularly
-informative, given that most effects in psychology are substantially
-smaller than d = 1. But by transparently reporting the effects one can
-detect and reject, based on the study design, researchers can
-communicate the information their study contributes.
+calculate the smallest effect size that one has sufficient power to
+detect. This approach is suggested here to provide researchers with a
+way to think about setting the SESOI for a study in a line of research
+where no quantifiable predictions are made, and where no previous
+studies have been performed. Observing statistical equivalence for a
+SESOI based on this approach allows one to reject effects that can be
+reliably detected with the given sample size. For example, a researcher
+who plans to perform a two-sided one-sample \emph{t}-test using an alpha
+of 5\%, based on data from 100 observations, has 90\% power to detect an
+effect of d = 0.364. Using equivalence bounds of d = -0.364 and d =
+0.364 in an equivalence test allows one to reject effects the study had
+high statistical power to detect. In this case, concluding equivalence
+would suggest one can reject effects that can reasonably be detected
+with N = 100. Whether or not this is interesting depends in part on how
+typical the chosen sample size is for a research area. One should not
+expect an equivalence test based on 15 observations testing against a
+SESOI of d = -1 and d = 1 to substantially increase our knowledge, given
+that most effects in psychology are substantially smaller than these
+bounds. But by transparently reporting the effects one can detect and
+reject, based on the study design, researchers can communicate the
+information their study contributes. When there are no previous studies
+to build on, and no theoretical predictions, specifying a SESOI based on
+the alpha level and a sample size can provide a starting point for a
+discussion on what a reasonable SESOI is. Nevertheless, the value of
+setting a SESOI based on this procedure is limited, and the equivalence
+test does not reject values that are theoretically interesting.
 
-\[t _ { L } = \frac { M _ { 1} - M _ { 2} - \Delta _ { L } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }, t _ { U } = \frac { M _ { 1} - M _ { 2} - \Delta _ { U } } { \sigma \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } } }
-\]
+\subsection{Example 1: Statistically Equivalent and Not Statistically
+Different}\label{example-1-statistically-equivalent-and-not-statistically-different}
 
-\subsection{Examples}\label{examples}
-
-\section{Equivalence test for
-Meta-analysis}\label{equivalence-test-for-meta-analysis}
-
-Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect
-sizes for gender differences in mathematics tests across the 7 million
-students in the US represent trivial differences, where a trivial
-difference is specified as an effect size smaller then d = 0.1. The
-present a table with Cohen's d and se is reproduced below:
-
-\begin{longtable}[]{@{}ll@{}}
-\toprule
-Grades & d + se\tabularnewline
-\midrule
-\endhead
-Grade 2 & 0.06 +/- 0.003\tabularnewline
-Grade 3 & 0.04 +/- 0.002\tabularnewline
-Grade 4 & -0.01 +/- 0.002\tabularnewline
-Grade 5 & -0.01 +/- 0.002\tabularnewline
-Grade 6 & -0.01 +/- 0.002\tabularnewline
-Grade 7 & -0.02 +/- 0.002\tabularnewline
-Grade 8 & -0.02 +/- 0.002\tabularnewline
-Grade 9 & -0.01 +/- 0.003\tabularnewline
-Grade 10 & 0.04 +/- 0.003\tabularnewline
-Grade 11 & 0.06 +/- 0.003\tabularnewline
-\bottomrule
-\end{longtable}
+Banerjee, Chatterjee, \& Sinha (2012) reported that participants who had
+been asked to describe an unethical deed from their past judged the room
+to be darker than participants who had been asked to describe an ethical
+deed (\(M_{unethical}= 4.71\), \({SD}_{unethical}= 0.85\),
+\(M_{ethical}=5.30\), \({SD}_{ethical}=0.97\), \(t(38)= 2.03\),
+\(p= 0.05\), \(d= 0.65\)). A close replication by Brandt, IJzerman, \&
+Blanken (2014) found no significant effect (\(t(98)=0.56\), \(p=0.57\),
+\(d=0.11\)). The smallest effect size that could have been statistically
+significant in the original study is
+\(d_{crit}= 2.02 \sqrt { \frac { 1} {20} + \frac { 1} {20} }=0.64\).
+Using this as our SESOI for a TOST with Welch's \emph{t}-test for
+independent samples --- resulting in equivalence bounds of
+\({\Delta}_L=-0.64\) and \({\Delta}_U=0.64\) --- we indeed find that the
+effect reported by the replication study is statistically equivalent,
+\(t(97.78)=-2.64\), \(p=0.00\).
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
 \caption{}
@@
 \begin{verbatim}
-## Using alpha = 0.05 the meta-analysis was significant, Z = 20, p = 2.753624e-89
-## Using alpha = 0.05 the equivalence test was significant, Z = -13.33333, p = 7.406413e-41TOST results:
+## Using alpha = 0.05 Welch's t-test was non-significant, t(97.78283) = 0.5649507, p = 0.5734012
+## Using alpha = 0.05 the equivalence test based on Welch's t-test  was significant, t(97.78283) = -2.638058, p = 0.004851125TOST results:
+##   t-value 1    p-value 1 t-value 2   p-value 2       df
+## 1   3.76796 0.0001407233 -2.638058 0.004851125 97.78283
+## 
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1  -0.6401696    0.6401696
+## 
+## Equivalence bounds (raw scores):
+##   low bound raw high bound raw
+## 1    -0.7302364      0.7302364
+## 
+## TOST confidence interval:
+##   Lower Limit 90% CI raw Upper Limit 90% CI raw
+## 1              -0.249788               0.507388
+\end{verbatim}
+
+\subsection{Example 2: Statistically Equivalent and Statistically
+Different}\label{example-2-statistically-equivalent-and-statistically-different}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{}
+\end{figure}
+
+\begin{verbatim}
+## Using alpha = 0.005 the meta-analysis was significant, Z = 20, p = 5.507248e-89
+## Using alpha = 0.005 the equivalence test was significant, Z = -13.33333, p = 7.406413e-41TOST results:
 ##   Z-value 1 p-value 1 Z-value 2    p-value 2
@@
 ## TOST confidence interval:
-##   Lower Limit 90% CI Upper Limit 90% CI
-## 1         0.05506544         0.06493456
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1         0.05227251         0.06772749
 \end{verbatim}
-
-We see that indeed, all effect size estimates are measured with such
-high precision, we can conclude that they fall within the equivalence
-bound of d = -0.1 and d = 0.1. However, note that all of the effects are
-also statistically significant - so, the the effects are statistically
-different from zero, and practically equivalent.
-
-\emph{Hyde, J. S., Lindberg, S. M., Linn, M. C., Ellis, A. B., \&
-Williams, C. C. (2008). Gender similarities characterize math
-performance. Science, 321(5888), 494-495.}
-
-~
-
-~
-
-\section{Independent Groups Equivalence
-Test}\label{independent-groups-equivalence-test}
-
-~
-
-\subsection{Independent Groups Student's Equivalence
-Test}\label{independent-groups-students-equivalence-test}
-
-Eskine (2013) showed that participants who had been exposed to organic
-food were substantially harsher in their moral judgments relative to
-those exposed to control (d = 0.81, 95\% CI: {[}0.19, 1.45{]}). A
-replication by Moery \& Calin-Jageman (2016, Study 2) did not observe a
-significant effect (Control: n = 95, M = 5.25, SD = 0.95, Organic Food:
-n = 89, M = 5.22, SD = 0.83). Following Simonsohn's (2015)
-recommendation the equivalence bound was set to the effect size the
-original study had 33\% power to detect (with n = 21 in each condition,
-this means the equivalence bound is d = 0.48, which equals a difference
-of 0.384 on a 7-point scale given the sample sizes and a pooled standard
-deviation of 0.894). Using a TOST equivalence test with alpha = 0.05,
-assuming equal variances, and equivalence bounds of d = -0.43 and d =
-0.43 is significant, t(182) = -2.69, p = 0.004. We can reject effects
-larger than d = 0.43.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-2.pdf}
 \caption{}
@@
 \begin{verbatim}
-## Using alpha = 0.05 Student's t-test was non-significant, t(182) = 0.2274761, p = 0.8203089
-## Using alpha = 0.05 the equivalence test based on Student's t-test was significant, t(182) = -2.684218, p = 0.003970479TOST results:
-##   t-value 1    p-value 1 t-value 2   p-value 2  df
-## 1   3.13917 0.0009885653 -2.684218 0.003970479 182
+## Using alpha = 0.005 the meta-analysis was significant, Z = 20, p = 5.507248e-89
+## Using alpha = 0.005 the equivalence test was significant, Z = -30, p = 4.906714e-198TOST results:
+##   Z-value 1 p-value 1 Z-value 2     p-value 2
+## 1        70         0       -30 4.906714e-198
 ## 
-## Equivalence bounds (raw scores):
-##   low bound raw high bound raw
-## 1        -0.384          0.384
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1        -0.1          0.1
 ## 
 ## TOST confidence interval:
-##   Lower Limit 90% CI raw Upper Limit 90% CI raw
-## 1             -0.1880364              0.2480364
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1         0.03484834         0.04515166
 \end{verbatim}
-
-\emph{Moery, E., \& Calin-Jageman, R. J. (2016). Direct and Conceptual
-Replications of Eskine (2013): Organic Food Exposure Has Little to No
-Effect on Moral Judgments and Prosocial Behavior. Social Psychological
-and Personality Science, 7(4), 312-319.
-\url{https://doi.org/10.1177/1948550616639649} }
-
-~
-
-\subsection{Independent Groups Welch's Equivalence
-Test}\label{independent-groups-welchs-equivalence-test}
-
-Deary, Thorpe, Wilson, Starr, and Whally (2003) report the IQ scores of
-79,376 children in Scotland (39,343 girls and 40,033 boys). The IQ score
-for girls (M = 100.64, SD = 14.1) girls and boys (100.48, SD = 14.9) was
-non-significant. With such a huge sample, we can examine whether the
-data is equivalent using very small equivalence bounds, such as -0.05
-and 0.05. Because sample sizes are unequal, and variances are unequal,
-Welch's t-test is used, and we can conclude IQ scores are indeed
-equivalent, based on equivalence bounds of -0.05 and 0.05.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-3.pdf}
 \caption{}
@@
 \begin{verbatim}
-## Using alpha = 0.05 Welch's t-test was non-significant, t(79260.94) = 1.554136, p = 0.1201559
-## Using alpha = 0.05 the equivalence test based on Welch's t-test  was significant, t(79260.94) = -5.490723, p = 2.007585e-08TOST results:
-##   t-value 1    p-value 1 t-value 2    p-value 2       df
-## 1  8.598995 4.092652e-18 -5.490723 2.007585e-08 79260.94
+## Using alpha = 0.005 the meta-analysis was significant, Z = -5, p = 5.733031e-07
+## Using alpha = 0.005 the equivalence test was significant, Z = 45, p = 0TOST results:
+##   Z-value 1 p-value 1 Z-value 2 p-value 2
+## 1        45         0       -55         0
 ## 
@@
 ##   low bound d high bound d
-## 1       -0.05         0.05
-## 
-## Equivalence bounds (raw scores):
-##   low bound raw high bound raw
-## 1    -0.7252758      0.7252758
+## 1        -0.1          0.1
 ## 
 ## TOST confidence interval:
-##   Lower Limit 90% CI raw Upper Limit 90% CI raw
-## 1           -0.009341433              0.3293414
-\end{verbatim}
-
-\emph{Deary, I. J., Thorpe, G., Wilson, V., Starr, J. M., \& Whalley, L.
-J. (2003). Population sex differences in IQ at age 11: The Scottish
-mental survey 1932. Intelligence, 31(6), 533-542.}
-
-~
-
-~
-
-\section{One-Sample Equivalence Test}\label{one-sample-equivalence-test}
-
-~
-
-Lakens, Semin, \& Foroni (2012) examined whether the color of Chinese
-ideographs (white vs.~black) would bias whether participants judged that
-the ideograph correctly translated positive, neutral, or negative words.
-In Study 4 the prediction was that participants would judge negative
-words to be correctly translated by black ideographs above guessing
-average, but no effects were predicted for translations in the other 5
-between subject conditions in the 2 (ideograph color: white vs.~black) x
-3 (word meaning: positive, neutral, negative) design. The table below is
-a summary of the data in all 6 conditions:
-
-\begin{longtable}[]{@{}lllllllll@{}}
-\toprule
-Color & Valence & M & \% & SD & t & df & p & d\tabularnewline
-\midrule
-\endhead
-White & Positive & 6.05 & 50.42 & 1.50 & 0.15 & 20 & .89 &
-.03\tabularnewline
-White & Negative & 6.68 & 55.67 & 1.70 & 1.75 & 18 & .10 &
-.40\tabularnewline
-White & Neutral & 5.95 & 49.58 & 1.40 & 0.87 & 19 & .87 &
-.04\tabularnewline
-Black & Positive & 6.45 & 53.75 & 1.91 & 1.06 & 19 & .30 &
-.24\tabularnewline
-Black & Negative & 6.95 & 57.92 & 1.15 & 3.71 & 19 & .00 &
-.80\tabularnewline
-Black & Neutral & 5.71 & 47.58 & 1.79 & 0.73 & 20 & .47 &
-.16\tabularnewline
-\bottomrule
-\end{longtable}
-
-With 19 to 21 participants in each between subject condition, the study
-had 80\% power to detect equivalence with equivalence bounds of -0.68 to
-d = 0.68.
-
-\begin{verbatim}
-## The required sample size to achieve 80 % power with equivalence bounds of -0.68 and 0.68 is 19
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1        -0.01515166       -0.004848341
 \end{verbatim}
@@
 \begin{verbatim}
-## [1] 18.52043
+## Using alpha = 0.005 the meta-analysis was significant, Z = -5, p = 5.733031e-07
+## Using alpha = 0.005 the equivalence test was significant, Z = 45, p = 0TOST results:
+##   Z-value 1 p-value 1 Z-value 2 p-value 2
+## 1        45         0       -55         0
+## 
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1        -0.1          0.1
+## 
+## TOST confidence interval:
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1        -0.01515166       -0.004848341
 \end{verbatim}
 
-When we perform 5 equivalence tests using equivalence bounds of -0.68
-and 0.68, using a Holm-Bonferroni sequential procedure to control error
-rates, we can conclude statistical equivalence for the data in row 1, 3,
-and 6, but the conclusion is undetermined for tests 2 and 4, where the
-data is neither significantly different from guessing average, nor
-statistically equivalent. For example, performing the test for the data
-in row 6:
+\begin{verbatim}
+## Using alpha = 0.005 the meta-analysis was significant, Z = -5, p = 5.733031e-07
+## Using alpha = 0.005 the equivalence test was significant, Z = 45, p = 0TOST results:
+##   Z-value 1 p-value 1 Z-value 2 p-value 2
+## 1        45         0       -55         0
+## 
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1        -0.1          0.1
+## 
+## TOST confidence interval:
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1        -0.01515166       -0.004848341
+\end{verbatim}
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-4.pdf}
 \caption{}
@@
 \begin{verbatim}
-## Using alpha = 0.05 the NHST one-sample t-test was non-significant, t(19) = -0.724536, p = 0.4775649
-## Using alpha = 0.05 the equivalence test was significant, t(19) = 2.316516, p = 0.01592727TOST results:
-##   t-value 1  p-value 1 t-value 2    p-value 2 df
-## 1  2.316516 0.01592727 -3.765588 0.0006543016 19
+## Using alpha = 0.005 the meta-analysis was significant, Z = -10, p = 1.523971e-23
+## Using alpha = 0.005 the equivalence test was significant, Z = 40, p = 0TOST results:
+##   Z-value 1 p-value 1 Z-value 2 p-value 2
+## 1        40         0       -60         0
 ## 
@@
 ##   low bound d high bound d
-## 1       -0.68         0.68
-## 
-## Equivalence bounds (raw scores):
-##   low bound raw high bound raw
-## 1       -1.2172         1.2172
+## 1        -0.1          0.1
 ## 
 ## TOST confidence interval:
-##   Lower Limit 90% CI raw Upper Limit 90% CI raw
-## 1             -0.9820961              0.4020961
-\end{verbatim}
-
-It is clear I should have been more tentative in my conclusions in this
-study. Not only can I not conclude equivalence in some of the
-conditions, the equivalence bound I had 80\% power to detect is very
-large, meaning the possibility that there are theoretically interesting
-but smaller effects remains.
-
-\emph{Lakens, D., Semin, G. R., \& Foroni, F. (2012). But for the bad,
-there would not be good: Grounding valence in brightness through shared
-relational structures. Journal of Experimental Psychology: General,
-141(3), 584-594. \url{https://doi.org/10.1037/a0026468} }
-
-~
-
-~
-
-\section{Equivalence test for
-correlations}\label{equivalence-test-for-correlations}
-
-~
-
-Olson, Fazio, and Hermann (2007) reported correlations between implicit
-and explicit measures of self-esteem, such as the IAT, Rosenberg's
-self-esteem scale, a feeling thermometer, and trait ratings. In Study 1
-71 participants completed the self-esteem measures. Because no
-equivalence bounds are mentioned, we can see which equivalence bounds
-the researchers would have 80\% power to detect.
-
-\begin{verbatim}
-## The required sample size to achieve 80 % power with equivalence bounds of -0.24 and 0.24 is 71 pairs
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1        -0.02515166        -0.01484834
 \end{verbatim}
@@
 \begin{verbatim}
-## [1] 70.05703
+## Using alpha = 0.005 the meta-analysis was significant, Z = -10, p = 1.523971e-23
+## Using alpha = 0.005 the equivalence test was significant, Z = 40, p = 0TOST results:
+##   Z-value 1 p-value 1 Z-value 2 p-value 2
+## 1        40         0       -60         0
+## 
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1        -0.1          0.1
+## 
+## TOST confidence interval:
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1        -0.02515166        -0.01484834
 \end{verbatim}
-
-With 71 pairs of observations between measures, the researchers have
-80\% power for equivalence bounds of r = -0.24 and r = 0.24.
-
-The correlations observed by Olson et al (2007), Study 1, are presented
-in the table below (significant correlations are flagged by an
-asteriks).
-
-\begin{longtable}[]{@{}lllll@{}}
-\toprule
-Measure & IAT & Rosenberg & Feeling thermometer & Trait
-ratings\tabularnewline
-\midrule
-\endhead
-IAT & - & -.12 & -.09 & -.06\tabularnewline
-Rosenberg & & - & .62* & .09\tabularnewline
-Feeling thermometer & & & - & .29*\tabularnewline
-Trait ratings & & & & -\tabularnewline
-\bottomrule
-\end{longtable}
-
-We can test each correlation, for example the correlation between the
-IAT and the Rosenberg self-esteem scale of -0.12, given 71 participants,
-and using equivalence bounds of r\_U = 0.24 and r\_L = -0.24.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-11-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-5.pdf}
 \caption{}
@@
 \begin{verbatim}
-## Using alpha = 0.05 the NHST t-test was non-significant, p = 0.6191585
-## Using alpha = 0.05 the equivalence test was non-significant, p = 0.06386793TOST results:
-##     p-value 1  p-value 2
-## 1 0.005971455 0.06386793
+## Using alpha = 0.005 the meta-analysis was significant, Z = -3.333333, p = 0.0008581207
+## Using alpha = 0.005 the equivalence test was significant, Z = 30, p = 4.906714e-198TOST results:
+##   Z-value 1     p-value 1 Z-value 2     p-value 2
+## 1        30 4.906714e-198 -36.66667 1.241408e-294
 ## 
-## Equivalence bounds (r):
-##   low bound r high bound r
-## 1       -0.24         0.24
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1        -0.1          0.1
+## 
+## TOST confidence interval:
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1        -0.01772749       -0.002272512
+\end{verbatim}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-6.pdf}
+\caption{}
+\end{figure}
+
+\begin{verbatim}
+## Using alpha = 0.005 the meta-analysis was significant, Z = 13.33333, p = 1.481283e-40
+## Using alpha = 0.005 the equivalence test was significant, Z = -20, p = 2.753624e-89TOST results:
+##   Z-value 1 p-value 1 Z-value 2    p-value 2
+## 1  46.66667         0       -20 2.753624e-89
+## 
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1        -0.1          0.1
+## 
+## TOST confidence interval:
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1         0.03227251         0.04772749
+\end{verbatim}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-7.pdf}
+\caption{}
+\end{figure}
+
+\begin{verbatim}
+## Using alpha = 0.005 the meta-analysis was significant, Z = 20, p = 5.507248e-89
+## Using alpha = 0.005 the equivalence test was significant, Z = -13.33333, p = 7.406413e-41TOST results:
+##   Z-value 1 p-value 1 Z-value 2    p-value 2
+## 1  53.33333         0 -13.33333 7.406413e-41
+## 
+## Equivalence bounds (Cohen's d):
+##   low bound d high bound d
+## 1        -0.1          0.1
+## 
+## TOST confidence interval:
+##   Lower Limit 99% CI Upper Limit 99% CI
+## 1         0.05227251         0.06772749
+\end{verbatim}
+
+Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect
+sizes for gender differences in mathematics tests across the 7 million
+students in the US represent trivial differences, where the authors
+specify a trivial difference as an effect size smaller than d = 0.1. For
+example, in grade 3 the difference is d = 0.04, with a standard error of
+0.002. When we perform equivalence tests on the meta-analytic effect
+sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005
+to correct for multiple comparisons), using equivalence bounds of d =
+-0.1 and d = 0.1, we see that indeed, all effect size estimates are
+measured with such high precision, we can conclude that they fall within
+the equivalence bound of d = -0.1 and d = 0.1, and can be considered
+trivially small, according to the bounds set by the authors. However,
+note that all of the effects are also statistically different from zero,
+as one might expect when there is no random assignment to conditions,
+and samples sizes are huge. This is one way in which the use of
+equivalence tests improve hypothesis testing procedures in science, by
+distinguishing between statistical significance and practical
+significance.
+
+\subsection{Example 3: Not Statistically Equivalent and Not
+Statistically
+Different}\label{example-3-not-statistically-equivalent-and-not-statistically-different}
+
+Moon \& Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999),
+conducted a study to investigate whether Asian-American women would
+perform better on a maths test when primed with their Asian identity.
+They observed a slightly reversed difference between the Asian primed
+group (\emph{n} = 53, \emph{M} = 0.46, \emph{SD} = 0.17) and the control
+(\emph{n} = 48, \emph{M} = 0.50, \emph{SD} - 0.18) which was not
+significant, d = -0.23, \emph{t}(96.62) = 1.15, \emph{p} = 0.26,
+two-sided). A non-significant null-hypothesis test leaves two possible
+conclusions:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  The data suggest the absence of a meaningful effect, and we can reject
+  effects large enough to be meaningful.
+\item
+  The data are not sensitive enough to tell us whether a meaningful
+  effect is present or absent.
+\end{enumerate}
+
+Let us imagine that we consult a teacher about the test, and discover
+that that grades for this test are set at every 6.25\% increase in
+correct answers (F = 0\% to 6.25\% \ldots{} A+ = 93.75\% to 100\%). We
+can decide that we are only interested in test differences so far as
+they correspond to at least a 1 grade point increase or decrease. Thus,
+our SESOI becomes a difference in raw scores of 6.25\%, or 0.0625.
+
+Using the TOSTtwo function in the TOSTER package in R, we can then
+calculate the TOST by inserting the summary statistics from Moon \&
+Roeder's (2014) data, assuming an alpha of 0.05, and using +/- the SESOI
+of 0.0625 as our equivalence bounds.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{}
+\end{figure}
+
+\begin{verbatim}
+## Using alpha = 0.05 Welch's t-test was non-significant, t(97.76702) = -1.07647, p = 0.2843664
+## Using alpha = 0.05 the equivalence test based on Welch's t-test  was non-significant, t(97.76702) = 0.7054677, p = 0.2410984TOST results:
+##   t-value 1 p-value 1 t-value 2   p-value 2       df
+## 1 0.7054677 0.2410984 -2.858407 0.002601883 97.76702
+## 
+## Equivalence bounds (raw scores):
+##   low bound raw high bound raw
+## 1       -0.0625         0.0625
 ## 
@@
 ##   Lower Limit 90% CI raw Upper Limit 90% CI raw
-## 1             -0.2538652              0.1384997
+## 1             -0.0960001              0.0204875
 \end{verbatim}
 
-We see that none of the correlations can be declared to be statistically
-equivalent. Instead, the tests yield undetermined outcomes: the
-correlations are not significantly different from 0, nor statistically
-equivalent.
+We find that the TOST is non-significant, t(97.77) = 0.71, p = 0.24.
+Thus, we cannot reject the possibility that the true effect really is
+larger than our SESOI.\footnote{In this example, the bounds of the TOST
+  was set based on a raw effect of 6.25\%. We could have converted this
+  raw effect into a standardized effect (Cohen's d) for the current
+  sample, and we would have gotten the exact same result, only then in
+  Cohen's d (equivalence bounds = +/- 0.36). Since this is a replication
+  of a previous study, we could also have chosen calculate the Cohen's d
+  that corresponds to the raw SESOI in the \emph{original} study, and
+  used this as the bounds for the replication. Now we would no longer be
+  answering a question about a raw effect independent of the samples.
+  Instead, we are asking whether we in the replication can reject
+  standardized effects larger than a certain standardized effect in the
+  original study. Since the standard deviations of the original study
+  were only slightly higher than in the replication in this case,
+  Cohen's d, and thus the bounds, would also be only sligthly smaller
+  (bounds = +/- 0.34). As standard deviations of the original study
+  compared to the replication becomes larger, the bounds become more
+  narrow.}
 
-\emph{Olson, M. A., Fazio, R. H., \& Hermann, A. D. (2007). Reporting
-tendencies underlie discrepancies between implicit and explicit measures
-of self-esteem. Psychological Science, 18(4), 287-291.}
+\subsection{Example 4: Statistically Inferior and Not Statistically
+Different}\label{example-4-statistically-inferior-and-not-statistically-different}
+
+Lynott et al. (2014) conducted a study to investigate the effect of
+being exposed to physically warm or cold stimuli on subsequent judgments
+related to interpersonal warmth and prosocial behavior (replicating
+Williams and Bargh, 2008). They observed that 51.2\% of participants who
+received a cold pack (n = 427) opted to receive a reward for themselves,
+while 57.9\% (n = 434) of participants who received a warm pack did the
+same. Calculating a z-test for the difference between the proportions,
+this effect is statistically significant (Diff = -0.067, z = -1.98, p =
+0.024). Having obtained a significant result, we are left with two
+alternatives:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  The difference is small enough that even though it is
+  \emph{statistically} significant, we can consider it
+  \emph{practically} equivalent to zero.
+\item
+  The difference is substantial enough that we cannot rule out that it
+  is also practically significant.
+\end{enumerate}
+
+In this case, it is not really clear what an objective cirterion for the
+smallest effect size of interest could be. However, since this is a
+replication, we could decide that we do not care about an effect if the
+difference is smaller than the smallest effect that the original study
+could detect. This essentially means that we use the critical z value
+(\textasciitilde{}1.96 in a two-tailed test with an alpha of 0.05) as
+our bounds. To figure out what difference corresponds to a critical z in
+the original study, we multiply critical z with the standard error:
+
+\begin{verbatim}
+## [1] 0.2510861
+\end{verbatim}
+
+Having determined the SESOI, we may also decide that since this is a
+replication, we are only interested in whether the effect of receiving a
+cold pack is significantly \emph{lower} than the smallest detectable
+effect of the original study. In this case, we can still use a TOST to
+analyse the data, but we only need to consider the test against the
+upper bound because an effect with CIs crossing the lower bound can
+still be \emph{lower} than the original effect.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
+\caption{}
+\end{figure}
+
+\begin{verbatim}
+## Using alpha = 0.05 Fishers exact z-test was non-significant, z = -1.925297, p = 0.05419224
+## Using alpha = 0.05 the equivalence test based on Fishers exact z-test was significant, z = 5.274128, p = 6.669446e-08
+## TOST results:
+##   z-value 1    p-value 1 z-value 2    p-value 2
+## 1  5.274128 6.669446e-08 -9.124721 3.596051e-20
+## 
+## Equivalence bounds:
+##    low bound high bound
+## 1 -0.2510861  0.2510861
+## 
+## TOST confidence interval:
+##   Lower Limit 90% CI Upper Limit 90% CI
+## 1         -0.1245121       -0.009780696
+\end{verbatim}
+
+We find that the test against the upper bound is significant, \(z =\)
+-9.12, \(p =\) \textless{} 0.001. Thus, we can conclude that the
+statistically non-significant effect is also statistically inferior to
+our SESOI. \#\# Example 5: Not Statistically Equivalent and
+Statistically Different Example 5: Correlations?
+
+Would be nice to have 1 bounds based on effect sizes you have power to
+detect. And one on just noticeable differences. This:
+\url{http://wiki.ihe.net/images/e/e5/Measurement_of_pain.pdf} suggests
+that this: Berry, H., Huskisson, E. C. Clin. Trials J. 1972, 9, 13. 12.
+Hardy, J. D., Wolff, H. G., Goodell, H. Pain Sensations and Reactions.
+Baltimore, 1952. shows that a VAS for pain can be divided into 20 just
+noticeable differences' Or we use
+
+\section{Adressing Trello cards}\label{adressing-trello-cards}
+
+For the example where the results are inconclusive, explain how adding
+more data, will, assuming the true effect is 0, lead to a smaller
+confidence interval, and thus will lead to an informative test. However,
+with increasing data, the effect could also slightly increase, to above
+or at the SESOI level, and then the effect will be significant and not
+equivalent. With massive amounts of data, many results will be
+significant and equivalent.
 
 \section{Discussion}\label{discussion}
+
+The given examples and figures XXX to YYY illustrate that equivalence
+testing is a relatively simple and flexible technique: As long as we can
+calculate a confidence interval around a mean effect, we can compare it
+to one or several bounds. In that sense, the result of an equivalence
+test can be obtained by mere visual inspection of the CI (Seaman \&
+Serlin, 1998; Tryon, 2001). This ease of use might tempt researchers to
+set equivalence bounds after looking at their data. For example, we
+showed above that Moon \& Roeder's replication of the identity-priming
+study by Shih et al. (1999; example 3) does not exclude effects that are
+large enough to affect a student's final grade on a test. However, if we
+look at the \(90 \%\) CI (figure XXX), we can see that the data would
+have been statistically equivalent if the bounds had been based on a
+slightly larger SESOI, e.g.~a mean difference of 0.1. It is important to
+point out that this strategy --- setting equivalence bounds \emph{based
+on the obtained results} --- may be considered as \enquote{boundary
+hacking}: Regardless of the underlying true effect, any finite CI will
+exclude some values, which means that this procedure will lead to
+capitalization on chance.
+
+In order to avoid fooling ourselves more often than \(\alpha \%\) of the
+time, equivalence bounds should be specified before results are known
+(ideally as part of a preregistration; cf.~CONSORT guidelines, Piaggio
+et al., 2006). A perhaps helpful guiding question may be: \enquote{Will
+testing against these bounds yield interesting information in the long
+run?} In most cases,
+
+\section{Raw versus standardized equivalence
+bounds}\label{raw-versus-standardized-equivalence-bounds}
+
+Equivalence bounds can be set in terms of standardized effect sizes, or
+raw differences. For example, the SESOI might be a raw mean difference
+of 0.5 points on a 7-point scale, or a Cohen's d of 0.5. The main
+difference is that equivalence bounds set in raw differences are
+independent of the standard deviation, while equivalence bounds set as
+standardized effects are dependent on the standard deviation (since they
+are calculated as {[}FORMULA (X1-X2)/SD{]}. The observed standard
+deviation randomly varies across samples. In practice, this means that
+when you use standardized differences as bounds (e.g., d = 0.35) the
+equivalence test depends on the standard deviation in the sample. The
+equivalence bound for a raw mean difference of 0.5 equals a standardized
+equivalence bound of d = 0.5 when the standard deviation is 1, but a
+standardized equivalence bound of d = 1 then the standard deviation is
+0.5.
+
+Both raw equivalence bounds and standardized equivalence bounds have
+specific benefits and limitations (for a discussion, see Baguley, 2009).
+When raw mean differences are meaningful and of theoretical interest, it
+makes sense to set equivalence bounds based on raw effect sizes. When
+the raw mean difference is of less theoretical importance, or different
+measures are used across research lines, it is often easier to set
+equivalence bounds based on standardized differences. Researchers should
+realize that equivalence bounds based on raw differences or standardized
+differences ask slightly different questions, and justify their choice
+for an equivalence bound. When setting equivalence bounds based on
+earlier research, such as in replication studies, equivalence bounds
+based on raw or standardized differences would ideally give the same
+result, and large differences in standard deviations between studies are
+as important to interpret as large differences in means.
 
@@
 
-\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}
+Lakens, D. (2017). Equivalence Tests: A Practical Primer for t Tests,
+Correlations, and Meta-Analyses. Social Psychological and Personality
+Science, 8(4), 355--362.
 

### 2017-11-08 NSEQ_Brandt_Example.Rmd

@@
 library(TOSTER)
+library(pwr)
 
@@
 
-# Calculate critical effect (d): smallest effect orig. study had the power to detect
-t.crit = qt(1-.05/2, (orig.N-2))
-d.crit = t.crit*sqrt((1/(orig.N/2)) + 1/(orig.N/2))
+# Calculate small telescopes effect (d) effect orig. study had 33% power to detect
+d.33 <- (pwr.t.test(n = orig.N/2, d = NULL, sig.level = 0.05, power = 0.33, type = "two.sample", alternative = "two.sided"))$d
 
-Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
+Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.33, high_eqbound_d = d.33, var.equal = FALSE)
 ```
@@
 
-As an example, Banerjee, Chatterjee, & Sinha (2012) reported that participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
+Banerjee, Chatterjee, & Sinha (2012) reported that participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
 A close replication by Brandt, IJzerman, & Blanken (2014) found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
-The smallest effect the original study could have detected is $d_{crit}= `r round(t.crit, 2)` \sqrt { \frac { 1} {`r orig.N/2`} + \frac { 1} {`r orig.N/2`} }=`r round(d.crit, 2)`$.
-Using this as our SESOI for a TOST with Welch's t-test for independent samples --- resulting in equivalence bounds of ${\Delta}_L=`r round(-d.crit, 2)`$ and ${\Delta}_U=`r round(d.crit, 2)`$ --- we indeed find that the effect reported by the replication study is equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.
-
-```{r, fig.width=6}
-TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
-```
-
-[^1]: This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect.
-This procedure will thus result in effect sizes that are substantially larger than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect.
+Following the small telescopes approach, we can calculate the effect size the original study had $33\%$ power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.
+If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.

### 2017-11-08 Dienes_example1.R
Mass deletion
* Deleted a large amount of redundant files, including html and pdf files generated by Rmd files

@@ -1,44 +1,0 @@
-# TOST on the values from Gibson, Losee, and Vitello example
-
-Gibson.m1 <- 0.59
-Gibson.m2 <- 0.53
-Gibson.sd1 <- 0.23
-Gibson.sd2 <- 0.23
-Gibson.n1 <- 52
-Gibson.n2 <- 54
-
-Shih.m1 <- 0.54
-Shih.n2 <- 0.43
-Shih.sd1 <- 0.17
-Shih.sd2 <- 0.16
-Shih.n1 <- 16
-Shih.n2 <- 14
-
-### Calculate subjective bound critera from original study data ###
-
-# Original had 33% Power to detect effect of d=0.57:
-library(pwr)
-pwr.t.test(power=0.33,sig.level=0.05,n=15,type="two.sample",alternative="two.sided")
-
-# Original study smallest detectable effect size
-# Critical t for original study was:
-crit_t = qt(1-.05/2, 29)
-# So d becomes:
-crit_d = crit_t*sqrt(1/16 + 1/14)
-
-### TOST for the different bounds ###
-
-# TOST for 33% Power rule:
-library(devtools)
-install_github("Lakens/TOSTER", force = TRUE)
-TOSTtwo(m1 = Gibson.m1, m2 = Gibson.m2, sd1 = Gibson.sd1, sd2 = Gibson.sd2, n1 = Gibson.n1, n2 = Gibson.n2, low_eqbound_d = -0.57, high_eqbound_d = 0.57, alpha = .05)
-
-# TOST for smallest detectable effect in original study
-TOSTtwo(m1 = 0.59, m2 = 0.53, sd1 = 0.23, sd2 = 0.23, n1 = 52, n2 = 54, low_eqbound_d = -crit_d, high_eqbound_d = crit_d, alpha = .05)
-
-#How much power did the replication have? 
-#For 33% power value
-powerTOSTtwo(alpha=0.05, N = Gibson.n1, low_eqbound_d=-0.57, high_eqbound_d=0.57)
-
-#For crit_d (note this will almost always be less strict than 33% threshold)
-powerTOSTtwo(alpha=0.05, N = Gibson.n1, low_eqbound_d=-crit_d, high_eqbound_d=crit_d)

### 2017-11-08 Dienes_example3.R

@@ -1,372 +1,0 @@
-###==========================================================================###
-library(TOSTER)
-library(pwr)
-###==========================================================================###
-
-
-## Dienes & McLatchie (2017) example 3: grounded cognition 
-## (Brandt et al. replicating Banerjee et al.)
-
-
-####################################################
-# Study 1
-####################################################
-
-# original study: Banerjee et al., study 1
-orig.m1 <- 4.71
-orig.m2 <- 5.3
-orig.sd1 <- 0.85
-orig.sd2 <- 0.97
-orig.d <- 0.65
-orig.t <- 2.03
-orig.N <- 40
-orig.n1 <- orig.N/2 #group size unknown, therefore equal n assumed
-orig.n2 <- orig.n1 #group size unknown, therefore equal n assumed
-orig.raw <- abs(orig.m2-orig.m1)
-orig.sdpooled <- sqrt((orig.sd1^2 + orig.sd2^2)/(orig.N-2))
-orig.se <- sqrt((orig.sd1^2/orig.n1)+(orig.sd2^2/orig.n2)) #standard error of *differences*
-orig.df <- orig.N-2
-
-# replication study: Brandt et al., study 1
-rep.m1 <- 4.79
-rep.m2 <- 4.66
-rep.sd1 <- 1.09
-rep.sd2 <- 1.19
-rep.d <- -0.11
-rep.n1 <- 50 # this is a guess, the paper only reports total N = 120
-rep.n2 <- 50 # see previous comment
-rep.N <- 120
-rep.raw <- abs(rep.m2-rep.m1)
-rep.sdpooled <- sqrt((rep.sd1^2 + rep.sd2^2)/(rep.N-2))
-rep.se <- sqrt((rep.sd1^2/rep.n1)+(rep.sd2^2/rep.n2)) #standard error of *differences*
-rep.df <- rep.N-2
-
-# Calculate critical effect (d and raw): smallest effect orig. study had the power to detect
-t.crit = qt(1-.05/2, (orig.N-2))
-d.crit = t.crit*sqrt((1/(orig.N/2)) + 1/(orig.N/2))
-raw.crit <- d.crit * orig.sdpooled # smallest raw effect orig. study could detect
-
-# Small telescopes: Calculate d for which orig. study had 33% power
-power <- pwr.t.test(n = orig.N/2, d = NULL, sig.level = 0.05, power = 0.33, type = "two.sample", alternative = "two.sided")
[…]

### 2017-11-08 Dienes_example5.R

@@ -1,453 +1,0 @@
-###==========================================================================###
-################ Daniel: TOSTtwo + BFs from standardised effects ###############
-###==========================================================================###
-
-#' TOST function for an independent t-test (Cohen's d)
-#' @param m1 mean of group 1
-#' @param m2 mean of group 2
-#' @param sd1 standard deviation of group 1
-#' @param sd2 standard deviation of group 2
-#' @param n1 sample size in group 1
-#' @param n2 sample size in group 2
-#' @param low_eqbound_d lower equivalence bounds (e.g., -0.5) expressed in standardized mean difference (Cohen's d)
-#' @param high_eqbound_d upper equivalence bounds (e.g., 0.5) expressed in standardized mean difference (Cohen's d)
-#' @param alpha alpha level (default = 0.05)
-#' @param var.equal logical variable indicating whether equal variances assumption is assumed to be TRUE or FALSE.  Defaults to FALSE.
-#' @return Returns TOST t-value 1, TOST p-value 1, TOST t-value 2, TOST p-value 2, degrees of freedom, low equivalence bound, high equivalence bound, low equivalence bound in Cohen's d, high equivalence bound in Cohen's d, Lower limit confidence interval TOST, Upper limit confidence interval TOST
-#' @importFrom stats pnorm pt qnorm qt
-#' @importFrom graphics abline plot points segments title
-#' @examples
-#' ## Eskine (2013) showed that participants who had been exposed to organic
-#' ## food were substantially harsher in their moral judgments relative to
-#' ## those exposed to control (d = 0.81, 95% CI: [0.19, 1.45]). A
-#' ## replication by Moery & Calin-Jageman (2016, Study 2) did not observe
[…]

### 2017-11-08 Example_Moon_Shih.R
Minor changes to multiple files

@@ -4,10 +4,82 @@
 
 library(TOSTER)
 library(pwr)
-source("Example from Neil Mclathie/2017.10.05 TOSTtwoBayes Raw with Cauchy+HalfCauchy.R")
-source("TOSTtwoBayes.R")
-source("tFromDescTwo.R")
-source("eqboundCritera.R")
+#source("Example from Neil Mclathie/2017.10.05 TOSTtwoBayes Raw with Cauchy+HalfCauchy.R")
+#source("TOSTtwoBayes.R")
+
+#### Create functions for calculating statistics ------------------------------
+
+# Create function for calculating t value, cohens d, pooled standard error, etc. from means, standard deviations and group n's.
+
+tFromDescTwo <- function(m1, m2, sd1, sd2, n1, n2, test = "welch") {
+  
+  output        <- list()       # Define the output as list
+  output$m1     <- m1           # Store input arguments in output
+  output$m2     <- m2
+  output$sd1    <- sd1
+  output$sd2    <- sd2
+  output$n1     <- n1
+  output$n2     <- n2
+  output$m.diff <- m1 - m2      # Calculate the mean difference
+  
+  
+  if (test == "student") {  
+    # Calculate se.diff, t and d based on a student t.test
+    print("student's t.test is performed and sample sizes are assumed to be equal")
+    output$test       <- "student"
+    output$df         <- n1 + n2 - 2  # Calculate degrees of freedom.
+    output$sd.pooled  <- sqrt(((( n1 - 1) * (sd1^2)) + (n2 - 1) * (sd2^2)) / ((n1 + n2) - 2 )) # Calculate the pooled SD
+    output$se.diff    <- output$sd.pooled * sqrt(1/n1 + 1/n2) # Calculate the SE of the difference.
+    
+  } else if (test == "welch") {  
+    # Calculate se.diff, t and d based on a welch t.test, and calculate sd.pooled
+    print("welch's t.test is performed and sample sizes are assumed to be unequal")
+    output$test       <- "welch"
+    output$df         <- (sd1^2 / n1 + sd2^2 /n2)^2 / (((sd1^2 / n1)^2 / (n1-1)) + ((sd2^2 / n2)^2 / (n2 - 1))) #degrees of freedom for Welch's t-test
+    output$sd.pooled  <- sqrt((sd1^2 + sd2^2) / 2)    # Calculate sd root mean squared for Welch's t-test
[…]

### 2017-11-08 TOST_simulator.R

@@ -13,12 +13,12 @@ sims <- list()
 # Group 1 parameters
 mean1 <- 0    # The mean effect
 std1  <- 10   # The standard deviation
-n1    <- 10   # The number of subjects in each group
+# make n random... n1    <- 10   # The number of subjects in each group
 
 # Group 2 parameters
 mean2 <- 0    # The mean effect
 std2  <- 10   # The standard deviation
-n2    <- 10   # The number of subjects in each group
+# make n random... n2    <- 10   # The number of subjects in each group
 
 # TOST
 alpha <- 0.05
@@ -28,13 +28,16 @@ tails <- 2 # 1 = one-tailed, 2 = two-tailed
 # Set simulation parameters
 
 nsims <- 10000  # Number of simulations
-bound.criterion <- "tost.current.power"  # Choose between "crit.val", "benchmark", "d33", "current.power", "tost.current.power"
+bound.criterion <- "specific.raw"  # Choose between "crit.val", "specific", "d33", "current.power", "tost.current.power"
 
 #  When using citerion "crit.val", the critical t value based on the parameters of the current sample are used as bounds
 crit.t <- qt(p = 1-(alpha/tails), df = n1+n2-2) # Calculate critical t
 
-#  When using criterion "benchmark", bounds = benhmark
-benchmark <- 0.5
+#  When using criterion "specific", bounds = specific
+specific <- 0.9
+
+#  When using criterion "specific.raw", bounds = specific.raw, recalculated to d for each test
+specific.raw <- 3
 
 #  When using criterion "d33", small telescops is applied, group n of previous study is random 10-100, and bounds are set to effect that previous study had tlscop.pwr power to detect. 
 tlscop.pwr <- 0.33
@@ -79,7 +82,7 @@ for (sim in 1:nsims) {
                            var.equal = ifelse(test = std1 == std2, yes = TRUE, no = FALSE), #  sets equal variance assumption based on whether the standard deviation in the populations match.
                            side.effects = FALSE)
     
-  } else if (bound.criterion == "benchmark") {  # Run TOST with benchmark value
+  } else if (bound.criterion == "specific") {  # Run TOST with specific value
     
     sims[[sim]] <- TOSTtwo(m1 = m1, 
                            m2 = m2, 
@@ -87,8 +90,24 @@ for (sim in 1:nsims) {
                            sd2 = sd2, 
                            n1 = n1, 
                            n2 = n2, 
-                           low_eqbound_d = -benchmark, 
[…]

### 2017-11-08 eqboundCritera.R

@@ -2,7 +2,7 @@
   # Calculates the effect size (Cohens d) that a study has % power (33% by default) to detect.
   # Calculates the the smallest detectable effect size of a study based on the degrees of freedom and p value (0.05 by default).
 
-eqboundCritera <- function(n1, n2, df, p = 0.05, power = 0.33, testtype = "two.sided", var.equal = FALSE) {
+eqboundCritera <- function(n1, n2, orig.df, p = 0.05, power = 0.33, testtype = "two.sided", var.equal = FALSE) {
   output <- list()  # Define output as list
   
   # Calculate d that study has 33% chance 
@@ -12,14 +12,13 @@ eqboundCritera <- function(n1, n2, df, p = 0.05, power = 0.33, testtype = "two.s
                                      sig.level   = p, 
                                      alternative = testtype)$d # Calculate effect size that study have 33% power to detect
   
-  output$dunb33 <- ( 1 - ( 3 / (4*df - 1) ) ) * output$d33     # Calculate d unbiased for the 33% power criterion
+  output$dunb33 <- ( 1 - ( 3 / (4*orig.df - 1) ) ) * output$d33     # Calculate d unbiased for the 33% power criterion
   
   
   ## Original study smallest detectable effect size ##
   
-  output$crit.t  <- qt(p = 1-p/2, df = df) # Critical t for original study
-  output$crit.d  <- crit.t * sqrt(1/(n1-1) + 1/(n2-1)) # Convert critical t to d
-  #output$crit.d2 <- 2*crit.t / sqrt(Shih.df) # Alternative formula. Does not produce the same results when n differs between groups
+  output$crit.t  <- qt(p = 1-p/2, df = orig.df) # Critical t for original study
+  output$crit.d  <- output$crit.t * sqrt(1/(n1-1) + 1/(n2-1)) # Convert critical t to d
   return(output)
 }
 
@@ -32,6 +31,6 @@ Shih.sd1 <-  0.17
 Shih.sd2 <-  0.20
 Shih.n1  <- 16
 Shih.n2  <- 16
-Shih <- tFromDescTwo(Shih.m1, Shih.m2, Shih.sd1, Shih.sd2, Shih.n1, Shih.n2, test = "Student")
+Shih <- tFromDescTwo(Shih.m1, Shih.m2, Shih.sd1, Shih.sd2, Shih.n1, Shih.n2, test = "student")
 
 eqboundCritera(Shih.n1, Shih.n2, Shih.df)

### 2017-11-08 examples_plotgrid.R

@@ -5,19 +5,22 @@
 library(purrr)
 library(gridExtra)
 library(ggplot2)
+library(grid)
+library(lattice)
 
 library("TOSTER")
-download.file(url = "https://raw.githubusercontent.com/Lakens/TOSTER/master/R/TOSTtwo.prop.R", destfile = "TOSTtwo.prop.R")
-source(file = "TOSTtwo.prop.R")
 
-plot.scaler = 2
+plot.scaler <- 1.8
+nhst.size <- 0.5
+tost.size <- 1
+point.size <- 2
 
 ####  Base plot ----------------------------------------
 
 baseplot <- ggplot(data.frame()) +
   scale_x_continuous(breaks=NULL) +
   coord_flip() + # flip coordinates (puts labels on y axis)
-  theme_classic() + # use a white background
+  theme_classic(base_size = 10) + # use a white background
   geom_hline(yintercept=0, lty=1) +
   theme(axis.title.y = element_blank(),
         axis.text.y = element_blank(),
@@ -25,16 +28,7 @@ baseplot <- ggplot(data.frame()) +
 
 ####  Meta analysis ----------------------------------------
 
-gr2 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr3 <- TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr4 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr5 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr6 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr7 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr8 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
[…]

### 2017-11-08 tFromDescTwo.R

@@ -38,7 +38,7 @@ tFromDescTwo <- function(m1, m2, sd1, sd2, n1, n2, test = "welch") {
   output$d    <- output$t * sqrt( 1 / output$n1 + 1 / output$n2 )    # calculate Cohens d (or Hedges g if "welch")
   output$dunb <- ( 1 - ( 3 / (4*output$df - 1) ) ) * output$d        # Calculate d unbiased (from Cumming 2012, p 294, eq 11.13)
   
-  return(output) # Return all items stored in 'output'
+  return(output) # Return all desired output statistics stored in 'output'
 }
 
 # Test

### 2017-11-08 twostproptwo.R
Deleted old proportion TOST function

@@ -1,74 +1,0 @@
-#' TOST function for two propotions (raw scores)
-#' @param proprop1 proportion of group 1
-#' @param proprop2 proportion of group 2
-#' @param n1 sample size in group 1
-#' @param n2 sample size in group 2
-#' @param margin equivalence bounds (e.g., 0.1) expressed in raw scale units (e.g., scalepoints)
-#' @param alpha alpha level (default = 0.05)
-#' @return Returns TOST z-value 1, TOST p-value 1, TOST z-value 2, TOST p-value 2, low equivalence bound, high equivalence bound, Lower limit confidence interval TOST, Upper limit confidence interval TOST
-#' @importFrom stats pnorm pt qnorm qt
-#' @importFrom graphics abline plot points segments title
-#' @examples
-#' ## Equivalence test for two independent proportions equal to .65 and .70, with 100 samples per group, an equivalence margin of .1 
-#' ## and alpha = 0.05.
-#' TOSTtwo.prop(prop1 = .65, prop2 = .70, n1 = 100, n2 = 100, margin = .1, alpha = .05)
-#' @section References:
-#' Tunes da Silva, G., Logan, B. R., & Klein, J. P. (2008). Methods for Equivalence and Noninferiority Testing. Biology of Blood Marrow Transplant, 15(1 Suppl), 120-127.
-#' Yin, G. (2012). Clinical Trial Design: Bayesian and Frequentist Adaptive Methods. Hoboken, New Jersey: John Wiley & Sons, Inc.
-#' @export
-
-TOSTtwo.prop <- function(prop1, prop2, n1, n2, margin, alpha) {
-  if(missing(alpha)) {
-    alpha <- 0.05
-  }
-  if(margin < 0) {
-    margin <- abs(margin)
-  }
-  
-  prop_dif <- prop1 - prop2
-  prop_se <- sqrt((prop1*(1-prop1))/n1 + (prop2*(1-prop2))/n2)
-  
-  ##defining z-statistic
-  z_lower_numerator <- prop_dif + margin
-  z_upper_numerator <- prop_dif - margin
-  z_lower <- z_lower_numerator/prop_se
-  z_upper <- z_upper_numerator/prop_se
-  
[…]

### 2017-11-08 NSEQ_Brandt_Example.R
Made .R scripts for all examples
* For sourcing in manuscript and plotgrid script
* Edit in .R script and source to avoid copy-paste errors when updating the code.
* Necessary for the OSF folder

@@ -1,0 +1,31 @@
+library(TOSTER)
+library(pwr)
+
+# original study: Banerjee et al., study 1
+orig.m1 <- 4.71
+orig.m2 <- 5.3
+orig.sd1 <- 0.85
+orig.sd2 <- 0.97
+orig.d <- 0.65
+orig.t <- 2.03
+orig.p <- 0.049
+orig.N <- 40 #group size unknown, therefore equal n assumed
+orig.df <- orig.N-2
+
+# replication study: Brandt et al., study 1
+rep.m1 <- 4.7857
+rep.m2 <- 4.6569
+rep.sd1 <- 1.0897
+rep.sd2 <- 1.1895
+rep.n1 <- 49
+rep.n2 <- 51
+rep.t <- 0.56
+rep.p <- .574
+rep.df <- rep.n1 + rep.n2 - 2
+rep.d <- 0.11
+
+
+# Calculate small telescopes effect (d) effect orig. study had 33% power to detect
+d.33 <- (pwr.t.test(n = orig.N/2, d = NULL, sig.level = 0.05, power = 0.33, type = "two.sample", alternative = "two.sided"))$d
+
+Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.33, high_eqbound_d = d.33, var.equal = FALSE)

### 2017-11-08 NSEQ_Lynott_example.R

@@ -1,0 +1,31 @@
+se <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
+crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05
+
+
+crit.diff <- crit.z * se  # Convert z to the critical difference
+
+
+
+library(TOSTER)  # Install TOSTER using devtools::install_github("Lakens/TOSTER"), or download TOSTtwo.prop.R from https://github.com/Lakens/TOSTER/tree/master/R and then source it. 
+
+
+#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
+lynott.data <- read.csv("NSEQ_Lynott_example_data.csv")
+lynott.data <- lynott.data[lynott.data$include1 == 1,]
+
+
+#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+n.hot  <- sum(lynott.data$packtype == 1)
+n.cold <- sum(lynott.data$packtype == 0)
+selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
+selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
+diff <- selfish.hot - selfish.cold
+
+#  We can now run the TOST!
+Lynott <- TOSTtwo.prop(prop1 = selfish.cold,
+                     prop2 = selfish.hot,
+                     n1    = n.cold,
+                     n2    = n.hot,
+                     low_eqbound  = -crit.diff,
+                     high_eqbound = crit.diff,
+                     alpha = 0.05)

### 2017-11-08 NSNEQ_Moon_Example.R

@@ -1,0 +1,40 @@
+#  Load the relevant datafile into R
+moon.data        <- read.csv("NSNEQ_Moon_Example_data.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)
+
+#  Set the bounds, in this case 1 grade point, which corresponds to 6.25%
+sesoi <- 0.0625
+
+#  Calculate the summary statistics from the data that we need in the TOST calculation below
+mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
+mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
+sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
+sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
+n.asian          <- sum(moon.data$identity_salience == 1)
+n.control        <- sum(moon.data$identity_salience == 2)
+
+df <- (sd.asian^2 / n.asian + sd.control^2 / n.control)^2 / (((sd.asian^2 / n.asian)^2 / (n.asian-1)) + ((sd.control^2 / n.control)^2 / (n.control - 1)))  # Degrees of freedom for Welch's t-test
+sample.sd <- sqrt((sd.asian^2 + sd.control^2) / 2)  # Calculate sd root mean squared for Welch's t-test
+sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate the se of the difference based on sd.pooled
+
+t <- (mean.asian - mean.control) / sample.se  # Welch t-test
+
+d <- (mean.asian - mean.control) / sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.control + n.control) - 2))
+
+p <- 2 * pt(t, df)
+
+sdpooled <- sqrt((((n.asian - 1)*(sd.asian^2)) + (n.control - 1)*(sd.control^2))/((n.asian+n.control)-2))
+d <- (mean.asian - mean.control) / sdpooled
+
+
+
+#  We can now run the TOST!
+#  Performing the TOST using the 'TOSTtwo.raw' function allows you to specify equivalence bounds in raw values
+Moon <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
+                            m2  = mean.control,
+                            sd1 = sd.asian,
+                            sd2 = sd.control,
+                            n1  = n.asian,
+                            n2  = n.control,
[…]

### 2017-11-08 SEQ_Hyde_Example.R

@@ -1,0 +1,12 @@
+
+# Hyde, Lindberg, Linn, Ellis, and Williams (2008) meta analysis
+gr2 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr3 <- TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr4 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr5 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr6 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr7 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr8 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr9 <- TOSTmeta(ES = -0.01, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr10 <- TOSTmeta(ES = 0.04, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+gr11 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)

### 2017-11-08 corr_Kahane_Example.R

@@ -1,0 +1,26 @@
+library(TOSTER)
+library(pwr)
+
+## Kahane, Everett, Earp, Farias, & Savulescu (2015)
+##
+## Study 4, 5.2.1. ii., p. 203: "There  was  no  relationship  between  perceived wrongness in 
+## personal dilemmas and in the ‘greater good’ dilemmas (r = -.04, p = .53): that is, people
+## who were more ‘utilitarian’ in the personal dilemmas were not more likely to be more truly
+## utilitarian in the other dilemmas, and vice versa. This lack of relationship held even when
+## controlling for primary psychopathy in a partial second order correlation (r = -.07, p = .32)."
+
+Kahane.data <- read.csv("Kahane_etal_study4_cleaned.csv", header = TRUE, sep = ",")
+Kahane.data <- as.data.frame(Kahane.data)
+
+# sample size
+Kahane.N <- min(sum(!is.na(Kahane.data$ClassicSacrificialDilemmas)), sum(!is.na(Kahane.data$GreaterGoodDilemmas)))
+
+# correlation b/w perceived wrongness in personal dilemmas and 'greater good' dilemmas
+Kahane.r.GG <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$estimate
+Kahane.p <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$p.value
+
+# correlation the study had 80% power to detect
+Kahane.r.80 <- (pwr.r.test(n = Kahane.N, r = NULL, sig.level = 0.05, power = 0.8, alternative = "two.sided"))$r
+
+# TOST with bounds set to the correlation the study had 80% power to detect
+Kahane <- TOSTr(n = Kahane.N, r = Kahane.r.GG, low_eqbound_r = -Kahane.r.80, high_eqbound_r = Kahane.r.80, alpha = 0.05)

### 2017-11-08 examples_plotgrid.R
Assigned plot in examples_plotgrid to object

@@ -118,4 +118,4 @@ lay <- rbind(c(1,2,5),
              c(3,4,5)) # Define the figure layout. Think of the c()s as a grid/ Plot 1-5 will be alloted the relative space in the grid.
 
 
-grid.arrange(brandtplot, moonplot, lynottplot, kahaneplot, metaplot, layout_matrix = lay)
+example.grid <- grid.arrange(brandtplot, moonplot, lynottplot, kahaneplot, metaplot, layout_matrix = lay)

### 2017-11-08 manuscript.Rmd
Major updates to manuscript
* Copied text from Google doc (08-11-17)
* Removed copy-pasted code in examples and figure generation
* Included source commands for generating example code and figures, to avoid errors when updating code outside of manuscript.

@@
 For example, we might predict that priming American Asian women with their Asian identity will increase their performance on a math test compared to women who are primed with their female identity.
-To be able to falsify this hypothesis, and design a study that allows for strong inferences [@platt_strong_1964], it is important to specify which test result would *disprove* the hypothesis.
+To be able to falsify this hypothesis, and design a study that allows for strong inferences (Platt, 1964), it is important to specify which test result would *disprove* the hypothesis.
 
@@
 However, this would ignore the fact that the human sex ratio of newborn boys to girls is not exactly 1:1, and we should not expect 50% of participants to be boys.
-On average, 103 boys are born for every 100 girls (CIA, 2017), so approximately 50.74% of applicants should be boys, and 49.26% should be girls (with 0.5074/0.4926≈1.03).
+On average, 103 boys are born for every 100 girls (CIA, 2017), so approximately 50.74% of applicants should be boys, and 49.26% should be girls (with $0.5074/0.4926 \approx 1.03$).
 If boys and girls were exactly equally likely to be brought into the program by their parents, we would not expect a difference of zero, but 1.5% more boys.
@@
 The researcher can for example test if the difference is smaller than -0.005, or larger than 0.035.
-This test against two bounds, with $H_0$ as a range rather than one value, is known as a *minimal effects test* (Murphy, Myors, & Wolach, 2014).(Murphy et al.
-2014)
+This test against two bounds, with $H_0$ as a range rather than one value, is known as a *minimal effects test* (Murphy, Myors, & Wolach, 2014).
 
@@
 As long as gender differences in the participation rate are not too extreme, there is no need to spend money to address the gender imbalance in the voluntary participation rate, and all money can be spent on the core training program.
-To test whether the gender difference in participants is *not large enough to matter, the researcher can perform an *equivalence test*.
-Equivalence tests examine whether the presence of gender differences that are large enough to matter can be rejected.
+To test whether the gender difference in participants is *not* large enough to matter, the researcher can perform an *equivalence test*.
+Equivalence tests examine whether values below and above a prespecified range can be rejected.
+Here, we would test if the presence of gender differences that are large enough to matter can be rejected.
 After an extensive discussion with experts, the researcher decides that as long as the difference in proportions is not larger than 6% the researcher can act as if the gender difference is too small to intervene.
-Given an expected true difference in the population of 0.015, the researcher will tests if the observed difference in the data is smaller than 0.075, and larger than -0.055.
+Given an expected true difference in the population of 0.015, the researcher will test if the observed difference in the data is smaller than -0.055, and larger than 0.075.
 If differences more extreme than both these boundary values can be rejected in two one-sided tests, the researcher will conclude statistical equivalence, and no money will be spent on addressing a gender difference in participation.
 
-One can conclude statistical equivalence whenever a 90% confidence interval[a][b][c][d][e][f] around the observed estimate does not contain the smallest effect size of interest (SESOI[g]).
+For an alpha level of 5%, we can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI).
+A 90% confidence interval is used because one-sided tests are performed, and it is not necessary to control for multiple comparisons when performing two one-sided tests because both one-sided tests need to be statistically significant to conclude equivalence.
+Note that although we will use an alpha of 5% in all examples because this alpha level was used in the papers we reanalyze, researchers should carefully decide upon a desired Type 1 error rate, and justify their choice (Lakens et al., 2017).
 Figure 1 illustrates the possible conclusions we can draw from null hypothesis, minimal effect, and equivalence tests.
+Although thinking about equivalence tests in terms of the values that can be rejected based on a confidence interval might be the easiest conceptualization, for any given study, you can also think of an equivalence test as determining whether effect sizes or test statistics are closer to zero than some critical, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
 
-Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, instead of a value of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017].
-In this article, we provide several examples of equivalence tests which illustrate the procedure.
+```{r echo=FALSE, fig.width=5, fig.height=7}
+#  Plot H0 and H1 regions for different significance tests
+source("H0H1_plot.R")
+```
+
+Figure caption: Figure 1.
+Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
+**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\Delta_{L}$ *or* larger than $\Delta_{U}$.
+**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
+**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\Delta_{L}$ *and* smaller than $\Delta_{U}$.
+**D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\Delta$.
+
+Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, rather than against 0), their use in psychology was limited, perhaps due to the lack of user-friendly software to perform the calculations in the past [@lakens_equivalence_2017].
+In this article, we provide several examples of equivalence tests which illustrate the procedure in free software (R, a spreadsheet, and in Jamovi).
 Furthermore, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.
@@
 Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI).
-The SESOI can sometimes be based on just noticeable differences, which can be objectively determined.
-Most often, however, it is a subjective decision that varies across individuals and time.
-The SESOI is ideally based on a cost-benefit analysis.
+Although in the examples here, we use symmetrical bounds around 0 (e.g., -0.3, 0.3), based on for example an explicit cost-benefit analysis, one might use asymmetric bounds (e.g., -0.2, 0.3).
+The SESOI can sometimes be determined objectively, for example based on just noticeable differences .
+Most often, however, it is a subjective decision that varies across researchers, fields, and time.
+In lieu of objective justifications, the SESOI should ideally be based on a cost-benefit analysis.
 Since both costs and benefits are necessarily subjective, the SESOI will depend on the researcher who designs the study.
@@
 The hypothesis was that a slightly redder skin signals greater attractiveness and physical health, and sending this signal to men yields an evolutionary advantage.
-This hypothesis requires that the increase in redness can be detected with the naked eye by men.
-They collected data from 22 women and showed that there was indeed an increase in redness of the facial skin of woman during their fertile period.
-However, this increase was not large enough to be noticeable with the naked eye by men, thus falsifying their hypothesis.
+This hypothesis presupposes that men can detect the increase in redness with the naked eye.
+Burriss and colleagues collected data from 22 women and showed that there was indeed an increase in redness of the facial skin during their fertile period.
+However, this increase was not large enough for men to detect with the naked eye, thus falsifying their hypothesis.
 Because the just noticeable difference in redness of the skin can be measured, it is possible to objectively establish the SESOI.
@@
 Another example of an objectively determined SESOI can be found in @button_minimal_2015 where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
+
+Norman, Sloan, and Wyrwich (2003) used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or d = 0.5.
+They make the argument that this bound of approximately half a standard deviation could be related to the limits of human discriminability.
+Therefore, even though a SESOI of d = 0.5 might seem quite large, it could be a plausible equivalence bound when examining effects that should be noticeable by single individuals.
 
@@
 First, researchers can use benchmarks.
-For example, one might set the SESOI to a standardized effect size of d = 0.5[h][i][j][k], which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988).
+For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988).
 Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small (Maxwell, Lau, & Howard, 2015).
 Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided.
+Based on a review of 112 meta-analyses, Weber and Popova (2012) conclude that setting a SESOI in communication research to a medium effect size (*r* = 0.3, or d = 0.5) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and Hemphill (2003) suggests that in psychology a SESOI of d = 0.5 would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
 
@@
 It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study.
-Simonsohn (2015) recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have 33% power to detect based on the sample size.
-This is called the 'small telescopes' approach.
-For example, consider a study where `r N` participants answered a question, which was analyzed with an one-sample t-test.
-For a two-sided test with an alpha of `r alpha` this test had 33% power to detect a d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`.
+Simonsohn (2015) recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have had 33% power to detect.
+With this so-called 'small telescopes' approach, the equivalence bounds are thus determined primarily based on the sample size in the original study.
+For example, consider a study in which `r N` participants answered a question, and the results were analyzed with a one-sample t-test.
+For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`.
+Another example of how previous research can be used to determine the SESOI can be found in Kordsmeyer and Penke (2017), who defined the SESOI based on the mean of effect sizes reported in the literature.
+Thus, they examined whether their replication study could reject effects as large or larger than had on average been reported in the literature.
+Given random variation and bias in the literature, a more conservative approach could be to use a lower bound of a confidence interval around the meta-analytic effect size estimate (cf.
+Perugini, Gallucci, & Constantini, 2014).
 
-Another justifiable choice for the SESOI we would like to propose here is to use the smallest observed effect size that could have been statistically significant in the original study.
-In other words, we decide that effects that could not have yielded a *p* < 0.05 in the original study (if this is your preferred alpha threshold) will not be considered meaningful in the replication either, even if effects of this size are statistically significant in the replication study.
+Another justifiable choice for the SESOI is to use the smallest observed effect size that could have been statistically significant in a previous study.
+In other words, we decide that effects that could not have yielded a $p < \alpha$ in an original study will not be considered meaningful in the replication study either, even if they are found to be statistically significant.
 Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
-This critical test value can also be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*.
-All observed effect sizes smaller than the critical effect size would not be statistically significant in an original study, given the alpha and sample size of that study.
+This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*[^footnote].
+All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
 By setting the SESOI to the critical effect size, an equivalence test can reject all observed effect sizes that could have been detected in an earlier study.
+
+[^footnote]: This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect (see Lenth, 2007).
+This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect.
 
@@
 
-Third[l][m], researchers can set the SESOI based on the resources they have available.
+Third, researchers can set the SESOI based on the sample size they are planning to collect.
 The amount of data you can collect limits the inferences you can make.
-Given a an alpha level and a sample size, researchers can calculate the smallest effect size that one has sufficient power to detect.[n] This approach is suggested here to provide researchers with a way to think about setting the SESOI for a study in a line of research where no quantifiable predictions are made, and where no previous studies have been performed.
-Observing statistical equivalence for a SESOI based on this approach allows one to reject effects that can be reliably detected with the given sample size.
+Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.
+This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
+In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size.
 For example, a researcher who plans to perform a two-sided one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364.
 Using equivalence bounds of d = -0.364 and d = 0.364 in an equivalence test allows one to reject effects the study had high statistical power to detect.
-In this case, concluding equivalence would suggest one can reject effects that can reasonably be detected with N = 100.
-Whether or not this is interesting depends in part on how typical the chosen sample size is for a research area.
-One should not expect an equivalence test based on 15 observations testing against a SESOI of d = -1 and d = 1 to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds.
-But by transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes.
-When there are no previous studies to build on, and no theoretical predictions, specifying a SESOI based on the alpha level and a sample size can provide a starting point for a discussion on what a reasonable SESOI is.
-Nevertheless, the value of setting a SESOI based on this procedure is limited, and the equivalence test does not reject values that are theoretically interesting.
-
+An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that can be reliably detected given a specific sample size?
+In this case, concluding equivalence would suggest one can reject effects that can reasonably be detected with a sample size of 100, and that future studies probably need larger sample sizes to examine a specific question.
+Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area.
+One should not expect an equivalence test based on 15 observations testing against a SESOI of d = -1 and d = 1 to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
+By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is.
 ```{r}
@@
 ```
+#Raw versus standardized equivalence bounds
+
+Equivalence bounds can be set in terms of standardized effect sizes, or raw differences.
+For example, the SESOI might be a raw mean difference of 0.5 points on a 7-point scale, or a Cohen’s d of 0.5.
+The main difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as [FORMULA (X1-X2)/SD].
+The observed standard deviation randomly varies across samples.
+In practice, this means that when you use standardized differences as bounds (e.g., d = 0.35) the equivalence test depends on the standard deviation in your the sample.
+The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of d = 0.5 when the standard deviation is 1, but a standardized equivalence bound of d = 1 when the standard deviation is 0.5.
+
+Both raw equivalence bounds and standardized equivalence bounds have specific benefits and limitations (for a discussion, see Baguley, 2009).
+When raw mean differences are meaningful and of theoretical interest, it makes sense to set equivalence bounds based on raw effect sizes.
+When the raw mean difference is of less theoretical importance, or different measures are used across research lines, it is often easier to set equivalence bounds based on standardized differences.
+Researchers should realize that equivalence bounds based on raw differences or standardized differences ask slightly different questions, and justify their choice for an equivalence bound.
+When setting equivalence bounds based on earlier research, such as in replication studies, equivalence bounds based on raw or standardized differences would ideally give the same result, and large differences in standard deviations between studies are as important to interpret as large differences in means.
 
@@
 ```{r include = FALSE}
-library(TOSTER)
+source("NSEQ_Brandt_Example.R")
 
-# original study: Banerjee et al., study 1
-orig.m1 <- 4.71
-orig.m2 <- 5.3
-orig.sd1 <- 0.85
-orig.sd2 <- 0.97
-orig.d <- 0.65
-orig.t <- 2.03
-orig.p <- 0.049
-orig.N <- 40 #group size unknown, therefore equal n assumed
-orig.df <- orig.N-2
-
-# replication study: Brandt et al., study 1
-rep.m1 <- 4.7857
-rep.m2 <- 4.6569
-rep.sd1 <- 1.0897
-rep.sd2 <- 1.1895
-rep.n1 <- 49
-rep.n2 <- 51
-rep.t <- 0.56
-rep.p <- .574
-rep.df <- rep.n1 + rep.n2 - 2
-rep.d <- 0.11
-
-# Calculate critical effect (d): smallest effect orig. study had the power to detect
-t.crit = qt(1-.05/2, (orig.N-2))
-d.crit = t.crit*sqrt((1/(orig.N/2)) + 1/(orig.N/2))
-
-Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
 ```
 
-Banerjee, Chatterjee, & Sinha (2012) reported that participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
-A close replication by Brandt, IJzerman, & Blanken (2014) found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
-The smallest effect size that could have been statistically significant in the original study is $d_{crit}= `r round(t.crit, 2)` \sqrt { \frac { 1} {`r orig.N/2`} + \frac { 1} {`r orig.N/2`} }=`r round(d.crit, 2)`$.
-Using this as our SESOI for a TOST with Welch's *t*-test for independent samples --- resulting in equivalence bounds of ${\Delta}_L=`r round(-d.crit, 2)`$ and ${\Delta}_U=`r round(d.crit, 2)`$ --- we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.
-
-```{r, fig.width=6}
-TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.crit, high_eqbound_d = d.crit, var.equal = FALSE)
-```
-
-[^1]: This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect[o].
-This procedure will thus result in equivalence bounds that are substantially larger than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect[p][q][r][s][t][u][v][w].
+Banerjee, Chatterjee, & Sinha (2012) reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
+A close replication by Brandt, IJzerman, & Blanken (2014) with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
+Following the small telescopes approach, we can calculate the effect size the original study had $33\%$ power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.
+If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.
 
@@
 ```{r, fig.width=6}
-library("TOSTER")
-TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = -0.01, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = 0.04, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
-TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005)
+source("SEQ_Hyde_Example.R")
 ```
@@
 For example, in grade 3 the difference is d = 0.04, with a standard error of 0.002.
-When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of d = -0.1 and d = 0.1, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1, and can be considered trivially small, according to the bounds set by the authors.
+When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of d = -0.1 and d = 0.1, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1, and can be considered trivially small, according to the bounds set by the authors (figure 2B).
 However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge.
-This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance[x].
+This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance.
 
 ## Example 3: Not Statistically Equivalent and Not Statistically Different
+```{r include=FALSE}
+source("NSNEQ_Moon_Example.R")
+```
 
@@
 
-Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by inserting the summary statistics from Moon & Roeder's (2014) data, assuming an alpha of 0.05, and using +/- the SESOI of 0.0625 as our equivalence bounds.[y][z][aa][ab][ac]
+We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using +/- the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
 
-```{r echo=FALSE}
+We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C).
+Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
+The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
+Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{lower}$ = -0.1 and $\Delta_{upper}$ = 0.1).
 
-#  We first load the relevant datafile into R
-moon.data        <- read.csv("NSNEQ_Moon_Example_data.csv")
-
-#  We then calculate the summary statistics from the data that we need in the TOST calculation below
-mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
-mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
-sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
-sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
-n.asian          <- sum(moon.data$identity_salience == 1)
-n.control        <- sum(moon.data$identity_salience == 2)
-
-#  We can now run the TOST!
-#  Performing the TOST using the 'TOSTtwo.raw' function allows us to specify equivalence bounds in raw values
-tost <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
-                                m2  = mean.control,
-                                sd1 = sd.asian,
-                                sd2 = sd.control,
-                                n1  = n.asian,
-                                n2  = n.control,
-                                low_eqbound  = -0.0625,
-                                high_eqbound =  0.0625,
-                                alpha = .05)
-```
-
-We find that the TOST is non-significant, t(`r round(tost$TOST_df, 2)`) = `r round(min(abs(c(tost$TOST_t1, tost$TOST_t2))), 2)`, p = `r round(max(tost$TOST_p1, tost$TOST_p2), 2)`.
-Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote]
-
-[^footnote]:[ad] In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
-We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d (equivalence bounds = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`).
-Since this is a replication of a previous study, we could also have chosen calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
+[^footnote]: In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
+We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
+Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
 Now we would no longer be answering a question about a raw effect independent of the samples.
 Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
-Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d, and thus the bounds, would also be only sligthly smaller (bounds = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`).
-As standard deviations of the original study compared to the replication becomes larger, the bounds become more narrow.
+Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
+As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.
 
-## Example 4: Statistically Inferior [ae][af]and Not Statistically Different
+## Example 4: Statistically Inferior and Not Statistically Different
+```{r include=FALSE}
+source("NSEQ_Lynott_example.R")
+```
 
@@
 (2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
-They observed that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same.
-Calculating a z-test for the difference between the proportions, this effect is statistically significant (Diff = -0.067, z = -1.98, p = 0.024[ag]).
-Having obtained a significant result, we are left with two alternatives:
+They observed that `r selfish.cold * 100`\% of participants who received a cold pack ($n_{cold}=`r n.cold`$) opted to receive a reward for themselves, while `r selfish.hot * 100`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
+In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r Lynott$dif * 100`$\%, $Z=`r Lynott$NHST_z`$, $p=`r Lynott$NHST_p`$).
+Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.
 
-1. The difference is small enough that even though it is *statistically* significant, we can consider it *practically* equivalent to zero.
-2. The difference is substantial enough that we cannot rule out that it is also practically significant.
-
-In this case, it is not really clear what an objective cirterion for the smallest effect size of interest could be.
+In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be.
 However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
-This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds.
+This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- round(qnorm(1-0.05/2),2)$).
 To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
 
-```{r}
-se         <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
-crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05
+Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study.
+In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{upper}$ because an effect with CIs crossing $\Delta_{lower}$ can still be *lower* than the original effect.
 
-crit.diff <- crit.z * se  # Convert z to the critical difference
+We find that the test against $\Delta_{upper}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D).
+Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.
 
-print(crit.diff)
+## Example 5: Statistically Equivalent and not Statistically Different
+```{r include = FALSE}
+source("corr_Kahane_Example.R")
 ```
 
-Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study.
-In this case, we can still use a TOST to analyse the data, but we only need to consider the test against the upper bound because an effect with CIs crossing the lower bound can still be *lower* than the original effect.
+In a paper published in 2015, Kahane and colleagues investigate moral dilemma vignettes in which participants have to decide whether or not they would sacrifice one person's life to save several other lives, or judge how morally admissible such actions are.
+Traditionally, greater endorsement for sacrificing a life to save others has been interpreted as a more "utilitarian" moral orientation, i.e.
+a stronger concern for the greater good (in total, fewer people lose their lives).
+In a number of studies, Kahane et al.
+contest this interpretation, for example by showing that greater endorsement for such "utilitarian" choices correlates with sub-clinical psychopathy, and, crucially, by comparing the traditionally used vignettes with a set of new ones that pit partial motivations against an impartial concern for the greater good (e.g.
+buying a new mobile phone vs.
+donating the money to save lives in a distant country, study 4).
 
-```{r}
+Kahane et al.
+find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$[^footnote]).
+They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
+206).
+This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had $80\%$ power to detect effects of $r = `r round(Kahane.r.80, 2)`$.
+Given bounds of $\Delta_{L} = `r -round(Kahane.r.80, 2)`$ and $\Delta_{U} = `r round(Kahane.r.80, 2)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(max(Kahane$TOST_p1, Kahane$TOST_p2), 3)`$.
+This means that values smaller than $r = `r -round(Kahane.r.80, 2)`$ and larger than $r = `r round(Kahane.r.80, 2)`$ can be rejected at an alpha level of $5\%$ and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.
 
-download.file(url = "https://raw.githubusercontent.com/Lakens/EquivalenceTestingForPsychologicalScience/master/TOSTtwo.prop.R", destfile = "TOSTtwo.prop.R")
-source(file = "TOSTtwo.prop.R")
+[^footnote]: Study 4 in Kahane et al.
+(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of only $N = `r Kahane.N`$.
 
-#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
-lynott.data <- read.csv("NSEQ_Lynott_example_data.csv")
-lynott.data <- lynott.data[lynott.data$include1 == 1,]
-
-#  We then calculate the summary statistics from the data that we need in the TOST calculation below
-n.hot            <- sum(lynott.data$packtype == 1)
-n.cold           <- sum(lynott.data$packtype == 0)
-selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
-selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
-
-#  We can now run the TOST!
-tost <- TOSTtwo.prop(prop1 = selfish.cold,
-                         prop2 = selfish.hot,
-                         n1        = n.cold,
-                         n2        = n.hot,
-                         low_eqbound  = -crit.diff,
-                         high_eqbound = crit.diff,
-                         alpha = 0.05)
-
+```{r include=FALSE}
+source("examples_plotgrid.R")
 ```
-
-We find that the test against the upper bound is `r ifelse(tost$TOST_p2 < tost$alpha, "significant", "non-significant")`, $z =$ `r round(tost$TOST_z2, 2)`, $p =$ `r ifelse(round(tost$TOST_p2, 3) >= 0.001, paste("=", round(tost$TOST_p2, 2)), "< 0.001")`.
-Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.
-## Example 5: Not Statistically Equivalent and Statistically Different
-Example 5: Correlations[ah]?
-
-Would be nice to have 1 bounds based on effect sizes you have power to detect.
-And one on just noticeable differences.
-This: http://wiki.ihe.net/images/e/e5/Measurement_of_pain.pdf suggests that  this: Berry, H., Huskisson, E.
-C.
-Clin.
-Trials J.
-1972, 9, 13.
-12. Hardy, J. D., Wolff, H. G., Goodell, H. Pain Sensations and
-Reactions. Baltimore, 1952. shows that a VAS for pain can be divided into 20 just noticeable differences' Or we use
+```{r echo=FALSE}
+plot(example.grid)
+```
 
@@
 The given examples and figures XXX to YYY illustrate that equivalence testing is a relatively simple and flexible technique: As long as we can calculate a confidence interval around a mean effect, we can compare it to one or several bounds.
-In that sense, the result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman[ai] & Serlin, 1998; Tryon, 2001).
-This ease of use might tempt researchers to set equivalence bounds after looking at their data.
-For example, we showed above that Moon & Roeder's replication of the identity-priming study by Shih et al.
+In that sense, the result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman & Serlin, 1998; Tryon, 2001).
+This ease of use might tempt researchers to set equivalence bounds *after* looking at their data.
+For instance, in example 3 we showed that Moon & Roeder's replication of the identity-priming study by Shih et al.
 (1999; example 3) does not exclude effects that are large enough to affect a student's final grade on a test.
-However, if we look at the $90 \%$ CI (figure XXX), we can see that the data would have been statistically equivalent if the bounds had been based on a slightly larger SESOI, e.g.
+However, if we look at the $90 \%$ CI (figure XXX), we can see that the data would have been statistically equivalent if the bounds had been set to  a slightly larger SESOI, e.g.
 a mean difference of 0.1.
@@
 CONSORT guidelines, Piaggio et al., 2006).
-A perhaps helpful guiding question may be: "Will testing against these bounds yield interesting information in the long run?" In most cases[aj],
+A perhaps helpful guiding question may be: "Will testing against these bounds yield interesting information in the long run?"
 
-#Raw[ak] versus standardized equivalence bounds
+#Justify your equivalence bounds
 
-Equivalence bounds can be set in terms of standardized effect sizes, or raw differences.
-For example, the SESOI might be a raw mean difference of 0.5 points on a 7-point scale, or a Cohen’s d of 0.5.
-The main difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as [FORMULA (X1-X2)/SD].
-The observed standard deviation randomly varies across samples.
-In practice, this means that when you use standardized differences as bounds (e.g., d = 0.35) the equivalence test depends on the standard deviation in the sample.
-The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of d = 0.5 when the standard deviation is 1, but a standardized equivalence bound of d = 1 then the standard deviation is 0.5.
+Like with any statistical test, the usefulness of the result an equivalence test gives us depends on the question we ask.
+The question manifests in the bounds we set: Is our data surprising if the true effect lies outside of the bounds?
+If we set very wide bounds, we ask if our data is surprising given that the true effect is very large[^footnote].
+For most psychological effects, the answer will usually be “yes,” which means that it may not be a very interesting question to ask.
+In order to obtain a meaningful result, we have to set the bounds based on meaningful criteria: Our SESOI should really be *the smallest effect we are interested in* for a given study.
+For this reason, researchers should provide a detailed justification of the choice for a SESOI.
+Some studies in the literature simply state the SESOI that is used, without providing a justification.
+For example, Brown and colleagues (2017) and Schumann and colleagues (2017)use a SESOI of Cohen’s d = +/- 0.5, but it is not clear why these bounds were chosen[^footnote].
+An equivalence test should answer a question, and justifying the SESOI is an essential part of asking the question.
+An equivalence test of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but the reason this SESOI was chosen should be explicitly justified.
 
-Both raw equivalence bounds and standardized equivalence bounds have specific benefits and limitations (for a discussion, see Baguley, 2009).
-When raw mean differences are meaningful and of theoretical interest, it makes sense to set equivalence bounds based on raw effect sizes.
-When the raw mean difference is of less theoretical importance, or different measures are used across research lines, it is often easier to set equivalence bounds based on standardized differences.
-Researchers should realize that equivalence bounds based on raw differences or standardized differences ask slightly different questions, and justify their choice for an equivalence bound.
-When setting equivalence bounds based on earlier research, such as in replication studies, equivalence bounds based on raw or standardized differences would ideally give the same result, and large differences in standard deviations between studies are as important to interpret as large differences in means.
+[^footnote]: To be precise, wide bounds set symmetrically around 0 ask if the data is surprising given that the true effect is very *extreme*, i.e.
+either very large (positive) or very small (negative).
+[^footnote]: In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.
 

### 2017-11-08 Lynott et al. (2014) - Potential error in statistics reporting.Rmd
Report document and related files for documenting error in Lynott et al. (2014)

---
title: "Lynott et al. (2014) - Potential error in statistics reporting"
author: "Peder M Isager"
date: "October 19, 2017"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Discovered potential minor error in reporting of omnibus statistics

Dear authors,

I am currently reanalyzing your 2014 replication of Williams and Bargh (2008) for an example. I discovered what I believe could be a minor error in the reported statistics for the omnibus test across labs. In the paper, the test is reported as follows: $\chi^2 (1) = 3.402$, $p = .065$, $OR = 0.77$, $95\% CI = [.58, 1.02]$.

However, when analyzing the omnibus_data_bargh.sav in R using the syntax_barghrep_forOSF.sps script, I get the following values for the omnibus:

![](Lynott_omnibus_SPSS_output.png)

In other words, the $\chi^2$ and $p$ value does not seem to correspond between the paper and the dataset on OSF. Note that the difference is minor. and in case of a mistake it should not change the overall conclusions. I did not do an exhaustive check of all statistics, but the $\chi^2$ tests from each individual test site appears to be correct at least (assuming no continuity correction).

I am planning to include your omnibus in an example for a paper (submission deadline Nov 13th). Could you possibly check whether I am correct by then?

I hope you find the notification useful, and should it turn out that this is somehow a mistake on my end, appologies in advance!

Kind regards, 
Peder M. Isager

### 2017-11-08 Bayes_functions/TOSTtwoBayes.R
Added copy of Bayes factor TOST functions

@@ -1,0 +1,167 @@
+#' TOST function for an independent t-test (Cohen's d)
+#' @param m1 mean of group 1
+#' @param m2 mean of group 2
+#' @param sd1 standard deviation of group 1
+#' @param sd2 standard deviation of group 2
+#' @param n1 sample size in group 1
+#' @param n2 sample size in group 2
+#' @param low_eqbound_d lower equivalence bounds (e.g., -0.5) expressed in standardized mean difference (Cohen's d)
+#' @param high_eqbound_d upper equivalence bounds (e.g., 0.5) expressed in standardized mean difference (Cohen's d)
+#' @param alpha alpha level (default = 0.05)
+#' @param var.equal logical variable indicating whether equal variances assumption is assumed to be TRUE or FALSE.  Defaults to FALSE.
+#' @return Returns TOST t-value 1, TOST p-value 1, TOST t-value 2, TOST p-value 2, degrees of freedom, low equivalence bound, high equivalence bound, low equivalence bound in Cohen's d, high equivalence bound in Cohen's d, Lower limit confidence interval TOST, Upper limit confidence interval TOST
+#' @importFrom stats pnorm pt qnorm qt
+#' @importFrom graphics abline plot points segments title
+#' @examples
+#' ## Eskine (2013) showed that participants who had been exposed to organic
+#' ## food were substantially harsher in their moral judgments relative to
+#' ## those exposed to control (d = 0.81, 95% CI: [0.19, 1.45]). A
+#' ## replication by Moery & Calin-Jageman (2016, Study 2) did not observe
+#' ## a significant effect (Control: n = 95, M = 5.25, SD = 0.95, Organic
+#' ## Food: n = 89, M = 5.22, SD = 0.83). Following Simonsohn's (2015)
+#' ## recommendation the equivalence bound was set to the effect size the
+#' ## original study had 33% power to detect (with n = 21 in each condition,
[…]

### 2017-11-08 zDeskProp.R

@@ -1,0 +1,24 @@
+#### Fisher's exact Z Test for proportions - from PASS Sample Size Software ch. 213 ####
+
+zDeskProp <- function(prop1, prop2, n1, n2) {
+  
+  diff <- prop1 - prop2  # Calculate difference
+  se <- sqrt( ( prop1 * ( 1 - prop1 ) ) / n1 + ( prop2 * ( 1 - prop2 ) ) / n2 )  # Calculate standard error based on Fisher z test
+  z = ( diff - 0 ) / se # Calculate z score of the difference
+  
+  output <- list()
+  output$diff <- diff
+  output$se <- se
+  output$z <- z
+  output$prop1 <- prop1
+  output$prop2 <- prop2
+  output$n1 <- n1
+  output$n2 <- n2
+  
+  return(output)
+}
+
+
+#### test ####
+
+zDeskProp(.512, .579, 422, 428)

### 2017-11-08 Example_Gibson_Shih.R
Added some old example scripts

@@ -1,0 +1,90 @@
+# TOST on the values from Gibson, Losee, and Vitiello example
+
+#### Setup ####
+
+library(TOSTER)
+library(pwr)
+source("Example from Neil Mclathie/2017.10.05 TOSTtwoBayes Raw with Cauchy+HalfCauchy.R")
+source("TOSTtwoBayes.R")
+source("tFromDescTwo.R")
+source("eqboundCritera.R")
+
+#### Data ####
+
+Gibson.m1  <-  0.63
+Gibson.m2  <-  0.51
+Gibson.sd1 <-  0.22
+Gibson.sd2 <-  0.23
+Gibson.n1  <- 40
+Gibson.n2  <- 43
+
+Gibson.se.pool <- sqrt((Gibson.sd1^2 + Gibson.sd2^2) / 2)
+
+Shih.m1  <-  0.54
+Shih.m2  <-  0.43
+Shih.sd1 <-  0.17
+Shih.sd2 <-  0.16
+Shih.n1  <- 16
+Shih.n2  <- 14
+
+Shih.se.pool <- sqrt( (Shih.sd1^2 + Shih.sd2^2) / 2 )
+Shih.t <- 2.02 # Stated in Shih paper, but is not reproduced by an independent sample t.test. 
+Shih.d <- Shih.t*sqrt(1/(16-1) + 1/(14-1))
+Shih.df <- Shih.n1+Shih.n2-2
+
+
+#### Calculate statistics to verify that the values from Dienes are reproducable ####
+
+Gibson <- tFromDescTwo(m1   = Gibson.m1, 
+                       m2   = Gibson.m2, 
+                       sd1  = Gibson.sd1, 
+                       sd2  = Gibson.sd2, 
+                       n1   = Gibson.n1, 
+                       n2   = Gibson.n2, 
+                       test = "student")
+
+Shih <- tFromDescTwo(m1   = Shih.m1, 
+                     m2   = Shih.m2, 
+                     sd1  = Shih.sd1,
+                     sd2  = Shih.sd2, 
+                     n1   = Shih.n1, 
+                     n2   = Shih.n2, 
+                     test = "student")
+
+#  In Dienes:
+  # For Shih, Pittinsky & Ambady:   t = 2.02, df = 29, p = 0.53 - Not replicated (t = 1.82, df = 28, p = 0.07)
+  # For Gibson, Losee and Vitiello: t = 2.40, df = 81, p = 0.02 - Replicated
+
[…]

### 2017-11-08 Example_Williams_Lynott.R

@@ -1,0 +1,66 @@
+#  TOST on the values from Lynott, Corker, Wortman, Lynott et al. (2014) example
+  # TOST test are calculated using difference between proportions as effect parameter
+
+#### Setup ####
+
+library(TOSTER)
+library(pwr)
+source("TOSTtwo.prop.R")
+source("zDeskProp.R")
+
+#### Data ####
+
+Williams.prop1 <- 0.75
+Williams.prop2 <- 0.46
+# Total n was 53. From Williams & Barge: Approximately half of the participants briefly held a hot pad, and the remaining participants held a cold pad.
+Williams.n1    <- 26 # Approximated
+Williams.n2    <- 27 # Approximated
+
+Lynott.prop1  <- 0.512
+Lynott.prop2  <- 0.579
+Lynott.n1     <- 427
+Lynott.n2     <- 434
+
+#### Calculate z for examples ####
+
+Williams <- zDeskProp(prop1 = Williams.prop1, 
+                      prop2 = Williams.prop2, 
+                      n1    = Williams.n1, 
+                      n2    = Williams.n2)
+
+Lynott <- zDeskProp(prop1 = Lynott.prop1, 
+                     prop2 = Lynott.prop2, 
+                     n1    = Lynott.n1, 
+                     n2    = Lynott.n2)
+
+(0.512/0.488)/(0.579/0.421)
+
+#### Calculate subjective equivalence bounds based on small telescopes and smallest detectable effect size ####
+
+#  Small telescopes
+
+h = 0.4174618
+# Solve equation in Wolfram using Lynott p1 as p1 and solve for p2
+sin(0.797399 - 0.5 * h)^2
+# Solve equation in Wolfram using Lynott p1 as p2 and solve for p2
+sin(0.5 * h + 0.797399)^2
+# Solve equation in Wolfram using Lynott p2 as p1 and solve for p2
+sin(0.864731 - 0.5 * h)^2
+# Solve equation in Wolfram using Lynott p2 as p2 and solve for p1
+sin(0.5 * h + 0.864731)^2
+
+
+pwr.2p2n.test(n1 = 26, n2 = 27, sig.level = 0.05, power = 0.33, alternative = "two.sided")
+pwr.2p.test(sig.level = 0.05, power = 0.33, alternative = "two.sided", n = 3333)
+
+
+#  Smallest detectable effect size
[…]

### 2017-11-08 Example_Gibson_Shih.R
Deleted some old example scripts

@@ -1,90 +1,0 @@
-# TOST on the values from Gibson, Losee, and Vitiello example
-
-#### Setup ####
-
-library(TOSTER)
-library(pwr)
-source("Example from Neil Mclathie/2017.10.05 TOSTtwoBayes Raw with Cauchy+HalfCauchy.R")
-source("TOSTtwoBayes.R")
-source("tFromDescTwo.R")
-source("eqboundCritera.R")
-
-#### Data ####
-
-Gibson.m1  <-  0.63
-Gibson.m2  <-  0.51
-Gibson.sd1 <-  0.22
-Gibson.sd2 <-  0.23
-Gibson.n1  <- 40
-Gibson.n2  <- 43
-
-Gibson.se.pool <- sqrt((Gibson.sd1^2 + Gibson.sd2^2) / 2)
-
-Shih.m1  <-  0.54
-Shih.m2  <-  0.43
-Shih.sd1 <-  0.17
-Shih.sd2 <-  0.16
-Shih.n1  <- 16
-Shih.n2  <- 14
-
-Shih.se.pool <- sqrt( (Shih.sd1^2 + Shih.sd2^2) / 2 )
-Shih.t <- 2.02 # Stated in Shih paper, but is not reproduced by an independent sample t.test. 
-Shih.d <- Shih.t*sqrt(1/(16-1) + 1/(14-1))
-Shih.df <- Shih.n1+Shih.n2-2
-
-
-#### Calculate statistics to verify that the values from Dienes are reproducable ####
-
-Gibson <- tFromDescTwo(m1   = Gibson.m1, 
-                       m2   = Gibson.m2, 
-                       sd1  = Gibson.sd1, 
-                       sd2  = Gibson.sd2, 
-                       n1   = Gibson.n1, 
-                       n2   = Gibson.n2, 
-                       test = "student")
-
-Shih <- tFromDescTwo(m1   = Shih.m1, 
-                     m2   = Shih.m2, 
-                     sd1  = Shih.sd1,
-                     sd2  = Shih.sd2, 
-                     n1   = Shih.n1, 
-                     n2   = Shih.n2, 
-                     test = "student")
-
-#  In Dienes:
-  # For Shih, Pittinsky & Ambady:   t = 2.02, df = 29, p = 0.53 - Not replicated (t = 1.82, df = 28, p = 0.07)
-  # For Gibson, Losee and Vitiello: t = 2.40, df = 81, p = 0.02 - Replicated
-
[…]

### 2017-11-08 Example_Moon_Shih.R

@@ -1,169 +1,0 @@
-##  Example from Dienes (2017): Moon & Roeder (2014) replication of Shih et al. (1999). 
-
-#### Setup ####
-
-library(TOSTER)
-library(pwr)
-#source("Example from Neil Mclathie/2017.10.05 TOSTtwoBayes Raw with Cauchy+HalfCauchy.R")
-#source("TOSTtwoBayes.R")
-
-#### Create functions for calculating statistics ------------------------------
-
-# Create function for calculating t value, cohens d, pooled standard error, etc. from means, standard deviations and group n's.
-
-tFromDescTwo <- function(m1, m2, sd1, sd2, n1, n2, test = "welch") {
-  
-  output        <- list()       # Define the output as list
-  output$m1     <- m1           # Store input arguments in output
-  output$m2     <- m2
-  output$sd1    <- sd1
-  output$sd2    <- sd2
-  output$n1     <- n1
-  output$n2     <- n2
-  output$m.diff <- m1 - m2      # Calculate the mean difference
-  
-  
-  if (test == "student") {  
-    # Calculate se.diff, t and d based on a student t.test
-    print("student's t.test is performed and sample sizes are assumed to be equal")
-    output$test       <- "student"
-    output$df         <- n1 + n2 - 2  # Calculate degrees of freedom.
-    output$sd.pooled  <- sqrt(((( n1 - 1) * (sd1^2)) + (n2 - 1) * (sd2^2)) / ((n1 + n2) - 2 )) # Calculate the pooled SD
-    output$se.diff    <- output$sd.pooled * sqrt(1/n1 + 1/n2) # Calculate the SE of the difference.
-    
-  } else if (test == "welch") {  
-    # Calculate se.diff, t and d based on a welch t.test, and calculate sd.pooled
-    print("welch's t.test is performed and sample sizes are assumed to be unequal")
-    output$test       <- "welch"
-    output$df         <- (sd1^2 / n1 + sd2^2 /n2)^2 / (((sd1^2 / n1)^2 / (n1-1)) + ((sd2^2 / n2)^2 / (n2 - 1))) #degrees of freedom for Welch's t-test
-    output$sd.pooled  <- sqrt((sd1^2 + sd2^2) / 2)    # Calculate sd root mean squared for Welch's t-test
-    output$se.diff    <- sqrt(sd1^2 / n1 + sd2^2 / n2)  # Calculate the se of the difference based on sd.pooled
-    
-    
-  } else {
[…]

### 2017-11-08 Example_Williams_Lynott.R

@@ -1,66 +1,0 @@
-#  TOST on the values from Lynott, Corker, Wortman, Lynott et al. (2014) example
-  # TOST test are calculated using difference between proportions as effect parameter
-
-#### Setup ####
-
-library(TOSTER)
-library(pwr)
-source("TOSTtwo.prop.R")
-source("zDeskProp.R")
-
-#### Data ####
-
-Williams.prop1 <- 0.75
-Williams.prop2 <- 0.46
-# Total n was 53. From Williams & Barge: Approximately half of the participants briefly held a hot pad, and the remaining participants held a cold pad.
-Williams.n1    <- 26 # Approximated
-Williams.n2    <- 27 # Approximated
-
-Lynott.prop1  <- 0.512
-Lynott.prop2  <- 0.579
-Lynott.n1     <- 427
-Lynott.n2     <- 434
-
-#### Calculate z for examples ####
-
-Williams <- zDeskProp(prop1 = Williams.prop1, 
-                      prop2 = Williams.prop2, 
-                      n1    = Williams.n1, 
-                      n2    = Williams.n2)
-
-Lynott <- zDeskProp(prop1 = Lynott.prop1, 
-                     prop2 = Lynott.prop2, 
-                     n1    = Lynott.n1, 
-                     n2    = Lynott.n2)
-
-(0.512/0.488)/(0.579/0.421)
-
-#### Calculate subjective equivalence bounds based on small telescopes and smallest detectable effect size ####
-
-#  Small telescopes
-
-h = 0.4174618
-# Solve equation in Wolfram using Lynott p1 as p1 and solve for p2
-sin(0.797399 - 0.5 * h)^2
-# Solve equation in Wolfram using Lynott p1 as p2 and solve for p2
-sin(0.5 * h + 0.797399)^2
-# Solve equation in Wolfram using Lynott p2 as p1 and solve for p2
-sin(0.864731 - 0.5 * h)^2
-# Solve equation in Wolfram using Lynott p2 as p2 and solve for p1
-sin(0.5 * h + 0.864731)^2
-
-
-pwr.2p2n.test(n1 = 26, n2 = 27, sig.level = 0.05, power = 0.33, alternative = "two.sided")
-pwr.2p.test(sig.level = 0.05, power = 0.33, alternative = "two.sided", n = 3333)
-
-
-#  Smallest detectable effect size
[…]

### 2017-11-08 manuscript.Rmd
Fixed formatting in example plotgrid
* Figure 2 now formatted in the style of figure 1
* Minor change in manuscript to supress metaanalysis output

@@
 
-```{r, fig.width=6}
+```{r, include=FALSE}
 source("SEQ_Hyde_Example.R")

### 2017-11-08 examples_plotgrid.R

@@ -13,7 +13,7 @@ library("TOSTER")
 plot.scaler <- 1.8
 nhst.size <- 0.5
 tost.size <- 1
-point.size <- 2
+point.size <- 1.5
 
 ####  Base plot ----------------------------------------
 
@@ -22,9 +22,11 @@ baseplot <- ggplot(data.frame()) +
   coord_flip() + # flip coordinates (puts labels on y axis)
   theme_classic(base_size = 10) + # use a white background
   geom_hline(yintercept=0, lty=1) +
-  theme(axis.title.y = element_blank(),
+  theme(plot.title = element_text(size = rel(1), face = "bold"),
+        axis.title.y = element_blank(),
         axis.text.y = element_blank(),
-        axis.ticks.y = element_blank())
+        axis.ticks.y = element_blank(),
+        axis.title.x = element_text(size=rel(0.7), lineheight = 0.5))
 
 ####  Meta analysis ----------------------------------------
 
@@ -54,12 +56,14 @@ metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
   geom_hline(yintercept=sesoi, lty=2) +
   geom_hline(yintercept=0, lty=1) +
   coord_flip() +  # flip coordinates (puts labels on y axis)
-  labs(title = "B - Example 2 (Hyde)", y = "Effect (Z)") +
+  labs(title = "B: Example 2 (Hyde et al.)", y = "Effect (Z)") +
   ylim(c(-sesoi*plot.scaler, sesoi*plot.scaler)) +
   theme_classic() + # use a white background
-  theme(axis.title.y = element_blank(),
+  theme(plot.title = element_text(size = rel(1), face = "bold"),
+        axis.title.y = element_blank(),
         axis.text.y = element_blank(),
-        axis.ticks.y = element_blank()) # use a white background
+        axis.ticks.y = element_blank(),
+        axis.title.x = element_text(size=rel(0.7), lineheight = 0.5))
 
 
 ####  Brandt ----------------------------------------
@@ -71,7 +75,7 @@ brandtplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Brandt$diff, ymin = Brandt$LL_CI_TOST, ymax = Brandt$UL_CI_TOST, size = tost.size, fatten = point.size) +
[…]

### 2017-11-09 manuscript.Rmd
Updated manuscript
* Copy-pasted latest version of Google doc (09-11-2017)
* Fixed minor issues and spelling errors

--- 2017-11-08 manuscript.Rmd
@@
 For an alpha level of 5%, we can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI).
-A 90% confidence interval is used because one-sided tests are performed, and it is not necessary to control for multiple comparisons when performing two one-sided tests because both one-sided tests need to be statistically significant to conclude equivalence.
+A 90% confidence interval is used because one-sided tests are performed, and it is not necessary to control for multiple comparisons when performing two one-sided tests because both tests need to be statistically significant to conclude equivalence.
 Note that although we will use an alpha of 5% in all examples because this alpha level was used in the papers we reanalyze, researchers should carefully decide upon a desired Type 1 error rate, and justify their choice (Lakens et al., 2017).
@@
 Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI).
-Although in the examples here, we use symmetrical bounds around 0 (e.g., -0.3, 0.3), based on for example an explicit cost-benefit analysis, one might use asymmetric bounds (e.g., -0.2, 0.3).
+Although we use symmetrical bounds around 0 in the examples we will discuss below (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is possible to use asymmetric bounds, for example based on an explicit cost-benefit analysis (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$).
 The SESOI can sometimes be determined objectively, for example based on just noticeable differences .
@@
 
-We distinguish between three categories of subjective justifications for SESOI.
+We distinguish between three categories of subjective justifications for the SESOI.
 First, researchers can use benchmarks.
@@
 The amount of data you can collect limits the inferences you can make.
-Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.
+Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect[^footnote].
 This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
 In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size.
-For example, a researcher who plans to perform a two-sided one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364.
-Using equivalence bounds of d = -0.364 and d = 0.364 in an equivalence test allows one to reject effects the study had high statistical power to detect.
-An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that can be reliably detected given a specific sample size?
-In this case, concluding equivalence would suggest one can reject effects that can reasonably be detected with a sample size of 100, and that future studies probably need larger sample sizes to examine a specific question.
+For example, a researcher who plans to perform a (two-sided) one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364.
+Using equivalence bounds of $\Delta_{L} = -0.364$ and $\Delta_{U} = 0.364$ in an equivalence test allows one to reject effects the study had high statistical power to detect.
+An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that could reliably be detected given a specific sample size?
+In this case, concluding equivalence would suggest that effects which could reasonably be detected with a sample size of 100 can be rejected, and that future studies probably need larger sample sizes to examine a specific question.
 Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area.
-One should not expect an equivalence test based on 15 observations testing against a SESOI of d = -1 and d = 1 to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
+An equivalence test based on 15 observations testing against a SESOI of d = 1 cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
 By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is.
@@
 ```
+
+[^footnote]: This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
+However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.
+
 #Raw versus standardized equivalence bounds
@@
 
+In the remainder of this article we will provide five detailed examples of equivalence tests performed on published studies.
+These concrete and easy to follow examples will illustrate all approaches to setting equivalence bounds discussed above, and demonstrate how to perform and report equivalence tests.
+The code to reproduce these analyses is available at XXXX.XXXX.
+
+```{r include=FALSE}
+source("examples_plotgrid.R")
+```
+```{r echo=FALSE}
+#  Plot the TOST examples
+plot(example.grid)
+```
+
+Figure caption: Figure 2.
+Effects with TOST CIs (thick lines) and NHST CIs (thin lines), NHST H0 (solid vertical line) and TOST H0 (dashed vertical lines) displayed.
+**A)** Example 1 - The mean difference.
+**B)** Example 2 - The effects included in the meta-analysis.
+**C)** Example 3 - The mean difference.
+**D)** - The proportion difference.
+
 ## Example 1: Statistically Equivalent and Not Statistically Different
@@
 Following the small telescopes approach, we can calculate the effect size the original study had $33\%$ power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.
-If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.
+If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$ (Figure 2A).
 
@@
 
-```{r, fig.width=6}
+```{r, include=FALSE}
 source("SEQ_Hyde_Example.R")
 ```
+
 Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1.
@@
 The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
-Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{lower}$ = -0.1 and $\Delta_{upper}$ = 0.1).
+Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
 
@@
 (2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
-They observed that `r selfish.cold * 100`\% of participants who received a cold pack ($n_{cold}=`r n.cold`$) opted to receive a reward for themselves, while `r selfish.hot * 100`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
-In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r Lynott$dif * 100`$\%, $Z=`r Lynott$NHST_z`$, $p=`r Lynott$NHST_p`$).
+They observed that `r round(selfish.cold,2)*100`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot,2)*100`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
+In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r round(Lynott$dif * 100)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$).
 Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.
@@
 However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
-This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- round(qnorm(1-0.05/2),2)$).
+This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- `r round(qnorm(1-0.05/2),2)`$).
 To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
@@
 Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study.
-In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{upper}$ because an effect with CIs crossing $\Delta_{lower}$ can still be *lower* than the original effect.
+In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{U}$ because an effect with CIs crossing $\Delta_{L}$ can still be *lower* than the original effect.
 
-We find that the test against $\Delta_{upper}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D).
+We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D).
 Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.
@@
 
-```{r include=FALSE}
-source("examples_plotgrid.R")
-```
-```{r echo=FALSE}
-plot(example.grid)
+## Box 1: Calculating an equivalence test in R
+
+If you want to calculate an equivalence test based on summary statistics, you can use the "TOSTER" package in the open-source programming language R.
+Using TOSTER requires very little experience with R, or programming in general.
+Below are the few simple steps required to calculate an equivalence test with TOSTER.
+
+1. Download, install, and open R.
+2. Run `install.packages("TOSTER")` in the console.
+3. Run `library(TOSTER)` in the console.
+4.
+The example below reproduces the result of example 1 in R.
+The parameters of the test are defined inside the parentheses, and you can run `?TOSTtwo` in the console to get information about them.
+Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code.
+Results and a plot will automatically be printed.
+```{r eval=FALSE, echo=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60)} 
+TOSTtwo(m1  = 4.7857, m2  = 4.6569,sd1 = 1.0897, sd2 = 1.1895, n1 = 49, n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05, var.equal = FALSE)
 ```
 
-#Adressing Trello cards
-
-For the example where the results are inconclusive, explain how adding more data, will, assuming the true effect is 0, lead to a smaller confidence interval, and thus will lead to an informative test.
-However, with increasing data, the effect could also slightly increase, to above or at the SESOI level, and then the effect will be significant and not equivalent.
-With massive amounts of data, many results will be significant and equivalent.
+5.
+The example above uses the `TOSTtwo()` function in the TOSTER package which calculates an equivalence test for an independent samples t-test.
+If you need to calculate TOST for a different type of test (e.g.
+a correlation), you can find documentation for `TOSTr()` and other functions of the TOSTER package by going to https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf in your web browser.
 
@@
 
-The given examples and figures XXX to YYY illustrate that equivalence testing is a relatively simple and flexible technique: As long as we can calculate a confidence interval around a mean effect, we can compare it to one or several bounds.
-In that sense, the result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman & Serlin, 1998; Tryon, 2001).
-This ease of use might tempt researchers to set equivalence bounds *after* looking at their data.
-For instance, in example 3 we showed that Moon & Roeder's replication of the identity-priming study by Shih et al.
-(1999; example 3) does not exclude effects that are large enough to affect a student's final grade on a test.
-However, if we look at the $90 \%$ CI (figure XXX), we can see that the data would have been statistically equivalent if the bounds had been set to  a slightly larger SESOI, e.g.
-a mean difference of 0.1.
-It is important to point out that this strategy --- setting equivalence bounds *based on the obtained results* --- may be considered as "boundary hacking": Regardless of the underlying true effect, any finite CI will exclude some values, which means that this procedure will lead to capitalization on chance.
+Equivalence testing is a relatively simple and flexible technique to falsify hypotheses.
+As long as we can calculate a confidence interval around a parameter estimate of interest, we can compare it to a smallest effect size of interest.
+The result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman & Serlin, 1998; Tryon, 2001), or by performing two one-sided tests.
 
-In order to avoid fooling ourselves more often than $\alpha \%$ of the time, equivalence bounds should be specified before results are known (ideally as part of a preregistration; cf.
+Equivalence bounds should be specified before results are known, ideally as part of a pre-registration (cf.
 CONSORT guidelines, Piaggio et al., 2006).
-A perhaps helpful guiding question may be: "Will testing against these bounds yield interesting information in the long run?"
+Researchers might be tempted to set equivalence bounds *after* looking at their data, but this "boundary hacking" will increase error rates.
 
-#Justify your equivalence bounds
+As with any statistical test, the usefulness of the result of an equivalence test depends on the question we ask.
+The question manifests itself in the bounds we set: Is our data surprising if the true effect lies outside of the bounds?
+If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003).
+Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)[^footnote].
+An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
 
-Like with any statistical test, the usefulness of the result an equivalence test gives us depends on the question we ask.
-The question manifests in the bounds we set: Is our data surprising if the true effect lies outside of the bounds?
-If we set very wide bounds, we ask if our data is surprising given that the true effect is very large[^footnote].
-For most psychological effects, the answer will usually be “yes,” which means that it may not be a very interesting question to ask.
-In order to obtain a meaningful result, we have to set the bounds based on meaningful criteria: Our SESOI should really be *the smallest effect we are interested in* for a given study.
-For this reason, researchers should provide a detailed justification of the choice for a SESOI.
-Some studies in the literature simply state the SESOI that is used, without providing a justification.
-For example, Brown and colleagues (2017) and Schumann and colleagues (2017)use a SESOI of Cohen’s d = +/- 0.5, but it is not clear why these bounds were chosen[^footnote].
-An equivalence test should answer a question, and justifying the SESOI is an essential part of asking the question.
-An equivalence test of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but the reason this SESOI was chosen should be explicitly justified.
+For many researchers, the biggest challenge in designing studies where they will perform both NHST and equivalence tests by default will be the need to specify the smallest effect size of interest.
+Psychological theories are usually too vague to derive precise predictions, and if there are no theoretical reference points, natural constraints, or prior studies a researcher can use to define the SESOI for a new hypothesis, any choice will be arbitrary to some extent.
+In some research lines researchers might simply use equivalence tests to reject consecutively smaller effect sizes by performing studies with increasingly larger sample sizes while controlling error rates, until no one is willing to invest the time and resources needed to examine the presence of even smaller effects.
+Nevertheless, it is important to realize that not specifying a SESOI in our research questions will severely hinder theoretical progress.
+While our initial choices may be arbitrary and subjective, our goal should be to use them as starting points to develop better and more falsifiable theories over time, and ultimately be able to make more precise predictions than we can today.
 
-[^footnote]: To be precise, wide bounds set symmetrically around 0 ask if the data is surprising given that the true effect is very *extreme*, i.e.
-either very large (positive) or very small (negative).
 [^footnote]: In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.

### 2017-11-09 examples_plotgrid.R
Minor formatting fix

@@ -50,15 +50,15 @@ sesoi <- 0.1
 
 metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
   scale_x_discrete(breaks=NULL) +
-  geom_pointrange(size = 0.6, fatten = point.size) + 
-  geom_pointrange(aes(ymin=li90, ymax=ui90), size = 0.5, fatten = point.size) +
+  geom_pointrange(size = nhst.size, fatten = point.size/2) + 
+  geom_pointrange(aes(ymin=li90, ymax=ui90), size = tost.size, fatten = point.size/2) +
   geom_hline(yintercept=-sesoi, lty=2) +
   geom_hline(yintercept=sesoi, lty=2) +
   geom_hline(yintercept=0, lty=1) +
   coord_flip() +  # flip coordinates (puts labels on y axis)
   labs(title = "B: Example 2 (Hyde et al.)", y = "Effect (Z)") +
   ylim(c(-sesoi*plot.scaler, sesoi*plot.scaler)) +
-  theme_classic() + # use a white background
+  theme_classic(base_size = 10) + # use a white background
   theme(plot.title = element_text(size = rel(1), face = "bold"),
         axis.title.y = element_blank(),
         axis.text.y = element_blank(),

### 2017-11-09 examples_plotgrid.R
Minor formatting fix

@@ -50,15 +50,15 @@ sesoi <- 0.1
 
 metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
   scale_x_discrete(breaks=NULL) +
-  geom_pointrange(size = 0.6, fatten = point.size) + 
-  geom_pointrange(aes(ymin=li90, ymax=ui90), size = 0.5, fatten = point.size) +
+  geom_pointrange(size = nhst.size, fatten = point.size) + 
+  geom_pointrange(aes(ymin=li90, ymax=ui90), size = tost.size, fatten = point.size) +
   geom_hline(yintercept=-sesoi, lty=2) +
   geom_hline(yintercept=sesoi, lty=2) +
   geom_hline(yintercept=0, lty=1) +
   coord_flip() +  # flip coordinates (puts labels on y axis)
   labs(title = "B: Example 2 (Hyde et al.)", y = "Effect (Z)") +
   ylim(c(-sesoi*plot.scaler, sesoi*plot.scaler)) +
-  theme_classic() + # use a white background
+  theme_classic(base_size = 10) + # use a white background
   theme(plot.title = element_text(size = rel(1), face = "bold"),
         axis.title.y = element_blank(),
         axis.text.y = element_blank(),

### 2017-11-09 manuscript.Rmd
Fixed the footnote issue (changed all footnotes to inline footnotes)

@@
 Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
-This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*[^footnote].
+This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*^[This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect (see Lenth, 2007).
+This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect.].
 All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
 By setting the SESOI to the critical effect size, an equivalence test can reject all observed effect sizes that could have been detected in an earlier study.
-
-[^footnote]: This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect (see Lenth, 2007).
-This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect.
 
@@
 The amount of data you can collect limits the inferences you can make.
-Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect[^footnote].
+Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
+However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.].
 This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
@@
 ```
-
-[^footnote]: This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
-However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.
 
@@
 We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C).
-Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
-Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
-The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
-Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
-
-[^footnote]: In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
+Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
 We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
@@
 Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
-As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.
+As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
+The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
+Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
 
@@
 Kahane et al.
-find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$[^footnote]).
+find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$^[Study 4 in Kahane et al.
+(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of only $N = `r Kahane.N`$.]).
 They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
@@
 This means that values smaller than $r = `r -round(Kahane.r.80, 2)`$ and larger than $r = `r round(Kahane.r.80, 2)`$ can be rejected at an alpha level of $5\%$ and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.
-
-[^footnote]: Study 4 in Kahane et al.
-(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of only $N = `r Kahane.N`$.
 
@@
 If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003).
-Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)[^footnote].
+Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.
+].
 An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
@@
 While our initial choices may be arbitrary and subjective, our goal should be to use them as starting points to develop better and more falsifiable theories over time, and ultimately be able to make more precise predictions than we can today.
-
-[^footnote]: In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.
 

### 2017-11-09 manuscript.Rmd
fixed rounding issues in the Lynott example

@@
 author_note: >
-  We would like to thank Courtney Soderbergh for creating the first version of the TOST function to test two independent proportions.
+  We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions.
 
@@
 
-However, even if the researcher would consider any difference in percentages smaller than $-0.5 \%$ or larger than $3.5 \%$ a rejection of the null hypothesis, not every difference is large enough to matter in practice.
+However, even if the researcher would consider any difference in percentages smaller than -0.5% or larger than 3.5% a rejection of the null hypothesis, not every difference is large enough to matter in practice.
 As long as gender differences in the participation rate are not too extreme, there is no need to spend money to address the gender imbalance in the voluntary participation rate, and all money can be spent on the core training program.
@@
 Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
-This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*^[This will typically, although not always, correspond to the effect size the study had $50\%$ power to detect (see Lenth, 2007).
-This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had $33\%$ power to detect.].
+This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*^[This will typically, although not always, correspond to the effect size the study had 50% power to detect (see Lenth, 2007).
+This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.].
 All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
@@
 A close replication by Brandt, IJzerman, & Blanken (2014) with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
-Following the small telescopes approach, we can calculate the effect size the original study had $33\%$ power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.
+Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.
 If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$ (Figure 2A).
@@
 (2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
-They observed that `r round(selfish.cold,2)*100`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot,2)*100`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
-In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r round(Lynott$dif * 100)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$).
+They observed that `r round(selfish.cold*100,2)`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot*100,2)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
+In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r round(Lynott$dif * 100, 2)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$).
 Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.
@@
 206).
-This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had $80\%$ power to detect effects of $r = `r round(Kahane.r.80, 2)`$.
+This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r round(Kahane.r.80, 2)`$.
 Given bounds of $\Delta_{L} = `r -round(Kahane.r.80, 2)`$ and $\Delta_{U} = `r round(Kahane.r.80, 2)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(max(Kahane$TOST_p1, Kahane$TOST_p2), 3)`$.
-This means that values smaller than $r = `r -round(Kahane.r.80, 2)`$ and larger than $r = `r round(Kahane.r.80, 2)`$ can be rejected at an alpha level of $5\%$ and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.
+This means that values smaller than $r = `r -round(Kahane.r.80, 2)`$ and larger than $r = `r round(Kahane.r.80, 2)`$ can be rejected at an alpha level of 5% and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.
 

### 2017-11-09 manuscript.Rmd
Fixed odd-looking spacing for variable name in inline math

@@
 They observed that `r round(selfish.cold*100,2)`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot*100,2)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
-In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r round(Lynott$dif * 100, 2)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$).
+In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r round(Lynott$dif * 100, 2)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$).
 Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.

### 2017-11-09 manuscript.Rmd
Fixed minor issue caused by earlier merge conflict

@@
 
-```{r echo=FALSE, fig.width=5, fig.height=7}
+```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap=paste("put your figure caption here")}
 #  Plot H0 and H1 regions for different significance tests
@@
 
-Figure caption: Figure 1.
-Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
+Figure 1: Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
 **A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\Delta_{L}$ *or* larger than $\Delta_{U}$.
@@
 The amount of data you can collect limits the inferences you can make.
-Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
-However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.].
+Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect[^footnote].
 This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
@@
 ```
+
+[^footnote]: This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
+However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.
 
@@
 ```
-```{r echo=FALSE}
+```{r echo=FALSE, fig.cap=paste("Example effects plotted with TOST CIs (thick lines) and NHST CIs (thin lines), NHST H0 (solid vertical line) and TOST H0 (dashed vertical lines) displayed.
+**A)** Example 1 - The mean difference.
+**B)** Example 2 - The effects included in the meta-analysis.
+**C)** Example 3 - The mean difference.
+**D)** Example 4 - The proportion difference.
+**E)** Example 5 - The pearson correlation")}
 #  Plot the TOST examples
@@
 ```
-
-Figure caption: Figure 2.
-Effects with TOST CIs (thick lines) and NHST CIs (thin lines), NHST H0 (solid vertical line) and TOST H0 (dashed vertical lines) displayed.
-**A)** Example 1 - The mean difference.
-**B)** Example 2 - The effects included in the meta-analysis.
-**C)** Example 3 - The mean difference.
-**D)** - The proportion difference.
 
@@
 We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C).
-Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
+Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
+The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
+Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
+
+[^footnote]: In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
 We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
@@
 Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
-As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
-Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
-The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
-Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
+As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.
 
@@
 If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003).
-Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.
-].
+Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)[^footnote].
 An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
@@
 While our initial choices may be arbitrary and subjective, our goal should be to use them as starting points to develop better and more falsifiable theories over time, and ultimately be able to make more precise predictions than we can today.
+
+[^footnote]: In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.
 

### 2017-11-09 manuscript.Rmd
Fixed caption for figure 1

@@
 
-```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap=paste("put your figure caption here")}
+```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
+**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
+**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
+**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
+**D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}
+
 #  Plot H0 and H1 regions for different significance tests
@@
 ```
-
-Figure 1: Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
-**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\Delta_{L}$ *or* larger than $\Delta_{U}$.
-**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
-**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\Delta_{L}$ *and* smaller than $\Delta_{U}$.
-**D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\Delta$.
 

### 2017-11-09 manuscript.Rmd
Changed the footnotes back to inline note (still need to fix pagebreak problem in footnotes), fixed +/- symbols

--- 2017-11-09 manuscript.Rmd
@@
 
-```{r echo=FALSE, fig.width=5, fig.height=7}
+```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
+**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
+**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
+**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
+**D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}
+
 #  Plot H0 and H1 regions for different significance tests
@@
 ```
-
-Figure caption: Figure 1.
-Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
-**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\Delta_{L}$ *or* larger than $\Delta_{U}$.
-**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
-**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\Delta_{L}$ *and* smaller than $\Delta_{U}$.
-**D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\Delta$.
 
@@
 Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.
-
 We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using +/- the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
-
 We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C).
 Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
-We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
+We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
 Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
@@
 Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
-Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
+Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
 As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
@@
 However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
-This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- `r round(qnorm(1-0.05/2),2)`$).
+This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r round(qnorm(1-0.05/2),2)`$).
 To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
@@
 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$^[Study 4 in Kahane et al.
-(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of only $N = `r Kahane.N`$.]).
+(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
 They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
@@
 If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003).
-Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.
-].
+Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.].
 An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.

### 2017-11-09 manuscript.Rmd
Minor formatting fix, placed periods before footnotes rather than after

@@
 Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
-This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*^[This will typically, although not always, correspond to the effect size the study had 50% power to detect (see Lenth, 2007).
-This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.].
-All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
+This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*.^[This will typically, although not always, correspond to the effect size the study had 50% power to detect (see Lenth, 2007).
+This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.] All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
 By setting the SESOI to the critical effect size, an equivalence test can reject all observed effect sizes that could have been detected in an earlier study.
@@
 The amount of data you can collect limits the inferences you can make.
-Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
-However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.].
-This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
+Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
+However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
 In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size.
@@
 Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.
-We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using +/- the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
+We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using $\pm$ the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
 We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C).
@@
 Equivalence bounds should be specified before results are known, ideally as part of a pre-registration (cf.
-CONSORT guidelines, Piaggio et al., 2006).
+Piaggio et al., 2006).
 Researchers might be tempted to set equivalence bounds *after* looking at their data, but this "boundary hacking" will increase error rates.
@@
 If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003).
-Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.].
-An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
+Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017).^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
 

### 2017-11-09 manuscript.Rmd
Fixed all citations (to do: change first citation in Kahane example)

@@
 For example, we might predict that priming American Asian women with their Asian identity will increase their performance on a math test compared to women who are primed with their female identity.
-To be able to falsify this hypothesis, and design a study that allows for strong inferences (Platt, 1964), it is important to specify which test result would *disprove* the hypothesis.
+To be able to falsify this hypothesis, and design a study that allows for strong inferences [@Platt1964], it is important to specify which test result would *disprove* the hypothesis.
 
@@
 However, this would ignore the fact that the human sex ratio of newborn boys to girls is not exactly 1:1, and we should not expect 50% of participants to be boys.
-On average, 103 boys are born for every 100 girls (CIA, 2017), so approximately 50.74% of applicants should be boys, and 49.26% should be girls (with $0.5074/0.4926 \approx 1.03$).
+On average, 103 boys are born for every 100 girls [@UnitedStates2016], so approximately 50.74% of applicants should be boys, and 49.26% should be girls (with $0.5074/0.4926 \approx 1.03$).
 If boys and girls were exactly equally likely to be brought into the program by their parents, we would not expect a difference of zero, but 1.5% more boys.
@@
 The researcher can for example test if the difference is smaller than -0.005, or larger than 0.035.
-This test against two bounds, with $H_0$ as a range rather than one value, is known as a *minimal effects test* (Murphy, Myors, & Wolach, 2014).
+This test against two bounds, with $H_0$ as a range rather than one value, is known as a *minimal effects test* [@Murphy2014].
 
@@
 A 90% confidence interval is used because one-sided tests are performed, and it is not necessary to control for multiple comparisons when performing two one-sided tests because both tests need to be statistically significant to conclude equivalence.
-Note that although we will use an alpha of 5% in all examples because this alpha level was used in the papers we reanalyze, researchers should carefully decide upon a desired Type 1 error rate, and justify their choice (Lakens et al., 2017).
+Note that although we will use an alpha of 5% in all examples because this alpha level was used in the papers we reanalyze, researchers should carefully decide upon a desired Type 1 error rate, and justify their choice [@Lakens2017b].
 Figure 1 illustrates the possible conclusions we can draw from null hypothesis, minimal effect, and equivalence tests.
@@
 
-Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, rather than against 0), their use in psychology was limited, perhaps due to the lack of user-friendly software to perform the calculations in the past [@lakens_equivalence_2017].
+Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, rather than against 0), their use in psychology was limited, perhaps due to the lack of user-friendly software to perform the calculations in the past [@Lakens2017a].
 In this article, we provide several examples of equivalence tests which illustrate the procedure in free software (R, a spreadsheet, and in Jamovi).
@@
 In such circumstances, the SESOI can be set based on just noticeable differences.
-For example, Burriss and colleagues [-@burriss_changes_2015] examined whether women displayed an increase in redness in the face during the fertile phase of their ovulatory cycle.
+For example, Burriss and colleagues [-@Burriss2015] examined whether women displayed an increase in redness in the face during the fertile phase of their ovulatory cycle.
 The hypothesis was that a slightly redder skin signals greater attractiveness and physical health, and sending this signal to men yields an evolutionary advantage.
@@
 
-Another example of an objectively determined SESOI can be found in @button_minimal_2015 where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
+Another example of an objectively determined SESOI can be found in @Button2015 where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
 
-Norman, Sloan, and Wyrwich (2003) used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or d = 0.5.
+@Norman2004 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or d = 0.5.
 They make the argument that this bound of approximately half a standard deviation could be related to the limits of human discriminability.
@@
 First, researchers can use benchmarks.
-For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988).
-Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small (Maxwell, Lau, & Howard, 2015).
+For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size [@Cohen1988].
+Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small [@Maxwell2015].
 Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided.
-Based on a review of 112 meta-analyses, Weber and Popova (2012) conclude that setting a SESOI in communication research to a medium effect size (*r* = 0.3, or d = 0.5) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and Hemphill (2003) suggests that in psychology a SESOI of d = 0.5 would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
+Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI in communication research to a medium effect size (*r* = 0.3, or d = 0.5) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that in psychology a SESOI of d = 0.5 would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
 
@@
 It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study.
-Simonsohn (2015) recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have had 33% power to detect.
+@Simonsohn2015 recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have had 33% power to detect.
 With this so-called 'small telescopes' approach, the equivalence bounds are thus determined primarily based on the sample size in the original study.
@@
 For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`.
-Another example of how previous research can be used to determine the SESOI can be found in Kordsmeyer and Penke (2017), who defined the SESOI based on the mean of effect sizes reported in the literature.
+Another example of how previous research can be used to determine the SESOI can be found in @Kordsmeyer2017, who defined the SESOI based on the mean of effect sizes reported in the literature.
 Thus, they examined whether their replication study could reject effects as large or larger than had on average been reported in the literature.
-Given random variation and bias in the literature, a more conservative approach could be to use a lower bound of a confidence interval around the meta-analytic effect size estimate (cf.
-Perugini, Gallucci, & Constantini, 2014).
+Given random variation and bias in the literature, a more conservative approach could be to use a lower bound of a confidence interval around the meta-analytic effect size estimate [cf.
+@Perugini2014].
 
@@
 Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
-This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*.^[This will typically, although not always, correspond to the effect size the study had 50% power to detect (see Lenth, 2007).
+This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*.^[This will typically, although not always, correspond to the effect size the study had 50% power to detect [see @Lenth2007].
 This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.] All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
@@
 Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
-However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
+However, @Meyners2012 explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
 In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size.
@@
 
-Both raw equivalence bounds and standardized equivalence bounds have specific benefits and limitations (for a discussion, see Baguley, 2009).
+Both raw equivalence bounds and standardized equivalence bounds have specific benefits and limitations [for a discussion, see @Baguley2009].
 When raw mean differences are meaningful and of theoretical interest, it makes sense to set equivalence bounds based on raw effect sizes.
@@
 
-Banerjee, Chatterjee, & Sinha (2012) reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
-A close replication by Brandt, IJzerman, & Blanken (2014) with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
+@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
+A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
 Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.
@@
 
-Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1.
+@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1.
 For example, in grade 3 the difference is d = 0.04, with a standard error of 0.002.
@@
 
-Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
+@Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
 They observed a slightly reversed difference between the Asian primed group (*n* = 53, *M* = 0.46, *SD* = 0.17) and the control (*n* = 48, *M* = 0.50, *SD* - 0.18) which was not significant, d = -0.23, *t*(96.62) = 1.15, *p* = 0.26, two-sided).
@@
 
-Lynott et al.
-(2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008).
+@Lynott2014 conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior [replicating @Williams2008].
 They observed that `r round(selfish.cold*100,2)`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot*100,2)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
@@
 In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{U}$ because an effect with CIs crossing $\Delta_{L}$ can still be *lower* than the original effect.
-
 We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D).
@@
 
-In a paper published in 2015, Kahane and colleagues investigate moral dilemma vignettes in which participants have to decide whether or not they would sacrifice one person's life to save several other lives, or judge how morally admissible such actions are.
+In a paper published in [-@Kahane2015], Kahane and colleagues investigate moral dilemma vignettes in which participants have to decide whether or not they would sacrifice one person's life to save several other lives, or judge how morally admissible such actions are.
 Traditionally, greater endorsement for sacrificing a life to save others has been interpreted as a more "utilitarian" moral orientation, i.e.
@@
 
-Kahane et al.
-find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$^[Study 4 in Kahane et al.
-(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
+@Kahane2015 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
 They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
@@
 As long as we can calculate a confidence interval around a parameter estimate of interest, we can compare it to a smallest effect size of interest.
-The result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman & Serlin, 1998; Tryon, 2001), or by performing two one-sided tests.
+The result of an equivalence test can be obtained by mere visual inspection of the CI [@Seaman1998; @Tryon2001], or by performing two one-sided tests.
 
-Equivalence bounds should be specified before results are known, ideally as part of a pre-registration (cf.
-Piaggio et al., 2006).
+Equivalence bounds should be specified before results are known, ideally as part of a pre-registration [cf.
+@Piaggio2006].
 Researchers might be tempted to set equivalence bounds *after* looking at their data, but this "boundary hacking" will increase error rates.
@@
 The question manifests itself in the bounds we set: Is our data surprising if the true effect lies outside of the bounds?
-If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003).
-Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017).^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
+If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003].
+Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2004], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
 
@@
 
-Button, K.
-S., Kounali, D., Thomas, L., Wiles, N.
-J., Peters, T.
-J., Welton, N.
-J., .
-Lewis, G.
-(2015).
-Minimal clinically important difference on the Beck Depression Inventory - II according to the patient's perspective.
-Psychological Medicine, 45(15), 3269-3279.
-https://doi.org/10.1017/S0033291715001270
-
-Burriss, R.
-P., Troscianko, J., Lovell, P.
-G., Fulford, A.
-J.
-C., Stevens, M., Quigley, R., .
-Rowland, H.
-M.
-(2015).
-Changes in Women's Facial Skin Color over the Ovulatory Cycle are Not Detectable by the Human Visual System.
-PLOS ONE, 10(7), e0130093.
-https://doi.org/10.1371/journal.pone.0130093
-
-Lakens, D. (2017). Equivalence Tests: A Practical Primer for t Tests, Correlations, and Meta-Analyses. Social Psychological and Personality Science, 8(4), 355–362.
-
 \setlength{\parindent}{-0.5in}

### 2017-11-10 manuscript.Rmd
Edits in example 3 and 4

--- 2017-11-09 manuscript.Rmd
@@
 They observed a slightly reversed difference between the Asian primed group (*n* = 53, *M* = 0.46, *SD* = 0.17) and the control (*n* = 48, *M* = 0.50, *SD* - 0.18) which was not significant, d = -0.23, *t*(96.62) = 1.15, *p* = 0.26, two-sided).
-A non-significant null-hypothesis test leaves two possible conclusions:
-
-1. The data suggest the absence of a meaningful effect, and we can reject effects large enough to be meaningful.
-2. The data are not sensitive enough to tell us whether a meaningful effect is present or absent.
-
-Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
+A non-significant null-hypothesis test leaves two possible conclusions, As an example, imagine we discover that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
 A+ = 93.75% to 100%).
@@
 Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.
+
 We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using +/- the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
+
 We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C).
-Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
-We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
+Thus, we cannot reject the possibility that the priming manipulation leads to a real change in test scores that is larger on average than one gradepoint.[^footnote] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
+The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
+Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
+
+[^footnote]: In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
+We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
 Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
@@
 Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
-Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
-As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
-Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
-The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
-Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
+Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
+As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.
 
@@
 In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r round(Lynott$dif * 100, 2)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$).
-Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.
 
-In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be.
-However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
-This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r round(qnorm(1-0.05/2),2)`$).
-To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
+In this case, it is not immediately clear what an objective criterion for the sESOI would be.
+However, as this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could potentially detect.
+This means that we will use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- `r round(qnorm(1-0.05/2),2)`$).
+To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error.
 
@@
 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$^[Study 4 in Kahane et al.
-(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
+(2015) had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of only $N = `r Kahane.N`$.]).
 They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
@@
 If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003).
-Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.].
+Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)[^footnote].
 An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.

### 2017-11-10 manuscript.Rmd
Made the text box an actual box, introducing a bunch of LaTeX packages in the head

--- 2017-11-09 manuscript.Rmd
@@
 wordcount         : "X"
+
+header-includes:
+  - \usepackage{float}
+  - \usepackage{framed}
+  - \usepackage{caption}
+  - \usepackage{setspace}
+  - \captionsetup[textbox]{name=Box,labelsep=period,labelfont=it}
+  - \newfloat{textbox}{thp}{lop}
+  - \floatname{textbox}{Box}
 
@@
 
-## Box 1: Calculating an equivalence test in R
-
-If you want to calculate an equivalence test based on summary statistics, you can use the "TOSTER" package in the open-source programming language R.
+\begin{textbox}
+\caption{Calculating an equivalence test in R}
+\begin{framed}
+{
+\small
+\setstretch{1}
+If you want to calculate an equivalence test based on summary statistics, you can use the ``TOSTER'' package in the open-source programming language R.
 Using TOSTER requires very little experience with R, or programming in general.
 Below are the few simple steps required to calculate an equivalence test with TOSTER.
-
-1. Download, install, and open R.
-2. Run `install.packages("TOSTER")` in the console.
-3. Run `library(TOSTER)` in the console.
-4.
-The example below reproduces the result of example 1 in R.
-The parameters of the test are defined inside the parentheses, and you can run `?TOSTtwo` in the console to get information about them.
+\begin{enumerate}
+\raggedright
+  \setlength{\parskip}{1pt}
+\item Download, install, and open R.
+\item Run \verb!install.packages("TOSTER")! in the console.
+\item Run \verb!library(TOSTER)! in the console.
+\item The example below reproduces the result of example 1 in R.
+The parameters of the test are defined inside the parentheses, and you can run \verb!?TOSTtwo!
+in the console to get information about them.
 Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code.
 Results and a plot will automatically be printed.
-```{r eval=FALSE, echo=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60)} 
-TOSTtwo(m1  = 4.7857, m2  = 4.6569,sd1 = 1.0897, sd2 = 1.1895, n1 = 49, n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05, var.equal = FALSE)
-```
-
-5.
-The example above uses the `TOSTtwo()` function in the TOSTER package which calculates an equivalence test for an independent samples t-test.
+\begin{verbatim}
+TOSTtwo(m1  = 4.7857, 
+        m2  = 4.6569,
+        sd1 = 1.0897, 
+        sd2 = 1.1895, 
+        n1 = 49, 
+        n2 = 51, 
+        low_eqbound_d = -0.6401696, 
+        high_eqbound_d = 0.6401696, 
+        alpha = 0.05, 
+        var.equal = FALSE)
+\end{verbatim}
+\item The example above uses the \verb!TOSTtwo()!
+function in the TOSTER package which calculates an equivalence test for an independent samples t-test.
 If you need to calculate TOST for a different type of test (e.g.
-a correlation), you can find documentation for `TOSTr()` and other functions of the TOSTER package by going to https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf in your web browser.
+a correlation), you can find documentation for \verb!TOSTr()!
+and other functions of the TOSTER package by going to {\sffamily \url{https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf}} in your web browser.
+\end{enumerate}}
+\end{framed} 
+\end{textbox}
 

### 2017-11-10 manuscript.Rmd
minor change to R code in Example 1

@@
 
-@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
-A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
+@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r round(orig.p, 3)`$, $d= `r orig.d`$).
+A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r round(rep.df, 2)`)=`r round(rep.t, 2)`$, $p=`r round(rep.p, 3)`$, $d=`r round(rep.d, 2)`$).
 Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.

### 2017-11-12 manuscript.Rmd
Changed plot order in Figure 1, changed all instances of rounding to printnum() or printp(); minor formatting fixes

--- 2017-11-10 manuscript.Rmd
@@
 ```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
-**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
-**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
-**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
+**A)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
+**B)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
+**C)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
 **D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}
@@
 
-@Norman2004 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or d = 0.5.
+@Norman2004 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or $d = 0.5$.
 They make the argument that this bound of approximately half a standard deviation could be related to the limits of human discriminability.
-Therefore, even though a SESOI of d = 0.5 might seem quite large, it could be a plausible equivalence bound when examining effects that should be noticeable by single individuals.
+Therefore, even though a SESOI of $d = 0.5$ might seem quite large, it could be a plausible equivalence bound when examining effects that should be noticeable by single individuals.
 
@@
 First, researchers can use benchmarks.
-For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size [@Cohen1988].
+For example, one might set the SESOI to a standardized effect size of $d = 0.5$, which would allow one to reject effect as large or larger than a 'medium' effect size [@Cohen1988].
 Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small [@Maxwell2015].
 Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided.
-Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI in communication research to a medium effect size (*r* = 0.3, or d = 0.5) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that in psychology a SESOI of d = 0.5 would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
+Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI in communication research to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that in psychology a SESOI of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
 
@@
 For example, consider a study in which `r N` participants answered a question, and the results were analyzed with a one-sample t-test.
-For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`.
+For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of $d = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d)`$.
 Another example of how previous research can be used to determine the SESOI can be found in @Kordsmeyer2017, who defined the SESOI based on the mean of effect sizes reported in the literature.
@@
 In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size.
-For example, a researcher who plans to perform a (two-sided) one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364.
+For example, a researcher who plans to perform a (two-sided) one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of $d = 0.364$.
 Using equivalence bounds of $\Delta_{L} = -0.364$ and $\Delta_{U} = 0.364$ in an equivalence test allows one to reject effects the study had high statistical power to detect.
@@
 Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area.
-An equivalence test based on 15 observations testing against a SESOI of d = 1 cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
+An equivalence test based on 15 observations testing against a SESOI of $d = 1$ cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
 By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is.
@@
 The observed standard deviation randomly varies across samples.
-In practice, this means that when you use standardized differences as bounds (e.g., d = 0.35) the equivalence test depends on the standard deviation in your the sample.
-The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of d = 0.5 when the standard deviation is 1, but a standardized equivalence bound of d = 1 when the standard deviation is 0.5.
+In practice, this means that when you use standardized differences as bounds (e.g., $d = 0.35$) the equivalence test depends on the standard deviation in your the sample.
+The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of $d = 0.5$ when the standard deviation is 1, but a standardized equivalence bound of $d = 1$ when the standard deviation is 0.5.
 
@@
 
-@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$).
-A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$).
-Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI.
-If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$ (Figure 2A).
+@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r printp(orig.p)`$, $d= `r printnum(orig.d)`$).
+A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
+Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r printnum(d.33)`$ --- and use this as our SESOI.
+If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2A).
 
@@
 
-@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1.
-For example, in grade 3 the difference is d = 0.04, with a standard error of 0.002.
-When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of d = -0.1 and d = 0.1, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1, and can be considered trivially small, according to the bounds set by the authors (figure 2B).
+@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than $d = 0.1$.
+For example, in grade 3 the difference is $d = 0.04$, with a standard error of 0.002.
+When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of $d = -0.1$ and $d = 0.1$, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of $d = -0.1$ and $d = 0.1$, and can be considered trivially small, according to the bounds set by the authors (figure 2B).
 However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge.
@@
 @Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
-They observed a slightly reversed difference between the Asian primed group (*n* = 53, *M* = 0.46, *SD* = 0.17) and the control (*n* = 48, *M* = 0.50, *SD* - 0.18) which was not significant, d = -0.23, *t*(96.62) = 1.15, *p* = 0.26, two-sided).
+They observed a slightly reversed difference between the Asian primed group ($n = 53$, $M = 0.46$, $SD = 0.17$) and the control ($n = 48$, $M = 0.50$, $SD = 0.18$) which was not significant, $d = -0.23$, $t(96.62) = 1.15$, $p = 0.26$ (two-sided).
 A non-significant null-hypothesis test leaves two possible conclusions:
@@
 We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using $\pm$ the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
-We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C).
+We find that the TOST is non-significant, t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`, p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))` (figure 2C).
 Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
-We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$.
+We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r printnum(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2))`)$.
 Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
@@
 Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
-Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$.
+Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r printnum(0.0625 / sqrt((0.17^2 + 0.20^2) / 2))`)$.
 As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
@@
 The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
-Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).
+Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$).
 
@@
 @Lynott2014 conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior [replicating @Williams2008].
-They observed that `r round(selfish.cold*100,2)`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot*100,2)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
-In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r round(Lynott$dif * 100, 2)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$).
+They observed that `r printnum(selfish.cold*100)`\% of participants who received a cold pack ($n_{cold}=`r n.cold`$) opted to receive a reward for themselves, while `r printnum(selfish.hot*100)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
+In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $Z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
 Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.
@@
 However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
-This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r round(qnorm(1-0.05/2),2)`$).
+This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
 To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
@@
 In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{U}$ because an effect with CIs crossing $\Delta_{L}$ can still be *lower* than the original effect.
-We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D).
+We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r printnum(Lynott$TOST_z2)`, $p `r printp(Lynott$TOST_p2)`$ (see figure 2D).
 Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.
@@
 
-@Kahane2015 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
-They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
+@Kahane2015 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$ ($N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
+They conclude that the classical vignettes fail to capture "true" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
 206).
-This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r round(Kahane.r.80, 2)`$.
-Given bounds of $\Delta_{L} = `r -round(Kahane.r.80, 2)`$ and $\Delta_{U} = `r round(Kahane.r.80, 2)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(max(Kahane$TOST_p1, Kahane$TOST_p2), 3)`$.
-This means that values smaller than $r = `r -round(Kahane.r.80, 2)`$ and larger than $r = `r round(Kahane.r.80, 2)`$ can be rejected at an alpha level of 5% and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.
+This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$.
+Given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$.
+This means that values smaller than $r = `r printnum(-Kahane.r.80)`$ and larger than $r = `r printnum(Kahane.r.80)`$ can be rejected at an alpha level of 5% and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.
 
 \begin{textbox}
-\caption{Calculating an equivalence test in R}
+\caption{Calculating an equivalence test in R.}
 \begin{framed}
@@
 If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003].
-Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2004], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
+Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2004], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
 

### 2017-11-12 manuscript.Rmd
Added Peder's latest changes (for examples 2-4)

@@
 
-Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
+As an example, imagine we discover that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
 A+ = 93.75% to 100%).
@@
 We find that the TOST is non-significant, t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`, p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))` (figure 2C).
-Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
-We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r printnum(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2))`)$.
+Thus, we cannot reject the possibility that the priming manipulation leads to a real change in test scores that is larger on average than one gradepoint.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
+We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r printnum(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2))`$).
 Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
@@
 Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
-Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r printnum(0.0625 / sqrt((0.17^2 + 0.20^2) / 2))`)$.
+Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r printnum(0.0625 / sqrt((0.17^2 + 0.20^2) / 2))`$).
 As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
@@
 In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $Z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
-Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.
 
-In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be.
-However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect.
-This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
-To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:
+In this case, it is not immediately clear what an objective criterion for the sESOI would be.
+However, as this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could potentially detect.
+This means that we will use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
+To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error.
 

### 2017-11-13 examples_plotgrid.R
Changed example 4 to inferioity test
* Done by removing lower equivalence bound in figure

@@ -99,7 +99,6 @@ source("NSEQ_Lynott_example.R")
 lynottplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Lynott$dif, ymin = Lynott$LL_CI_ZTEST, ymax = Lynott$UL_CI_ZTEST, size = nhst.size, fatten = point.size) +
   annotate(geom = "pointrange", x = 0.5, y = Lynott$dif, ymin = Lynott$LL_CI_TOST, ymax = Lynott$UL_CI_TOST, size = tost.size, fatten = point.size) +
-  geom_hline(yintercept=Lynott$low_eqbound, lty=2) +
   geom_hline(yintercept=Lynott$high_eqbound, lty=2) +
   ylim(c(Lynott$low_eqbound*plot.scaler, Lynott$high_eqbound*plot.scaler)) +
   labs(title = "D: Example 4 (Lynott et al.)", y = "Effect (proportion difference)")

### 2017-11-13 examples_plotgrid.R
Renamed all R scripts for the examples & changed the file names in the plot script and the manuscript accordingly

@@ -30,7 +30,7 @@ baseplot <- ggplot(data.frame()) +
 
 ####  Meta analysis ----------------------------------------
 
-source("SEQ_Hyde_Example.R")
+source("Example2_Hyde-et-al.R")
 
 grades <- list(gr2, gr3, gr4, gr5, gr6, gr7, gr8, gr9, gr10, gr11)
 
@@ -68,7 +68,7 @@ metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
 
 ####  Brandt ----------------------------------------
 
-source("NSEQ_Brandt_Example.R")
+source("Example1_Brandt-et-al_study1.R")
 
 brandtplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Brandt$diff, ymin = Brandt$LL_CI_TTEST, ymax = Brandt$UL_CI_TTEST, size = nhst.size, fatten = point.size) +
@@ -80,21 +80,21 @@ brandtplot = baseplot +
 
 #### Moon ----------------------------------------
 
-source("NSNEQ_Moon_Example.R")
+source("Example3_MoonRoeder.R")
 
 moonplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Moon$diff, ymin = Moon$LL_CI_TTEST, ymax = Moon$UL_CI_TTEST, size = nhst.size, fatten = point.size) +
   annotate(geom = "pointrange", x = 0.5, y = Moon$diff, ymin = Moon$LL_CI_TOST, ymax = Moon$UL_CI_TOST, size = tost.size, fatten = point.size) +
   geom_hline(yintercept=Moon$low_eqbound, lty=2) +
   geom_hline(yintercept=Moon$high_eqbound, lty=2) +
-  labs(title = "C: Example 3 (Moon et al.)", y = "Effect (mean difference)") +
+  labs(title = "C: Example 3 (Moon & Roeder)", y = "Effect (mean difference)") +
   ylim(c(Moon$low_eqbound*plot.scaler, Moon$high_eqbound*plot.scaler))
 
 #### Lynott ----------------------------------------
 
 
 
-source("NSEQ_Lynott_example.R")
+source("Example4_Lynott-et-al.R")
 
 lynottplot = baseplot +
[…]

### 2017-11-13 manuscript.Rmd
fixed weird "HEAD" insertions from earlier merge conflict; changed example 1 to report means and SDs from replication study rather than original

@@
 ```{r include = FALSE}
-source("NSEQ_Brandt_Example.R")
+source("Example1_Brandt-et-al_study1.R")
 
@@
 
-@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r printp(orig.p)`$, $d= `r printnum(orig.d)`$).
-A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
+@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($t(`r orig.df`)= `r orig.t`$, $p= `r printp(orig.p)`$, $d= `r printnum(orig.d)`$).
+A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($M_{unethical}= `r rep.m1`$, ${SD}_{unethical}= `r rep.sd1`$, $M_{ethical}=`r rep.m2`$, ${SD}_{ethical}=`r rep.sd2`$, $t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
 Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r printnum(d.33)`$ --- and use this as our SESOI.
@@
 ```{r, include=FALSE}
-source("SEQ_Hyde_Example.R")
+source("Example2_Hyde-et-al.R")
 ```
@@
 ```{r include=FALSE}
-source("NSNEQ_Moon_Example.R")
+source("Example3_MoonRoeder.R")
 ```
@@
 ```{r include=FALSE}
-source("NSEQ_Lynott_example.R")
+source("Example4_Lynott-et-al.R")
 ```
@@
 However, as this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could potentially detect.
-This means that we will use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
+This means that we will use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bound ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
 To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error.
@@
 ```{r include = FALSE}
-source("corr_Kahane_Example.R")
+source("Example5_Kahane-et-al-study4.R")
 ```

### 2017-11-15 manuscript.Rmd
Corrected an earlier error where I had mistaken the reference Norman et al., 2003, for Norman et al., 2004. Changed it back to the (correct) 2003 reference, both in Zotero and in all citations in the manuscript.

@@
 
-Another example of an objectively determined SESOI can be found in @Button2015 where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
+Another example of an objectively determined SESOI can be found in @Button2015 where the minimal clinically important difference on the Beck Depression Inventory -- II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
 
-@Norman2004 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or $d = 0.5$.
+@Norman2003 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or $d = 0.5$.
 They make the argument that this bound of approximately half a standard deviation could be related to the limits of human discriminability.
@@
 If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003].
-Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2004], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
+Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2003], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
 

### 2017-11-15 examples_plotgrid.R
Changed plot order to match examples
* A becomes B
* B becomes C
* C becomes A

@@ -56,7 +56,7 @@ metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
   geom_hline(yintercept=sesoi, lty=2) +
   geom_hline(yintercept=0, lty=1) +
   coord_flip() +  # flip coordinates (puts labels on y axis)
-  labs(title = "B: Example 2 (Hyde et al.)", y = "Effect (Z)") +
+  labs(title = "C: Example 3 (Hyde et al.)", y = "Effect (Z)") +
   ylim(c(-sesoi*plot.scaler, sesoi*plot.scaler)) +
   theme_classic(base_size = 10) + # use a white background
   theme(plot.title = element_text(size = rel(1), face = "bold"),
@@ -75,7 +75,7 @@ brandtplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Brandt$diff, ymin = Brandt$LL_CI_TOST, ymax = Brandt$UL_CI_TOST, size = tost.size, fatten = point.size) +
   geom_hline(yintercept=Brandt$low_eqbound, lty=2) +
   geom_hline(yintercept=Brandt$high_eqbound, lty=2) +
-  labs(title = "A: Example 1 (Brandt et al.)", y = "Effect (mean difference)") +
+  labs(title = "B: Example 2 (Brandt et al.)", y = "Effect (mean difference)") +
   ylim(c(Brandt$low_eqbound*plot.scaler, Brandt$high_eqbound*plot.scaler))
 
 #### Moon ----------------------------------------
@@ -87,7 +87,7 @@ moonplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Moon$diff, ymin = Moon$LL_CI_TOST, ymax = Moon$UL_CI_TOST, size = tost.size, fatten = point.size) +
   geom_hline(yintercept=Moon$low_eqbound, lty=2) +
   geom_hline(yintercept=Moon$high_eqbound, lty=2) +
-  labs(title = "C: Example 3 (Moon & Roeder)", y = "Effect (mean difference)") +
+  labs(title = "A: Example 1 (Moon & Roeder)", y = "Effect (mean difference)") +
   ylim(c(Moon$low_eqbound*plot.scaler, Moon$high_eqbound*plot.scaler))
 
 #### Lynott ----------------------------------------
@@ -121,4 +121,4 @@ lay <- rbind(c(1,2,5),
[…]

### 2017-11-15 Example1_MoonRoeder.R
Changed example order
* Changed example 1 to 2, 2 to 3, and 3 to 1
* Filenames changed
* Scripts edited to source new filenames
* Jamovi scripts updated to newest version (works with Jamovi 0.8.1.7 and TOSTER 0.2.4)

@@ -1,5 +1,5 @@
 #  Load the relevant datafile into R
-moon.data        <- read.csv("NSNEQ_Moon_Example_data.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)
+moon.data        <- read.csv("Example1_MoonRoeder_data-cleaned.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)
 
 #  Set the bounds, in this case 1 grade point, which corresponds to 6.25%
 sesoi <- 0.0625

### 2017-11-15 Example2_Brandt-et-al_study1.R

@@ -1,7 +1,7 @@
 library(TOSTER)
 library(pwr)
 
-Brandt.data <- read.csv("Example1_Brandt-et-al-study1_data-cleaned.csv", header = TRUE, sep = ",")
+Brandt.data <- read.csv("Example2_Brandt-et-al_study1_data-cleaned.csv", header = TRUE, sep = ",")
 Brandt.data <- as.data.frame(Brandt.data)
 
 # original study: Banerjee et al., study 1

### 2017-11-15 examples_plotgrid.R

@@ -30,7 +30,7 @@ baseplot <- ggplot(data.frame()) +
 
 ####  Meta analysis ----------------------------------------
 
-source("Example2_Hyde-et-al.R")
+source("Example3_Hyde-et-al.R")
 
 grades <- list(gr2, gr3, gr4, gr5, gr6, gr7, gr8, gr9, gr10, gr11)
 
@@ -68,7 +68,7 @@ metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
 
 ####  Brandt ----------------------------------------
 
-source("Example1_Brandt-et-al_study1.R")
+source("Example2_Brandt-et-al_study1.R")
 
 brandtplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Brandt$diff, ymin = Brandt$LL_CI_TTEST, ymax = Brandt$UL_CI_TTEST, size = nhst.size, fatten = point.size) +
@@ -80,7 +80,7 @@ brandtplot = baseplot +
 
 #### Moon ----------------------------------------
 
-source("Example3_MoonRoeder.R")
+source("Example1_MoonRoeder.R")
 
 moonplot = baseplot +
   annotate(geom = "pointrange", x = 0.5, y = Moon$diff, ymin = Moon$LL_CI_TTEST, ymax = Moon$UL_CI_TTEST, size = nhst.size, fatten = point.size) +

### 2017-11-16 Example1_MoonRoeder.R
Minor cleanup in example code

@@ -13,16 +13,12 @@ n.asian          <- sum(moon.data$identity_salience == 1)
 n.control        <- sum(moon.data$identity_salience == 2)
 
 df <- (sd.asian^2 / n.asian + sd.control^2 / n.control)^2 / (((sd.asian^2 / n.asian)^2 / (n.asian-1)) + ((sd.control^2 / n.control)^2 / (n.control - 1)))  # Degrees of freedom for Welch's t-test
-sample.sd <- sqrt((sd.asian^2 + sd.control^2) / 2)  # Calculate sd root mean squared for Welch's t-test
-sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate the se of the difference based on sd.pooled
 
 t <- (mean.asian - mean.control) / sample.se  # Welch t-test
 
-d <- (mean.asian - mean.control) / sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.control + n.control) - 2))
-
 p <- 2 * pt(t, df)
 
-sdpooled <- sqrt((((n.asian - 1)*(sd.asian^2)) + (n.control - 1)*(sd.control^2))/((n.asian+n.control)-2))
+sdpooled <- sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.asian + n.control) - 2))
 d <- (mean.asian - mean.control) / sdpooled
 
 

### 2017-11-16 Example3_Hyde-et-al.R

@@ -1,12 +1,20 @@
 
 # Hyde, Lindberg, Linn, Ellis, and Williams (2008) meta analysis
-gr2 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr3 <- TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr4 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr5 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr6 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr7 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr8 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr9 <- TOSTmeta(ES = -0.01, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr10 <- TOSTmeta(ES = 0.04, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
-gr11 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-0.1, high_eqbound_d=0.1, alpha=0.005, plot = FALSE)
+
+sesoi <- 0.1
+alpha <- 0.005
+
+gr2  <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr3  <- TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
[…]

### 2017-11-16 Example4_Lynott-et-al.R

@@ -10,7 +10,7 @@ library(TOSTER)  # Install TOSTER using devtools::install_github("Lakens/TOSTER"
 
 
 #  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
-lynott.data <- read.csv("NSEQ_Lynott_example_data.csv")
+lynott.data <- read.csv("Example4_Lynott-et-al_data-cleaned.csv")
 lynott.data <- lynott.data[lynott.data$include1 == 1,]
 
 

### 2017-11-16 Example1_MoonRoeder.R
Minor cleanup in example code

@@ -18,11 +18,9 @@ sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate 
 
 t <- (mean.asian - mean.control) / sample.se  # Welch t-test
 
-d <- (mean.asian - mean.control) / sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.control + n.control) - 2))
-
 p <- 2 * pt(t, df)
 
-sdpooled <- sqrt((((n.asian - 1)*(sd.asian^2)) + (n.control - 1)*(sd.control^2))/((n.asian+n.control)-2))
+sdpooled <- sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.asian + n.control) - 2))
 d <- (mean.asian - mean.control) / sdpooled
 
 

### 2017-11-16 Example1_MoonRoeder.R
Fixed bug in example introduced in previous commit

@@ -11,6 +11,8 @@ sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
 sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
 n.asian          <- sum(moon.data$identity_salience == 1)
 n.control        <- sum(moon.data$identity_salience == 2)
+sample.sd <- sqrt((sd.asian^2 + sd.control^2) / 2)  # Calculate sd root mean squared for Welch's t-test
+sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate the se of the difference based on sd.pooled
 
 df <- (sd.asian^2 / n.asian + sd.control^2 / n.control)^2 / (((sd.asian^2 / n.asian)^2 / (n.asian-1)) + ((sd.control^2 / n.control)^2 / (n.control - 1)))  # Degrees of freedom for Welch's t-test
 

### 2017-11-16 manuscript.Rmd
Major manuscript update - v.1.0
* Copied in all text from Google doc (16-11-2017 15pm)
* Kept the Rmarkdown formatting of the text box
* Fixed minor syntax error

--- 2017-11-13 manuscript.Rmd
@@
 abstract: >
-  Psychologists  need to be able to test for the absence of an effect.
-Using the Two-One-Sided Tests (TOST) procedure, researchers can easily test whether the observed effects are too small to be meaningful.
-By specifying a smallest effect size of interest (SESOI) researchers test whether observed effects are surprisingly closer to zero, assuming there was an effect the consider meaningful.
+  Psychologists must be able to test both for the presence of an effect and for the the absence of an effect.
+In addition to testing against zero, researchers can use the Two-One-Sided Tests (TOST) procedure to test for *equivalence* and reject the presence of a meaningful effect.
+TOST can be used to determine if  an observed effect is surprisingly small, given that a true effect at least as large as the SESOI exists.
 We explain a range of approaches to determine the SESOI in psychological science, and provide detailed examples of how equivalence tests should be performed and reported.
+Equivalence tests enable researchers to falsify predictions about the presence, and declare the absence, of meaningful effects.
 
@@
 Psychologists should be able to falsify predictions.
-A common prediction in psychological research is that an effect exists in the population that differs from zero.
+A common prediction in psychological research is that an effect that differs from zero exists in the population.
 For example, we might predict that priming American Asian women with their Asian identity will increase their performance on a math test compared to women who are primed with their female identity.
-To be able to falsify this hypothesis, and design a study that allows for strong inferences [@Platt1964], it is important to specify which test result would *disprove* the hypothesis.
+To be able to design a study that allows for strong inferences [@Platt1964], it is important to specify which test result would *falsify* this hypothesis.
 
-An equivalence test can be used to test whether the observed effect is surprisingly small, assuming a meaningful effect exists in the population.
-The test is a simple variation of the widely used null hypothesis significance tests (NHST).
+Equivalence testing can be used to test whether an observed effect is surprisingly small, assuming a meaningful effect exists in the population.
+The test is a simple variation of widely used null hypothesis significance tests (NHST).
 To understand the idea behind equivalence tests, it is useful to realize the null hypothesis we test against can be any numerical value.
-When we compare two groups, we often test whether the difference between these groups is zero, but we may sometimes want to test against other values than zero.
+When we compare two groups, we often test whether the difference between these groups is zero (see Figure 1A), but we may sometimes want to test against other values than zero.
 Imagine a researcher who is interested in voluntary participation in a national program to train young infants' motor skills.
 The researcher wants to test whether more boys than girls are brought into the program by their parents.
-One could test whether the difference between boys and girls is zero.
-However, this would ignore the fact that the human sex ratio of newborn boys to girls is not exactly 1:1, and we should not expect 50% of participants to be boys.
-On average, 103 boys are born for every 100 girls [@UnitedStates2016], so approximately 50.74% of applicants should be boys, and 49.26% should be girls (with $0.5074/0.4926 \approx 1.03$).
-If boys and girls were exactly equally likely to be brought into the program by their parents, we would not expect a difference of zero, but 1.5% more boys.
-This can be used as a null hypothesis: Rather than testing against a difference in proportions of 0, the researcher tests against 0.015.
+Because the human sex ratio of newborn boys to girls is not exactly 1:1, we should not expect 50% of participants to be boys.
+On average, 103 boys are born for every 100 girls [@UnitedStates2016], so approximately 50.74% of applicants should be boys, and 49.26% should be girls.
+If boys and girls were exactly equally likely to be brought into the program by their parents, we would not expect a difference of zero, but 50.74% - 49.26% = 1.5% more boys.
+Rather than testing against a null hypothesis of 0 difference, the researcher tests against a null hypothesis of 0.015.
 
-Even though 0.015 might be a more sensible value for a null hypothesis, our researcher knows that the true sex ratio in their population will differ slightly from that in the entire world.
-Rather than testing a point null hypothesis (setting $H_0$ to one specific value), the researcher decides to define a range of values around the difference in proportions of 0.015 that can be considered trivially small, even when the true ratio in the population is not exactly 0, or even 0.015.
-The researcher can for example test if the difference is smaller than -0.005, or larger than 0.035.
-This test against two bounds, with $H_0$ as a range rather than one value, is known as a *minimal effects test* [@Murphy2014].
+Alternatively, the researcher could decide that even if the true ratio in the population is not exactly 0.015, the null hypothesis should consist of a range of values around the difference in proportions of 0.015 that can be considered trivially small.
+The researcher could for example test if the difference is smaller than -0.005, or larger than 0.035.
+This test against two bounds, with $H_0$ being a range rather than one value (see Figure 1B), is known as a *minimal effects test* [@Murphy2014].
 
-However, even if the researcher would consider any difference in percentages smaller than -0.5% or larger than 3.5% a rejection of the null hypothesis, not every difference is large enough to matter in practice.
-As long as gender differences in the participation rate are not too extreme, there is no need to spend money to address the gender imbalance in the voluntary participation rate, and all money can be spent on the core training program.
-To test whether the gender difference in participants is *not* large enough to matter, the researcher can perform an *equivalence test*.
-Equivalence tests examine whether values below and above a prespecified range can be rejected.
-Here, we would test if the presence of gender differences that are large enough to matter can be rejected.
-After an extensive discussion with experts, the researcher decides that as long as the difference in proportions is not larger than 6% the researcher can act as if the gender difference is too small to intervene.
-Given an expected true difference in the population of 0.015, the researcher will test if the observed difference in the data is smaller than -0.055, and larger than 0.075.
-If differences more extreme than both these boundary values can be rejected in two one-sided tests, the researcher will conclude statistical equivalence, and no money will be spent on addressing a gender difference in participation.
+*Equivalence tests* can be seen as the opposite of minimal effects tests: They examine whether the presence of effects that are large enough to matter can be *rejected* (see Figure 1C).
+In our example, the researcher can perform an equivalence test to test whether the gender difference in participants is *not* large enough to be considered meaningful: the *smallest effect size of interest* (SESOI).
+After an extensive discussion with experts, the researcher decides that as long as the difference in proportions is not more extreme than 6%, the gender difference is too small to care about.
+Given an expected true difference in the population of 0.015, the researcher will test if the observed difference falls outside the boundary values of -0.055 and 0.075.
+If differences more extreme than these boundary values can be rejected in two one-sided tests (TOST), the researcher will conclude statistical equivalence, the gender difference will be considered trivially small, and no money will be spent on addressing a gender difference in participation.
 
-For an alpha level of 5%, we can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI).
-A 90% confidence interval is used because one-sided tests are performed, and it is not necessary to control for multiple comparisons when performing two one-sided tests because both tests need to be statistically significant to conclude equivalence.
-Note that although we will use an alpha of 5% in all examples because this alpha level was used in the papers we reanalyze, researchers should carefully decide upon a desired Type 1 error rate, and justify their choice [@Lakens2017b].
-Figure 1 illustrates the possible conclusions we can draw from null hypothesis, minimal effect, and equivalence tests.
-Although thinking about equivalence tests in terms of the values that can be rejected based on a confidence interval might be the easiest conceptualization, for any given study, you can also think of an equivalence test as determining whether effect sizes or test statistics are closer to zero than some critical, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
+In any one-sided test, for an alpha level of 5%, we can reject $H_0$ when the 90% confidence interval around the observed estimate is in the predicted direction, and does not contain the value tested against (e.g., 0).
+In TOST, the first one-sided test is used to test against values smaller than the lower bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper bound ($\Delta_{U}$).
+It is not necessary to control for multiple tests when using the TOST procedure because both tests need to be statistically significant to conclude equivalence.
+This means that when reporting a TOST result, authors can suffice by reporting the one-sided test with the smaller test parameter (e.g.
+$t$ or $r$) and thus the larger $p$-value.
+Statistical equivalence can be concluded when the largest of the two $p$-values is smaller than alpha.
+
+Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze (but we recommend to set and justify the Type 1 error rate in original research based on the goals researchers have [@Lakens2017b].
+It might be easiest to think of an equivalence test as checking whether the confidence interval excludes the SESOI, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
 
 ```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
-**A)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
-**B)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
-**C)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
+**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
+**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
+**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
 **D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}
@@
 
-Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, rather than against 0), their use in psychology was limited, perhaps due to the lack of user-friendly software to perform the calculations in the past [@Lakens2017a].
-In this article, we provide several examples of equivalence tests which illustrate the procedure in free software (R, a spreadsheet, and in Jamovi).
-Furthermore, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.
+In this article we provide several reproducible examples of equivalence tests which illustrate the procedure in easy-to-use free software (R, jamovi, and a spreadsheet).
+But first, we discuss different approaches to determining the SESOI for psychological research, which determines the statistical question an equivalence test answers.
 
@@
 
-Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI).
-Although we use symmetrical bounds around 0 in the examples we will discuss below (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is possible to use asymmetric bounds, for example based on an explicit cost-benefit analysis (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$).
-The SESOI can sometimes be determined objectively, for example based on just noticeable differences .
-Most often, however, it is a subjective decision that varies across researchers, fields, and time.
+TOST is performed against equivalence bounds that are considered the smallest effect size of interest.
+The SESOI can sometimes be determined objectively, for example based on just noticeable differences.
 In lieu of objective justifications, the SESOI should ideally be based on a cost-benefit analysis.
-Since both costs and benefits are necessarily subjective, the SESOI will depend on the researcher who designs the study.
-The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base.
+Since both costs and benefits are necessarily subjective, the SESOI will vary across researchers, fields, and time.
+The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects more extreme than the specified equivalence bounds contributes to our knowledge base.
 A SESOI should be chosen such that inferences based on it answer a meaningful question.
+Although we use bounds that are symmetric around 0 in all examples in this manuscript for equivalence tests (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is also possible to use asymmetric bounds, for example based on an explicit cost-benefit analysis (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$).
 
-##Objective Justifications of a SESOI
+##Objective Justification of a SESOI
 
@@
 However, this increase was not large enough for men to detect with the naked eye, thus falsifying their hypothesis.
-Because the just noticeable difference in redness of the skin can be measured, it is possible to objectively establish the SESOI.
+Because the just noticeable difference in redness of the skin can be measured, it is possible to establish an objective SESOI.
 
-Another example of an objectively determined SESOI can be found in @Button2015 where the minimal clinically important difference on the Beck Depression Inventory - II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
+Another example of an objectively determined SESOI can be found in @Button2015, where the minimal clinically important difference on the Beck Depression Inventory -- II was determined by asking 1039 patients when they subjectively felt less depressed (i.e., when they personally noticed an improvement) and relating this to the corresponding difference score on the depression inventory.
+Similarly, @Norman2003 proposed there is a surprisingly consistent minimally important difference of half a standard deviation, or $d = 0.5$, in health outcomes.
 
-@Norman2004 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or $d = 0.5$.
-They make the argument that this bound of approximately half a standard deviation could be related to the limits of human discriminability.
-Therefore, even though a SESOI of $d = 0.5$ might seem quite large, it could be a plausible equivalence bound when examining effects that should be noticeable by single individuals.
-
-##Subjective justifications of a SESOI
+##Subjective Justification of a SESOI
 
@@
 First, researchers can use benchmarks.
-For example, one might set the SESOI to a standardized effect size of $d = 0.5$, which would allow one to reject effect as large or larger than a 'medium' effect size [@Cohen1988].
-Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small [@Maxwell2015].
+For example, one might set the SESOI to a standardized effect size of $d = 0.5$, which would allow rejecting effects more extreme than a 'medium' effect size [@Cohen1988].
+Similarly, effect sizes smaller than a Cohen's $d$ of 0.1 are sometimes considered trivially small [@Maxwell2015].
 Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided.
-Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI in communication research to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that in psychology a SESOI of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
+Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that a SESOI of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
 
@@
 N <- 100
-pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d
 ```
@@
 Second, researchers can determine the SESOI based on previous work in the literature.
-Ideally, researchers would specify their SESOI in studies they report, but this is not yet common practice.
+Ideally, researchers would always specify their SESOI in studies they report, but this is not yet common practice.
 It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study.
@@
 Thus, they examined whether their replication study could reject effects as large or larger than had on average been reported in the literature.
-Given random variation and bias in the literature, a more conservative approach could be to use a lower bound of a confidence interval around the meta-analytic effect size estimate [cf.
+Given random variation and bias in the literature, a more conservative approach could be to use the lower end of a confidence interval around the meta-analytic effect size estimate [cf.
 @Perugini2014].
 
-Another justifiable choice for the SESOI is to use the smallest observed effect size that could have been statistically significant in a previous study.
-In other words, we decide that effects that could not have yielded a $p < \alpha$ in an original study will not be considered meaningful in the replication study either, even if they are found to be statistically significant.
+Another justifiable choice when basing the SESOI on previous work is to use the smallest observed effect size that could have been statistically significant in a previous study.
+In other words, we decide that effects that could not have yielded a $p < \alpha$ in an original study will not be considered meaningful in the replication study either, even if they were found to be statistically significant.
 Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
-This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as a *critical effect size*.^[This will typically, although not always, correspond to the effect size the study had 50% power to detect [see @Lenth2007].
-This procedure will thus result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.] All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
+This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as the *critical effect size*.^[This will typically, although not always, correspond to the effect size the study had 50% power to detect [see @Lenth2007].
+This procedure will result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.] All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
 By setting the SESOI to the critical effect size, an equivalence test can reject all observed effect sizes that could have been detected in an earlier study.
-
-```{r, results="hide"}
-library(TOSTER)
-powerTOSTone(alpha=0.05, statistical_power=0.9, low_eqbound_d=-0.33, high_eqbound_d=0.33)
-```
 
@@
 The amount of data you can collect limits the inferences you can make.
-Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST.
-However, @Meyners2012 explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
+Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a $p$-value *larger than* 0.05 in a traditional NHST.
+However, @Meyners2012 explains that this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and error rates are not controlled accurately.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
 In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size.
-For example, a researcher who plans to perform a (two-sided) one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of $d = 0.364$.
-Using equivalence bounds of $\Delta_{L} = -0.364$ and $\Delta_{U} = 0.364$ in an equivalence test allows one to reject effects the study had high statistical power to detect.
-An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that could reliably be detected given a specific sample size?
+For example, a researcher who plans to perform a one-sample $t$-test using an alpha of `r alpha*100`% (two-sided), based on data from `r N` observations, has 90% power to detect an effect of $d = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$.
+Using equivalence bounds of $\Delta_{L} = `r printnum(-pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ and $\Delta_{U} = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ in an equivalence test allows to reject effects the study had high statistical power to detect.
+
+Importantly, an equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that could reliably be detected given a specific sample size?
 In this case, concluding equivalence would suggest that effects which could reasonably be detected with a sample size of 100 can be rejected, and that future studies probably need larger sample sizes to examine a specific question.
 Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area.
-An equivalence test based on 15 observations testing against a SESOI of $d = 1$ cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
-By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is.
-```{r}
-p <- 0.05
-N <- 100
-crit_d <- abs(qt(p/2, (N*2)-2))/sqrt(N/2)
-```
+An equivalence test based on 15 observations rejecting a SESOI of $d = 1$ cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
+By transparently reporting the effects that can be detected and rejected based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what the size of a reasonable SESOI may be.
 
@@
 
-Equivalence bounds can be set in terms of standardized effect sizes, or raw differences.
-For example, the SESOI might be a raw mean difference of 0.5 points on a 7-point scale, or a Cohen’s d of 0.5.
-The main difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as [FORMULA (X1-X2)/SD].
+The SESOI, and thus the equivalence bounds, can be set in terms of standardized effect sizes (e.g.
+a Cohen’s $d$ of 0.5) or as a raw mean difference (e.g.
+0.5 points on a 7-point scale).
+The key difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as $\frac {X _ { 1} - X _ { 2} } {S D}$.
 The observed standard deviation randomly varies across samples.
-In practice, this means that when you use standardized differences as bounds (e.g., $d = 0.35$) the equivalence test depends on the standard deviation in your the sample.
+In practice, this means that when using standardized differences as bounds (e.g., $d = 0.35$), the equivalence test depends on the standard deviation in the sample.
 The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of $d = 0.5$ when the standard deviation is 1, but a standardized equivalence bound of $d = 1$ when the standard deviation is 0.5.
@@
 When the raw mean difference is of less theoretical importance, or different measures are used across research lines, it is often easier to set equivalence bounds based on standardized differences.
-Researchers should realize that equivalence bounds based on raw differences or standardized differences ask slightly different questions, and justify their choice for an equivalence bound.
+Researchers should realize that equivalence bounds based on raw differences or standardized differences ask slightly different questions, and justify their choice for either option.
 When setting equivalence bounds based on earlier research, such as in replication studies, equivalence bounds based on raw or standardized differences would ideally give the same result, and large differences in standard deviations between studies are as important to interpret as large differences in means.
@@
 These concrete and easy to follow examples will illustrate all approaches to setting equivalence bounds discussed above, and demonstrate how to perform and report equivalence tests.
-The code to reproduce these analyses is available at XXXX.XXXX.
+The code to reproduce these analyses is available at <https://osf.io/qamc6/>.
 
@@
 ```
-```{r echo=FALSE, fig.cap=paste("Example effects plotted with TOST CIs (thick lines) and NHST CIs (thin lines), NHST H0 (solid vertical line) and TOST H0 (dashed vertical lines) displayed.
-**A)** Example 1 - The mean difference.
-**B)** Example 2 - The effects included in the meta-analysis.
-**C)** Example 3 - The mean difference.
-**D)** Example 4 - The proportion difference.
-**E)** Example 5 - The pearson correlation")}
+```{r echo=FALSE, fig.cap=paste("Example effects plotted with 90% TOST CIs (thick lines) and 95% NHST CIs (thin lines), the NHST null hypothesis (solid vertical line) and the equivalence bounds (dashed vertical lines) displayed.
+**A)** Example 1 - Mean Difference.
+**B)** Example 2 - Mean Difference.
+**C)** Example 3 - Meta-analytic Effect Size.
+**D)** Example 4 - Proportion Difference.
+**E)** Example 5 - Pearson Correlation")}
 # Plot the TOST examples
@@
 
-## Example 1: Statistically Equivalent and Not Statistically Different
+## Example 1: Not Statistically Equivalent and Not Statistically Different
+```{r include=FALSE}
+source("Example1_MoonRoeder.R")
+```
+
+@Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
+In contrast to the original study, they found a negative difference between the Asian primed group ($n = `r n.asian`$, $M = `r printnum(mean.asian)`$, $SD = `r printnum(sd.asian)`$) and the control ($n = `r n.control`$, $M = `r printnum(mean.control)`$, $SD = `r printnum(sd.control)`$) which was not significant, $d = `r printnum(d)`$, $t(96.62) = `r printnum(t)`$, $p = `r printp(p)`$, two-sided).
+The non-significant null-hypothesis test does not allow us to distinguish between the absence of a meaningful effect and data that are not sensitive enough to tell us whether a meaningful effect is present or absent.
+
+In order to distinguish between these possibilities, we can define what a “meaningful effect” (SESOI) would be and use it as the bounds for an equivalence test.
+If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25% ...
+A+ = 93.75% to 100%) we could decide that we are only interested in test differences that correspond to an increase or decrease of at least 1 grade point.
+Thus, our SESOI becomes a difference in raw scores of `r printnum(sesoi * 100)`\%, or `r printnum(sesoi, digits = 4)`.
+We can then perform the TOST procedure for a two-sample Welch’s *t*-test, using $\pm$ the SESOI of `r printnum(sesoi, digits = 4)` as the equivalence bounds ($\Delta$).
+The TOST is non-significant, $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (Figure 2A).
+Thus, we cannot reject a true effect as large or larger than 6.25%.^[In this example, the bounds of the TOST were based on a raw effect of 6.25%.
+We could have converted this raw effect into a standardized effect (Cohen's $d$) for the current sample, and we would have gotten the exact same result ($\Delta = \pm `r printnum(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2))`)$.
+Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's $d$ that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
+Now we would no longer be answering a question about a raw effect independent of the samples, but rather if the replication can reject standardized effects larger than a certain standardized effect in the original.
+Since the standard deviations of the original study were larger than in the replication in this case, Cohen's $d$ (and thus the bounds) would be narrower ($\Delta = \pm `r printnum(sesoi / sqrt((sd.asian^2 + sd.control^2) / 2))`)$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
+The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
+Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$).
+
+## Example 2: Statistically Equivalent and Not Statistically Different
 
 ```{r include = FALSE}
-source("Example1_Brandt-et-al_study1.R")
+source("Example2_Brandt-et-al_study1.R")
 
@@
 A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($M_{unethical}= `r rep.m1`$, ${SD}_{unethical}= `r rep.sd1`$, $M_{ethical}=`r rep.m2`$, ${SD}_{ethical}=`r rep.sd2`$, $t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
-Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r printnum(d.33)`$ --- and use this as our SESOI.
-If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2A).
+Using the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r printnum(d.33)`$ --- and use this as the SESOI.
+A TOST procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
 
-## Example 2: Statistically Equivalent and Statistically Different
+## Example 3: Statistically Equivalent and Statistically Different
 
 ```{r, include=FALSE}
-source("Example2_Hyde-et-al.R")
+source("Example3_Hyde-et-al.R")
 ```
 
-@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than $d = 0.1$.
+@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as effect sizes smaller than $d = printnum(sesoi)$.
 For example, in grade 3 the difference is $d = 0.04$, with a standard error of 0.002.
-When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of $d = -0.1$ and $d = 0.1$, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of $d = -0.1$ and $d = 0.1$, and can be considered trivially small, according to the bounds set by the authors (figure 2B).
-However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge.
-This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance.
-
-## Example 3: Not Statistically Equivalent and Not Statistically Different
-```{r include=FALSE}
-source("Example3_MoonRoeder.R")
-```
-
-@Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
-They observed a slightly reversed difference between the Asian primed group ($n = 53$, $M = 0.46$, $SD = 0.17$) and the control ($n = 48$, $M = 0.50$, $SD = 0.18$) which was not significant, $d = -0.23$, $t(96.62) = 1.15$, $p = 0.26$ (two-sided).
-A non-significant null-hypothesis test leaves two possible conclusions:
-
-1. The data suggest the absence of a meaningful effect, and we can reject effects large enough to be meaningful.
-2. The data are not sensitive enough to tell us whether a meaningful effect is present or absent.
-
-As an example, imagine we discover that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ...
-A+ = 93.75% to 100%).
-We can decide that we are only interested in test differences so far as they correspond to at least a 1 grade point increase or decrease.
-Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.
-We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using $\pm$ the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
-We find that the TOST is non-significant, t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`, p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))` (figure 2C).
-Thus, we cannot reject the possibility that the priming manipulation leads to a real change in test scores that is larger on average than one gradepoint.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%.
-We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r printnum(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2))`$).
-Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
-Now we would no longer be answering a question about a raw effect independent of the samples.
-Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study.
-Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r printnum(0.0625 / sqrt((0.17^2 + 0.20^2) / 2))`$).
-As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
-Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
-The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
-Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$).
+When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of printp(alpha) to correct for multiple comparisons), using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that indeed, all effect size estimates are measured with such high precision that they are statistically equivalent and can be considered trivially small, according to the authors’ definition (Figure 2C).
+In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printnum(max(gr3$TOST_p1, gr3$TOST_p2))`$ However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large.
+For grade 3, $Z = `r printnum(gr3.z)`$, $p = `r printnum(gr3.p)`$.
+This is one way in which the use of equivalence tests improves hypothesis testing procedures in science, by allowing researchers to distinguish between statistical significance and practical significance.
 
@@
 In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $Z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
+Had the authors planned to perform a NHST *and* an equivalence test, the latter would allow one to distinguish between an inconclusive outcome, or a statistically equivalent result.
 
-In this case, it is not immediately clear what an objective criterion for the sESOI would be.
-However, as this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could potentially detect.
-This means that we will use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bound ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
+Since this is a replication, we could decide that we are interested in whether the observed effect size is smaller than the smallest effect size that the original study could have detected.
+This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
 To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error.
 
-Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study.
-In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{U}$ because an effect with CIs crossing $\Delta_{L}$ can still be *lower* than the original effect.
-We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r printnum(Lynott$TOST_z2)`, $p `r printp(Lynott$TOST_p2)`$ (see figure 2D).
-Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.
+Since the original study had a clear directional hypothesis,  in this replication study we are only interested in whether receiving a warm pack *increases* the proportion of people who choose a gift for a friend.
+This means we can test for inferiority (see Figure 1D) and conclude the absence of a meaningful effect if the observed effect size is reliably *lower* than the SESOI.
+We find that we can reject effects larger than  $\Delta_{U}$, $z = `r printnum(Lynott$TOST_z2)`$, $p `r printp(Lynott$TOST_p2)`$ (see Figure 2D).
+Thus, we can conclude that the statistically non-significant effect is also statistically smaller than our SESOI.
 
@@
 
-In a paper published in [-@Kahane2015], Kahane and colleagues investigate moral dilemma vignettes in which participants have to decide whether or not they would sacrifice one person's life to save several other lives, or judge how morally admissible such actions are.
-Traditionally, greater endorsement for sacrificing a life to save others has been interpreted as a more "utilitarian" moral orientation, i.e.
-a stronger concern for the greater good (in total, fewer people lose their lives).
-In a number of studies, Kahane et al.
-contest this interpretation, for example by showing that greater endorsement for such "utilitarian" choices correlates with sub-clinical psychopathy, and, crucially, by comparing the traditionally used vignettes with a set of new ones that pit partial motivations against an impartial concern for the greater good (e.g.
+Kahane and colleagues [-@Kahane2015] investigate how responses to moral dilemmas in which participants have to decide whether or not they would sacrifice one person's life to save several other lives relate to other indicators of moral orientation.
+Traditionally, greater endorsement for sacrificing one life to save several others has been interpreted as a more "utilitarian" moral orientation (i.e.
+a stronger concern for the greater good).
+Kahane et al.
+contest this interpretation in a number of studies.
+In Study 4, they compare the traditionally used dilemmas with a new set of new ones in which the sacrifice for the greater good is not another person’s life, but something the participant would have a partial interest in(e.g.
 buying a new mobile phone vs.
-donating the money to save lives in a distant country, study 4).
+donating the money to save lives in a distant country) and find no significant correlation between the two ($r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$), $N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
+The authors conclude that “a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely” (p.
+206), given the statistical power their study had to detect meaningful effect sizes.
 
-@Kahane2015 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$ ($N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
-They conclude that the classical vignettes fail to capture "true" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p.
-206).
-This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$.
-Given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$.
-This means that values smaller than $r = `r printnum(-Kahane.r.80)`$ and larger than $r = `r printnum(Kahane.r.80)`$ can be rejected at an alpha level of 5% and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.
+This interpretation can be formalized by performing an equivalence test for correlations, where equivalence bounds are set to an effect size the study had reasonable power to detect.
+With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$.
+Performing the TOST procedure given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$.
+This means that values smaller than $r = `r printnum(-Kahane.r.80)`$ and larger than $r = `r printnum(Kahane.r.80)`$ can be rejected at an alpha level of 5%.
+If other researchers are interested in examining possible smaller effects, they can design studies with a larger sample size that can detect, or reject, smaller effect sizes.
 
-\begin{textbox}
-\caption{Calculating an equivalence test in R.}
-\begin{framed}
-{
-\small
-\setstretch{1}
-If you want to calculate an equivalence test based on summary statistics, you can use the ``TOSTER'' package in the open-source programming language R.
+\begin{textbox} \caption{Calculating an equivalence test in R.} \begin{framed} { \small \setstretch{1} If you want to calculate an equivalence test based on summary statistics, you can use the ``TOSTER'' package in the open-source programming language R.
 Using TOSTER requires very little experience with R, or programming in general.
 Below are the few simple steps required to calculate an equivalence test with TOSTER.
-\begin{enumerate}
-\raggedright
-  \setlength{\parskip}{1pt}
-\item Download, install, and open R.
-\item Run \verb!install.packages("TOSTER")! in the console.
-\item Run \verb!library(TOSTER)! in the console.
+\begin{enumerate} \raggedright \setlength{\parskip}{1pt} \item Download, install, and open R.
+\item Run \verb!install.packages("TOSTER")!
+in the console.
+\item Run \verb!library(TOSTER)!
+in the console.
 \item The example below reproduces the result of example 1 in R.
@@
 Results and a plot will automatically be printed.
-\begin{verbatim}
-TOSTtwo(m1  = 4.7857, 
-        m2  = 4.6569,
-        sd1 = 1.0897, 
-        sd2 = 1.1895, 
-        n1 = 49, 
-        n2 = 51, 
-        low_eqbound_d = -0.6401696, 
-        high_eqbound_d = 0.6401696, 
-        alpha = 0.05, 
-        var.equal = FALSE)
-\end{verbatim}
-\item The example above uses the \verb!TOSTtwo()!
+\begin{verbatim} TOSTtwo(m1 = 4.7857, m2 = 4.6569, sd1 = 1.0897, sd2 = 1.1895, n1 = 49, n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05, var.equal = FALSE) \end{verbatim} \item The example above uses the \verb!TOSTtwo()!
 function in the TOSTER package which calculates an equivalence test for an independent samples t-test.
@@
 and other functions of the TOSTER package by going to {\sffamily \url{https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf}} in your web browser.
-\end{enumerate}}
-\end{framed} 
-\end{textbox}
+\end{enumerate}} \end{framed} \end{textbox}
 
@@
 Equivalence testing is a relatively simple and flexible technique to falsify hypotheses.
-As long as we can calculate a confidence interval around a parameter estimate of interest, we can compare it to a smallest effect size of interest.
-The result of an equivalence test can be obtained by mere visual inspection of the CI [@Seaman1998; @Tryon2001], or by performing two one-sided tests.
-
-Equivalence bounds should be specified before results are known, ideally as part of a pre-registration [cf.
-@Piaggio2006].
-Researchers might be tempted to set equivalence bounds *after* looking at their data, but this "boundary hacking" will increase error rates.
+As long as we can calculate a confidence interval around a parameter estimate, we can compare it to a smallest effect size of interest.
+The result of an equivalence test can be obtained by mere visual inspection of the CI [@Seaman1998; @Tryon2001], or by performing two one-sided tests (the TOST procedure).
 
@@
 The question manifests itself in the bounds we set: Is our data surprising if the true effect lies outside of the bounds?
-If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003].
-Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2004], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
+If we test against very wide bounds, observing statistical equivalence can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003].
+Examining the papers citing @Lakens2017a, we found that some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s $d$ of “medium” size was chosen.] An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2003], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
+Equivalence bounds should be specified before results are known, ideally as part of a pre-registration [cf.
+@Piaggio2006].
+Researchers might be tempted to set equivalence bounds *after* looking at their data, but like any other results-dependent decision, this "boundary hacking" will increase error rates.
 
-For many researchers, the biggest challenge in designing studies where they will perform both NHST and equivalence tests by default will be the need to specify the smallest effect size of interest.
-Psychological theories are usually too vague to derive precise predictions, and if there are no theoretical reference points, natural constraints, or prior studies a researcher can use to define the SESOI for a new hypothesis, any choice will be arbitrary to some extent.
+We recommend that researchers by default perform both a null hypothesis significance test and an equivalence test on their data.
+The biggest challenge for researchers will be specifying the smallest effect size of interest.
+Psychological theories are usually too vague to derive precise predictions, and if there are no theoretical reference points, natural constraints, or prior studies a researcher can use to define the SESOI for a hypothesis test, any choice will be somewhat arbitrary.
 In some research lines researchers might simply use equivalence tests to reject consecutively smaller effect sizes by performing studies with increasingly larger sample sizes while controlling error rates, until no one is willing to invest the time and resources needed to examine the presence of even smaller effects.
-Nevertheless, it is important to realize that not specifying a SESOI in our research questions will severely hinder theoretical progress.
-While our initial choices may be arbitrary and subjective, our goal should be to use them as starting points to develop better and more falsifiable theories over time, and ultimately be able to make more precise predictions than we can today.
+Nevertheless, it is important to realize that not specifying a SESOI for our research questions at all will severely hinder theoretical progress [@Morey2016].
+While our initial choices of a smallest effect size of interest may be arbitrary, incorporating equivalence tests in our statistical toolbox will in time contribute to better --- and more falsifiable --- theories.
 

### 2017-11-16 manuscript.Rmd
fixed/updated the text box

@@
 
-\begin{textbox} \caption{Calculating an equivalence test in R.} \begin{framed} { \small \setstretch{1} If you want to calculate an equivalence test based on summary statistics, you can use the ``TOSTER'' package in the open-source programming language R.
-Using TOSTER requires very little experience with R, or programming in general.
-Below are the few simple steps required to calculate an equivalence test with TOSTER.
-\begin{enumerate} \raggedright \setlength{\parskip}{1pt} \item Download, install, and open R.
-\item Run \verb!install.packages("TOSTER")!
-in the console.
-\item Run \verb!library(TOSTER)!
-in the console.
-\item The example below reproduces the result of example 1 in R.
-The parameters of the test are defined inside the parentheses, and you can run \verb!?TOSTtwo!
-in the console to get information about them.
+\begin{textbox} 
+\caption{Calculating an equivalence test in R.} 
+\begin{framed} 
+{\sffamily \small \setstretch{1}
+Equivalence tests can be performed  based on summary statistics (e.g., means, standard deviations, and sample sizes) with the ``TOSTER'' package in the open-source programming language R.
+Using TOSTER in R requires no more programming experience than reproducing three lines of code.
+\vspace{2mm}
+
+The code below reproduces the result of Example 2 in R, which can be typed into R or RStudio.
+The parameters of the test are defined inside the parentheses.
 Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code.
 Results and a plot will automatically be printed.
-\begin{verbatim} TOSTtwo(m1 = 4.7857, m2 = 4.6569, sd1 = 1.0897, sd2 = 1.1895, n1 = 49, n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05, var.equal = FALSE) \end{verbatim} \item The example above uses the \verb!TOSTtwo()!
-function in the TOSTER package which calculates an equivalence test for an independent samples t-test.
-If you need to calculate TOST for a different type of test (e.g.
-a correlation), you can find documentation for \verb!TOSTr()!
-and other functions of the TOSTER package by going to {\sffamily \url{https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf}} in your web browser.
-\end{enumerate}} \end{framed} \end{textbox}
+A help file provides more detailed information.
+
+\begin{verbatim}
+install.packages("TOSTER")
+
+library(TOSTER)
+
+TOSTtwo(m1 = 4.7857, m2 = 4.6569, sd1 = 1.0897, sd2 = 1.1895, n1 = 49, 
+n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05, 
+var.equal = FALSE)
+\end{verbatim}
+}
+\end{framed} 
+\end{textbox}
 

### 2017-11-16 examples_plotgrid.R
Removed library(grid)

@@ -5,7 +5,6 @@
 library(purrr)
 library(gridExtra)
 library(ggplot2)
-library(grid)
 library(lattice)
 
 library("TOSTER")

### 2017-11-16 manuscript.Rmd
Minor edit

@@
 ```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
-**A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
-**B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
-**C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
+**A)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
+**B)**  Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
+**C)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
 **D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}
@@
 **D)** Example 4 - Proportion Difference.
-**E)** Example 5 - Pearson Correlation")}
+**E)** Example 5 - Pearson Correlation.")}
 # Plot the TOST examples
@@
 
-@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as effect sizes smaller than $d = printnum(sesoi)$.
+@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as effect sizes smaller than $d = `r printnum(sesoi)`$.
 For example, in grade 3 the difference is $d = 0.04$, with a standard error of 0.002.
-When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of printp(alpha) to correct for multiple comparisons), using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that indeed, all effect size estimates are measured with such high precision that they are statistically equivalent and can be considered trivially small, according to the authors’ definition (Figure 2C).
+When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of `r printp(alpha)` to correct for multiple comparisons), using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that indeed, all effect size estimates are measured with such high precision that they are statistically equivalent and can be considered trivially small, according to the authors’ definition (Figure 2C).
 In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printnum(max(gr3$TOST_p1, gr3$TOST_p2))`$ However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large.
-For grade 3, $Z = `r printnum(gr3.z)`$, $p = `r printnum(gr3.p)`$.
+For grade 3, $z = `r printnum(gr3.z)`$, $p = `r printnum(gr3.p)`$.
 This is one way in which the use of equivalence tests improves hypothesis testing procedures in science, by allowing researchers to distinguish between statistical significance and practical significance.

### 2017-11-16 manuscript.Rmd
Minor edit

@@
 When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of `r printp(alpha)` to correct for multiple comparisons), using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that indeed, all effect size estimates are measured with such high precision that they are statistically equivalent and can be considered trivially small, according to the authors’ definition (Figure 2C).
-In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printnum(max(gr3$TOST_p1, gr3$TOST_p2))`$ However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large.
+In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printnum(max(gr3$TOST_p1, gr3$TOST_p2))`$.
+However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large.
 For grade 3, $z = `r printnum(gr3.z)`$, $p = `r printnum(gr3.p)`$.

### 2017-11-16 manuscript.Rmd
Added author contributions.

@@
   We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions.
+  DL: Article conceptualization, programming TOSTER package and spreadsheet, main contributor to the introduction section and, discussion section.
+  AS and PI: Main contributors to examples and figures.
+  All authors contributed to manuscript conceptualization, editing, review for submission, and additional software development.
+AS and PI contributed equally to the manuscript, and order was determined by executing the following commands in R:
+  \newline
+  `set.seed(7998976/5271)`
+  \newline
+  `x <- sample(c("Anne", "Peder"), 1)`
+  \newline
+  `print(paste("The winner is", x, "!"))`
+  \newline
 

### 2017-11-16 Example2_Brandt-et-al_study1.R
add folder structure

@@ -1,35 +1,0 @@
-library(TOSTER)
-library(pwr)
-
-Brandt.data <- read.csv("Example2_Brandt-et-al_study1_data-cleaned.csv", header = TRUE, sep = ",")
-Brandt.data <- as.data.frame(Brandt.data)
-
-# original study: Banerjee et al., study 1
-orig.m1 <- 4.71
-orig.m2 <- 5.3
-orig.sd1 <- 0.85
-orig.sd2 <- 0.97
-orig.d <- 0.65
-orig.t <- 2.03
-orig.p <- 0.049
-orig.N <- 40 #group size unknown, therefore equal n assumed
-orig.df <- orig.N-2
-
-# replication study: Brandt et al., study 1
-rep.m1 <- mean(Brandt.data$WellLitSca[Brandt.data$ExpCond==-1])
-rep.m2 <- mean(Brandt.data$WellLitSca[Brandt.data$ExpCond==1])
-rep.sd1 <- sd(Brandt.data$WellLitSca[Brandt.data$ExpCond==-1])
-rep.sd2 <- sd(Brandt.data$WellLitSca[Brandt.data$ExpCond==1])
-rep.n1 <- sum(Brandt.data$ExpCond==-1)
-rep.n2 <- sum(Brandt.data$ExpCond==1)
-rep.test <- t.test(Brandt.data$WellLitSca[Brandt.data$ExpCond==-1],Brandt.data$WellLitSca[Brandt.data$ExpCond==1], var.equal=FALSE, paired=FALSE)
-rep.t <- rep.test$statistic
-rep.p <- rep.test$p.value
-rep.df <- rep.test$parameter
-rep.d <- (rep.m1 - rep.m2)/sqrt(((rep.n1-1)*rep.sd1^2 + (rep.n2-1)*rep.sd2^2)/(rep.n1+rep.n2-2))
-
-
-# Calculate small telescopes effect (d) effect orig. study had 33% power to detect
-d.33 <- (pwr.t.test(n = orig.N/2, d = NULL, sig.level = 0.05, power = 0.33, type = "two.sample", alternative = "two.sided"))$d
-
-Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.33, high_eqbound_d = d.33, var.equal = FALSE)

### 2017-11-16 Example4_Lynott-et-al.R

@@ -1,31 +1,0 @@
-se <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
-crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05
-
-
-crit.diff <- crit.z * se  # Convert z to the critical difference
-
-
-
-library(TOSTER)  # Install TOSTER using devtools::install_github("Lakens/TOSTER"), or download TOSTtwo.prop.R from https://github.com/Lakens/TOSTER/tree/master/R and then source it. 
-
-
-#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
-lynott.data <- read.csv("Example4_Lynott-et-al_data-cleaned.csv")
-lynott.data <- lynott.data[lynott.data$include1 == 1,]
-
-
-#  We then calculate the summary statistics from the data that we need in the TOST calculation below
-n.hot  <- sum(lynott.data$packtype == 1)
-n.cold <- sum(lynott.data$packtype == 0)
-selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
-selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
-diff <- selfish.hot - selfish.cold
-
-#  We can now run the TOST!
-Lynott <- TOSTtwo.prop(prop1 = selfish.cold,
-                     prop2 = selfish.hot,
-                     n1    = n.cold,
-                     n2    = n.hot,
-                     low_eqbound  = -crit.diff,
-                     high_eqbound = crit.diff,
-                     alpha = 0.05)

### 2017-11-16 Manuscript/manuscript.Rmd
add manuscript folder

--- 2017-11-16 manuscript.Rmd
@@
 These concrete and easy to follow examples will illustrate all approaches to setting equivalence bounds discussed above, and demonstrate how to perform and report equivalence tests.
-The code to reproduce these analyses is available at <https://osf.io/qamc6/>.
+The code to reproduce these analyses is available at <https://osf.io/qamc6/>^[The OSF project contains all necessary files to reproduce examples 1--5 in R, jamovi [@jamovi2017], and in a spreadsheet; as well as a preprint of this manuscript.
+The manuscript, including the figures and statistical analyses, was created using R version 3.4.2 [@RCoreTeam2017] and RStudio version 1.1.383 [@RStudioTeam2016], with the packages `TOSTER` version 0.3 [@Lakens2017c], `pwr` version 1.2--1 [@Champely2017], `rmarkdown` version 1.7[@Allaire2017], `knitr` version 1.17 [@Xie2017], `papaja` version 0.1.0.9492 [@Aust2017], `ggplot2` version 2.2.1 [@Wickham2009], `gridExtra` version 2.3 [@Auguie2017], `lattice` version 0.20--35 [@Sarkar2008], and `purrr` version 0.2.4 [@Henry2017].].
 

### 2017-11-16 Bayes_functions/Example from Neil Mclathie/2017.10.05 TOSTtwoBayes Raw with Cauchy+HalfCauchy.R

@@ -1,209 +1,0 @@
-TOSTtwo.raw.nm<-function(m1,m2,sd1,sd2,n1,n2,low_eqbound, high_eqbound, alpha, var.equal, prior_dist, effect_prior, se_prior, df_prior)
-  {
-  if(missing(alpha)) {
-    alpha<-0.05
-  }
-  if(missing(var.equal)) {
-    var.equal<-FALSE
-  }
-  if(var.equal==TRUE) {
-    sdpooled<-sqrt((((n1 - 1)*(sd1^2)) + (n2 - 1)*(sd2^2))/((n1+n2)-2)) #calculate sd pooled
-    t1<-((m1-m2)-low_eqbound)/(sdpooled*sqrt(1/n1 + 1/n2))
-    degree_f<-n1+n2-2
-    p1<-pt(t1, degree_f, lower.tail=FALSE)
-    t2<-((m1-m2)-high_eqbound)/(sdpooled*sqrt(1/n1 + 1/n2))
-    p2<-pt(t2, degree_f, lower.tail=TRUE)
-    LL90<-(m1-m2)-qt(1-alpha, degree_f)*(sdpooled*sqrt(1/n1 + 1/n2))
-    UL90<-(m1-m2)+qt(1-alpha, degree_f)*(sdpooled*sqrt(1/n1 + 1/n2))
-    LL95<-(m1-m2)-qt(1-(alpha/2), degree_f)*(sdpooled*sqrt(1/n1 + 1/n2))
-    UL95<-(m1-m2)+qt(1-(alpha/2), degree_f)*(sdpooled*sqrt(1/n1 + 1/n2))
-    t<-(m1-m2)/(sdpooled*sqrt(1/n1 + 1/n2))
-    pttest<-2*pt(-abs(t), df=degree_f)
-  } else {
-    sdpooled<-sqrt((sd1^2 + sd2^2)/2) #calculate sd root mean squared for Welch's t-test
-    t1<-((m1-m2)-low_eqbound)/sqrt(sd1^2/n1 + sd2^2/n2) #welch's t-test lower bound
-    degree_f<-(sd1^2/n1+sd2^2/n2)^2/(((sd1^2/n1)^2/(n1-1))+((sd2^2/n2)^2/(n2-1))) #degrees of freedom for Welch's t-test
-    p1<-pt(t1, degree_f, lower.tail=FALSE) #p-value for Welch's TOST t-test
[…]

### 2017-11-16 Bayes_functions/TOSTtwoBayes.R

@@ -1,167 +1,0 @@
-#' TOST function for an independent t-test (Cohen's d)
-#' @param m1 mean of group 1
-#' @param m2 mean of group 2
-#' @param sd1 standard deviation of group 1
-#' @param sd2 standard deviation of group 2
-#' @param n1 sample size in group 1
-#' @param n2 sample size in group 2
-#' @param low_eqbound_d lower equivalence bounds (e.g., -0.5) expressed in standardized mean difference (Cohen's d)
-#' @param high_eqbound_d upper equivalence bounds (e.g., 0.5) expressed in standardized mean difference (Cohen's d)
-#' @param alpha alpha level (default = 0.05)
-#' @param var.equal logical variable indicating whether equal variances assumption is assumed to be TRUE or FALSE.  Defaults to FALSE.
-#' @return Returns TOST t-value 1, TOST p-value 1, TOST t-value 2, TOST p-value 2, degrees of freedom, low equivalence bound, high equivalence bound, low equivalence bound in Cohen's d, high equivalence bound in Cohen's d, Lower limit confidence interval TOST, Upper limit confidence interval TOST
-#' @importFrom stats pnorm pt qnorm qt
-#' @importFrom graphics abline plot points segments title
-#' @examples
-#' ## Eskine (2013) showed that participants who had been exposed to organic
-#' ## food were substantially harsher in their moral judgments relative to
-#' ## those exposed to control (d = 0.81, 95% CI: [0.19, 1.45]). A
-#' ## replication by Moery & Calin-Jageman (2016, Study 2) did not observe
-#' ## a significant effect (Control: n = 95, M = 5.25, SD = 0.95, Organic
-#' ## Food: n = 89, M = 5.22, SD = 0.83). Following Simonsohn's (2015)
-#' ## recommendation the equivalence bound was set to the effect size the
-#' ## original study had 33% power to detect (with n = 21 in each condition,
[…]

### 2017-11-16 Manuscript/Example3_Hyde-et-al.R

@@ -1,0 +1,20 @@
+
+# Hyde, Lindberg, Linn, Ellis, and Williams (2008) meta analysis
+
+sesoi <- 0.1
+alpha <- 0.005
+
+gr2  <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr3  <- TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr4 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr5 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr6 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr7 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr8 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr9 <- TOSTmeta(ES = -0.01, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr10 <- TOSTmeta(ES = 0.04, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+gr11 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
+
+
+gr3.z <- (gr3$ES/0.002)  # Calculate Z value for grade 3
+gr3.p <- 2*pnorm(-abs(gr3.z))  # Calculate p value for grade 3

### 2017-11-16 Manuscript/Example4_Lynott-et-al.R

@@ -1,0 +1,31 @@
+se <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
+crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05
+
+
+crit.diff <- crit.z * se  # Convert z to the critical difference
+
+
+
+library(TOSTER)  # Install TOSTER using devtools::install_github("Lakens/TOSTER"), or download TOSTtwo.prop.R from https://github.com/Lakens/TOSTER/tree/master/R and then source it. 
+
+
+#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
+lynott.data <- read.csv("Example4_Lynott-et-al_data-cleaned.csv")
+lynott.data <- lynott.data[lynott.data$include1 == 1,]
+
+
+#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+n.hot  <- sum(lynott.data$packtype == 1)
+n.cold <- sum(lynott.data$packtype == 0)
+selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
+selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
+diff <- selfish.hot - selfish.cold
+
+#  We can now run the TOST!
+Lynott <- TOSTtwo.prop(prop1 = selfish.cold,
+                     prop2 = selfish.hot,
+                     n1    = n.cold,
+                     n2    = n.hot,
+                     low_eqbound  = -crit.diff,
+                     high_eqbound = crit.diff,
+                     alpha = 0.05)

### 2017-11-16 Manuscript/Example5_Kahane-et-al-study4.R

@@ -1,0 +1,26 @@
+library(TOSTER)
+library(pwr)
+
+## Kahane, Everett, Earp, Farias, & Savulescu (2015)
+##
+## Study 4, 5.2.1. ii., p. 203: "There  was  no  relationship  between  perceived wrongness in 
+## personal dilemmas and in the ‘greater good’ dilemmas (r = -.04, p = .53): that is, people
+## who were more ‘utilitarian’ in the personal dilemmas were not more likely to be more truly
+## utilitarian in the other dilemmas, and vice versa. This lack of relationship held even when
+## controlling for primary psychopathy in a partial second order correlation (r = -.07, p = .32)."
+
+Kahane.data <- read.csv("Example5_Kahane-et-al-study4_data-cleaned.csv", header = TRUE, sep = ",")
+Kahane.data <- as.data.frame(Kahane.data)
+
+# sample size
+Kahane.N <- min(sum(!is.na(Kahane.data$ClassicSacrificialDilemmas)), sum(!is.na(Kahane.data$GreaterGoodDilemmas)))
+
+# correlation b/w perceived wrongness in personal dilemmas and 'greater good' dilemmas
+Kahane.r.GG <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$estimate
+Kahane.p <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$p.value
+
+# correlation the study had 80% power to detect
+Kahane.r.80 <- (pwr.r.test(n = Kahane.N, r = NULL, sig.level = 0.05, power = 0.8, alternative = "two.sided"))$r
+
+# TOST with bounds set to the correlation the study had 80% power to detect
+Kahane <- TOSTr(n = Kahane.N, r = Kahane.r.GG, low_eqbound_r = -Kahane.r.80, high_eqbound_r = Kahane.r.80, alpha = 0.05)

### 2017-11-16 Manuscript/H0H1_plot.R

@@ -1,0 +1,67 @@
+####  Setup ----------------------------------------
+library(ggplot2)
+library(gridExtra)
+
+plotheight <- 0.9
+lowerbound <- -0.2
+upperbound <- 0.2
+df <- data.frame()
+
+####  Base plot ------------------------------------
+baseplot <-   ggplot(df) +
+  scale_y_continuous(limits = c(0,plotheight+0.02), breaks=NULL) + # no y-axis will be displayed
+  theme_classic() +
+  theme(plot.title = element_text(size = rel(1), face = "bold"), #font size & appearance for plot titles
+        axis.title.y = element_blank(), #remove title of y-axis
+        axis.title.x = element_text(size=rel(0.7), lineheight = 0.5), #font size for x-axis label
+        plot.margin=unit(c(0.5,0.8,0.5,0.8),"cm")) #add padding around each plot to make them look nicer when combined; margin order: top, right, bottom, left
+
+#NHST plot
+NHSTplot <- baseplot +  
+  scale_x_continuous(limits = c(-0.85,0.85), breaks=c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8),
+                     labels=c("", "", "", "", "0", "", "", "", ""), name = "effect size") +
+  ggtitle("A: Classic NHST (two-sided)") +
+  annotate("segment", x = 0, xend = 0, y = plotheight-plotheight/2, yend = -Inf) + #vertical line at x=0 (H0)
+  annotate("text", size = rel(3.5), x=0, y = plotheight-plotheight/20, parse=TRUE, label="H[0]", hjust = 0.3) + #label for point null (H0)
+  annotate("segment", x = 0, xend = 0, y = plotheight-plotheight/6, yend=plotheight-plotheight/2.3, 
+           arrow = arrow(type = "closed", length=unit(2, "mm"))) + #arrow pointing from H0 label to H0 line
+  annotate("text", size = rel(3.5), x=-0.45, y=plotheight/3, parse=TRUE, label="H[1]") + #label for lower area (H1)
+  annotate("text", size = rel(3.5), x=0.45, y=plotheight/3, parse=TRUE, label="H[1]") #label for upper area (H1)
+
+#minimal effects test plot
+METplot <-   baseplot +  
[…]

### 2017-11-16 app.R

@@ -1,198 +1,0 @@
-#To do: Scale y-axis p-curve
-#       Include Likelihood H1/H0
-
-library(shiny)
-library(shinythemes)
-ui <- fluidPage(theme=shinytheme("flatly"),
-                titlePanel("Distribution of Cohen's d, p-values, and power curves for an independent two-tailed t-test"),
-                sidebarLayout(
-                  sidebarPanel(numericInput("N", "Participants per group:", 50, min = 1, max = 1000),
-                               sliderInput("d", label = HTML("Cohen's d (&delta;)"), min = 0, max = 2, value = 0.5, step= 0.01),
-                               sliderInput("p_upper", "alpha, or p-value (upper limit):", min = 0, max = 1, value = 0.05, step= 0.005),
-                               uiOutput("p_low"),
-                               h4(textOutput("pow0")),br(),
-                               h4(textOutput("pow1")),br(),
-                               h4(textOutput("pow2")),br(),
-                               h4("The three other plots indicate power for a range of alpha levels (top right), sample sizes per group (bottom left), and effect sizes (bottom right). The bottom right figure illustrates the point that when the true effect size of a study is unknown, the power of a study is best thought of as a curve, not as a single value."),br()
-                  ),
-                  mainPanel(
-                    plotOutput("plot_d", width = "1004px", height = "500px"),
-                    splitLayout(style = "border: 1px solid silver:", cellWidths = c(500,500),cellHeights = c(800,800), 
-                                plotOutput("pdf"),
-                                plotOutput("cdf")
-                    ),
-                    splitLayout(style = "border: 1px solid silver:", cellWidths = c(500,500), 
[…]

### 2017-11-17 Leftovers/Box1_Equivalence_test_in_R.Rmd
update folder structure

---
title: "Box 1"
author: "Peder M Isager"
date: "November 8, 2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Box 1: Calculating an equivalence test in R

If you want to calculate an equivalence test based on summary statistics, you can use the "TOSTER" package in the open-source progamming language R. Using TOSTER requires very little eperience with R, or programming in general. Below are the few simple steps required to calculate an equivalence test with TOSTER.

1. Download, install, and open R.
2. Run `install.packages("TOSTER")` in the console.
3. Run `library(TOSTER)` in the console.
4. The example below reproduces the result of example 1 in R. The parameters of the test are defined inside the parantheses, and you can run `?TOSTtwo` in the console to get information about them. Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code. Results and a plot will automatially be printed.
```{r eval=FALSE} 
TOSTtwo(m1  = 4.7857, m2  = 4.6569, sd1 = 1.0897, sd2 = 1.1895, n1 = 49, n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05, var.equal = FALSE)
```

5. The example above uses the `TOSTtwo()` function in the TOSTER package which calculates an equivalence test for an independent samples t-test. If you need to calculate TOST for a different type of test (e.g. a correlation), you can find documentation for `TOSTr()` and other functions of the TOSTER package by going to https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf in your web browser.

### 2017-11-17 Leftovers/NHST_and_TOST.Rmd
update folder structure

--- 2017-10-18 NHST_and_TOST.Rmd
@@
 output:
+  pdf_document: default
+  html_notebook: default
   html_document: default
-  html_notebook: default
-  pdf_document: default
 editor_options: 

### 2017-11-17 Leftovers/NSEQ_Lynott_example.Rmd
update folder structure

---
title: "SNEQ_Lynott_example"
author: "Peder M Isager"
date: "October 17, 2017"
output: html_document
---

## Example 4: Statistically Inferior and Not Statistically Different
```{r include=FALSE}
se <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05

crit.diff <- crit.z * se  # Convert z to the critical difference

library(TOSTER)  # Install TOSTER using devtools::install_github("Lakens/TOSTER"), or download TOSTtwo.prop.R from https://github.com/Lakens/TOSTER/tree/master/R and then source it.

#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
lynott.data <- read.csv("NSEQ_Lynott_example_data.csv")
lynott.data <- lynott.data[lynott.data$include1 == 1,]

#  We then calculate the summary statistics from the data that we need in the TOST calculation below
n.hot            <- sum(lynott.data$packtype == 1)
n.cold           <- sum(lynott.data$packtype == 0)
selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
diff <- selfish.hot - selfish.cold

#  We can now run the TOST!
tost <- TOSTtwo.prop(prop1 = selfish.cold,
                         prop2 = selfish.hot,
                         n1        = n.cold,
                         n2        = n.hot,
                         low_eqbound  = -crit.diff,
                         high_eqbound = crit.diff,
                         alpha = 0.05)

```

Lynott et al. (2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008). They observed that `r round(selfish.cold,2)*100`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot,2)*100`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same. In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r round(diff,2)`$, $Z=`r round(tost$NHST_z,2)`$, $p=`r round(tost$NHST_p,2)`$). Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.

In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be. However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect. This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- `r round(qnorm(1-0.05/2),2)`$). To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:

Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study. In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{upper}$ because an effect with CIs crossing $\Delta_{lower}$ can still be *lower* than the original effect.

We find that the test against $\Delta_{upper}$ is `r ifelse(tost$TOST_p2 < tost$alpha, "significant", "non-significant")`, $z =$ `r round(tost$TOST_z2, 2)`, $p =$ `r ifelse(round(tost$TOST_p2, 3) >= 0.001, paste("=", round(tost$TOST_p2, 2)), "< 0.001")` (see figure 2D). Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.

### 2017-11-17 Leftovers/NSNEQ_Moon_Example.Rmd
update folder structure

---
title: "NSNEQ_Moon_Example"
author: "Peder M Isager"
date: "2017 M10 16"
output:
  html_document: default
  pdf_document: default
---

```{r include=FALSE}

#  Load the relevant datafile into R
moon.data        <- read.csv("NSNEQ_Moon_Example_data.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)

#  Set the bounds, in this case 1 grade point, which corresponds to 6.25%
sesoi <- 0.0625

#  Calculate the summary statistics from the data that we need in the TOST calculation below
mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
n.asian          <- sum(moon.data$identity_salience == 1)
n.control        <- sum(moon.data$identity_salience == 2)

df <- (sd.asian^2 / n.asian + sd.control^2 / n.control)^2 / (((sd.asian^2 / n.asian)^2 / (n.asian-1)) + ((sd.control^2 / n.control)^2 / (n.control - 1)))  # Degrees of freedom for Welch's t-test
sample.sd <- sqrt((sd.asian^2 + sd.control^2) / 2)  # Calculate sd root mean squared for Welch's t-test
sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate the se of the difference based on sd.pooled

t <- (mean.asian - mean.control) / sample.se  # Welch t-test

d <- (mean.asian - mean.control) / sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.control + n.control) - 2))

p <- 2 * pt(t, df)

sdpooled <- sqrt((((n.asian - 1)*(sd.asian^2)) + (n.control - 1)*(sd.control^2))/((n.asian+n.control)-2))
d <- (mean.asian - mean.control) / sdpooled

#  We can now run the TOST!
#  Performing the TOST using the 'TOSTtwo.raw' function allows you to specify equivalence bounds in raw values
tost <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
                                m2  = mean.control,
                                sd1 = sd.asian,
                                sd2 = sd.control,
                                n1  = n.asian,
                                n2  = n.control,
                                low_eqbound  = -sesoi,
                                high_eqbound =  sesoi,
                                alpha = .05)

```

Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity. They observed a slightly reversed difference between the Asian primed group (*n* = 53, *M* = 0.46, *SD* = 0.17) and the control (*n* = 48, *M* = 0.50, *SD* - 0.18) which was not significant, d = -0.23, *t*(96.62) = 1.15, *p* = 0.26, two-sided). A non-significant null-hypothesis test leaves two possible conclusions:

1. The data suggest the absence of a meaningful effect, and we can reject effects large enough to be meaningful.
2. The data are not sensitive enough to tell us whether a meaningful effect is present or absent.

Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%). We can decide that we are only interested in test differences so far as they correspond to at least a 1 grade point increase or decrease. Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.

We then calculate the TOST for a two-sample Welch's t-test, assuming an alpha of 0.05 (two-sided), and using +/- the SESOI of 0.0625 as our equivalence bounds ($\Delta$).

We find that the TOST is non-significant, t(`r round(tost$TOST_df, 2)`) = `r round(min(abs(c(tost$TOST_t1, tost$TOST_t2))), 2)`, p = `r round(max(tost$TOST_p1, tost$TOST_p2), 2)` (figure 2C). Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions. Further studies are needed, which can be analyzed using a (small-scale) meta-analysis. The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both. Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{lower}$ = -0.1 and $\Delta_{upper}$ = 0.1).

[^footnote]: In this example, the bounds of the TOST was set based on a raw effect of 6.25%. We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$. Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication. Now we would no longer be answering a question about a raw effect independent of the samples. Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study. Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$. As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.

### 2017-11-17 manuscript.Rmd
update folder structure

@@
 ---
-
+getwd()
 ```{r include = FALSE}
@@
 source("H0H1_plot.R")
+getwd()
 ```

### 2017-11-17 Example1_MoonRoeder.R

@@ -1,0 +1,38 @@
+#  Load the relevant datafile into R
+moon.data        <- read.csv("Example1_MoonRoeder_data-cleaned.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)
+
+#  Set the bounds, in this case 1 grade point, which corresponds to 6.25%
+sesoi <- 0.0625
+
+#  Calculate the summary statistics from the data that we need in the TOST calculation below
+mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
+mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
+sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
+sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
+n.asian          <- sum(moon.data$identity_salience == 1)
+n.control        <- sum(moon.data$identity_salience == 2)
+sample.sd <- sqrt((sd.asian^2 + sd.control^2) / 2)  # Calculate sd root mean squared for Welch's t-test
+sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate the se of the difference based on sd.pooled
+
+df <- (sd.asian^2 / n.asian + sd.control^2 / n.control)^2 / (((sd.asian^2 / n.asian)^2 / (n.asian-1)) + ((sd.control^2 / n.control)^2 / (n.control - 1)))  # Degrees of freedom for Welch's t-test
+
+t <- (mean.asian - mean.control) / sample.se  # Welch t-test
+
+p <- 2 * pt(t, df)
+
+sdpooled <- sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.asian + n.control) - 2))
+d <- (mean.asian - mean.control) / sdpooled
+
+
+
+#  We can now run the TOST!
+#  Performing the TOST using the 'TOSTtwo.raw' function allows you to specify equivalence bounds in raw values
+Moon <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
+                            m2  = mean.control,
+                            sd1 = sd.asian,
+                            sd2 = sd.control,
+                            n1  = n.asian,
+                            n2  = n.control,
+                            low_eqbound  = -sesoi,
+                            high_eqbound =  sesoi,
+                            alpha = .05)

### 2017-11-17 Example2_Brandt-et-al_study1.R

@@ -1,0 +1,35 @@
+library(TOSTER)
+library(pwr)
+
+Brandt.data <- read.csv("Example2_Brandt-et-al_study1_data-cleaned.csv", header = TRUE, sep = ",")
+Brandt.data <- as.data.frame(Brandt.data)
+
+# original study: Banerjee et al., study 1
+orig.m1 <- 4.71
+orig.m2 <- 5.3
+orig.sd1 <- 0.85
+orig.sd2 <- 0.97
+orig.d <- 0.65
+orig.t <- 2.03
+orig.p <- 0.049
+orig.N <- 40 #group size unknown, therefore equal n assumed
+orig.df <- orig.N-2
+
+# replication study: Brandt et al., study 1
+rep.m1 <- mean(Brandt.data$WellLitSca[Brandt.data$ExpCond==-1])
+rep.m2 <- mean(Brandt.data$WellLitSca[Brandt.data$ExpCond==1])
+rep.sd1 <- sd(Brandt.data$WellLitSca[Brandt.data$ExpCond==-1])
+rep.sd2 <- sd(Brandt.data$WellLitSca[Brandt.data$ExpCond==1])
+rep.n1 <- sum(Brandt.data$ExpCond==-1)
+rep.n2 <- sum(Brandt.data$ExpCond==1)
+rep.test <- t.test(Brandt.data$WellLitSca[Brandt.data$ExpCond==-1],Brandt.data$WellLitSca[Brandt.data$ExpCond==1], var.equal=FALSE, paired=FALSE)
+rep.t <- rep.test$statistic
+rep.p <- rep.test$p.value
+rep.df <- rep.test$parameter
+rep.d <- (rep.m1 - rep.m2)/sqrt(((rep.n1-1)*rep.sd1^2 + (rep.n2-1)*rep.sd2^2)/(rep.n1+rep.n2-2))
+
+
+# Calculate small telescopes effect (d) effect orig. study had 33% power to detect
+d.33 <- (pwr.t.test(n = orig.N/2, d = NULL, sig.level = 0.05, power = 0.33, type = "two.sample", alternative = "two.sided"))$d
+
+Brandt <- TOSTtwo(m1=rep.m1, m2=rep.m2, sd1=rep.sd1, sd2 = rep.sd2, n1 = rep.n1, n2=rep.n2, low_eqbound_d = -d.33, high_eqbound_d = d.33, var.equal = FALSE)

### 2017-11-17 Leftovers/TOST_simulator.R

@@ -1,0 +1,201 @@
+
+####  Setup ----------------------------------------
+
+library(TOSTER)
+library(purrr)
+download.file(url = "https://raw.githubusercontent.com/pederisager/TOSTER/TOST-modification/R/TOSTtwo.R", destfile = "TOSTtwo.prop.R")
+source(file = "TOSTtwo.prop.R") #  Source modified TOSTtwo from pederisager Fork in order to get necessary values from TOST
+library(ggplot2)
+
+sims <- list()
+
+
+# Group 1 parameters
+mean1 <- 0    # The mean effect
+std1  <- 10   # The standard deviation
+# make n random... n1    <- 10   # The number of subjects in each group
+
+# Group 2 parameters
+mean2 <- 0    # The mean effect
+std2  <- 10   # The standard deviation
+# make n random... n2    <- 10   # The number of subjects in each group
+
+# TOST
+alpha <- 0.05
+tails <- 2 # 1 = one-tailed, 2 = two-tailed
+
+
+# Set simulation parameters
+
+nsims <- 10000  # Number of simulations
+bound.criterion <- "specific.raw"  # Choose between "crit.val", "specific", "d33", "current.power", "tost.current.power"
+
+#  When using citerion "crit.val", the critical t value based on the parameters of the current sample are used as bounds
+crit.t <- qt(p = 1-(alpha/tails), df = n1+n2-2) # Calculate critical t
+
+#  When using criterion "specific", bounds = specific
+specific <- 0.9
+
+#  When using criterion "specific.raw", bounds = specific.raw, recalculated to d for each test
+specific.raw <- 3
+
+#  When using criterion "d33", small telescops is applied, group n of previous study is random 10-100, and bounds are set to effect that previous study had tlscop.pwr power to detect. 
+tlscop.pwr <- 0.33
+
+#  When using criterion "current.power", effect that current study had  pwr.thrsh power to detect are used as bounds
+pwr.thrsh <- 0.80
+crnt.pwr  <- pwr::pwr.t2n.test(n1 = n1,
+                               n2 = n2, 
+                               power = pwr.thrsh, 
+                               sig.level = alpha, 
+                               alternative = "two.sided")$d # Calculate effect size that original study have 33% power to detect
+
+# When using criterion "tost.current.power", effect is set to bounds that current study has 80% power for
+tost.pwr <- 0.80
+crnt.tost.pwr <- powerTOSTtwo(alpha = alpha, statistical_power = tost.pwr, N = n1)[2]
+
+
+
[…]

### 2017-11-17 Leftovers/eqboundCritera.R

@@ -1,0 +1,36 @@
+#  Function which calculates the values used to determine subjective equivalence bounds for TOST tests. 
+  # Calculates the effect size (Cohens d) that a study has % power (33% by default) to detect.
+  # Calculates the the smallest detectable effect size of a study based on the degrees of freedom and p value (0.05 by default).
+
+eqboundCritera <- function(n1, n2, orig.df, p = 0.05, power = 0.33, testtype = "two.sided", var.equal = FALSE) {
+  output <- list()  # Define output as list
+  
+  # Calculate d that study has 33% chance 
+  output$d33    <- pwr::pwr.t2n.test(n1          = n1, 
+                                     n2          = n2, 
+                                     power       = power, 
+                                     sig.level   = p, 
+                                     alternative = testtype)$d # Calculate effect size that study have 33% power to detect
+  
+  output$dunb33 <- ( 1 - ( 3 / (4*orig.df - 1) ) ) * output$d33     # Calculate d unbiased for the 33% power criterion
+  
+  
+  ## Original study smallest detectable effect size ##
+  
+  output$crit.t  <- qt(p = 1-p/2, df = orig.df) # Critical t for original study
+  output$crit.d  <- output$crit.t * sqrt(1/(n1-1) + 1/(n2-1)) # Convert critical t to d
+  return(output)
+}
+
+# Test
+
+source(file = "tFromDescTwo.R")
+Shih.m1  <-  0.54
+Shih.m2  <-  0.49
+Shih.sd1 <-  0.17
+Shih.sd2 <-  0.20
+Shih.n1  <- 16
+Shih.n2  <- 16
+Shih <- tFromDescTwo(Shih.m1, Shih.m2, Shih.sd1, Shih.sd2, Shih.n1, Shih.n2, test = "student")
+
+eqboundCritera(Shih.n1, Shih.n2, Shih.df)

### 2017-11-17 Leftovers/tFromDescTwo.R

@@ -1,0 +1,52 @@
+# Function for calculating t value, cohens d, pooled standard error, etc. from means, standard deviations and group n's.
+
+tFromDescTwo <- function(m1, m2, sd1, sd2, n1, n2, test = "welch") {
+  
+  output        <- list()       # Define the output as list
+  output$m1     <- m1           # Store input arguments in output
+  output$m2     <- m2
+  output$sd1    <- sd1
+  output$sd2    <- sd2
+  output$n1     <- n1
+  output$n2     <- n2
+  output$m.diff <- m1 - m2      # Calculate the mean difference
+
+  
+  if (test == "student") {  
+    # Calculate se.diff, t and d based on a student t.test
+    print("student's t.test is performed and sample sizes are assumed to be equal")
+    output$test       <- "student"
+    output$df         <- n1 + n2 - 2  # Calculate degrees of freedom.
+    output$sd.pooled  <- sqrt( ( ( ( n1 - 1 ) * ( sd1^2 ) ) + ( n2 - 1 ) * ( sd2^2 ) ) / ( ( n1 + n2 ) - 2 ) ) # Calculate the pooled SD
+    output$se.diff    <- output$sd.pooled * sqrt( 1/n1 + 1/n2 ) # Calculate the SE of the difference.
+    
+  } else if (test == "welch") {  
+    # Calculate se.diff, t and d based on a welch t.test, and calculate sd.pooled
+    print("welch's t.test is performed and sample sizes are assumed to be unequal")
+    output$test       <- "welch"
+    output$df         <- ( sd1^2 / n1 + sd2^2 /n2 )^2 / ( ( ( sd1^2 / n1 )^2 / ( n1-1 ) ) + ( ( sd2^2 / n2 )^2 / ( n2 - 1 ) ) ) #degrees of freedom for Welch's t-test
+    output$sd.pooled  <- sqrt( ( sd1^2 + sd2^2 ) / 2 )    # Calculate sd root mean squared for Welch's t-test
+    output$se.diff    <- sqrt( sd1^2 / n1 + sd2^2 / n2 )  # Calculate the se of the difference based on sd.pooled
+    
+    
+  } else {
+    stop("The 'test' argument must be set to either 'student'or 'welch'", call. = FALSE)
+  }
+  
[…]

### 2017-11-17 Manuscript/Example1_MoonRoeder.R

@@ -1,38 +1,0 @@
-#  Load the relevant datafile into R
-moon.data        <- read.csv("Example1_MoonRoeder_data-cleaned.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)
-
-#  Set the bounds, in this case 1 grade point, which corresponds to 6.25%
-sesoi <- 0.0625
-
-#  Calculate the summary statistics from the data that we need in the TOST calculation below
-mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
-mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
-sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
-sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
-n.asian          <- sum(moon.data$identity_salience == 1)
-n.control        <- sum(moon.data$identity_salience == 2)
-sample.sd <- sqrt((sd.asian^2 + sd.control^2) / 2)  # Calculate sd root mean squared for Welch's t-test
-sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate the se of the difference based on sd.pooled
-
-df <- (sd.asian^2 / n.asian + sd.control^2 / n.control)^2 / (((sd.asian^2 / n.asian)^2 / (n.asian-1)) + ((sd.control^2 / n.control)^2 / (n.control - 1)))  # Degrees of freedom for Welch's t-test
-
-t <- (mean.asian - mean.control) / sample.se  # Welch t-test
-
-p <- 2 * pt(t, df)
-
-sdpooled <- sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.asian + n.control) - 2))
-d <- (mean.asian - mean.control) / sdpooled
-
-
-
-#  We can now run the TOST!
-#  Performing the TOST using the 'TOSTtwo.raw' function allows you to specify equivalence bounds in raw values
-Moon <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
-                            m2  = mean.control,
-                            sd1 = sd.asian,
-                            sd2 = sd.control,
-                            n1  = n.asian,
-                            n2  = n.control,
-                            low_eqbound  = -sesoi,
-                            high_eqbound =  sesoi,
-                            alpha = .05)

### 2017-11-17 Manuscript/Example3_Hyde-et-al.R

@@ -1,20 +1,0 @@
-
-# Hyde, Lindberg, Linn, Ellis, and Williams (2008) meta analysis
-
-sesoi <- 0.1
-alpha <- 0.005
-
-gr2  <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr3  <- TOSTmeta(ES = 0.04, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr4 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr5 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr6 <- TOSTmeta(ES = -0.01, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr7 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr8 <- TOSTmeta(ES = -0.02, se = 0.002, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr9 <- TOSTmeta(ES = -0.01, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr10 <- TOSTmeta(ES = 0.04, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-gr11 <- TOSTmeta(ES = 0.06, se = 0.003, low_eqbound_d=-sesoi, high_eqbound_d=sesoi, alpha=alpha, plot = FALSE)
-
-
-gr3.z <- (gr3$ES/0.002)  # Calculate Z value for grade 3
-gr3.p <- 2*pnorm(-abs(gr3.z))  # Calculate p value for grade 3

### 2017-11-17 Manuscript/Example5_Kahane-et-al-study4.R

@@ -1,26 +1,0 @@
-library(TOSTER)
-library(pwr)
-
-## Kahane, Everett, Earp, Farias, & Savulescu (2015)
-##
-## Study 4, 5.2.1. ii., p. 203: "There  was  no  relationship  between  perceived wrongness in 
-## personal dilemmas and in the ‘greater good’ dilemmas (r = -.04, p = .53): that is, people
-## who were more ‘utilitarian’ in the personal dilemmas were not more likely to be more truly
-## utilitarian in the other dilemmas, and vice versa. This lack of relationship held even when
-## controlling for primary psychopathy in a partial second order correlation (r = -.07, p = .32)."
-
-Kahane.data <- read.csv("Example5_Kahane-et-al-study4_data-cleaned.csv", header = TRUE, sep = ",")
-Kahane.data <- as.data.frame(Kahane.data)
-
-# sample size
-Kahane.N <- min(sum(!is.na(Kahane.data$ClassicSacrificialDilemmas)), sum(!is.na(Kahane.data$GreaterGoodDilemmas)))
-
-# correlation b/w perceived wrongness in personal dilemmas and 'greater good' dilemmas
-Kahane.r.GG <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$estimate
-Kahane.p <- cor.test(Kahane.data$ClassicSacrificialDilemmas, Kahane.data$GreaterGoodDilemmas, method = "pearson")$p.value
-
-# correlation the study had 80% power to detect
-Kahane.r.80 <- (pwr.r.test(n = Kahane.N, r = NULL, sig.level = 0.05, power = 0.8, alternative = "two.sided"))$r
-
-# TOST with bounds set to the correlation the study had 80% power to detect
-Kahane <- TOSTr(n = Kahane.N, r = Kahane.r.GG, low_eqbound_r = -Kahane.r.80, high_eqbound_r = Kahane.r.80, alpha = 0.05)

### 2017-11-17 Manuscript/H0H1_plot.R

@@ -1,67 +1,0 @@
-####  Setup ----------------------------------------
-library(ggplot2)
-library(gridExtra)
-
-plotheight <- 0.9
-lowerbound <- -0.2
-upperbound <- 0.2
-df <- data.frame()
-
-####  Base plot ------------------------------------
-baseplot <-   ggplot(df) +
-  scale_y_continuous(limits = c(0,plotheight+0.02), breaks=NULL) + # no y-axis will be displayed
-  theme_classic() +
-  theme(plot.title = element_text(size = rel(1), face = "bold"), #font size & appearance for plot titles
-        axis.title.y = element_blank(), #remove title of y-axis
-        axis.title.x = element_text(size=rel(0.7), lineheight = 0.5), #font size for x-axis label
-        plot.margin=unit(c(0.5,0.8,0.5,0.8),"cm")) #add padding around each plot to make them look nicer when combined; margin order: top, right, bottom, left
-
-#NHST plot
-NHSTplot <- baseplot +  
-  scale_x_continuous(limits = c(-0.85,0.85), breaks=c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8),
-                     labels=c("", "", "", "", "0", "", "", "", ""), name = "effect size") +
-  ggtitle("A: Classic NHST (two-sided)") +
-  annotate("segment", x = 0, xend = 0, y = plotheight-plotheight/2, yend = -Inf) + #vertical line at x=0 (H0)
-  annotate("text", size = rel(3.5), x=0, y = plotheight-plotheight/20, parse=TRUE, label="H[0]", hjust = 0.3) + #label for point null (H0)
-  annotate("segment", x = 0, xend = 0, y = plotheight-plotheight/6, yend=plotheight-plotheight/2.3, 
-           arrow = arrow(type = "closed", length=unit(2, "mm"))) + #arrow pointing from H0 label to H0 line
-  annotate("text", size = rel(3.5), x=-0.45, y=plotheight/3, parse=TRUE, label="H[1]") + #label for lower area (H1)
-  annotate("text", size = rel(3.5), x=0.45, y=plotheight/3, parse=TRUE, label="H[1]") #label for upper area (H1)
-
-#minimal effects test plot
-METplot <-   baseplot +  
[…]

### 2017-11-17 Manuscript/examples_plotgrid.R

@@ -1,123 +1,0 @@
-#  Plot example effects for AMPPS equivalence paper
-
-####  Setup ----------------------------------------
-
-library(purrr)
-library(gridExtra)
-library(ggplot2)
-library(lattice)
-
-library("TOSTER")
-
-plot.scaler <- 1.8
-nhst.size <- 0.5
-tost.size <- 1
-point.size <- 1.5
-
-####  Base plot ----------------------------------------
-
-baseplot <- ggplot(data.frame()) +
-  scale_x_continuous(breaks=NULL) +
-  coord_flip() + # flip coordinates (puts labels on y axis)
-  theme_classic(base_size = 10) + # use a white background
-  geom_hline(yintercept=0, lty=1) +
-  theme(plot.title = element_text(size = rel(1), face = "bold"),
-        axis.title.y = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.title.x = element_text(size=rel(0.7), lineheight = 0.5))
-
-####  Meta analysis ----------------------------------------
-
-source("Example3_Hyde-et-al.R")
-
-grades <- list(gr2, gr3, gr4, gr5, gr6, gr7, gr8, gr9, gr10, gr11)
-
-es <- map_dbl(grades, 1)
-l_eqb <- map_dbl(grades, 7)
-u_eqb <- map_dbl(grades, 8)
-li90 <- map_dbl(grades, 9)
-ui90 <- map_dbl(grades, 10)
-li95 <- map_dbl(grades, 11)
-ui95 <- map_dbl(grades, 12)
-label <- as.factor(paste0("grade ", 2:11))
-
-df <- data.frame(es, li90, ui90, label)
-
-sesoi <- 0.1
-
-
-metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
-  scale_x_discrete(breaks=NULL) +
-  geom_pointrange(size = nhst.size, fatten = point.size/2) + 
-  geom_pointrange(aes(ymin=li90, ymax=ui90), size = tost.size, fatten = point.size/2) +
-  geom_hline(yintercept=-sesoi, lty=2) +
-  geom_hline(yintercept=sesoi, lty=2) +
-  geom_hline(yintercept=0, lty=1) +
-  coord_flip() +  # flip coordinates (puts labels on y axis)
[…]

### 2017-11-17 TOST_simulator.R

@@ -1,201 +1,0 @@
-
-####  Setup ----------------------------------------
-
-library(TOSTER)
-library(purrr)
-download.file(url = "https://raw.githubusercontent.com/pederisager/TOSTER/TOST-modification/R/TOSTtwo.R", destfile = "TOSTtwo.prop.R")
-source(file = "TOSTtwo.prop.R") #  Source modified TOSTtwo from pederisager Fork in order to get necessary values from TOST
-library(ggplot2)
-
-sims <- list()
-
-
-# Group 1 parameters
-mean1 <- 0    # The mean effect
-std1  <- 10   # The standard deviation
-# make n random... n1    <- 10   # The number of subjects in each group
-
-# Group 2 parameters
-mean2 <- 0    # The mean effect
-std2  <- 10   # The standard deviation
-# make n random... n2    <- 10   # The number of subjects in each group
-
-# TOST
-alpha <- 0.05
-tails <- 2 # 1 = one-tailed, 2 = two-tailed
-
-
-# Set simulation parameters
-
-nsims <- 10000  # Number of simulations
-bound.criterion <- "specific.raw"  # Choose between "crit.val", "specific", "d33", "current.power", "tost.current.power"
-
-#  When using citerion "crit.val", the critical t value based on the parameters of the current sample are used as bounds
-crit.t <- qt(p = 1-(alpha/tails), df = n1+n2-2) # Calculate critical t
-
-#  When using criterion "specific", bounds = specific
-specific <- 0.9
-
-#  When using criterion "specific.raw", bounds = specific.raw, recalculated to d for each test
-specific.raw <- 3
-
-#  When using criterion "d33", small telescops is applied, group n of previous study is random 10-100, and bounds are set to effect that previous study had tlscop.pwr power to detect. 
-tlscop.pwr <- 0.33
-
-#  When using criterion "current.power", effect that current study had  pwr.thrsh power to detect are used as bounds
-pwr.thrsh <- 0.80
-crnt.pwr  <- pwr::pwr.t2n.test(n1 = n1,
-                               n2 = n2, 
-                               power = pwr.thrsh, 
-                               sig.level = alpha, 
-                               alternative = "two.sided")$d # Calculate effect size that original study have 33% power to detect
-
-# When using criterion "tost.current.power", effect is set to bounds that current study has 80% power for
-tost.pwr <- 0.80
-crnt.tost.pwr <- powerTOSTtwo(alpha = alpha, statistical_power = tost.pwr, N = n1)[2]
-
-
-
[…]

### 2017-11-17 eqboundCritera.R

@@ -1,36 +1,0 @@
-#  Function which calculates the values used to determine subjective equivalence bounds for TOST tests. 
-  # Calculates the effect size (Cohens d) that a study has % power (33% by default) to detect.
-  # Calculates the the smallest detectable effect size of a study based on the degrees of freedom and p value (0.05 by default).
-
-eqboundCritera <- function(n1, n2, orig.df, p = 0.05, power = 0.33, testtype = "two.sided", var.equal = FALSE) {
-  output <- list()  # Define output as list
-  
-  # Calculate d that study has 33% chance 
-  output$d33    <- pwr::pwr.t2n.test(n1          = n1, 
-                                     n2          = n2, 
-                                     power       = power, 
-                                     sig.level   = p, 
-                                     alternative = testtype)$d # Calculate effect size that study have 33% power to detect
-  
-  output$dunb33 <- ( 1 - ( 3 / (4*orig.df - 1) ) ) * output$d33     # Calculate d unbiased for the 33% power criterion
-  
-  
-  ## Original study smallest detectable effect size ##
-  
-  output$crit.t  <- qt(p = 1-p/2, df = orig.df) # Critical t for original study
-  output$crit.d  <- output$crit.t * sqrt(1/(n1-1) + 1/(n2-1)) # Convert critical t to d
-  return(output)
-}
-
-# Test
-
-source(file = "tFromDescTwo.R")
-Shih.m1  <-  0.54
-Shih.m2  <-  0.49
-Shih.sd1 <-  0.17
-Shih.sd2 <-  0.20
-Shih.n1  <- 16
-Shih.n2  <- 16
-Shih <- tFromDescTwo(Shih.m1, Shih.m2, Shih.sd1, Shih.sd2, Shih.n1, Shih.n2, test = "student")
-
-eqboundCritera(Shih.n1, Shih.n2, Shih.df)

### 2017-11-17 examples_plotgrid.R

@@ -1,0 +1,123 @@
+#  Plot example effects for AMPPS equivalence paper
+
+####  Setup ----------------------------------------
+
+library(purrr)
+library(gridExtra)
+library(ggplot2)
+library(lattice)
+
+library("TOSTER")
+
+plot.scaler <- 1.8
+nhst.size <- 0.5
+tost.size <- 1
+point.size <- 1.5
+
+####  Base plot ----------------------------------------
+
+baseplot <- ggplot(data.frame()) +
+  scale_x_continuous(breaks=NULL) +
+  coord_flip() + # flip coordinates (puts labels on y axis)
+  theme_classic(base_size = 10) + # use a white background
+  geom_hline(yintercept=0, lty=1) +
+  theme(plot.title = element_text(size = rel(1), face = "bold"),
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.x = element_text(size=rel(0.7), lineheight = 0.5))
+
+####  Meta analysis ----------------------------------------
+
+source("Example3_Hyde-et-al.R")
+
+grades <- list(gr2, gr3, gr4, gr5, gr6, gr7, gr8, gr9, gr10, gr11)
+
+es <- map_dbl(grades, 1)
+l_eqb <- map_dbl(grades, 7)
+u_eqb <- map_dbl(grades, 8)
+li90 <- map_dbl(grades, 9)
+ui90 <- map_dbl(grades, 10)
+li95 <- map_dbl(grades, 11)
+ui95 <- map_dbl(grades, 12)
+label <- as.factor(paste0("grade ", 2:11))
+
+df <- data.frame(es, li90, ui90, label)
+
+sesoi <- 0.1
+
+
+metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
+  scale_x_discrete(breaks=NULL) +
+  geom_pointrange(size = nhst.size, fatten = point.size/2) + 
+  geom_pointrange(aes(ymin=li90, ymax=ui90), size = tost.size, fatten = point.size/2) +
+  geom_hline(yintercept=-sesoi, lty=2) +
+  geom_hline(yintercept=sesoi, lty=2) +
+  geom_hline(yintercept=0, lty=1) +
+  coord_flip() +  # flip coordinates (puts labels on y axis)
[…]

### 2017-11-17 zDeskProp.R

@@ -1,24 +1,0 @@
-#### Fisher's exact Z Test for proportions - from PASS Sample Size Software ch. 213 ####
-
-zDeskProp <- function(prop1, prop2, n1, n2) {
-  
-  diff <- prop1 - prop2  # Calculate difference
-  se <- sqrt( ( prop1 * ( 1 - prop1 ) ) / n1 + ( prop2 * ( 1 - prop2 ) ) / n2 )  # Calculate standard error based on Fisher z test
-  z = ( diff - 0 ) / se # Calculate z score of the difference
-  
-  output <- list()
-  output$diff <- diff
-  output$se <- se
-  output$z <- z
-  output$prop1 <- prop1
-  output$prop2 <- prop2
-  output$n1 <- n1
-  output$n2 <- n2
-  
-  return(output)
-}
-
-
-#### test ####
-
-zDeskProp(.512, .579, 422, 428)

### 2017-11-17 manuscript.Rmd
update manuscript (remove bib loading, add grant)

@@
 author_note: >
+  This work was funded by VIDI grant 452-17-013.
 We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions.
+Authorship contributions:
   DL: Article conceptualization, programming TOSTER package and spreadsheet, main contributor to the introduction section and, discussion section.
@@
 library("papaja")
-
-# papaja:::fetch_zotero_refs(
-#   x = "1786197"
-#   , bib_name = "equivalence.bib"
-#   , API_key = "F7JGNtCxZGT6q87Eeg2zNWiP"
-#   , lib_type = "group"
-# )
 ```

### 2017-11-17 Leftovers/tFromDescTwo.R
clean up

@@ -1,52 +1,0 @@
-# Function for calculating t value, cohens d, pooled standard error, etc. from means, standard deviations and group n's.
-
-tFromDescTwo <- function(m1, m2, sd1, sd2, n1, n2, test = "welch") {
-  
-  output        <- list()       # Define the output as list
-  output$m1     <- m1           # Store input arguments in output
-  output$m2     <- m2
-  output$sd1    <- sd1
-  output$sd2    <- sd2
-  output$n1     <- n1
-  output$n2     <- n2
-  output$m.diff <- m1 - m2      # Calculate the mean difference
-
-  
-  if (test == "student") {  
-    # Calculate se.diff, t and d based on a student t.test
-    print("student's t.test is performed and sample sizes are assumed to be equal")
-    output$test       <- "student"
-    output$df         <- n1 + n2 - 2  # Calculate degrees of freedom.
-    output$sd.pooled  <- sqrt( ( ( ( n1 - 1 ) * ( sd1^2 ) ) + ( n2 - 1 ) * ( sd2^2 ) ) / ( ( n1 + n2 ) - 2 ) ) # Calculate the pooled SD
-    output$se.diff    <- output$sd.pooled * sqrt( 1/n1 + 1/n2 ) # Calculate the SE of the difference.
-    
-  } else if (test == "welch") {  
-    # Calculate se.diff, t and d based on a welch t.test, and calculate sd.pooled
-    print("welch's t.test is performed and sample sizes are assumed to be unequal")
-    output$test       <- "welch"
-    output$df         <- ( sd1^2 / n1 + sd2^2 /n2 )^2 / ( ( ( sd1^2 / n1 )^2 / ( n1-1 ) ) + ( ( sd2^2 / n2 )^2 / ( n2 - 1 ) ) ) #degrees of freedom for Welch's t-test
-    output$sd.pooled  <- sqrt( ( sd1^2 + sd2^2 ) / 2 )    # Calculate sd root mean squared for Welch's t-test
-    output$se.diff    <- sqrt( sd1^2 / n1 + sd2^2 / n2 )  # Calculate the se of the difference based on sd.pooled
-    
-    
-  } else {
-    stop("The 'test' argument must be set to either 'student'or 'welch'", call. = FALSE)
-  }
-  
[…]

### 2017-11-17 Example1_MoonRoeder.R
Added library(TOSTER) to .R examples
* Minor comment edits
* Minor formatting edits

@@ -1,32 +1,30 @@
+# Load TOSTER into R environment 
+library(TOSTER)
+
 #  Load the relevant datafile into R
-moon.data        <- read.csv("Example1_MoonRoeder_data-cleaned.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)
+moon.data <- read.csv("Example1_MoonRoeder_data-cleaned.csv")  # Data can be found at https://osf.io/rwv5s/ (final_dataset_139_to_send.sav)
 
 #  Set the bounds, in this case 1 grade point, which corresponds to 6.25%
 sesoi <- 0.0625
 
 #  Calculate the summary statistics from the data that we need in the TOST calculation below
-mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
+mean.asian <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
 mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
-sd.asian         <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
-sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
-n.asian          <- sum(moon.data$identity_salience == 1)
-n.control        <- sum(moon.data$identity_salience == 2)
+sd.asian <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
+sd.control <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
+n.asian <- sum(moon.data$identity_salience == 1)
+n.control <- sum(moon.data$identity_salience == 2)
 sample.sd <- sqrt((sd.asian^2 + sd.control^2) / 2)  # Calculate sd root mean squared for Welch's t-test
 sample.se <- sqrt(sd.asian^2 / n.asian + sd.control^2 / n.control)  # Calculate the se of the difference based on sd.pooled
 
 df <- (sd.asian^2 / n.asian + sd.control^2 / n.control)^2 / (((sd.asian^2 / n.asian)^2 / (n.asian-1)) + ((sd.control^2 / n.control)^2 / (n.control - 1)))  # Degrees of freedom for Welch's t-test
-
 t <- (mean.asian - mean.control) / sample.se  # Welch t-test
-
 p <- 2 * pt(t, df)
-
 sdpooled <- sqrt((((n.asian - 1) * (sd.asian^2)) + (n.control - 1) * (sd.control^2)) / ((n.asian + n.control) - 2))
[…]

### 2017-11-17 Example4_Lynott-et-al.R

@@ -1,27 +1,23 @@
-se <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error
-crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05
+# Load TOSTER
+library(TOSTER)
 
+# Calculate the smallest effect sizer of interest (crit.diff), used for setting the bounds.
+se <- sqrt((0.75 * (1 - 0.75)) / 26 + (0.46 * (1 - 0.46)) / 27)  # Calculate the (unpooled) standard error.
+crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05.
+crit.diff <- crit.z * se  # Convert z to the critical difference.
 
-crit.diff <- crit.z * se  # Convert z to the critical difference
-
-
-
-library(TOSTER)  # Install TOSTER using devtools::install_github("Lakens/TOSTER"), or download TOSTtwo.prop.R from https://github.com/Lakens/TOSTER/tree/master/R and then source it. 
-
-
-#  We first load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
+#  Load the data and exclude cases to only include cases that did not meet the authors' exclusion criteria.
 lynott.data <- read.csv("Example4_Lynott-et-al_data-cleaned.csv")
 lynott.data <- lynott.data[lynott.data$include1 == 1,]
 
-
-#  We then calculate the summary statistics from the data that we need in the TOST calculation below
+#  Calculate the summary statistics from the data that we need in the TOST calculation below.
 n.hot  <- sum(lynott.data$packtype == 1)
 n.cold <- sum(lynott.data$packtype == 0)
 selfish.hot  <- sum(lynott.data$selfish[lynott.data$packtype == 1] == 1, na.rm = TRUE) / n.hot
 selfish.cold <- sum(lynott.data$selfish[lynott.data$packtype == 0] == 1, na.rm = TRUE) / n.cold
 diff <- selfish.hot - selfish.cold
 
-#  We can now run the TOST!
+#  Calculate equivalence test for two proportions.
 Lynott <- TOSTtwo.prop(prop1 = selfish.cold,
                      prop2 = selfish.hot,
                      n1    = n.cold,

### 2017-11-17 manuscript.Rmd
new version manuscript

@@
 ---
-title             : "Equivalence Testing for Psychological Science"
-shorttitle        : "Equivalence Testing"
+title             : "Equivalence Testing for Psychological Research: A Tutorial"
+shorttitle        : "Equivalence Testing Tutorial"
 
@@
   Psychologists must be able to test both for the presence of an effect and for the the absence of an effect.
-In addition to testing against zero, researchers can use the Two-One-Sided Tests (TOST) procedure to test for *equivalence* and reject the presence of a meaningful effect.
+In addition to testing against zero, researchers can use the Two-One-Sided Tests (TOST) procedure to test for *equivalence* and reject the presence of a smallest effect size of interest (SESOI).
 TOST can be used to determine if an observed effect is surprisingly small, given that a true effect at least as large as the SESOI exists.
 We explain a range of approaches to determine the SESOI in psychological science, and provide detailed examples of how equivalence tests should be performed and reported.
-Equivalence tests enable researchers to falsify predictions about the presence, and declare the absence, of meaningful effects.
+Equivalence tests are an important extension of statistical tools psychologist currently use, and enable researchers to falsify predictions about the presence, and declare the absence, of meaningful effects.
 
-keywords          : "Equivalence Testing, NHST, power, TOST"
+keywords          : "Equivalence Testing, NHST, power, TOST, falsification."
 wordcount         : "X"
@@
 Alternatively, the researcher could decide that even if the true ratio in the population is not exactly 0.015, the null hypothesis should consist of a range of values around the difference in proportions of 0.015 that can be considered trivially small.
-The researcher could for example test if the difference is smaller than -0.005, or larger than 0.035.
+The researcher could for example test if the difference is smaller than -0.005, and larger than 0.035.
 This test against two bounds, with $H_0$ being a range rather than one value (see Figure 1B), is known as a *minimal effects test* [@Murphy2014].
 
-*Equivalence tests* can be seen as the opposite of minimal effects tests: They examine whether the presence of effects that are large enough to matter can be *rejected* (see Figure 1C).
-In our example, the researcher can perform an equivalence test to test whether the gender difference in participants is *not* large enough to be considered meaningful: the *smallest effect size of interest* (SESOI).
+*Equivalence tests* can be seen as the opposite of minimal effects tests: They examine whether the presence of effects that are large enough to be considered meaningful can be *rejected* (see Figure 1C).
+In our example, the researcher can perform an equivalence test to examine whether the gender difference in participants is *not* as large or larger than the *smallest effect size of interest* (SESOI).
 After an extensive discussion with experts, the researcher decides that as long as the difference in proportions is not more extreme than 6%, the gender difference is too small to care about.
-Given an expected true difference in the population of 0.015, the researcher will test if the observed difference falls outside the boundary values of -0.055 and 0.075.
+Given an expected true difference in the population of 0.015, the researcher will test if the observed difference falls outside the boundary values (or *equivalence bounds*) of -0.055 and 0.075.
 If differences more extreme than these boundary values can be rejected in two one-sided tests (TOST), the researcher will conclude statistical equivalence, the gender difference will be considered trivially small, and no money will be spent on addressing a gender difference in participation.
@@
 In any one-sided test, for an alpha level of 5%, we can reject $H_0$ when the 90% confidence interval around the observed estimate is in the predicted direction, and does not contain the value tested against (e.g., 0).
-In TOST, the first one-sided test is used to test against values smaller than the lower bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper bound ($\Delta_{U}$).
+In TOST, the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
 It is not necessary to control for multiple tests when using the TOST procedure because both tests need to be statistically significant to conclude equivalence.
@@
 
-Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze (but we recommend to set and justify the Type 1 error rate in original research based on the goals researchers have [@Lakens2017b].
+Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze (but we recommend to justify the Type 1 error rate in original research based on substantive arguments [@Lakens2017b].
 It might be easiest to think of an equivalence test as checking whether the confidence interval excludes the SESOI, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
@@
 
-In this article we provide several reproducible examples of equivalence tests which illustrate the procedure in easy-to-use free software (R, jamovi, and a spreadsheet).
+In this article we provide five reproducible examples of equivalence tests which illustrate the procedure in easy-to-use free software (R, jamovi, and a spreadsheet).
 But first, we discuss different approaches to determining the SESOI for psychological research, which determines the statistical question an equivalence test answers.
@@
 A SESOI should be chosen such that inferences based on it answer a meaningful question.
-Although we use bounds that are symmetric around 0 in all examples in this manuscript for equivalence tests (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is also possible to use asymmetric bounds, for example based on an explicit cost-benefit analysis (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$).
+Although we use bounds that are symmetric around 0 in all examples in this manuscript for equivalence tests (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is also possible to use asymmetric bounds (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$).
 
@@
 
-Second, researchers can determine the SESOI based on previous work in the literature.
-Ideally, researchers would always specify their SESOI in studies they report, but this is not yet common practice.
-It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study.
-@Simonsohn2015 recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have had 33% power to detect.
+Second, the SESOI can be based on published studies in the literature.
+Ideally, researchers who publish novel research would always specify their smallest effect size of interest, but this is not yet common practice.
+It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful when examining the same hypothesis.
+@Simonsohn2015 recently proposed to set the SESOI to the effect size that an earlier study would have had 33% power to detect.
 With this so-called 'small telescopes' approach, the equivalence bounds are thus determined primarily based on the sample size in the original study.
 For example, consider a study in which `r N` participants answered a question, and the results were analyzed with a one-sample t-test.
-For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of $d = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d)`$.
+For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of $d = `r printnum(pwr.t.test(n = N, sig.level = alpha, power = 0.33,type = "one.sample", alternative = "two.sided")$d)`$.
 Another example of how previous research can be used to determine the SESOI can be found in @Kordsmeyer2017, who defined the SESOI based on the mean of effect sizes reported in the literature.
@@
 
-Another justifiable choice when basing the SESOI on previous work is to use the smallest observed effect size that could have been statistically significant in a previous study.
+Another justifiable choice when choosing the SESOI based on earlier work is to use the smallest observed effect size that could have been statistically significant in a previous study.
 In other words, we decide that effects that could not have yielded a $p < \alpha$ in an original study will not be considered meaningful in the replication study either, even if they were found to be statistically significant.
@@
 This critical test value can be transformed to a standardized effect size (e.g., $d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }$), which can thus be interpreted as the *critical effect size*.^[This will typically, although not always, correspond to the effect size the study had 50% power to detect [see @Lenth2007].
-This procedure will result in equivalence bounds that are substantially wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.] All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
+This procedure will result in equivalence bounds that are wider than the ones obtained using the small telescopes approach, which gives the effect size a study had 33% power to detect.] All observed effect sizes smaller than the critical effect size would not have been statistically significant in the original study, given the alpha and sample size of that study.
 By setting the SESOI to the critical effect size, an equivalence test can reject all observed effect sizes that could have been detected in an earlier study.
@@
 An equivalence test based on 15 observations rejecting a SESOI of $d = 1$ cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
-By transparently reporting the effects that can be detected and rejected based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what the size of a reasonable SESOI may be.
+By transparently reporting the effects that can be detected and rejected based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI may be.
 
@@
 0.5 points on a 7-point scale).
-The key difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as $\frac {X _ { 1} - X _ { 2} } {S D}$.
+The key difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as $\frac {X _ { 1} - X _ { 2} } {S D}$).
 The observed standard deviation randomly varies across samples.
@@
 A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($M_{unethical}= `r rep.m1`$, ${SD}_{unethical}= `r rep.sd1`$, $M_{ethical}=`r rep.m2`$, ${SD}_{ethical}=`r rep.sd2`$, $t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
-Using the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r printnum(d.33)`$ --- and use this as the SESOI.
+Using the small telescopes approach, we can calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$) and use this as the SESOI.
 A TOST procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
@@
 
-@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as effect sizes smaller than $d = `r printnum(sesoi)`$.
+@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as absolute effect sizes smaller than $d = `r printnum(sesoi)`$.
 For example, in grade 3 the difference is $d = 0.04$, with a standard error of 0.002.
-When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of `r printp(alpha)` to correct for multiple comparisons), using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that indeed, all effect size estimates are measured with such high precision that they are statistically equivalent and can be considered trivially small, according to the authors’ definition (Figure 2C).
+When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of `r printp(alpha)` to correct for multiple comparisons) and using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that all effect size estimates are measured with such high precision that they are statistically equivalent, and can be considered trivially small according to the authors’ definition (Figure 2C).
 In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printnum(max(gr3$TOST_p1, gr3$TOST_p2))`$.
 However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large.
-For grade 3, $z = `r printnum(gr3.z)`$, $p = `r printnum(gr3.p)`$.
-This is one way in which the use of equivalence tests improves hypothesis testing procedures in science, by allowing researchers to distinguish between statistical significance and practical significance.
+(e.g., for grade 3, $z = `r printnum(gr3.z)`$, *p* < 0.001.
+We see how equivalence tests allow researchers to distinguish between statistical significance and practical significance, which demonstrates how equivalence tests improve hypothesis testing procedures in psychological research.
 
@@
 This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
-To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error.
+To calculate which difference corresponds to a critical z value in the original study, we multiply critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diffe)`$ or more.
 
@@
 This means we can test for inferiority (see Figure 1D) and conclude the absence of a meaningful effect if the observed effect size is reliably *lower* than the SESOI.
-We find that we can reject effects larger than  $\Delta_{U}$, $z = `r printnum(Lynott$TOST_z2)`$, $p `r printp(Lynott$TOST_p2)`$ (see Figure 2D).
+We find that we can reject effects larger than  $\Delta_{U} = `r printnum(crit.diff)`$, $z = `r printnum(Lynott$TOST_z2)`$, $p `r printp(Lynott$TOST_p2)`$ (see Figure 2D).
 Thus, we can conclude that the statistically non-significant effect is also statistically smaller than our SESOI.
@@
 contest this interpretation in a number of studies.
-In Study 4, they compare the traditionally used dilemmas with a new set of new ones in which the sacrifice for the greater good is not another person’s life, but something the participant would have a partial interest in(e.g.
+In Study 4, they compare the traditionally used dilemmas with a set of new delimmas in which the sacrifice for the greater good is not another person’s life, but something the participant would have a partial interest in (e.g.
 buying a new mobile phone vs.
-donating the money to save lives in a distant country) and find no significant correlation between the two ($r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$), $N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]).
+donating the money to save lives in a distant country).
+The authors find no significant correlation between the two set of dilemmas, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$, $N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.].
 The authors conclude that “a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely” (p.
@@
 With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$.
-Performing the TOST procedure given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$.
+Performing the TOST procedure given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, the data is statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$.
 This means that values smaller than $r = `r printnum(-Kahane.r.80)`$ and larger than $r = `r printnum(Kahane.r.80)`$ can be rejected at an alpha level of 5%.
-If other researchers are interested in examining possible smaller effects, they can design studies with a larger sample size that can detect, or reject, smaller effect sizes.
+If other researchers are interested in examining the presence of smaller effect size, they can design studies with a larger sample size.
 
@@
 
-Equivalence testing is a relatively simple and flexible technique to falsify hypotheses.
-As long as we can calculate a confidence interval around a parameter estimate, we can compare it to a smallest effect size of interest.
+Equivalence testing is a simple statistical technique to reject the presence of a smallest effect size of interest.
+As long as we can calculate a confidence interval around a parameter estimate, we can compare it to the SESOI.
 The result of an equivalence test can be obtained by mere visual inspection of the CI [@Seaman1998; @Tryon2001], or by performing two one-sided tests (the TOST procedure).
@@
 As with any statistical test, the usefulness of the result of an equivalence test depends on the question we ask.
-The question manifests itself in the bounds we set: Is our data surprising if the true effect lies outside of the bounds?
+The question manifests itself in the bounds we set: Is our data surprisingly close to zero, assuming a true effect exists that is as large as our smallest effect size of interest?
 If we test against very wide bounds, observing statistical equivalence can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003].
-Examining the papers citing @Lakens2017a, we found that some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s $d$ of “medium” size was chosen.] An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2003], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
+Examining  papers citing @Lakens2017a, we found that some researchers state, but do not justify, the SESOI used in the equivalence test [e.g., @Brown2017; @Schumann2017].
+An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2003], but researchers should always explain why they chose a SESOI.
+It makes little sense to report a statistical test, without explaining why one would want to answer this question to begin with.
 Equivalence bounds should be specified before results are known, ideally as part of a pre-registration [cf.
@@
 We recommend that researchers by default perform both a null hypothesis significance test and an equivalence test on their data.
-The biggest challenge for researchers will be specifying the smallest effect size of interest.
+The biggest challenge for researchers will be to specify the smallest effect size of interest.
 Psychological theories are usually too vague to derive precise predictions, and if there are no theoretical reference points, natural constraints, or prior studies a researcher can use to define the SESOI for a hypothesis test, any choice will be somewhat arbitrary.
@@
 Nevertheless, it is important to realize that not specifying a SESOI for our research questions at all will severely hinder theoretical progress [@Morey2016].
-While our initial choices of a smallest effect size of interest may be arbitrary, incorporating equivalence tests in our statistical toolbox will in time contribute to better --- and more falsifiable --- theories.
+Incorporating equivalence tests in our statistical toolbox will in time contribute to better --- and more falsifiable --- theories.
 

### 2017-11-17 manuscript.Rmd
minor fix

@@
 It is not necessary to control for multiple tests when using the TOST procedure because both tests need to be statistically significant to conclude equivalence.
-This means that when reporting a TOST result, authors can suffice by reporting the one-sided test with the smaller test parameter (e.g.
+This means that when reporting a TOST result, it suffices to reporting the one-sided test with the smaller test parameter (e.g.
 $t$ or $r$) and thus the larger $p$-value.
@@
 
-Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze (but we recommend to justify the Type 1 error rate in original research based on substantive arguments [@Lakens2017b].
+Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze [but we recommend to justify the Type 1 error rate in original research based on substantive arguments, @Lakens2017b].
 It might be easiest to think of an equivalence test as checking whether the confidence interval excludes the SESOI, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
@@
 This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
-To calculate which difference corresponds to a critical z value in the original study, we multiply critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diffe)`$ or more.
+To calculate which difference corresponds to a critical z value in the original study, we multiply critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
 

### 2017-11-17 examples_plotgrid.R
Reduced the size of metaplot points to make the CIs more visible.

@@ -49,8 +49,8 @@ sesoi <- 0.1
 
 metaplot <- ggplot(data=df, aes(x=label, y=es, ymin=li95, ymax=ui95)) +
   scale_x_discrete(breaks=NULL) +
-  geom_pointrange(size = nhst.size, fatten = point.size/2) + 
-  geom_pointrange(aes(ymin=li90, ymax=ui90), size = tost.size, fatten = point.size/2) +
+  geom_pointrange(size = nhst.size/2, fatten = point.size/5) + 
+  geom_pointrange(aes(ymin=li90, ymax=ui90), size = tost.size/2, fatten = point.size/5) +
   geom_hline(yintercept=-sesoi, lty=2) +
   geom_hline(yintercept=sesoi, lty=2) +
   geom_hline(yintercept=0, lty=1) +

### 2017-11-17 manuscript.Rmd
figures now in text, hooray!

@@
 
-figsintext        : no
+figsintext        : yes
 figurelist        : no
@@
 ---
-getwd()
+
 ```{r include = FALSE}
@@
 
-```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
+```{r echo=FALSE, fig.width=3, fig.height=6, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
 **A)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
@@
 source("H0H1_plot.R")
-getwd()
 ```
@@
 ```
-```{r echo=FALSE, fig.cap=paste("Example effects plotted with 90% TOST CIs (thick lines) and 95% NHST CIs (thin lines), the NHST null hypothesis (solid vertical line) and the equivalence bounds (dashed vertical lines) displayed.
+```{r echo=FALSE, fig.width=7, fig.height=4, fig.cap=paste("Example effects plotted with 90% TOST CIs (thick lines) and 95% NHST CIs (thin lines), the NHST null hypothesis (solid vertical line) and the equivalence bounds (dashed vertical lines) displayed.
 **A)** Example 1 - Mean Difference.

### 2017-11-17 manuscript.Rmd
Edit to footnote 4

--- 2017-11-17 manuscript.Rmd
@@
 The TOST is non-significant, $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (Figure 2A).
-Thus, we cannot reject a true effect as large or larger than 6.25%.^[In this example, the bounds of the TOST were based on a raw effect of 6.25%.
-We could have converted this raw effect into a standardized effect (Cohen's $d$) for the current sample, and we would have gotten the exact same result ($\Delta = \pm `r printnum(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2))`)$.
-Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's $d$ that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication.
-Now we would no longer be answering a question about a raw effect independent of the samples, but rather if the replication can reject standardized effects larger than a certain standardized effect in the original.
-Since the standard deviations of the original study were larger than in the replication in this case, Cohen's $d$ (and thus the bounds) would be narrower ($\Delta = \pm `r printnum(sesoi / sqrt((sd.asian^2 + sd.control^2) / 2))`)$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+Thus, we cannot reject a true effect as large or larger than 6.25%.^[Because this is a replication of a study, it would have been reasonable to assume that we only want reject effects in the same direction as the effect in the original study.
+After all, much smaller effects may indicate non-replication, but so does large effects in the opposite direction.
+In that case, we would perform an inferiority test (see figure 1D) against $\Delta_{U}$.
+As can be seen in figure 2A, the $90\% CIs$ does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
 Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.

### 2017-11-17 manuscript.Rmd
Changed caption of Fig. 1 to say "Tests if ... can be rejected"

--- 2017-11-17 manuscript.Rmd
@@
 ```{r echo=FALSE, fig.width=3, fig.height=6, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
-**A)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0.
-**B)**  Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$.
-**C)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$.
-**D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}
+**A)** NHST: Tests if the hypothesis ($H_0$) that an effect is equal to 0 can be rejected.
+**B)**  Minimal effects test: Tests if the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$ can be rejected.
+**C)** Equivalence test: Tests if the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$ can be rejected.
+**D)** Inferiority test: Tests if the hypothesis ($H_0$) that an effect is larger than $\\Delta$ can be rejected."}
 

### 2017-11-17 manuscript.Rmd
Changed the author notes to thank Dermot, Katie, Jim, and Brian, and changed Peder's and Anne's initials in the author contributions

@@
     affiliation   : "1"
-  - name          : "Peder Isager"
+  - name          : "Peder M. Isager"
     affiliation   : "1"
@@
 
+note: "\\clearpage"
+
 author_note: >
   This work was funded by VIDI grant 452-17-013.
-We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions.
-Authorship contributions:
-  DL: Article conceptualization, programming TOSTER package and spreadsheet, main contributor to the introduction section and, discussion section.
-  AS and PI: Main contributors to examples and figures.
+We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions, Dermot Lynott and Katie Corker for their helpful responses to our inquiries about @Lynott2014, and Jim Everett and Brian Earp for their helpful responses and for providing the raw data for study 4 in Kahane et al.
+[-@Kahane2015].
+  \newline
+  Author contributions: 
+  DL: Article conceptualization, programming TOSTER package and spreadsheet, main contributor to the introduction section and discussion section.
+  AMS and PMI: Main contributors to examples and figures.
   All authors contributed to manuscript conceptualization, editing, review for submission, and additional software development.
-AS and PI contributed equally to the manuscript, and order was determined by executing the following commands in R:
+AMS and PMI contributed equally to the manuscript, and order was determined by executing the following commands in R:
   \newline

### 2017-11-17 manuscript.Rmd
Minor edit

--- 2017-11-17 manuscript.Rmd
@@
 In that case, we would perform an inferiority test (see figure 1D) against $\Delta_{U}$.
-As can be seen in figure 2A, the $90\% CIs$ does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+As can be seen in figure 2A, the 90\% CIs does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
 Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.

### 2017-11-17 H0H1_plot.R
tiny change of label height in the plots

@@ -36,9 +36,9 @@ METplot <-   baseplot +
   annotate("segment", x = lowerbound, xend = lowerbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for lower bound
   annotate("segment", x = upperbound, xend = upperbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for upper bound
   annotate("rect", xmin = -0.2, xmax = 0.2, ymin = -Inf, ymax = 0.9, fill = "red", alpha = .2, color = NA) + #shading for H0 area
-  annotate("text", size = rel(3.5), x=-0.6, y=plotheight/2, parse=TRUE, label="H[1]") + #label for lower area (H1)
-  annotate("text", size = rel(3.5), x=0.6, y=plotheight/2, parse=TRUE, label="H[1]") + #label for upper area (H1)
-  annotate("text", size = rel(3.5), x=-0, y=plotheight/2, parse=TRUE, label="H[0]", hjust = 0.3) #label for minimal effects area (H1)
+  annotate("text", size = rel(3.5), x=-0.6, y=plotheight/3, parse=TRUE, label="H[1]") + #label for lower area (H1)
+  annotate("text", size = rel(3.5), x=0.6, y=plotheight/3, parse=TRUE, label="H[1]") + #label for upper area (H1)
+  annotate("text", size = rel(3.5), x=-0, y=plotheight/3, parse=TRUE, label="H[0]", hjust = 0.3) #label for minimal effects area (H1)
 
 #equivalence test plot
 eqplot <- baseplot +  
@@ -49,9 +49,9 @@ eqplot <- baseplot +
   annotate("segment", x = upperbound, xend = upperbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for upper bound
   annotate("rect", xmin = -Inf, xmax = lowerbound, ymin = -Inf, ymax = plotheight, fill = "red", alpha = .2, color = NA) + #shading for lower area
   annotate("rect", xmin = upperbound, xmax = Inf, ymin = -Inf, ymax = plotheight, fill = "red", alpha = .2, color = NA) + #shading for upper area
-  annotate("text", size = rel(3.5), x=-0.6, y=plotheight/2, parse=TRUE, label="H[0]") + #label for lower area (H0)
[…]

### 2017-11-17 manuscript.Rmd
tiny formatting change Fig 1

@@
 
-```{r echo=FALSE, fig.width=3, fig.height=6, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
+```{r echo=FALSE, fig.width=4, fig.height=6.2, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
 **A)** NHST: Tests if the hypothesis ($H_0$) that an effect is equal to 0 can be rejected.

### 2017-11-17 H0H1_plot.R

@@ -36,9 +36,9 @@ METplot <-   baseplot +
   annotate("segment", x = lowerbound, xend = lowerbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for lower bound
   annotate("segment", x = upperbound, xend = upperbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for upper bound
   annotate("rect", xmin = -0.2, xmax = 0.2, ymin = -Inf, ymax = 0.9, fill = "red", alpha = .2, color = NA) + #shading for H0 area
-  annotate("text", size = rel(3.5), x=-0.6, y=plotheight/3, parse=TRUE, label="H[1]") + #label for lower area (H1)
-  annotate("text", size = rel(3.5), x=0.6, y=plotheight/3, parse=TRUE, label="H[1]") + #label for upper area (H1)
-  annotate("text", size = rel(3.5), x=-0, y=plotheight/3, parse=TRUE, label="H[0]", hjust = 0.3) #label for minimal effects area (H1)
+  annotate("text", size = rel(3.5), x=-0.6, y=plotheight/2.5, parse=TRUE, label="H[1]") + #label for lower area (H1)
+  annotate("text", size = rel(3.5), x=0.6, y=plotheight/2.5, parse=TRUE, label="H[1]") + #label for upper area (H1)
+  annotate("text", size = rel(3.5), x=-0, y=plotheight/2.5, parse=TRUE, label="H[0]", hjust = 0.3) #label for minimal effects area (H1)
 
 #equivalence test plot
 eqplot <- baseplot +  
@@ -49,9 +49,9 @@ eqplot <- baseplot +
   annotate("segment", x = upperbound, xend = upperbound, y = plotheight, yend = -Inf, linetype = "dashed") + #dashed line for upper bound
   annotate("rect", xmin = -Inf, xmax = lowerbound, ymin = -Inf, ymax = plotheight, fill = "red", alpha = .2, color = NA) + #shading for lower area
   annotate("rect", xmin = upperbound, xmax = Inf, ymin = -Inf, ymax = plotheight, fill = "red", alpha = .2, color = NA) + #shading for upper area
-  annotate("text", size = rel(3.5), x=-0.6, y=plotheight/3, parse=TRUE, label="H[0]") + #label for lower area (H0)
[…]

### 2017-11-17 manuscript.Rmd
fixed references and software citations

@@
 
-bibliography      : ["equivalence.bib"]
+bibliography      : ["equivalence.bib","r-references.bib"]
 
@@
 library("papaja")
+library("bookdown")
+library("rmarkdown")
+library("knitr")
+library("TOSTER")
+library("purrr")
 ```
@@
 
+```{r include=FALSE}
+r_refs(file = "r-references.bib")
+my_citation <- cite_r(file = "r-references.bib")
+```
+
 #Raw versus standardized equivalence bounds
@@
 The code to reproduce these analyses is available at <https://osf.io/qamc6/>^[The OSF project contains all necessary files to reproduce examples 1--5 in R, jamovi [@jamovi2017], and in a spreadsheet; as well as a preprint of this manuscript.
-The manuscript, including the figures and statistical analyses, was created using R version 3.4.2 [@RCoreTeam2017] and RStudio version 1.1.383 [@RStudioTeam2016], with the packages `TOSTER` version 0.3 [@Lakens2017c], `pwr` version 1.2--1 [@Champely2017], `rmarkdown` version 1.7[@Allaire2017], `knitr` version 1.17 [@Xie2017], `papaja` version 0.1.0.9492 [@Aust2017], `ggplot2` version 2.2.1 [@Wickham2009], `gridExtra` version 2.3 [@Auguie2017], `lattice` version 0.20--35 [@Sarkar2008], and `purrr` version 0.2.4 [@Henry2017].].
+The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`].
 

### 2017-11-17 manuscript.Rmd
(hopefully) final reference fixes

@@
 library("purrr")
+library("lattice")
 ```
@@
 The code to reproduce these analyses is available at <https://osf.io/qamc6/>^[The OSF project contains all necessary files to reproduce examples 1--5 in R, jamovi [@jamovi2017], and in a spreadsheet; as well as a preprint of this manuscript.
-The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`].
+The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`.].
 

### 2017-11-17 H0H1_plot.R
fig 1 and 2 high res save

@@ -63,5 +63,6 @@ infplot <-   baseplot +
   annotate("text", size = rel(3.5), x=0.6, y=plotheight/2.5, parse=TRUE, label="H[0]") + #label for upper area (H0)
   annotate("text", size = rel(3.5), x=-0.35, y=plotheight/2.5, parse=TRUE, label="H[1]") #label for lower area (H1)
 
+#tiff(file="NHSTplot.tiff",width=1400,height=2000, units = "px", res = 300)
 grid.arrange(NHSTplot, METplot, eqplot, infplot, ncol = 1) #combine plots in one column (all stacked)
-
+#dev.off()

### 2017-11-17 examples_plotgrid.R

@@ -119,5 +119,6 @@ kahaneplot = baseplot +
 lay <- rbind(c(1,2,5),
              c(3,4,5)) # Define the figure layout. Think of the c()s as a grid/ Plot 1-5 will be alloted the relative space in the grid.
 
-
+#tiff(file="example_grid.tiff",width=2000,height=1400, units = "px", res = 300)
 example.grid <- grid.arrange(moonplot, brandtplot, lynottplot, kahaneplot, metaplot, layout_matrix = lay)
+#dev.off()

### 2017-11-17 manuscript.Rmd
tiny fix to minimal effects test paragraph
Added wordcount (4969)

@@
 keywords          : "Equivalence Testing, NHST, power, TOST, falsification."
-wordcount         : "X"
+wordcount         : "4969"
 
@@
 Alternatively, the researcher could decide that even if the true ratio in the population is not exactly 0.015, the null hypothesis should consist of a range of values around the difference in proportions of 0.015 that can be considered trivially small.
-The researcher could for example test if the difference is smaller than -0.005, and larger than 0.035.
+The researcher could for example test if the difference is smaller than -0.005, or larger than 0.035.
 This test against two bounds, with $H_0$ being a range rather than one value (see Figure 1B), is known as a *minimal effects test* [@Murphy2014].

### 2017-11-17 manuscript.Rmd
Minor edits in abstract

@@
 abstract: >
-  Psychologists must be able to test both for the presence of an effect and for the the absence of an effect.
-In addition to testing against zero, researchers can use the Two-One-Sided Tests (TOST) procedure to test for *equivalence* and reject the presence of a smallest effect size of interest (SESOI).
+  Psychologists must be able to test both for the presence of an effect and for the absence of an effect.
+In addition to testing against zero, researchers can use the Two One-Sided Tests (TOST) procedure to test for *equivalence* and reject the presence of a smallest effect size of interest (SESOI).
 TOST can be used to determine if an observed effect is surprisingly small, given that a true effect at least as large as the SESOI exists.

### 2017-11-17 manuscript.Rmd
Minor edits in intro

@@
 To understand the idea behind equivalence tests, it is useful to realize the null hypothesis we test against can be any numerical value.
-When we compare two groups, we often test whether the difference between these groups is zero (see Figure 1A), but we may sometimes want to test against other values than zero.
+When we compare two groups, we often test whether we can reject that the difference between these groups is zero (see Figure 1A), but we may sometimes want to reject other values than zero.
 Imagine a researcher who is interested in voluntary participation in a national program to train young infants' motor skills.

### 2017-11-17 manuscript.Rmd
minor edit

@@
 It is not necessary to control for multiple tests when using the TOST procedure because both tests need to be statistically significant to conclude equivalence.
-This means that when reporting a TOST result, it suffices to reporting the one-sided test with the smaller test parameter (e.g.
+This means that when reporting a TOST result, it suffices to report the one-sided test with the smaller test parameter (e.g.
 $t$ or $r$) and thus the larger $p$-value.
@@
 @Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
-In contrast to the original study, they found a negative difference between the Asian primed group ($n = `r n.asian`$, $M = `r printnum(mean.asian)`$, $SD = `r printnum(sd.asian)`$) and the control ($n = `r n.control`$, $M = `r printnum(mean.control)`$, $SD = `r printnum(sd.control)`$) which was not significant, $d = `r printnum(d)`$, $t(96.62) = `r printnum(t)`$, $p = `r printp(p)`$, two-sided).
+In contrast to the original study, they found a negative difference between the Asian primed group ($n = `r n.asian`$, $M = `r printnum(mean.asian)`$, $SD = `r printnum(sd.asian)`$) and the control ($n = `r n.control`$, $M = `r printnum(mean.control)`$, $SD = `r printnum(sd.control)`$) which was not significant, $d = `r printnum(d)`$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(t)`$, $p = `r printp(p)`$, two-sided).
 The non-significant null-hypothesis test does not allow us to distinguish between the absence of a meaningful effect and data that are not sensitive enough to tell us whether a meaningful effect is present or absent.
@@
 The TOST is non-significant, $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (Figure 2A).
-Thus, we cannot reject a true effect as large or larger than 6.25%.^[Because this is a replication of a study, it would have been reasonable to assume that we only want reject effects in the same direction as the effect in the original study.
-After all, much smaller effects may indicate non-replication, but so does large effects in the opposite direction.
+Thus, we cannot reject a true effect as large or larger than 6.25%.^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
+After all, much smaller effects may indicate non-replication, but so do large effects in the opposite direction.
 In that case, we would perform an inferiority test (see figure 1D) against $\Delta_{U}$.
-As can be seen in figure 2A, the 90\% CIs does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+As can be seen in figure 2A, the 90\% CI does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
 Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
@@
 When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of `r printp(alpha)` to correct for multiple comparisons) and using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that all effect size estimates are measured with such high precision that they are statistically equivalent, and can be considered trivially small according to the authors’ definition (Figure 2C).
-In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printnum(max(gr3$TOST_p1, gr3$TOST_p2))`$.
-However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large.
-(e.g., for grade 3, $z = `r printnum(gr3.z)`$, *p* < 0.001.
+In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printp(max(gr3$TOST_p1, gr3$TOST_p2))`$.
+However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large (e.g., for grade 3, $z = `r printnum(gr3.z)`$, $p < `r printp(gr3.p`)$).
 We see how equivalence tests allow researchers to distinguish between statistical significance and practical significance, which demonstrates how equivalence tests improve hypothesis testing procedures in psychological research.
@@
 This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
-To calculate which difference corresponds to a critical z value in the original study, we multiply critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
+To calculate which difference corresponds to a critical z value in the original study, we multiply the critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
 

### 2017-11-17 manuscript.Rmd
minor edit

@@
 In this case, concluding equivalence would suggest that effects which could reasonably be detected with a sample size of 100 can be rejected, and that future studies probably need larger sample sizes to examine a specific question.
-Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area.
+Whether or not the answer such an equivalence test provides is interesting depends on the sample size used in the study, and in part on how typical the chosen sample size is for the research area.
 An equivalence test based on 15 observations rejecting a SESOI of $d = 1$ cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.

### 2017-11-17 manuscript.Rmd

@@
 This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
-To calculate which difference corresponds to a critical z value in the original study, we multiply the critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
+To calculate which difference corresponds to a critical z value in the original study, we multiply the critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find that the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
 

### 2017-11-17 manuscript.Rmd

@@
 When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of `r printp(alpha)` to correct for multiple comparisons) and using equivalence bounds of $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$, we see that all effect size estimates are measured with such high precision that they are statistically equivalent, and can be considered trivially small according to the authors’ definition (Figure 2C).
-In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p = `r printp(max(gr3$TOST_p1, gr3$TOST_p2))`$.
-However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large (e.g., for grade 3, $z = `r printnum(gr3.z)`$, $p < `r printp(gr3.p`)$).
+In the case of grade 3, $z = `r printnum(max(gr3$TOST_Z1, gr3$TOST_Z2))`$, $p `r printp(max(gr3$TOST_p1, gr3$TOST_p2))`$.
+However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large (e.g., for grade 3, $z = `r printnum(gr3.z)`$, $p `r printp(gr3.p)`$).
 We see how equivalence tests allow researchers to distinguish between statistical significance and practical significance, which demonstrates how equivalence tests improve hypothesis testing procedures in psychological research.

### 2017-11-17 manuscript.Rmd
minor edit

@@
 If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25% ...
-A+ = 93.75% to 100%) we could decide that we are only interested in test differences that correspond to an increase or decrease of at least 1 grade point.
+A+ = 93.75% to 100%) we could decide that we are only interested in test score differences that correspond to an increase or decrease of at least 1 grade point.
 Thus, our SESOI becomes a difference in raw scores of `r printnum(sesoi * 100)`\%, or `r printnum(sesoi, digits = 4)`.

### 2017-11-17 manuscript.Rmd
minor edit

@@
 Thus, our SESOI becomes a difference in raw scores of `r printnum(sesoi * 100)`\%, or `r printnum(sesoi, digits = 4)`.
-We can then perform the TOST procedure for a two-sample Welch’s *t*-test, using $\pm$ the SESOI of `r printnum(sesoi, digits = 4)` as the equivalence bounds ($\Delta$).
+We can then perform an equivalence test for a two-sample Welch’s *t*-test, with equivalence bounds of $\pm$ the SESOI of `r printnum(sesoi, digits = 4)`.
 The TOST is non-significant, $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (Figure 2A).
@@
 Since this is a replication, we could decide that we are interested in whether the observed effect size is smaller than the smallest effect size that the original study could have detected.
-This means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
-To calculate which difference corresponds to a critical z value in the original study, we multiply the critical z value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find that the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
+This means that we use the critical *z* value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
+To calculate which difference corresponds to a critical *z* value in the original study, we multiply the critical *z* value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find that the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
 

### 2017-11-17 manuscript.Rmd
tiny fix

@@
 # Justifying the Smallest Effect Size of Interest
-
 TOST is performed against equivalence bounds that are considered the smallest effect size of interest.
@@
 ##Objective Justification of a SESOI
-
 An objectively determined SESOI should be based on quantifiable theoretical predictions, such as computational models.
@@
 ##Subjective Justification of a SESOI
-
 We distinguish between three categories of subjective justifications for the SESOI.
@@
 
-#Raw versus standardized equivalence bounds
-
+##Raw versus standardized equivalence bounds
 The SESOI, and thus the equivalence bounds, can be set in terms of standardized effect sizes (e.g.
@@
 ```
-
 @Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
@@
 ## Example 2: Statistically Equivalent and Not Statistically Different
-
 ```{r include = FALSE}
@@
 ```
-
 @Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($t(`r orig.df`)= `r orig.t`$, $p= `r printp(orig.p)`$, $d= `r printnum(orig.d)`$).
@@
 ## Example 3: Statistically Equivalent and Statistically Different
-
 ```{r, include=FALSE}
@@
 ```
-
 @Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as absolute effect sizes smaller than $d = `r printnum(sesoi)`$.
@@
 ```
-
 @Lynott2014 conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior [replicating @Williams2008].
 They observed that `r printnum(selfish.cold*100)`\% of participants who received a cold pack ($n_{cold}=`r n.cold`$) opted to receive a reward for themselves, while `r printnum(selfish.hot*100)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
-In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $Z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
+In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
 Had the authors planned to perform a NHST *and* an equivalence test, the latter would allow one to distinguish between an inconclusive outcome, or a statistically equivalent result.
@@
 ```
-
 Kahane and colleagues [-@Kahane2015] investigate how responses to moral dilemmas in which participants have to decide whether or not they would sacrifice one person's life to save several other lives relate to other indicators of moral orientation.
@@
 # Discussion
-
 Equivalence testing is a simple statistical technique to reject the presence of a smallest effect size of interest.

### 2017-11-17 Example 2 (Brandt et al.)/Example2_simplified.R
add simple examples

@@ -1,0 +1,23 @@
+## Example 2 in three easy steps: t-test for two independent groups 
+
+# Every line that starts with a "#" is a comment, and contains information for you.
+
+# 1. Install the TOSTER package. NB! You only need to do this once.
+install.packages("TOSTER")
+
+# 2. Load the TOSTER package into your R environment. NB! You only need to do this once every time you open R.
+library(TOSTER)
+
+# 3. Run TOST with the numbers from example 2.
+# Performing the TOST using the 'TOSTtwo' function allows you to specify equivalence bounds in Cohen's d.
+# If you want to use this for your own data, simply edit the numbers in the function below to match your data.
+
+TOSTER::TOSTtwo(m1  = 4.785714,   # Mean of group 1
+                m2  = 4.656863,  # Mean of group 2
+                sd1 = 1.089725,  # Standard deviation of group 1
+                sd2 = 1.189497,  # Standard deviation of group 2
+                n1  = 49,  # Number of subjects in group 1
+                n2  = 51,  # Number of subjects in group 2
+                low_eqbound  = -0.4929019,  # Value for the lower equivalence bound
+                high_eqbound =  0.4929019,  # Value for the higher equivalence bound
+                alpha = .05)  # Alpha level for TOST and NHST. 

### 2017-11-17 Example 3 (Hyde et al.)/Example3_simplified.R

@@ -1,0 +1,20 @@
+## Example 3 in three easy steps: meta-analysis on Cohen's d
+
+# Every line that starts with a "#" is a comment, and contains information for you.
+
+# 1. Install the TOSTER package. NB! You only need to do this once.
+install.packages("TOSTER")
+
+# 2. Load the TOSTER package into your R environment. NB! You only need to do this once every time you open R.
+library(TOSTER)
+
+# 3. Run TOST with the numbers from example 3, for Grade 3.
+# Performing the TOST using the 'TOSTmeta' function allows you to specify equivalence bounds in Cohen's d.
+# If you want to use this for your own data, simply edit the numbers in the function below to match your data.
+
+TOSTER::TOSTmeta(ES = 0.04, 
+                 se = 0.002, 
+                 low_eqbound_d = -0.1, 
+                 high_eqbound_d = 0.1, 
+                 alpha = 0.05, 
+                 plot = TRUE)

### 2017-11-17 Example 4 (Lynott et al.)/Example4_simplified.R

@@ -1,0 +1,21 @@
+## Example 4 in three easy steps: comparing two independent proportions
+
+# Every line that starts with a "#" is a comment, and contains information for you.
+
+# 1. Install the TOSTER package. NB! You only need to do this once.
+install.packages("TOSTER")
+
+# 2. Load the TOSTER package into your R environment. NB! You only need to do this once every time you open R.
+library(TOSTER)
+
+# 3. Run TOST with the numbers from example 4.
+# Performing the TOST using the 'TOSTtwo.prop' function allows you to specify equivalence bounds in proportions.
+# If you want to use this for your own data, simply edit the numbers in the function below to match your data.
+
+TOSTER::TOSTtwo.prop(prop1 = 0.5074257,
+                     prop2 = 0.5745721,
+                     n1    = 404,
+                     n2    = 409,
+                     low_eqbound  = -0.2510861,
+                     high_eqbound = 0.2510861,
+                     alpha = 0.05)

### 2017-11-17 Example 5 (Kahane et al.)/Example5_simplified.R

@@ -1,0 +1,19 @@
+## Example 5 in three easy steps: correlations 
+
+# Every line that starts with a "#" is a comment, and contains information for you.
+
+# 1. Install the TOSTER package. NB! You only need to do this once.
+install.packages("TOSTER")
+
+# 2. Load the TOSTER package into your R environment. NB! You only need to do this once every time you open R.
+library(TOSTER)
+
+# 3. Run TOST with the numbers from example 5.
+# Performing the TOST using the 'TOSTr' function allows you to specify equivalence bounds for a correlation.
+# If you want to use this for your own data, simply edit the numbers in the function below to match your data.
+
+TOSTER::TOSTr(n = 231, 
+              r = 0.04206585, 
+              low_eqbound_r = -0.1830961, 
+              high_eqbound_r = 0.1830961, 
+              alpha = 0.05)

### 2017-11-18 Example 1 (Moon et al.)/Example1_simplified.R
Minor edits to comments and TOST functions

@@ -1,9 +1,9 @@
 
 
-## Example 1 in three easy steps: t-test for two independent groups 
+## Example 1 in three easy steps: t-test for two independent groups.
+
 
 # Every line that starts with a "#" is a comment, and contains information for you.
-
 
 # 1. Install the TOSTER package. NB! You only need to do this once.
 install.packages("TOSTER")
@@ -15,12 +15,12 @@ library(TOSTER)
 # Performing the TOST using the 'TOSTtwo.raw' function allows you to specify equivalence bounds in raw values.
 # If you want to use this for your own data, simply edit the numbers in the function below to match your data.
 
-TOSTER::TOSTtwo.raw(m1  = 0.458513,   # Mean of group 1
-                    m2  = 0.4962693,  # Mean of group 2
-                    sd1 = 0.1749011,  # Standard deviation of group 1
-                    sd2 = 0.1770452,  # Standard deviation of group 2
-                    n1  = 53,  # Number of subjects in group 1
-                    n2  = 48,  # Number of subjects in group 2
-                    low_eqbound  = -0.0625,  # Value for the lower equivalence bound
-                    high_eqbound =  0.0625,  # Value for the higher equivalence bound
-                    alpha = .05)  # Alpha level for TOST and NHST. 
+TOSTtwo.raw(m1  = 0.458513,   # Mean of group 1
+            m2  = 0.4962693,  # Mean of group 2
+            sd1 = 0.1749011,  # Standard deviation of group 1
+            sd2 = 0.1770452,  # Standard deviation of group 2
+            n1  = 53,  # Number of subjects in group 1
+            n2  = 48,  # Number of subjects in group 2
+            low_eqbound  = -0.0625,  # Value for the lower equivalence bound
+            high_eqbound =  0.0625,  # Value for the higher equivalence bound
+            alpha = .05)  # Alpha level for TOST and NHST

### 2017-11-18 Example 2 (Brandt et al.)/Example2_simplified.R

@@ -1,4 +1,7 @@
-## Example 2 in three easy steps: t-test for two independent groups 
+
+
+## Example 2 in three easy steps: t-test for two independent groups.
+
 
 # Every line that starts with a "#" is a comment, and contains information for you.
 
@@ -12,12 +15,12 @@ library(TOSTER)
 # Performing the TOST using the 'TOSTtwo' function allows you to specify equivalence bounds in Cohen's d.
 # If you want to use this for your own data, simply edit the numbers in the function below to match your data.
 
-TOSTER::TOSTtwo(m1  = 4.785714,   # Mean of group 1
-                m2  = 4.656863,  # Mean of group 2
-                sd1 = 1.089725,  # Standard deviation of group 1
-                sd2 = 1.189497,  # Standard deviation of group 2
-                n1  = 49,  # Number of subjects in group 1
-                n2  = 51,  # Number of subjects in group 2
-                low_eqbound  = -0.4929019,  # Value for the lower equivalence bound
-                high_eqbound =  0.4929019,  # Value for the higher equivalence bound
-                alpha = .05)  # Alpha level for TOST and NHST. 
+TOSTtwo(m1  = 4.785714,  # Mean of group 1
+        m2  = 4.656863,  # Mean of group 2
+        sd1 = 1.089725,  # Standard deviation of group 1
+        sd2 = 1.189497,  # Standard deviation of group 2
+        n1  = 49,  # Number of subjects in group 1
+        n2  = 51,  # Number of subjects in group 2
+        low_eqbound  = -0.4929019,  # Value for the lower equivalence bound
+        high_eqbound =  0.4929019,  # Value for the higher equivalence bound
+        alpha = .05)  # Alpha level for TOST and NHST

### 2017-11-18 Example 3 (Hyde et al.)/Example3_simplified.R

@@ -1,4 +1,7 @@
-## Example 3 in three easy steps: meta-analysis on Cohen's d
+
+
+## Example 3 in three easy steps: meta-analysis on Cohen's d.
+
 
 # Every line that starts with a "#" is a comment, and contains information for you.
 
@@ -12,9 +15,8 @@ library(TOSTER)
 # Performing the TOST using the 'TOSTmeta' function allows you to specify equivalence bounds in Cohen's d.
 # If you want to use this for your own data, simply edit the numbers in the function below to match your data.
 
-TOSTER::TOSTmeta(ES = 0.04, 
-                 se = 0.002, 
-                 low_eqbound_d = -0.1, 
-                 high_eqbound_d = 0.1, 
-                 alpha = 0.05, 
-                 plot = TRUE)
+TOSTmeta(ES = 0.04,  # Meta-analytic effect size
+         se = 0.002,  # Standard deviation
+         low_eqbound_d = -0.1,  # Value for the lower equivalence bound
+         high_eqbound_d = 0.1,  # Value for the higher equivalence bound
+         alpha = 0.05)  # Alpha level for TOST and NHST

### 2017-11-18 Example 4 (Lynott et al.)/Example4_simplified.R

@@ -1,4 +1,5 @@
-## Example 4 in three easy steps: comparing two independent proportions
+## Example 4 in three easy steps: comparing two independent proportions.
+
 
 # Every line that starts with a "#" is a comment, and contains information for you.
 
@@ -12,10 +13,10 @@ library(TOSTER)
 # Performing the TOST using the 'TOSTtwo.prop' function allows you to specify equivalence bounds in proportions.
 # If you want to use this for your own data, simply edit the numbers in the function below to match your data.
 
-TOSTER::TOSTtwo.prop(prop1 = 0.5074257,
-                     prop2 = 0.5745721,
-                     n1    = 404,
-                     n2    = 409,
-                     low_eqbound  = -0.2510861,
-                     high_eqbound = 0.2510861,
-                     alpha = 0.05)
+TOSTtwo.prop(prop1 = 0.5074257,  # Proportion of group 1
+             prop2 = 0.5745721,  # Proportion of group 2
+             n1    = 404,  # Number of subjects in group 1
+             n2    = 409,  # Number of subjects in group 2
+             low_eqbound  = -0.2510861,  # Value for the lower equivalence bound
+             high_eqbound = 0.2510861,  # Value for the lower equivalence bound
+             alpha = 0.05)  # Alpha level for TOST and NHST

### 2017-11-18 Example 5 (Kahane et al.)/Example5_simplified.R

@@ -1,4 +1,4 @@
-## Example 5 in three easy steps: correlations 
+## Example 5 in three easy steps: correlations.
 
 # Every line that starts with a "#" is a comment, and contains information for you.
 
@@ -12,8 +12,8 @@ library(TOSTER)
 # Performing the TOST using the 'TOSTr' function allows you to specify equivalence bounds for a correlation.
 # If you want to use this for your own data, simply edit the numbers in the function below to match your data.
 
-TOSTER::TOSTr(n = 231, 
-              r = 0.04206585, 
-              low_eqbound_r = -0.1830961, 
-              high_eqbound_r = 0.1830961, 
-              alpha = 0.05)
+TOSTr(n = 231,  # Number of pairs of observations
+      r = 0.04206585,  # Correlation effect size
+      low_eqbound_r = -0.1830961,  # Value for the lower equivalence bound
+      high_eqbound_r = 0.1830961,  # Value for the higher equivalence bound
+      alpha = 0.05)  # Alpha level for TOST and NHST

### 2018-02-07 manuscript.Rmd
Manuscript updated to latest version in Google doc (07-02-2018 14:15).

@@
   This work was funded by VIDI grant 452-17-013.
-We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions, Dermot Lynott and Katie Corker for their helpful responses to our inquiries about @Lynott2014, and Jim Everett and Brian Earp for their helpful responses and for providing the raw data for study 4 in Kahane et al.
-[-@Kahane2015].
+We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions, Dermot Lynott and Katie Corker for their helpful responses to our inquiries about @Lynott2014, Jim Everett and Brian Earp for their helpful responses and for providing the raw data for study 4 in Kahane et al.
+[-@Kahane2015], and Neil McLatchie and Jan Vanhove for valuable comments on an earlier version of this manuscript.
   \newline
@@
 We explain a range of approaches to determine the SESOI in psychological science, and provide detailed examples of how equivalence tests should be performed and reported.
-Equivalence tests are an important extension of statistical tools psychologist currently use, and enable researchers to falsify predictions about the presence, and declare the absence, of meaningful effects.
+Equivalence tests are an important extension of statistical tools psychologists currently use, and enable researchers to falsify predictions about the presence, and declare the absence, of meaningful effects.
 
 keywords          : "Equivalence Testing, NHST, power, TOST, falsification."
-wordcount         : "4969"
+wordcount         : "5516"
 
@@
 
-Equivalence testing can be used to test whether an observed effect is surprisingly small, assuming a meaningful effect exists in the population.
+Equivalence testing can be used to test whether an observed effect is surprisingly small, assuming a meaningful effect exists in the population (see for example [@Goertzen2010]; [@Quertemont2011]; [@Meyners2012]; [@Rogers1993]).
 The test is a simple variation of widely used null hypothesis significance tests (NHST).
@@
 In our example, the researcher can perform an equivalence test to examine whether the gender difference in participants is *not* as large or larger than the *smallest effect size of interest* (SESOI).
-After an extensive discussion with experts, the researcher decides that as long as the difference in proportions is not more extreme than 6%, the gender difference is too small to care about.
+After an extensive discussion with experts, the researcher decides that as long as the difference in proportions does not deviate from the population gender ratio by more than 6%, the gender difference is too small to care about.
 Given an expected true difference in the population of 0.015, the researcher will test if the observed difference falls outside the boundary values (or *equivalence bounds*) of -0.055 and 0.075.
@@
 In TOST, the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
-It is not necessary to control for multiple tests when using the TOST procedure because both tests need to be statistically significant to conclude equivalence.
-This means that when reporting a TOST result, it suffices to report the one-sided test with the smaller test parameter (e.g.
+Even though the TOST procedure consists of two one-sided tests, it is not necessary to control for multiple testing because both tests need to be statistically significant to conclude equivalence.
+Consequently, when reporting a TOST result, it suffices to report the one-sided test with the smaller test parameter (e.g.
 $t$ or $r$) and thus the larger $p$-value.
@@
 Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze [but we recommend to justify the Type 1 error rate in original research based on substantive arguments, @Lakens2017b].
-It might be easiest to think of an equivalence test as checking whether the confidence interval excludes the SESOI, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
+It may be easiest to think of an equivalence test as checking whether the confidence interval excludes the equivalence bounds, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
 
@@
 In this article we provide five reproducible examples of equivalence tests which illustrate the procedure in easy-to-use free software (R, jamovi, and a spreadsheet).
-But first, we discuss different approaches to determining the SESOI for psychological research, which determines the statistical question an equivalence test answers.
+But first, we discuss different approaches to determining the smallest effect size of interest for psychological research, which determines the statistical question an equivalence test answers.
 
 # Justifying the Smallest Effect Size of Interest
-TOST is performed against equivalence bounds that are considered the smallest effect size of interest.
-The SESOI can sometimes be determined objectively, for example based on just noticeable differences.
+The Two One-Sided Test procedure is performed against equivalence bounds that are considered the smallest effect size of interest.
+The smallest effect size of interest (SESOI) can sometimes be determined objectively (for example based on just noticeable differences, as we will explain below).
 In lieu of objective justifications, the SESOI should ideally be based on a cost-benefit analysis.
@@
 
-##Objective Justification of a SESOI
-An objectively determined SESOI should be based on quantifiable theoretical predictions, such as computational models.
+##Objective Justification of a Smallest Effect Size of Interest
+An objectively determined smallest effect size of interest (SESOI) should be based on quantifiable theoretical predictions, such as computational models.
 Sometimes, the only theoretical prediction is that an effect should be noticeable.
@@
 
-##Subjective Justification of a SESOI
-We distinguish between three categories of subjective justifications for the SESOI.
+##Subjective Justification of a smallest effect size of interest
+We distinguish between three categories of subjective justifications to determine the smallest effect size of interest.
 First, researchers can use benchmarks.
-For example, one might set the SESOI to a standardized effect size of $d = 0.5$, which would allow rejecting effects more extreme than a 'medium' effect size [@Cohen1988].
+For example, one might set the smallest effect size of interest to a standardized effect size of $d = 0.5$, which would allow rejecting effects more extreme than a 'medium' effect size [@Cohen1988].
 Similarly, effect sizes smaller than a Cohen's $d$ of 0.1 are sometimes considered trivially small [@Maxwell2015].
-Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided.
-Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that a SESOI of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
+Relying on a benchmark is the weakest possible justification of a smallest effect size of interest, and should be avoided.
+Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a smallest effect size of interest (SESOI) to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting only the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that a smallest effect size of interest of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
 
@@
 
-Second, the SESOI can be based on published studies in the literature.
+Second, the SESOI can be based on related published studies in the literature.
 Ideally, researchers who publish novel research would always specify their smallest effect size of interest, but this is not yet common practice.
@@
 In other words, we decide that effects that could not have yielded a $p < \alpha$ in an original study will not be considered meaningful in the replication study either, even if they were found to be statistically significant.
+The assumption here is that authors were interested in observing a significant effect, and thus were not interested in observed effect sizes that could not have yielded a significant effect.
+It might be likely that the original authors did not consider which effect sizes they could detect at all, or that they were interested in smaller effects, but gambled on observing an especially large effect in the sample purely due to random variation.
+Even then, observed effect sizes that could not have been significant in an earlier study might be a justifiable starting point when building on earlier research that did not specify a smallest effect size of interest.
 Based only on the alpha level and the sample size, we can calculate the critical test value (e.g., *t*, *F*, *Z*).
@@
 
-Third, researchers can set the SESOI based on the sample size they are planning to collect.
-The amount of data you can collect limits the inferences you can make.
-Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a $p$-value *larger than* 0.05 in a traditional NHST.
-However, @Meyners2012 explains that this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and error rates are not controlled accurately.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size.
-In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size.
-For example, a researcher who plans to perform a one-sample $t$-test using an alpha of `r alpha*100`% (two-sided), based on data from `r N` observations, has 90% power to detect an effect of $d = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$.
-Using equivalence bounds of $\Delta_{L} = `r printnum(-pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ and $\Delta_{U} = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ in an equivalence test allows to reject effects the study had high statistical power to detect.
+Third, the SESOI can be determined based on a *resource* question.
+Not all psychological theories make quantifiable predictions (possibly related to the strong reliance on null-hypothesis significance testing in some fields, see [@Meehl1978]).
+Researchers often have more precise ideas of the amount of data that they can afford to collect, that other researchers in their field commonly collect.
+The amount of data that is collected limits the inferences you can make.
+Given the alpha level and the planned sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a $p$-value *larger than* 0.05 in a traditional NHST.
+However, @Meyners2012 explains that this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and error rates are not controlled accurately.]
 
-Importantly, an equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that could reliably be detected given a specific sample size?
-In this case, concluding equivalence would suggest that effects which could reasonably be detected with a sample size of 100 can be rejected, and that future studies probably need larger sample sizes to examine a specific question.
-Whether or not the answer such an equivalence test provides is interesting depends on the sample size used in the study, and in part on how typical the chosen sample size is for the research area.
-An equivalence test based on 15 observations rejecting a SESOI of $d = 1$ cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected.
-By transparently reporting the effects that can be detected and rejected based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI may be.
+An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions).
+When theories only allow for directional predictions, but do not predict effects of a specific size, , the sample size justification can be used to determine which effects can be studied reliably, and thus which effects would be interesting to reject.
+For example, imagine a research line where almost all studies examined a hypothesis by performing a one-sample $t$-test on sample sizes smaller than `r N`.
+A one-sample $t$-test using an alpha of `r alpha*100`% (two-sided) with `r N` observations has 90% power to detect an effect of $d = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$.
+Concluding equivalence in a test using equivalence bounds of $\Delta_{L} = `r printnum(-pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ and $\Delta_{U} = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ would suggest that effects as large or larger than those which published studies were sensitive enough to detect can be rejected.
+Such a study does not test a theoretical prediction, but it contributes to the literature by suggesting much larger sample sizes are needed than `r N` observations.
+
+When there is no previous literature on a topic, researchers can justify their sample size based on reasonable resource limitations, which are specific to scientific disciplines and research questions (and can be expected to change over time).
+Where 22 patients with a rare neurological disorder might be the largest sample a single researcher can collect, a study on mTurk might easily collect hundreds or thousands of observations.
+Thus, whether or not the resource question that is answered by an equivalence test is interesting must be evaluated by peers, preferably when the study design is submitted to an ethical review board or as a Registered Report.
+
+Additional subjective justifications for a smallest effect size of interest are possible.
+For example, the Food and Drug Authorization has set equivalence bounds for bioequivalence studies, taking the decision out of the hands of individual researchers [@Senn2007].
+We have provided several approaches to help researchers justify equivalence bounds, but want to repeat the warning by @Rogers1993, p.
+564: 'The thoughtless application of "cookbook" prescriptions is ill-advised, regardless of whether the goal is to establish a difference or to establish equivalency between treatments'.
+By transparently reporting and adequately justifying the smallest effect size of interest, researchers can communicate the information their study contributes to the literature, and provide a starting point for a discussion about what a reasonable SESOI may be.
 
@@
 ```
-
 ##Raw versus standardized equivalence bounds
@@
 plot(example.grid)
-```
 
@@
 @Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
-In contrast to the original study, they found a negative difference between the Asian primed group ($n = `r n.asian`$, $M = `r printnum(mean.asian)`$, $SD = `r printnum(sd.asian)`$) and the control ($n = `r n.control`$, $M = `r printnum(mean.control)`$, $SD = `r printnum(sd.control)`$) which was not significant, $d = `r printnum(d)`$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(t)`$, $p = `r printp(p)`$, two-sided).
+In contrast to the original study, they found a negative difference between the Asian primed group ($n = `r n.asian`$, $M = `r printnum(mean.asian)`$, $SD = `r printnum(sd.asian)`$) and the control ($n = `r n.control`$, $M = `r printnum(mean.control)`$, $SD = `r printnum(sd.control)`$) which was not significant, $d = `r printnum(d)`$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(t)`$, $p = `r printp(p)`$, two-sided)^[Fractional degrees of freedom in t-tests are a result of using Welch’s t-test instead of Student’s t-test, which is the recommended default when sample sizes are unequal between groups [@Delacre2017].
+When using the TOSTER package in R, this is achieved by setting “var.equal = FALSE”.].
 The non-significant null-hypothesis test does not allow us to distinguish between the absence of a meaningful effect and data that are not sensitive enough to tell us whether a meaningful effect is present or absent.
@@
 We can then perform an equivalence test for a two-sample Welch’s *t*-test, with equivalence bounds of $\pm$ the SESOI of `r printnum(sesoi, digits = 4)`.
-The TOST is non-significant, $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (Figure 2A).
-Thus, we cannot reject a true effect as large or larger than 6.25%.^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
+The two-one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$.
+While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r sesoi`, the test against $\Delta_{L}$ shows we can *not* reject effects as large or larger than `r -sesoi`.
+The TOST is non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
 After all, much smaller effects may indicate non-replication, but so do large effects in the opposite direction.
 In that case, we would perform an inferiority test (see figure 1D) against $\Delta_{U}$.
-As can be seen in figure 2A, the 90\% CI does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+As can be seen in Figure 2A, the 90\% CI does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
 Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
@@
 A TOST procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
+Following a Neyman-Pearson approach, this means we can reject the hypothesis that the true effect is smaller than $d = `r printnum(-d.33)`$ or larger than $d = `r printnum(d.33)`$, and we can act as if the effect size falls within the equivalence bounds, with the understanding that based on our chosen alpha level, we will not be wrong more than 5% of the time.
 
@@
 contest this interpretation in a number of studies.
-In Study 4, they compare the traditionally used dilemmas with a set of new delimmas in which the sacrifice for the greater good is not another person’s life, but something the participant would have a partial interest in (e.g.
+In Study 4, they compare the traditionally used dilemmas with a set of new dilemmas in which the sacrifice for the greater good is not another person’s life, but something the participant would have a partial interest in (e.g.
 buying a new mobile phone vs.
@@
 An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2003], but researchers should always explain why they chose a SESOI.
-It makes little sense to report a statistical test, without explaining why one would want to answer this question to begin with.
+It makes little sense to report a statistical test without explaining why one would want to answer this question to begin with.
+
 Equivalence bounds should be specified before results are known, ideally as part of a pre-registration [cf.
 @Piaggio2006].
-Researchers might be tempted to set equivalence bounds *after* looking at their data, but like any other results-dependent decision, this "boundary hacking" will increase error rates.
+In the most extreme case, a researcher can always choose equivalence bounds wide enough for the test to yield a 'statistically equivalent' result [for a discussion, see @Weber2012].
+However, fixed error rates are no longer valid when bounds are determined based on the observed data.
+The value of reporting an equivalence test is determined by the strength of the justification of the equivalence bounds.
+If bounds are chosen based on the observed data, an equivalence test becomes meaningless.
+We have proposed several ways to justify equivalence bounds, but in the end these discussions must happen among peers, and the best place for these discussions is during peer-review of proposals for Registered Reports.
 
-We recommend that researchers by default perform both a null hypothesis significance test and an equivalence test on their data.
+As with null-hypothesis significance tests, equivalence tests interpreted from a Neyman-Pearson perspective on statistical inferences are used to guide the actions of researchers while controlling error rates.
+Research lines sometimes require dichotomous choices.
+For example, a researcher might want to decide to discontinue a research line when an effect is too small to be considered meaningful.
+Equivalence tests based on carefully chosen equivalence bounds and error rates can be used as a rule to govern our behavior [@Neyman1933].
+An equivalence test and a null-hypothesis significance test examine two different hypotheses, and can therefore be used in concert [@Weber2012].
+We recommend that researchers by default perform both a null hypothesis significance test *and* an equivalence test on their data, as long as they can justify a smallest effect size of interest, in order to improve the falsifiability of predictions in psychological science.
+
 The biggest challenge for researchers will be to specify the smallest effect size of interest.

### 2018-02-07 manuscript.Rmd
Included include=false to supress the warning message from library(pwr)
Revision final changes

@@
 In addition to testing against zero, researchers can use the Two One-Sided Tests (TOST) procedure to test for *equivalence* and reject the presence of a smallest effect size of interest (SESOI).
-TOST can be used to determine if an observed effect is surprisingly small, given that a true effect at least as large as the SESOI exists.
+The two-one-sided tests procedure can be used to determine if an observed effect is surprisingly small, given that a true effect at least as large as the SESOI exists.
 We explain a range of approaches to determine the SESOI in psychological science, and provide detailed examples of how equivalence tests should be performed and reported.
@@
 
-keywords          : "Equivalence Testing, NHST, power, TOST, falsification."
-wordcount         : "5516"
+keywords          : "Equivalence Testing, null-hypothesis significance test, power, TOST, falsification."
+wordcount         : "5545"
 
@@
 Equivalence testing can be used to test whether an observed effect is surprisingly small, assuming a meaningful effect exists in the population (see for example [@Goertzen2010]; [@Quertemont2011]; [@Meyners2012]; [@Rogers1993]).
-The test is a simple variation of widely used null hypothesis significance tests (NHST).
+The test is a simple variation of widely used null hypothesis significance tests.
 To understand the idea behind equivalence tests, it is useful to realize the null hypothesis we test against can be any numerical value.
@@
 *Equivalence tests* can be seen as the opposite of minimal effects tests: They examine whether the presence of effects that are large enough to be considered meaningful can be *rejected* (see Figure 1C).
+Note that when a meaningful effect can be rejected, it does not imply there is no effect at all.
 In our example, the researcher can perform an equivalence test to examine whether the gender difference in participants is *not* as large or larger than the *smallest effect size of interest* (SESOI).
@@
 In any one-sided test, for an alpha level of 5%, we can reject $H_0$ when the 90% confidence interval around the observed estimate is in the predicted direction, and does not contain the value tested against (e.g., 0).
-In TOST, the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
+In the two-one-sided tests procedure the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
 Even though the TOST procedure consists of two one-sided tests, it is not necessary to control for multiple testing because both tests need to be statistically significant to conclude equivalence.
-Consequently, when reporting a TOST result, it suffices to report the one-sided test with the smaller test parameter (e.g.
+Consequently, when reporting an equivalence test, it suffices to report the one-sided test with the smaller test parameter (e.g.
 $t$ or $r$) and thus the larger $p$-value.
@@
 Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze [but we recommend to justify the Type 1 error rate in original research based on substantive arguments, @Lakens2017b].
-It may be easiest to think of an equivalence test as checking whether the confidence interval excludes the equivalence bounds, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a NHST is larger than some *p*-value bound.
+It may be easiest to think of an equivalence test as checking whether the confidence interval excludes the equivalence bounds, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a null-hypothesis significance test is larger than some *p*-value bound.
 
 ```{r echo=FALSE, fig.width=4, fig.height=6.2, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests.
-**A)** NHST: Tests if the hypothesis ($H_0$) that an effect is equal to 0 can be rejected.
+**A)** Null-Hypothesis Significance Test: Tests if the hypothesis ($H_0$) that an effect is equal to 0 can be rejected.
 **B)**  Minimal effects test: Tests if the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$ can be rejected.
@@
 
-```{r, results="hide"}
+```{r, results="hide", include=FALSE}
 library(pwr)
@@
 The amount of data that is collected limits the inferences you can make.
-Given the alpha level and the planned sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a $p$-value *larger than* 0.05 in a traditional NHST.
+Given the alpha level and the planned sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a $p$-value *larger than* 0.05 in a traditional null-hypothesis significance test.
 However, @Meyners2012 explains that this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and error rates are not controlled accurately.]
@@
 An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions).
-When theories only allow for directional predictions, but do not predict effects of a specific size, , the sample size justification can be used to determine which effects can be studied reliably, and thus which effects would be interesting to reject.
+When theories only allow for directional predictions, but do not predict effects of a specific size, the sample size justification can be used to determine which effects can be studied reliably, and thus which effects would be interesting to reject.
 For example, imagine a research line where almost all studies examined a hypothesis by performing a one-sample $t$-test on sample sizes smaller than `r N`.
@@
 ```
-```{r echo=FALSE, fig.width=7, fig.height=4, fig.cap=paste("Example effects plotted with 90% TOST CIs (thick lines) and 95% NHST CIs (thin lines), the NHST null hypothesis (solid vertical line) and the equivalence bounds (dashed vertical lines) displayed.
+```{r echo=FALSE, fig.width=7, fig.height=4, fig.cap=paste("Example effects plotted with 90% two one-sided tests confidence intervals (thick lines) and 95% null-hypothesis significance test confidence intervals (thin lines), the null hypothesis (solid vertical line) and the equivalence bounds (dashed vertical lines) displayed.
 **A)** Example 1 - Mean Difference.
@@
 While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r sesoi`, the test against $\Delta_{L}$ shows we can *not* reject effects as large or larger than `r -sesoi`.
-The TOST is non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
+The equivalence test is non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
 After all, much smaller effects may indicate non-replication, but so do large effects in the opposite direction.
@@
 A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($M_{unethical}= `r rep.m1`$, ${SD}_{unethical}= `r rep.sd1`$, $M_{ethical}=`r rep.m2`$, ${SD}_{ethical}=`r rep.sd2`$, $t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
-Using the small telescopes approach, we can calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$) and use this as the SESOI.
-A TOST procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
+Using the small telescopes approach, we can calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$) and use this as the equivalence bounds.
+The two one-sided tests procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
 Following a Neyman-Pearson approach, this means we can reject the hypothesis that the true effect is smaller than $d = `r printnum(-d.33)`$ or larger than $d = `r printnum(d.33)`$, and we can act as if the effect size falls within the equivalence bounds, with the understanding that based on our chosen alpha level, we will not be wrong more than 5% of the time.
@@
 In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
-Had the authors planned to perform a NHST *and* an equivalence test, the latter would allow one to distinguish between an inconclusive outcome, or a statistically equivalent result.
+Had the authors planned to perform a null-hypothesis significance test *and* an equivalence test, the latter would allow one to distinguish between an inconclusive outcome, or a statistically equivalent result.
 
@@
 With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$.
-Performing the TOST procedure given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, the data is statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$.
+Performing the two one-sided tests procedure given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, the data is statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$.
 This means that values smaller than $r = `r printnum(-Kahane.r.80)`$ and larger than $r = `r printnum(Kahane.r.80)`$ can be rejected at an alpha level of 5%.
@@
 As long as we can calculate a confidence interval around a parameter estimate, we can compare it to the SESOI.
-The result of an equivalence test can be obtained by mere visual inspection of the CI [@Seaman1998; @Tryon2001], or by performing two one-sided tests (the TOST procedure).
+The result of an equivalence test can be obtained by mere visual inspection of the CI [@Seaman1998; @Tryon2001], or by performing two one-sided tests (known as the TOST procedure).
 

### 2018-02-07 manuscript.Rmd
minor spelling changes

@@
 In addition to testing against zero, researchers can use the Two One-Sided Tests (TOST) procedure to test for *equivalence* and reject the presence of a smallest effect size of interest (SESOI).
-The two-one-sided tests procedure can be used to determine if an observed effect is surprisingly small, given that a true effect at least as large as the SESOI exists.
+The two one-sided tests procedure can be used to determine if an observed effect is surprisingly small, given that a true effect at least as large as the SESOI exists.
 We explain a range of approaches to determine the SESOI in psychological science, and provide detailed examples of how equivalence tests should be performed and reported.
@@
 In any one-sided test, for an alpha level of 5%, we can reject $H_0$ when the 90% confidence interval around the observed estimate is in the predicted direction, and does not contain the value tested against (e.g., 0).
-In the two-one-sided tests procedure the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
+In the two one-sided tests procedure the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
 Even though the TOST procedure consists of two one-sided tests, it is not necessary to control for multiple testing because both tests need to be statistically significant to conclude equivalence.
@@
 We can then perform an equivalence test for a two-sample Welch’s *t*-test, with equivalence bounds of $\pm$ the SESOI of `r printnum(sesoi, digits = 4)`.
-The two-one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$.
+The two one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$.
 While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r sesoi`, the test against $\Delta_{L}$ shows we can *not* reject effects as large or larger than `r -sesoi`.

### 2018-02-10 manuscript.Rmd
Adressed editor comments 1, 2, 3, and 9.

@@
 
-author_note: >
-  This work was funded by VIDI grant 452-17-013.
-We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions, Dermot Lynott and Katie Corker for their helpful responses to our inquiries about @Lynott2014, Jim Everett and Brian Earp for their helpful responses and for providing the raw data for study 4 in Kahane et al.
-[-@Kahane2015], and Neil McLatchie and Jan Vanhove for valuable comments on an earlier version of this manuscript.
-  \newline
-  Author contributions: 
-  DL: Article conceptualization, programming TOSTER package and spreadsheet, main contributor to the introduction section and discussion section.
-  AMS and PMI: Main contributors to examples and figures.
-  All authors contributed to manuscript conceptualization, editing, review for submission, and additional software development.
-AMS and PMI contributed equally to the manuscript, and order was determined by executing the following commands in R:
-  \newline
-  `set.seed(7998976/5271)`
-  \newline
-  `x <- sample(c("Anne", "Peder"), 1)`
-  \newline
-  `print(paste("The winner is", x, "!"))`
-  \newline
+author_note:
 
@@
 Note that in this paper we will control the Type 1 error rate at 5% for all examples, mirroring the studies we reanalyze [but we recommend to justify the Type 1 error rate in original research based on substantive arguments, @Lakens2017b].
-It may be easiest to think of an equivalence test as checking whether the confidence interval excludes the equivalence bounds, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a null-hypothesis significance test is larger than some *p*-value bound.
+It may be easiest to think of an equivalence test as checking whether the entire 90% confidence interval falls entirely between the upper and lower equivalence bounds, but for any given study an equivalence test could also be conceptualized as determining whether effect sizes or test statistics are closer to zero than some critical value, or even whether the *p*-value for a null-hypothesis significance test is larger than some *p*-value bound.
 
@@
 Additional subjective justifications for a smallest effect size of interest are possible.
-For example, the Food and Drug Authorization has set equivalence bounds for bioequivalence studies, taking the decision out of the hands of individual researchers [@Senn2007].
+For example, the Food and Drug Administration has set equivalence bounds for bioequivalence studies, taking the decision out of the hands of individual researchers [@Senn2007].
 We have provided several approaches to help researchers justify equivalence bounds, but want to repeat the warning by @Rogers1993, p.
@@
 These concrete and easy to follow examples will illustrate all approaches to setting equivalence bounds discussed above, and demonstrate how to perform and report equivalence tests.
-The code to reproduce these analyses is available at <https://osf.io/qamc6/>^[The OSF project contains all necessary files to reproduce examples 1--5 in R, jamovi [@jamovi2017], and in a spreadsheet; as well as a preprint of this manuscript.
-The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`.].
 
@@
 plot(example.grid)
-
+```
 ## Example 1: Not Statistically Equivalent and Not Statistically Different
@@
 This means that values smaller than $r = `r printnum(-Kahane.r.80)`$ and larger than $r = `r printnum(Kahane.r.80)`$ can be rejected at an alpha level of 5%.
-If other researchers are interested in examining the presence of smaller effect size, they can design studies with a larger sample size.
+If other researchers are interested in examining the presence of a smaller effect size, they can design studies with a larger sample size.
 
@@
 
+##Disclosures
+
+###Data, materials, and online resources
+The code to reproduce these analyses is available at <https://osf.io/qamc6/>^[The OSF project contains all necessary files to reproduce examples 1--5 in R, jamovi [@jamovi2017], and in a spreadsheet; as well as a preprint of this manuscript.
+The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`.].
+
+###Conflicts of Interest
+The authors declare that they have no conflicts of interest with respect to the authorship or the publication of this article.
+
+###Author Contributions
+DL: Article conceptualization, programming TOSTER package and spreadsheet, main contributor to the introduction section and discussion section.
+AMS and PMI: Main contributors to examples and figures.
+All authors contributed to manuscript conceptualization, editing, review for submission, and additional software development.
+AMS and PMI contributed equally to the manuscript, and order was determined by executing the following commands in R:
+  \newline
+  `set.seed(7998976/5271)`
+  \newline
+  `x <- sample(c("Anne", "Peder"), 1)`
+  \newline
+  `print(paste("The winner is", x, "!"))`
+  \newline
+
+###Acknowledgments
+This work was funded by VIDI grant 452-17-013.
+We would like to thank Courtney Soderberg for creating the first version of the TOST function to test two independent proportions, Dermot Lynott and Katie Corker for their helpful responses to our inquiries about @Lynott2014, Jim Everett and Brian Earp for their helpful responses and for providing the raw data for study 4 in Kahane et al.
+[-@Kahane2015], and Neil McLatchie and Jan Vanhove for valuable comments on an earlier version of this manuscript.
+
+###Prior versions
+A preprint version of this manuscript prior to peer-review can be found on PsyArxiv: https://psyarxiv.com/v3zkt/ (version 1).
+It is however imprecise in its explanation of how to justify the SESOI based on available resources.
+The updated preprint version (version 2) is identical in content to the published manuscript.
+
 \newpage

### 2018-02-10 manuscript.Rmd
Adressed editor comment 4 * Did not add emphasis to example 3 because in this case, the authors have already specified a SESOI.

@@
 The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects more extreme than the specified equivalence bounds contributes to our knowledge base.
+The SESOI is thus independent of the outcome of the study, and should be determined before looking at the data.
 A SESOI should be chosen such that inferences based on it answer a meaningful question.
@@
 In order to distinguish between these possibilities, we can define what a “meaningful effect” (SESOI) would be and use it as the bounds for an equivalence test.
+In reality, we should do this before looking at the results of the study.
 If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25% ...
@@
 A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($M_{unethical}= `r rep.m1`$, ${SD}_{unethical}= `r rep.sd1`$, $M_{ethical}=`r rep.m2`$, ${SD}_{ethical}=`r rep.sd2`$, $t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
-Using the small telescopes approach, we can calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$) and use this as the equivalence bounds.
+Using the small telescopes approach (a priori of looking at the results), we can calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$) and use this as the equivalence bounds.
 The two one-sided tests procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
@@
 
-Since this is a replication, we could decide that we are interested in whether the observed effect size is smaller than the smallest effect size that the original study could have detected.
+Since this is a replication, we could decide in advance that we are interested in whether the observed effect size is smaller than the smallest effect size that the original study could have detected.
 This means that we use the critical *z* value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
@@
 This interpretation can be formalized by performing an equivalence test for correlations, where equivalence bounds are set to an effect size the study had reasonable power to detect.
+Again, these bounds should be determined and preregistered prior to running the study.
 With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$.

### 2018-02-10 manuscript.Rmd
Adressed editor comment 5 * Also made minor revisions elsewhere in the same paragraph.

@@
 The two one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$.
-While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r sesoi`, the test against $\Delta_{L}$ shows we can *not* reject effects as large or larger than `r -sesoi`.
-The equivalence test is non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
+While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r printnum(sesoi, digits = 4)`, the test against $\Delta_{L}$ shows we can *not* reject effects as extreme or more extreme than `r printnum(-sesoi, digits = 4)`.
+The equivalence test is therefore non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
 After all, much smaller effects may indicate non-replication, but so do large effects in the opposite direction.

### 2018-02-10 manuscript.Rmd
Adressed reviewer comment 6 * Moved the description to the end of the intro, after some of the necessary concepts are mentioned, but before specific SESOI justification discussion begins.

@@
 Note that when a meaningful effect can be rejected, it does not imply there is no effect at all.
-In our example, the researcher can perform an equivalence test to examine whether the gender difference in participants is *not* as large or larger than the *smallest effect size of interest* (SESOI).
+In our example, the researcher can perform an equivalence test to examine whether the gender difference in participants is *not* as extreme or more extreme than the *smallest effect size of interest* (SESOI).
 After an extensive discussion with experts, the researcher decides that as long as the difference in proportions does not deviate from the population gender ratio by more than 6%, the gender difference is too small to care about.
@@
 Statistical equivalence can be concluded when the largest of the two $p$-values is smaller than alpha.
+If results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
+Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
+The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
+Analogous to the large sample sizes needed to detect small effects, as the SESOI becomes smaller, the equivalence bounds become narrower, and the sample size needed to obtain a sufficiently narrow confidence interval increases.
 
@@
 In that case, we would perform an inferiority test (see figure 1D) against $\Delta_{U}$.
-As can be seen in Figure 2A, the 90\% CI does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions.
-Further studies are needed, which can be analyzed using a (small-scale) meta-analysis.
-The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both.
-Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$).
+As can be seen in Figure 2A, the 90\% CI does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.
+However, this must be decided before running the tests to prevent loss of error control.]
 

### 2018-02-10 manuscript.Rmd
Adressed editor comment 7

@@
 Using the small telescopes approach (a priori of looking at the results), we can calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$) and use this as the equivalence bounds.
-The two one-sided tests procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
+The two one-sided tests procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, because the larger of the two p values was less than .05, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
 Following a Neyman-Pearson approach, this means we can reject the hypothesis that the true effect is smaller than $d = `r printnum(-d.33)`$ or larger than $d = `r printnum(d.33)`$, and we can act as if the effect size falls within the equivalence bounds, with the understanding that based on our chosen alpha level, we will not be wrong more than 5% of the time.

### 2018-02-10 manuscript.Rmd
Added *practical* in italics for emphasis

@@
 However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large (e.g., for grade 3, $z = `r printnum(gr3.z)`$, $p `r printp(gr3.p)`$).
-We see how equivalence tests allow researchers to distinguish between statistical significance and practical significance, which demonstrates how equivalence tests improve hypothesis testing procedures in psychological research.
+We see how equivalence tests allow researchers to distinguish between statistical significance and *practical* significance, which demonstrates how equivalence tests improve hypothesis testing procedures in psychological research.
 

### 2018-02-11 manuscript.Rmd
Final changes R2

--- 2018-02-10 manuscript.Rmd
@@
 
-In order to distinguish between these possibilities, we can define what a “meaningful effect” (SESOI) would be and use it as the bounds for an equivalence test.
-In reality, we should do this before looking at the results of the study.
+In order to distinguish between these possibilities, we can define what a “meaningful effect” would be and use it as the bounds for an equivalence test (remember these bounds should normally be specified before looking at the data).
 If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25% ...
@@
 A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($M_{unethical}= `r rep.m1`$, ${SD}_{unethical}= `r rep.sd1`$, $M_{ethical}=`r rep.m2`$, ${SD}_{ethical}=`r rep.sd2`$, $t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
-Using the small telescopes approach (a priori of looking at the results), we can calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$) and use this as the equivalence bounds.
+We can chose (before looking at the data) to use the small telescopes approach, calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$), and use this effect size as the equivalence bounds.
 The two one-sided tests procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, because the larger of the two p values was less than .05, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
@@
 However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions and samples sizes are very large (e.g., for grade 3, $z = `r printnum(gr3.z)`$, $p `r printp(gr3.p)`$).
-We see how equivalence tests allow researchers to distinguish between statistical significance and practical significance, which demonstrates how equivalence tests improve hypothesis testing procedures in psychological research.
+We see how equivalence tests allow researchers to distinguish between *statistical* significance and *practical* significance, which demonstrates how equivalence tests improve hypothesis testing procedures in psychological research.
 
@@
 
-This interpretation can be formalized by performing an equivalence test for correlations, where equivalence bounds are set to an effect size the study had reasonable power to detect.
-Again, these bounds should be determined and preregistered prior to running the study.
+This interpretation can be formalized by performing an equivalence test for correlations, where equivalence bounds are set to an effect size the study had reasonable power to detect (as decided before looking at the data).
 With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$.
@@
 
+\newpage
+
 ##Disclosures
@@
 The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`.].
+  \newline
 
@@
 The authors declare that they have no conflicts of interest with respect to the authorship or the publication of this article.
+  \newline
 
@@
 [-@Kahane2015], and Neil McLatchie and Jan Vanhove for valuable comments on an earlier version of this manuscript.
+  \newline
 
@@
 A preprint version of this manuscript prior to peer-review can be found on PsyArxiv: https://psyarxiv.com/v3zkt/ (version 1).
-It is however imprecise in its explanation of how to justify the SESOI based on available resources.
+We have updated our recommendations on how to justify the SESOI based on available resources.
 The updated preprint version (version 2) is identical in content to the published manuscript.

### 2018-02-11 manuscript.Rmd
final

@@
 author: 
-  - name          : "Daniel Lakens"
+  - name          : "Daniël Lakens"
     affiliation   : "1"
@@
 
-Equivalence testing can be used to test whether an observed effect is surprisingly small, assuming a meaningful effect exists in the population (see for example [@Goertzen2010]; [@Quertemont2011]; [@Meyners2012]; [@Rogers1993]).
+Equivalence testing can be used to test whether an observed effect is surprisingly small, assuming a meaningful effect exists in the population [see for example @Goertzen2010; @Quertemont2011; @Meyners2012; @Rogers1993].
 The test is a simple variation of widely used null hypothesis significance tests.
@@
 In any one-sided test, for an alpha level of 5%, we can reject $H_0$ when the 90% confidence interval around the observed estimate is in the predicted direction, and does not contain the value tested against (e.g., 0).
-In the two one-sided tests procedure the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
-Even though the TOST procedure consists of two one-sided tests, it is not necessary to control for multiple testing because both tests need to be statistically significant to conclude equivalence.
+In the two one-sided tests procedure, the first one-sided test is used to test against values smaller than the lower equivalence bound ($\Delta_{L}$), and the second one-sided test is used to test against values larger than the upper equivalence bound ($\Delta_{U}$).
+Even though the TOST procedure consists of two one-sided tests, it is not necessary to control for multiple comparisons because both tests need to be statistically significant to conclude equivalence.
 Consequently, when reporting an equivalence test, it suffices to report the one-sided test with the smaller test parameter (e.g.
@@
 
-In this article we provide five reproducible examples of equivalence tests which illustrate the procedure in easy-to-use free software (R, jamovi, and a spreadsheet).
+In this article we provide five reproducible examples of equivalence tests which illustrate the procedure in free, easy-to-use software (R, jamovi, and a spreadsheet).
 But first, we discuss different approaches to determining the smallest effect size of interest for psychological research, which determines the statistical question an equivalence test answers.
@@
 A SESOI should be chosen such that inferences based on it answer a meaningful question.
-Although we use bounds that are symmetric around 0 in all examples in this manuscript for equivalence tests (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is also possible to use asymmetric bounds (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$).
+Although we use bounds that are symmetric around 0 for all equivalence tests in this manuscript (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is also possible to use asymmetric bounds (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$).
 
@@
 
-##Subjective Justification of a smallest effect size of interest
+##Subjective Justification of a Smallest Effect Size of Interest
 We distinguish between three categories of subjective justifications to determine the smallest effect size of interest.
@@
 Relying on a benchmark is the weakest possible justification of a smallest effect size of interest, and should be avoided.
-Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a smallest effect size of interest (SESOI) to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting only the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that a smallest effect size of interest of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
+Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a smallest effect size of interest (SESOI) to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting only the upper 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that a smallest effect size of interest of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.
 
@@
 @Simonsohn2015 recently proposed to set the SESOI to the effect size that an earlier study would have had 33% power to detect.
-With this so-called 'small telescopes' approach, the equivalence bounds are thus determined primarily based on the sample size in the original study.
+With this so-called 'small telescopes' approach, the equivalence bounds are thus primarily based on the sample size in the original study.
 For example, consider a study in which `r N` participants answered a question, and the results were analyzed with a one-sample t-test.
@@
 
-Another justifiable choice when choosing the SESOI based on earlier work is to use the smallest observed effect size that could have been statistically significant in a previous study.
+Another justifiable option when choosing the SESOI based on earlier work is to use the smallest observed effect size that could have been statistically significant in a previous study.
 In other words, we decide that effects that could not have yielded a $p < \alpha$ in an original study will not be considered meaningful in the replication study either, even if they were found to be statistically significant.
@@
 Third, the SESOI can be determined based on a *resource* question.
-Not all psychological theories make quantifiable predictions (possibly related to the strong reliance on null-hypothesis significance testing in some fields, see [@Meehl1978]).
-Researchers often have more precise ideas of the amount of data that they can afford to collect, that other researchers in their field commonly collect.
+Not all psychological theories make quantifiable predictions [possibly related to the strong reliance on null-hypothesis significance testing in some fields, see @Meehl1978].
+Researchers often have more precise ideas of the amount of data that they can afford to collect, or that other researchers in their field commonly collect.
 The amount of data that is collected limits the inferences you can make.
@@
 Concluding equivalence in a test using equivalence bounds of $\Delta_{L} = `r printnum(-pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ and $\Delta_{U} = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.90,type="one.sample",alternative="two.sided")$d)`$ would suggest that effects as large or larger than those which published studies were sensitive enough to detect can be rejected.
-Such a study does not test a theoretical prediction, but it contributes to the literature by suggesting much larger sample sizes are needed than `r N` observations.
+Such a study does not test a theoretical prediction, but it contributes to the literature by suggesting much larger sample sizes than `r N` observations are needed.
 
@@
 ```
-##Raw versus standardized equivalence bounds
+##Raw Versus Standardized Equivalence Bounds
 The SESOI, and thus the equivalence bounds, can be set in terms of standardized effect sizes (e.g.
@@
 ```
+
 ## Example 1: Not Statistically Equivalent and Not Statistically Different
@@
 @Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity.
-In contrast to the original study, they found a negative difference between the Asian primed group ($n = `r n.asian`$, $M = `r printnum(mean.asian)`$, $SD = `r printnum(sd.asian)`$) and the control ($n = `r n.control`$, $M = `r printnum(mean.control)`$, $SD = `r printnum(sd.control)`$) which was not significant, $d = `r printnum(d)`$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(t)`$, $p = `r printp(p)`$, two-sided)^[Fractional degrees of freedom in t-tests are a result of using Welch’s t-test instead of Student’s t-test, which is the recommended default when sample sizes are unequal between groups [@Delacre2017].
+In contrast to the original study, they found a negative difference between the Asian primed group ($n = `r n.asian`$, $M = `r printnum(mean.asian)`$, $SD = `r printnum(sd.asian)`$) and the control ($n = `r n.control`$, $M = `r printnum(mean.control)`$, $SD = `r printnum(sd.control)`$) which was not significant, $d = `r printnum(d)`$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(t)`$, $p = `r printp(p)`$, two-sided^[Fractional degrees of freedom in *t*-tests are a result of using Welch’s *t*-test instead of Student’s *t*-test, which is the recommended default when sample sizes are unequal between groups [@Delacre2017].
 When using the TOSTER package in R, this is achieved by setting “var.equal = FALSE”.].
@@
 
-In order to distinguish between these possibilities, we can define what a “meaningful effect” would be and use it as the bounds for an equivalence test (remember these bounds should normally be specified before looking at the data).
-If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25% ...
+In order to distinguish between these possibilities, we can define what a “meaningful effect” would be and use it as the bounds for an equivalence test (remember that these bounds should normally be specified before looking at the data).
+If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25%, ...
 A+ = 93.75% to 100%) we could decide that we are only interested in test score differences that correspond to an increase or decrease of at least 1 grade point.
@@
 We can then perform an equivalence test for a two-sample Welch’s *t*-test, with equivalence bounds of $\pm$ the SESOI of `r printnum(sesoi, digits = 4)`.
-The two one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$.
+The two one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for the test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$.
 While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r printnum(sesoi, digits = 4)`, the test against $\Delta_{L}$ shows we can *not* reject effects as extreme or more extreme than `r printnum(-sesoi, digits = 4)`.
-The equivalence test is therefore non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
+The equivalence test is therefore non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the higher *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study.
 After all, much smaller effects may indicate non-replication, but so do large effects in the opposite direction.
@@
 A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($M_{unethical}= `r rep.m1`$, ${SD}_{unethical}= `r rep.sd1`$, $M_{ethical}=`r rep.m2`$, ${SD}_{ethical}=`r rep.sd2`$, $t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$).
-We can chose (before looking at the data) to use the small telescopes approach, calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$), and use this effect size as the equivalence bounds.
+We can choose (before looking at the data) to use the small telescopes approach, calculate the effect size the original study had 33% power to detect ($d = `r printnum(d.33)`$), and use this effect size as the equivalence bounds.
 The two one-sided tests procedure for Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$ reveals that the effect observed in the replication study is statistically equivalent, because the larger of the two p values was less than .05, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2B).
@@
 ```
-@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as absolute effect sizes smaller than $d = `r printnum(sesoi)`$.
+@Hyde2008 reported that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, which the authors specify as absolute effect sizes smaller than $d = `r printnum(sesoi)`$.
 For example, in grade 3 the difference is $d = 0.04$, with a standard error of 0.002.
@@
 They observed that `r printnum(selfish.cold*100)`\% of participants who received a cold pack ($n_{cold}=`r n.cold`$) opted to receive a reward for themselves, while `r printnum(selfish.hot*100)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same.
-In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
+In a *z*-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$).
 Had the authors planned to perform a null-hypothesis significance test *and* an equivalence test, the latter would allow one to distinguish between an inconclusive outcome, or a statistically equivalent result.
@@
 This means that we use the critical *z* value (~1.96 in a two-tailed test with an alpha of 0.05) as equivalence bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$).
-To calculate which difference corresponds to a critical *z* value in the original study, we multiply the critical *z* value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find that the original study could observe a significant effect for differences of $`r printnum(crit.diff)`$ or more.
-
+To calculate which difference corresponds to a critical *z* value in the original study, we multiply the critical *z* value with the standard error ($`r printnum(crit.z)`$ * $`r printnum(se)`$), and find that the original study could have observed a significant effect for differences of $`r printnum(crit.diff)`$ or more.
 Since the original study had a clear directional hypothesis,  in this replication study we are only interested in whether receiving a warm pack *increases* the proportion of people who choose a gift for a friend.
@@
 ```
-Kahane and colleagues [-@Kahane2015] investigate how responses to moral dilemmas in which participants have to decide whether or not they would sacrifice one person's life to save several other lives relate to other indicators of moral orientation.
+Kahane and colleagues [-@Kahane2015] investigated how responses to moral dilemmas in which participants have to decide whether or not they would sacrifice one person's life to save several other lives relate to other indicators of moral orientation.
 Traditionally, greater endorsement for sacrificing one life to save several others has been interpreted as a more "utilitarian" moral orientation (i.e.
@@
 donating the money to save lives in a distant country).
-The authors find no significant correlation between the two set of dilemmas, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$, $N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.].
+The authors find no significant correlation between the two sets of dilemmas, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$, $N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.].
 The authors conclude that “a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely” (p.
@@
 
-Equivalence bounds should be specified before results are known, ideally as part of a pre-registration [cf.
+Equivalence bounds should be specified before results are known, ideally as part of a preregistration [cf.
 @Piaggio2006].
-In the most extreme case, a researcher can always choose equivalence bounds wide enough for the test to yield a 'statistically equivalent' result [for a discussion, see @Weber2012].
+In the most extreme case, a researcher can always first look at the data, and then choose equivalence bounds wide enough for the test to yield a 'statistically equivalent' result [for a discussion, see @Weber2012].
 However, fixed error rates are no longer valid when bounds are determined based on the observed data.
@@
 Psychological theories are usually too vague to derive precise predictions, and if there are no theoretical reference points, natural constraints, or prior studies a researcher can use to define the SESOI for a hypothesis test, any choice will be somewhat arbitrary.
-In some research lines researchers might simply use equivalence tests to reject consecutively smaller effect sizes by performing studies with increasingly larger sample sizes while controlling error rates, until no one is willing to invest the time and resources needed to examine the presence of even smaller effects.
+In some research lines, researchers might use equivalence tests to simply reject consecutively smaller effect sizes by performing studies with increasingly large sample sizes while controlling error rates, until no one is willing to invest the time and resources needed to examine the presence of even smaller effects.
 Nevertheless, it is important to realize that not specifying a SESOI for our research questions at all will severely hinder theoretical progress [@Morey2016].
@@
 ###Data, materials, and online resources
-The code to reproduce these analyses is available at <https://osf.io/qamc6/>^[The OSF project contains all necessary files to reproduce examples 1--5 in R, jamovi [@jamovi2017], and in a spreadsheet; as well as a preprint of this manuscript.
-The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`.].
+The code to reproduce these analyses is available at <https://osf.io/qamc6/>.
+The OSF project contains all necessary files to reproduce examples 1--5 in R, jamovi [@jamovi2017], and in a spreadsheet; as well as a preprint of this manuscript.
+The manuscript, including the figures and statistical analyses, was created using RStudio [1.1.383, @RStudioTeam2016] and `r my_citation`.
   \newline

### 2018-03-01 manuscript_filesforproofing/manuscript_texforproofing_20180301.tex
Made a new folder containing manuscript files for proofing, which contains: - the unchanged PDF - a newly knitted .docx which I manually tweaked (added the textbox which doesn't knit to Word automatically, made sure footnote labels are superscript, etc.) - the two figures as PDF files with changed names (now "Figure_1" and "Figure_2") - the .tex file which I tweaked so that it contains the changed names and paths of the two figures

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Equivalence Testing for Psychological Research: A Tutorial},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Equivalence Testing for Psychological Research: A Tutorial}

  \shorttitle{Equivalence Testing Tutorial}

  \author{Daniel Lakens\textsuperscript{1}, Anne M. Scheel\textsuperscript{1}, \& Peder M. Isager\textsuperscript{1}}

  \def\affdep{{"", "", ""}}%
  \def\affcity{{"", "", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Eindhoven University of Technology  }

 % If no author_note is defined give only author information if available
      \newcounter{author}
                              \authornote{
            Correspondence concerning this article should be addressed to Daniel Lakens, Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands. E-mail: \href{mailto:D.Lakens@tue.nl}{\nolinkurl{D.Lakens@tue.nl}}
          }

  \note{\clearpage}

  \abstract{Psychologists must be able to test both for the presence of an effect
and for the absence of an effect. In addition to testing against zero,
researchers can use the Two One-Sided Tests (TOST) procedure to test for
\emph{equivalence} and reject the presence of a smallest effect size of
interest (SESOI). The two one-sided tests procedure can be used to
determine if an observed effect is surprisingly small, given that a true
effect at least as large as the SESOI exists. We explain a range of
approaches to determine the SESOI in psychological science, and provide
detailed examples of how equivalence tests should be performed and
reported. Equivalence tests are an important extension of statistical
tools psychologists currently use, and enable researchers to falsify
predictions about the presence, and declare the absence, of meaningful
effects.}
  \keywords{Equivalence Testing, null-hypothesis significance test, power, TOST,
falsification. \\

    \indent Word count: 5545
  }

  \usepackage{float}
  \usepackage{framed}
  \usepackage{caption}
  \usepackage{setspace}
  \captionsetup[textbox]{name=Box,labelsep=period,labelfont=it}
  \newfloat{textbox}{thp}{lop}
  \floatname{textbox}{Box}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

Psychologists should be able to falsify predictions. A common prediction
in psychological research is that an effect that differs from zero
exists in the population. For example, we might predict that priming
American Asian women with their Asian identity will increase their
performance on a math test compared to women who are primed with their
female identity. To be able to design a study that allows for strong
inferences (Platt, 1964), it is important to specify which test result
would \emph{falsify} this hypothesis.

Equivalence testing can be used to test whether an observed effect is
surprisingly small, assuming a meaningful effect exists in the
population (see for example Goertzen \& Cribbie, 2010; Meyners, 2012;
Quertemont, 2011; Rogers, Howard, \& Vessey, 1993). The test is a simple
variation of widely used null hypothesis significance tests. To
understand the idea behind equivalence tests, it is useful to realize
the null hypothesis we test against can be any numerical value. When we
compare two groups, we often test whether we can reject that the
difference between these groups is zero (see Figure 1A), but we may
sometimes want to reject other values than zero. Imagine a researcher
who is interested in voluntary participation in a national program to
train young infants' motor skills. The researcher wants to test whether
more boys than girls are brought into the program by their parents.
Because the human sex ratio of newborn boys to girls is not exactly 1:1,
we should not expect 50\% of participants to be boys. On average, 103
boys are born for every 100 girls (United States \& Central Intelligence
Agency, 2016), so approximately 50.74\% of applicants should be boys,
and 49.26\% should be girls. If boys and girls were exactly equally
likely to be brought into the program by their parents, we would not
expect a difference of zero, but 50.74\% - 49.26\% = 1.5\% more boys.
Rather than testing against a null hypothesis of 0 difference, the
researcher tests against a null hypothesis of 0.015.

Alternatively, the researcher could decide that even if the true ratio
in the population is not exactly 0.015, the null hypothesis should
consist of a range of values around the difference in proportions of
0.015 that can be considered trivially small. The researcher could for
example test if the difference is smaller than -0.005, or larger than
0.035. This test against two bounds, with \(H_0\) being a range rather
than one value (see Figure 1B), is known as a \emph{minimal effects
test} (Murphy, Myors, \& Wolach, 2014).

\emph{Equivalence tests} can be seen as the opposite of minimal effects
tests: They examine whether the presence of effects that are large
enough to be considered meaningful can be \emph{rejected} (see Figure
1C). Note that when a meaningful effect can be rejected, it does not
imply there is no effect at all. In our example, the researcher can
perform an equivalence test to examine whether the gender difference in
participants is \emph{not} as extreme or more extreme than the
\emph{smallest effect size of interest} (SESOI). After an extensive
discussion with experts, the researcher decides that as long as the
difference in proportions does not deviate from the population gender
ratio by more than 6\%, the gender difference is too small to care
about. Given an expected true difference in the population of 0.015, the
researcher will test if the observed difference falls outside the
boundary values (or \emph{equivalence bounds}) of -0.055 and 0.075. If
differences more extreme than these boundary values can be rejected in
two one-sided tests (TOST), the researcher will conclude statistical
equivalence, the gender difference will be considered trivially small,
and no money will be spent on addressing a gender difference in
participation.

In any one-sided test, for an alpha level of 5\%, we can reject \(H_0\)
when the 90\% confidence interval around the observed estimate is in the
predicted direction, and does not contain the value tested against
(e.g., 0). In the two one-sided tests procedure, the first one-sided
test is used to test against values smaller than the lower equivalence
bound (\(\Delta_{L}\)), and the second one-sided test is used to test
against values larger than the upper equivalence bound (\(\Delta_{U}\)).
Even though the TOST procedure consists of two one-sided tests, it is
not necessary to control for multiple comparisons because both tests
need to be statistically significant to conclude equivalence.
Consequently, when reporting an equivalence test, it suffices to report
the one-sided test with the smaller test parameter (e.g. \(t\) or \(r\))
and thus the larger \(p\)-value. Statistical equivalence can be
concluded when the largest of the two \(p\)-values is smaller than
alpha. If results are neither statistically different from zero nor
statistically equivalent, there is insufficient data to draw
conclusions. Further studies are needed, which can be analyzed using a
(small-scale) meta-analysis. The additional data will narrow the
confidence interval around the observed effect, allowing us to reject
the null, reject the SESOI, or both. Analogous to the large sample sizes
needed to detect small effects, as the SESOI becomes smaller, the
equivalence bounds become narrower, and the sample size needed to obtain
a sufficiently narrow confidence interval increases.

Note that in this paper we will control the Type 1 error rate at 5\% for
all examples, mirroring the studies we reanalyze (but we recommend to
justify the Type 1 error rate in original research based on substantive
arguments, Lakens et al., 2017). It may be easiest to think of an
equivalence test as checking whether the entire 90\% confidence interval
falls entirely between the upper and lower equivalence bounds, but for
any given study an equivalence test could also be conceptualized as
determining whether effect sizes or test statistics are closer to zero
than some critical value, or even whether the \emph{p}-value for a
null-hypothesis significance test is larger than some \emph{p}-value
bound.

\begin{figure}
\centering
\includegraphics{Figure_1.pdf}
\caption{\label{fig:unnamed-chunk-2}Illustration of null hypotheses
(\(H_0\)) and alternative hypotheses (\(H_1\)) for different types of
significance tests. \textbf{A)} Null-Hypothesis Significance Test: Tests
if the hypothesis (\(H_0\)) that an effect is equal to 0 can be
rejected. \textbf{B)} Minimal effects test: Tests if the hypothesis
(\(H_0\)) that an effect is larger than \(\Delta_{L}\) \emph{and}
smaller than \(\Delta_{U}\) can be rejected. \textbf{C)} Equivalence
test: Tests if the hypothesis (\(H_0\)) that an effect is smaller than
\(\Delta_{L}\) \emph{or} larger than \(\Delta_{U}\) can be rejected.
\textbf{D)} Inferiority test: Tests if the hypothesis (\(H_0\)) that an
effect is larger than \(\Delta\) can be rejected.}
\end{figure}

In this article we provide five reproducible examples of equivalence
tests which illustrate the procedure in free, easy-to-use software (R,
jamovi, and a spreadsheet). But first, we discuss different approaches
to determining the smallest effect size of interest for psychological
research, which determines the statistical question an equivalence test
answers.

\section{Justifying the Smallest Effect Size of
Interest}\label{justifying-the-smallest-effect-size-of-interest}

The Two One-Sided Test procedure is performed against equivalence bounds
that are considered the smallest effect size of interest. The smallest
effect size of interest (SESOI) can sometimes be determined objectively
(for example based on just noticeable differences, as we will explain
below). In lieu of objective justifications, the SESOI should ideally be
based on a cost-benefit analysis. Since both costs and benefits are
necessarily subjective, the SESOI will vary across researchers, fields,
and time. The goal of setting a SESOI is to clearly justify why
designing a study that has a high probability of rejecting effects more
extreme than the specified equivalence bounds contributes to our
knowledge base. The SESOI is thus independent of the outcome of the
study, and should be determined before looking at the data. A SESOI
should be chosen such that inferences based on it answer a meaningful
question. Although we use bounds that are symmetric around 0 for all
equivalence tests in this manuscript (e.g., \(\Delta_{L} = -0.3\),
\(\Delta_{U} = 0.3\)), it is also possible to use asymmetric bounds
(e.g., \(\Delta_{L} = -0.2\), \(\Delta_{U} = 0.3\)).

\subsection{Objective Justification of a Smallest Effect Size of
Interest}\label{objective-justification-of-a-smallest-effect-size-of-interest}

An objectively determined smallest effect size of interest (SESOI)
should be based on quantifiable theoretical predictions, such as
computational models. Sometimes, the only theoretical prediction is that
an effect should be noticeable. In such circumstances, the SESOI can be
set based on just noticeable differences. For example, Burriss and
colleagues (2015) examined whether women displayed an increase in
redness in the face during the fertile phase of their ovulatory cycle.
The hypothesis was that a slightly redder skin signals greater
attractiveness and physical health, and sending this signal to men
yields an evolutionary advantage. This hypothesis presupposes that men
can detect the increase in redness with the naked eye. Burriss and
colleagues collected data from 22 women and showed that there was indeed
an increase in redness of the facial skin during their fertile period.
However, this increase was not large enough for men to detect with the
naked eye, thus falsifying their hypothesis. Because the just noticeable
difference in redness of the skin can be measured, it is possible to
establish an objective SESOI.

Another example of an objectively determined SESOI can be found in
Button et al. (2015), where the minimal clinically important difference
on the Beck Depression Inventory -- II was determined by asking 1039
patients when they subjectively felt less depressed (i.e., when they
personally noticed an improvement) and relating this to the
corresponding difference score on the depression inventory. Similarly,
Norman, Sloan, and Wyrwich (2003) proposed there is a surprisingly
consistent minimally important difference of half a standard deviation,
or \(d = 0.5\), in health outcomes.

\subsection{Subjective Justification of a Smallest Effect Size of
Interest}\label{subjective-justification-of-a-smallest-effect-size-of-interest}

We distinguish between three categories of subjective justifications to
determine the smallest effect size of interest. First, researchers can
use benchmarks. For example, one might set the smallest effect size of
interest to a standardized effect size of \(d = 0.5\), which would allow
rejecting effects more extreme than a \enquote{medium} effect size
(Cohen, 1988). Similarly, effect sizes smaller than a Cohen's \(d\) of
0.1 are sometimes considered trivially small (Maxwell, Lau, \& Howard,
2015). Relying on a benchmark is the weakest possible justification of a
smallest effect size of interest, and should be avoided. Based on a
review of 112 meta-analyses, Weber and Popova (2012) conclude that
setting a smallest effect size of interest (SESOI) to a medium effect
size (\(r = 0.3\), or \(d = 0.5\)) corresponds to rejecting only the
upper 25\% of effect sizes reported in communications research, and
Hemphill (2003) suggests that a smallest effect size of interest of
\(d = 0.5\) would imply rejecting effects as large as the upper 33\% of
effect sizes reported in the psychological literature.

Second, the SESOI can be based on related published studies in the
literature. Ideally, researchers who publish novel research would always
specify their smallest effect size of interest, but this is not yet
common practice. It is thus up to researchers who build on earlier work
to decide which effect size is too small to be meaningful when examining
the same hypothesis. Simonsohn (2015) recently proposed to set the SESOI
to the effect size that an earlier study would have had 33\% power to
detect. With this so-called \enquote{small telescopes} approach, the
equivalence bounds are thus primarily based on the sample size in the
original study. For example, consider a study in which 100 participants
answered a question, and the results were analyzed with a one-sample
t-test. For a two-sided test with an alpha of 0.05, this test had 33\%
power to detect an effect of \(d = 0.15\). Another example of how
previous research can be used to determine the SESOI can be found in
Kordsmeyer and Penke (2017), who defined the SESOI based on the mean of
effect sizes reported in the literature. Thus, they examined whether
their replication study could reject effects as large or larger than had
on average been reported in the literature. Given random variation and
bias in the literature, a more conservative approach could be to use the
lower end of a confidence interval around the meta-analytic effect size
estimate (cf. Perugini, Gallucci, \& Costantini, 2014).

Another justifiable option when choosing the SESOI based on earlier work
is to use the smallest observed effect size that could have been
statistically significant in a previous study. In other words, we decide
that effects that could not have yielded a \(p < \alpha\) in an original
study will not be considered meaningful in the replication study either,
even if they were found to be statistically significant. The assumption
here is that authors were interested in observing a significant effect,
and thus were not interested in observed effect sizes that could not
have yielded a significant effect. It might be likely that the original
authors did not consider which effect sizes they could detect at all, or
that they were interested in smaller effects, but gambled on observing
an especially large effect in the sample purely due to random variation.
Even then, observed effect sizes that could not have been significant in
an earlier study might be a justifiable starting point when building on
earlier research that did not specify a smallest effect size of
interest. Based only on the alpha level and the sample size, we can
calculate the critical test value (e.g., \emph{t}, \emph{F}, \emph{Z}).
This critical test value can be transformed to a standardized effect
size (e.g.,
\(d = t \sqrt { \frac { 1} { n _ { 1} } + \frac { 1} { n _ { 2} } }\)),
which can thus be interpreted as the \emph{critical effect
size}.\footnote{This will typically, although not always, correspond to
  the effect size the study had 50\% power to detect (see Lenth, 2007).
  This procedure will result in equivalence bounds that are wider than
  the ones obtained using the small telescopes approach, which gives the
  effect size a study had 33\% power to detect.} All observed effect
sizes smaller than the critical effect size would not have been
statistically significant in the original study, given the alpha and
sample size of that study. By setting the SESOI to the critical effect
size, an equivalence test can reject all observed effect sizes that
could have been detected in an earlier study.

Third, the SESOI can be determined based on a \emph{resource} question.
Not all psychological theories make quantifiable predictions (possibly
related to the strong reliance on null-hypothesis significance testing
in some fields, see Meehl, 1978). Researchers often have more precise
ideas of the amount of data that they can afford to collect, or that
other researchers in their field commonly collect. The amount of data
that is collected limits the inferences you can make. Given the alpha
level and the planned sample size, researchers can calculate the
smallest effect size that they have sufficient power to
detect.\footnote{This approach is conceptually very similar to the
  \emph{power approach}, where the effect size you had 95\% power to
  detect is calculated, and the presence of effects more extreme than
  this value is rejected after observing a \(p\)-value \emph{larger
  than} 0.05 in a traditional null-hypothesis significance test.
  However, Meyners (2012) explains that this approach is not recommended
  (even though it is common) because it ignores the possibility that
  effects are significant \emph{and} equivalent, and error rates are not
  controlled accurately.}

An equivalence test based on this approach does not answer any
theoretical question (after all, the equivalence bounds are not based on
any theoretical predictions). When theories only allow for directional
predictions, but do not predict effects of a specific size, the sample
size justification can be used to determine which effects can be studied
reliably, and thus which effects would be interesting to reject. For
example, imagine a research line where almost all studies examined a
hypothesis by performing a one-sample \(t\)-test on sample sizes smaller
than 100. A one-sample \(t\)-test using an alpha of 5\% (two-sided) with
100 observations has 90\% power to detect an effect of \(d = 0.33\).
Concluding equivalence in a test using equivalence bounds of
\(\Delta_{L} = -0.33\) and \(\Delta_{U} = 0.33\) would suggest that
effects as large or larger than those which published studies were
sensitive enough to detect can be rejected. Such a study does not test a
theoretical prediction, but it contributes to the literature by
suggesting much larger sample sizes than 100 observations are needed.

When there is no previous literature on a topic, researchers can justify
their sample size based on reasonable resource limitations, which are
specific to scientific disciplines and research questions (and can be
expected to change over time). Where 22 patients with a rare
neurological disorder might be the largest sample a single researcher
can collect, a study on mTurk might easily collect hundreds or thousands
of observations. Thus, whether or not the resource question that is
answered by an equivalence test is interesting must be evaluated by
peers, preferably when the study design is submitted to an ethical
review board or as a Registered Report.

Additional subjective justifications for a smallest effect size of
interest are possible. For example, the Food and Drug Administration has
set equivalence bounds for bioequivalence studies, taking the decision
out of the hands of individual researchers (Senn, 2007). We have
provided several approaches to help researchers justify equivalence
bounds, but want to repeat the warning by Rogers et al. (1993), p.~564:
\enquote{The thoughtless application of \enquote{cookbook} prescriptions
is ill-advised, regardless of whether the goal is to establish a
difference or to establish equivalency between treatments}. By
transparently reporting and adequately justifying the smallest effect
size of interest, researchers can communicate the information their
study contributes to the literature, and provide a starting point for a
discussion about what a reasonable SESOI may be.

\subsection{Raw Versus Standardized Equivalence
Bounds}\label{raw-versus-standardized-equivalence-bounds}

The SESOI, and thus the equivalence bounds, can be set in terms of
standardized effect sizes (e.g.~a Cohen's \(d\) of 0.5) or as a raw mean
difference (e.g.~0.5 points on a 7-point scale). The key difference is
that equivalence bounds set in raw differences are independent of the
standard deviation, while equivalence bounds set as standardized effects
are dependent on the standard deviation (since they are calculated as
\(\frac {X _ { 1} - X _ { 2} } {S D}\)). The observed standard deviation
randomly varies across samples. In practice, this means that when using
standardized differences as bounds (e.g., \(d = 0.35\)), the equivalence
test depends on the standard deviation in the sample. The equivalence
bound for a raw mean difference of 0.5 equals a standardized equivalence
bound of \(d = 0.5\) when the standard deviation is 1, but a
standardized equivalence bound of \(d = 1\) when the standard deviation
is 0.5.

Both raw equivalence bounds and standardized equivalence bounds have
specific benefits and limitations (for a discussion, see Baguley, 2009).
When raw mean differences are meaningful and of theoretical interest, it
makes sense to set equivalence bounds based on raw effect sizes. When
the raw mean difference is of less theoretical importance, or different
measures are used across research lines, it is often easier to set
equivalence bounds based on standardized differences. Researchers should
realize that equivalence bounds based on raw differences or standardized
differences ask slightly different questions, and justify their choice
for either option. When setting equivalence bounds based on earlier
research, such as in replication studies, equivalence bounds based on
raw or standardized differences would ideally give the same result, and
large differences in standard deviations between studies are as
important to interpret as large differences in means.

In the remainder of this article we will provide five detailed examples
of equivalence tests performed on published studies. These concrete and
easy to follow examples will illustrate all approaches to setting
equivalence bounds discussed above, and demonstrate how to perform and
report equivalence tests.

\begin{figure}
\centering
\includegraphics{Figure_2.pdf}
\caption{\label{fig:unnamed-chunk-6}Example effects plotted with 90\% two
one-sided tests confidence intervals (thick lines) and 95\%
null-hypothesis significance test confidence intervals (thin lines), the
null hypothesis (solid vertical line) and the equivalence bounds (dashed
vertical lines) displayed. \textbf{A)} Example 1 - Mean Difference.
\textbf{B)} Example 2 - Mean Difference. \textbf{C)} Example 3 -
Meta-analytic Effect Size. \textbf{D)} Example 4 - Proportion
Difference. \textbf{E)} Example 5 - Pearson Correlation.}
\end{figure}

\subsection{Example 1: Not Statistically Equivalent and Not
Statistically
Different}\label{example-1-not-statistically-equivalent-and-not-statistically-different}

Moon and Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999),
conducted a study to investigate whether Asian-American women would
perform better on a maths test when primed with their Asian identity. In
contrast to the original study, they found a negative difference between
the Asian primed group (\(n = 53\), \(M = 0.46\), \(SD = 0.17\)) and the
control (\(n = 48\), \(M = 0.50\), \(SD = 0.18\)) which was not
significant, \(d = -0.21\), \(t(97.77) = -1.08\), \(p = .284\),
two-sided\footnote{Fractional degrees of freedom in \emph{t}-tests are a
  result of using Welch's \emph{t}-test instead of Student's
  \emph{t}-test, which is the recommended default when sample sizes are
  unequal between groups (Delacre, Lakens, \& Leys, 2017). When using
  the TOSTER package in R, this is achieved by setting
  \enquote{var.equal = FALSE}.}. The non-significant null-hypothesis
test does not allow us to distinguish between the absence of a
meaningful effect and data that are not sensitive enough to tell us
whether a meaningful effect is present or absent.

In order to distinguish between these possibilities, we can define what
a \enquote{meaningful effect} would be and use it as the bounds for an
equivalence test (remember that these bounds should normally be
specified before looking at the data). If grades for this test were set
at every 6.25\% increase in correct answers (F = 0\% to 6.25\%, \ldots{}
A+ = 93.75\% to 100\%) we could decide that we are only interested in
test score differences that correspond to an increase or decrease of at
least 1 grade point. Thus, our SESOI becomes a difference in raw scores
of 6.25\%, or 0.0625. We can then perform an equivalence test for a
two-sample Welch's \emph{t}-test, with equivalence bounds of \(\pm\) the
SESOI of 0.0625. The two one-sided-tests procedure consists of two
one-sided tests, and we calculate \emph{t}-values and \emph{p}-values
both for the test against \(\Delta_{L}\), \(t(97.77) = 0.71\),
\(p = .241\), and \(\Delta_{U}\), \(t(97.77) = -2.86\), \(p = .003\).
While the \emph{t}-test against \(\Delta_{U}\) indicates that we can
reject differences at least as extreme as 0.0625, the test against
\(\Delta_{L}\) shows we can \emph{not} reject effects as extreme or more
extreme than -0.0625. The equivalence test is therefore non-significant,
which means we cannot reject a true effect as large or larger than
6.25\% (Figure 2A), \(t(97.77) = 0.71\), \(p = .241\) (note we typically
only report the one-sided test yielding the higher \emph{p}-value in the
result section).\footnote{Because this is a replication of a study, it
  would have been reasonable to assume that we only want to reject
  effects in the same direction as the effect in the original study.
  After all, much smaller effects may indicate non-replication, but so
  do large effects in the opposite direction. In that case, we would
  perform an inferiority test (see figure 1D) against \(\Delta_{U}\). As
  can be seen in Figure 2A, the 90\% CI does not overlap with
  \(\Delta_{U}\), so we can reject effects as large or larger than
  \(0.0625\). However, this must be decided before running the tests to
  prevent loss of error control.}

\subsection{Example 2: Statistically Equivalent and Not Statistically
Different}\label{example-2-statistically-equivalent-and-not-statistically-different}

Banerjee, Chatterjee, and Sinha (2012) reported that 40 participants who
had been asked to describe an unethical deed from their past judged the
room to be darker than participants who had been asked to describe an
ethical deed (\(t(38)= 2.03\), \(p= .049\), \(d= 0.65\)). A close
replication by Brandt, IJzerman, and Blanken (2014) with 100
participants found no significant effect (\(M_{unethical}= 4.79\),
\({SD}_{unethical}= 1.09\), \(M_{ethical}=4.66\),
\({SD}_{ethical}=1.19\), \(t(97.78)=0.57\), \(p=.573\), \(d=0.11\)). We
can choose (before looking at the data) to use the small telescopes
approach, calculate the effect size the original study had 33\% power to
detect (\(d = 0.49\)), and use this effect size as the equivalence
bounds. The two one-sided tests procedure for Welch's \emph{t}-test for
independent samples and equivalence bounds of \({\Delta}_L=-0.49\) and
\({\Delta}_U=0.49\) reveals that the effect observed in the replication
study is statistically equivalent, because the larger of the two p
values was less than .05, \(t(97.78)=-1.90\), \(p=.030\) (Figure 2B).
Following a Neyman-Pearson approach, this means we can reject the
hypothesis that the true effect is smaller than \(d = -0.49\) or larger
than \(d = 0.49\), and we can act as if the effect size falls within the
equivalence bounds, with the understanding that based on our chosen
alpha level, we will not be wrong more than 5\% of the time.

\subsection{Example 3: Statistically Equivalent and Statistically
Different}\label{example-3-statistically-equivalent-and-statistically-different}

Hyde, Lindberg, Linn, Ellis, and Williams (2008) reported that effect
sizes for gender differences in mathematics tests across the 7 million
students in the US represent trivial differences, which the authors
specify as absolute effect sizes smaller than \(d = 0.10\). For example,
in grade 3 the difference is \(d = 0.04\), with a standard error of
0.002. When we perform equivalence tests on the meta-analytic effect
sizes of IQ difference for grades 2 to 11 (using an alpha level of .005
to correct for multiple comparisons) and using equivalence bounds of
\(\Delta_{L} = -0.1\) and \(\Delta_{U} = 0.1\), we see that all effect
size estimates are measured with such high precision that they are
statistically equivalent, and can be considered trivially small
according to the authors' definition (Figure 2C). In the case of grade
3, \(z = 70.00\), \(p < .001\). However, note that all of the effects
are also statistically different from zero, as one might expect when
there is no random assignment to conditions and samples sizes are very
large (e.g., for grade 3, \(z = 20.00\), \(p < .001\)). We see how
equivalence tests allow researchers to distinguish between
\emph{statistical} significance and \emph{practical} significance, which
demonstrates how equivalence tests improve hypothesis testing procedures
in psychological research.

\subsection{Example 4: Statistically Inferior and Not Statistically
Different}\label{example-4-statistically-inferior-and-not-statistically-different}

Lynott et al. (2014) conducted a study to investigate the effect of
being exposed to physically warm or cold stimuli on subsequent judgments
related to interpersonal warmth and prosocial behavior (replicating L.
E. Williams \& Bargh, 2008). They observed that 50.74\% of participants
who received a cold pack (\(n_{cold}=404\)) opted to receive a reward
for themselves, while 57.46\% of participants who received a warm pack
(\(n_{hot}=409\)) did the same. In a \emph{z}-test for the difference
between the proportions, this effect is not statistically significant
(\(\mathit{Diff}=-6.71\)\%, \(z=-1.93\), \(p=.054\)). Had the authors
planned to perform a null-hypothesis significance test \emph{and} an
equivalence test, the latter would allow one to distinguish between an
inconclusive outcome, or a statistically equivalent result.

Since this is a replication, we could decide in advance that we are
interested in whether the observed effect size is smaller than the
smallest effect size that the original study could have detected. This
means that we use the critical \emph{z} value (\textasciitilde{}1.96 in
a two-tailed test with an alpha of 0.05) as equivalence bounds
(\(\Delta = \pm 1.96\)). To calculate which difference corresponds to a
critical \emph{z} value in the original study, we multiply the critical
\emph{z} value with the standard error (\(1.96\) * \(0.13\)), and find
that the original study could have observed a significant effect for
differences of \(0.25\) or more. Since the original study had a clear
directional hypothesis, in this replication study we are only interested
in whether receiving a warm pack \emph{increases} the proportion of
people who choose a gift for a friend. This means we can test for
inferiority (see Figure 1D) and conclude the absence of a meaningful
effect if the observed effect size is reliably \emph{lower} than the
SESOI. We find that we can reject effects larger than
\(\Delta_{U} = 0.25\), \(z = -9.12\), \(p < .001\) (see Figure 2D).
Thus, we can conclude that the statistically non-significant effect is
also statistically smaller than our SESOI.

\subsection{Example 5: Statistically Equivalent and not Statistically
Different}\label{example-5-statistically-equivalent-and-not-statistically-different}

Kahane and colleagues (2015) investigated how responses to moral
dilemmas in which participants have to decide whether or not they would
sacrifice one person's life to save several other lives relate to other
indicators of moral orientation. Traditionally, greater endorsement for
sacrificing one life to save several others has been interpreted as a
more \enquote{utilitarian} moral orientation (i.e.~a stronger concern
for the greater good). Kahane et al. contest this interpretation in a
number of studies. In Study 4, they compare the traditionally used
dilemmas with a set of new dilemmas in which the sacrifice for the
greater good is not another person's life, but something the participant
would have a partial interest in (e.g.~buying a new mobile phone
vs.~donating the money to save lives in a distant country). The authors
find no significant correlation between the two sets of dilemmas,
\(r(229) = -0.04\), \(p = .525\), \(N = 231\)\footnote{Study 4 in Kahane
  et al. (2015) had a final sample size of \(N = 232\), but due to
  missing data in 1 case, the correlation reported here is based on a
  sample of \(N = 231\).}. The authors conclude that \enquote{a robust
association between \enquote{utilitarian} judgment and genuine concern
for the greater good seems extremely unlikely} (p.~206), given the
statistical power their study had to detect meaningful effect sizes.

This interpretation can be formalized by performing an equivalence test
for correlations, where equivalence bounds are set to an effect size the
study had reasonable power to detect (as decided before looking at the
data). With 231 participants, the study had 80\% power to detect effects
of \(r = 0.18\). Performing the two one-sided tests procedure given
bounds of \(\Delta_{L} = -0.18\) and \(\Delta_{U} = 0.18\), the data is
statistically equivalent, \(r(229) = -0.04\), \(p = .015\). This means
that values smaller than \(r = -0.18\) and larger than \(r = 0.18\) can
be rejected at an alpha level of 5\%. If other researchers are
interested in examining the presence of a smaller effect size, they can
design studies with a larger sample size.

\begin{textbox}
\caption{Calculating an equivalence test in R.}
\begin{framed}
{\sffamily \small \setstretch{1}
Equivalence tests can be performed  based on summary statistics (e.g., means, standard deviations, and sample sizes) with the ``TOSTER'' package in the open-source programming language R. Using TOSTER in R requires no more programming experience than reproducing three lines of code.
\vspace{2mm}

The code below reproduces the result of Example 2 in R, which can be typed into R or RStudio. The parameters of the test are defined inside the parentheses. Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code. Results and a plot will automatically be printed. A help file provides more detailed information.

\begin{verbatim}
install.packages("TOSTER")

library(TOSTER)

TOSTtwo(m1 = 4.7857, m2 = 4.6569, sd1 = 1.0897, sd2 = 1.1895, n1 = 49,
n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05,
var.equal = FALSE)
\end{verbatim}
}
\end{framed}
\end{textbox}

\section{Discussion}\label{discussion}

Equivalence testing is a simple statistical technique to reject the
presence of a smallest effect size of interest. As long as we can
calculate a confidence interval around a parameter estimate, we can
compare it to the SESOI. The result of an equivalence test can be
obtained by mere visual inspection of the CI (Seaman \& Serlin, 1998;
Tryon, 2001), or by performing two one-sided tests (known as the TOST
procedure).

As with any statistical test, the usefulness of the result of an
equivalence test depends on the question we ask. The question manifests
itself in the bounds we set: Is our data surprisingly close to zero,
assuming a true effect exists that is as large as our smallest effect
size of interest? If we test against very wide bounds, observing
statistical equivalence can hardly be considered surprising, given that
most effects in psychology are small to medium (Hemphill, 2003).
Examining papers citing Lakens (2017a), we found that some researchers
state, but do not justify, the SESOI used in the equivalence test (e.g.,
Brown, Rodriguez, Gretak, \& Berry, 2017; Schumann, Klein, Douglas, \&
Hewstone, 2017). An equivalence test using a SESOI of \(d = 0.5\) might
very well answer a question the researchers are interested in (for one
possible justification based on minimally important differences, see
Norman et al., 2003), but researchers should always explain why they
chose a SESOI. It makes little sense to report a statistical test
without explaining why one would want to answer this question to begin
with.

Equivalence bounds should be specified before results are known, ideally
as part of a preregistration (cf. Piaggio et al., 2006). In the most
extreme case, a researcher can always first look at the data, and then
choose equivalence bounds wide enough for the test to yield a
\enquote{statistically equivalent} result (for a discussion, see Weber
\& Popova, 2012). However, fixed error rates are no longer valid when
bounds are determined based on the observed data. The value of reporting
an equivalence test is determined by the strength of the justification
of the equivalence bounds. If bounds are chosen based on the observed
data, an equivalence test becomes meaningless. We have proposed several
ways to justify equivalence bounds, but in the end these discussions
must happen among peers, and the best place for these discussions is
during peer-review of proposals for Registered Reports.

As with null-hypothesis significance tests, equivalence tests
interpreted from a Neyman-Pearson perspective on statistical inferences
are used to guide the actions of researchers while controlling error
rates. Research lines sometimes require dichotomous choices. For
example, a researcher might want to decide to discontinue a research
line when an effect is too small to be considered meaningful.
Equivalence tests based on carefully chosen equivalence bounds and error
rates can be used as a rule to govern our behavior (Neyman \& Pearson,
1933). An equivalence test and a null-hypothesis significance test
examine two different hypotheses, and can therefore be used in concert
(Weber \& Popova, 2012). We recommend that researchers by default
perform both a null hypothesis significance test \emph{and} an
equivalence test on their data, as long as they can justify a smallest
effect size of interest, in order to improve the falsifiability of
predictions in psychological science.

The biggest challenge for researchers will be to specify the smallest
effect size of interest. Psychological theories are usually too vague to
derive precise predictions, and if there are no theoretical reference
points, natural constraints, or prior studies a researcher can use to
define the SESOI for a hypothesis test, any choice will be somewhat
arbitrary. In some research lines, researchers might use equivalence
tests to simply reject consecutively smaller effect sizes by performing
studies with increasingly large sample sizes while controlling error
rates, until no one is willing to invest the time and resources needed
to examine the presence of even smaller effects. Nevertheless, it is
important to realize that not specifying a SESOI for our research
questions at all will severely hinder theoretical progress (Morey \&
Lakens, 2016). Incorporating equivalence tests in our statistical
toolbox will in time contribute to better --- and more falsifiable ---
theories.

\newpage

\subsection{Disclosures}\label{disclosures}

\subsubsection{Data, materials, and online
resources}\label{data-materials-and-online-resources}

The code to reproduce these analyses is available at
\url{https://osf.io/qamc6/}. The OSF project contains all necessary
files to reproduce examples 1--5 in R, jamovi (jamovi project, 2017),
and in a spreadsheet; as well as a preprint of this manuscript. The
manuscript, including the figures and statistical analyses, was created
using RStudio (1.1.383, RStudio Team, 2016) and R (3.4.2, R Core Team,
2017) and the R-packages \emph{bookdown} (0.5, Xie, 2016),
\emph{ggplot2} (2.2.1, Wickham, 2009), \emph{gridExtra} (2.3, Auguie,
2017), \emph{knitr} (1.17, Xie, 2015), \emph{lattice} (0.20.35, Sarkar,
2008), \emph{papaja} (0.1.0.9492, Aust \& Barth, 2017), \emph{purrr}
(0.2.4, Henry \& Wickham, 2017), \emph{pwr} (1.2.1, Champely, 2017),
\emph{rmarkdown} (1.7, Allaire et al., 2017), and \emph{TOSTER} (0.3,
Lakens, 2017b). \newline

\subsubsection{Conflicts of Interest}\label{conflicts-of-interest}

The authors declare that they have no conflicts of interest with respect
to the authorship or the publication of this article. \newline

\subsubsection{Author Contributions}\label{author-contributions}

DL: Article conceptualization, programming TOSTER package and
spreadsheet, main contributor to the introduction section and discussion
section. AMS and PMI: Main contributors to examples and figures. All
authors contributed to manuscript conceptualization, editing, review for
submission, and additional software development. AMS and PMI contributed
equally to the manuscript, and order was determined by executing the
following commands in R: \newline
 \texttt{set.seed(7998976/5271)} \newline
 \texttt{x\ \textless{}-\ sample(c("Anne",\ "Peder"),\ 1)} \newline
 \texttt{print(paste("The\ winner\ is",\ x,\ "!"))} \newline

\subsubsection{Acknowledgments}\label{acknowledgments}

This work was funded by VIDI grant 452-17-013. We would like to thank
Courtney Soderberg for creating the first version of the TOST function
to test two independent proportions, Dermot Lynott and Katie Corker for
their helpful responses to our inquiries about Lynott et al. (2014), Jim
Everett and Brian Earp for their helpful responses and for providing the
raw data for study 4 in Kahane et al. (2015), and Neil McLatchie and Jan
Vanhove for valuable comments on an earlier version of this manuscript.
\newline

\subsubsection{Prior versions}\label{prior-versions}

A preprint version of this manuscript prior to peer-review can be found
on PsyArxiv: \url{https://psyarxiv.com/v3zkt/} (version 1). We have
updated our recommendations on how to justify the SESOI based on
available resources. The updated preprint version (version 2) is
identical in content to the published manuscript.

\newpage

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-R-rmarkdown}{}
Allaire, J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., \ldots{} Chang, W. (2017). \emph{rmarkdown: Dynamic documents for
R}. Retrieved from \url{https://CRAN.R-project.org/package=rmarkdown}

\hypertarget{ref-R-gridExtra}{}
Auguie, B. (2017). \emph{gridExtra: Miscellaneous functions for ``grid''
graphics}. Retrieved from
\url{https://CRAN.R-project.org/package=gridExtra}

\hypertarget{ref-R-papaja}{}
Aust, F., \& Barth, M. (2017). \emph{papaja: Create APA manuscripts with
R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-Baguley2009}{}
Baguley, T. (2009). Standardized or simple effect size: What should be
reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617.
doi:\href{https://doi.org/10.1348/000712608X377117}{10.1348/000712608X377117}

\hypertarget{ref-Banerjee2012}{}
Banerjee, P., Chatterjee, P., \& Sinha, J. (2012). Is It Light or Dark?
Recalling Moral Behavior Changes Perception of Brightness.
\emph{Psychological Science}, \emph{23}(4), 407--409.
doi:\href{https://doi.org/10.1177/0956797611432497}{10.1177/0956797611432497}

\hypertarget{ref-Brandt2014}{}
Brandt, M. J., IJzerman, H., \& Blanken, I. (2014). Does Recalling Moral
Behavior Change the Perception of Brightness?: A Replication and
Meta-Analysis of Banerjee, Chatterjee, and Sinha (2012). \emph{Social
Psychology}, \emph{45}(3), 246--252.
doi:\href{https://doi.org/10.1027/1864-9335/a000191}{10.1027/1864-9335/a000191}

\hypertarget{ref-Brown2017}{}
Brown, M., Rodriguez, D. N., Gretak, A. P., \& Berry, M. A. (2017).
Preliminary Evidence for How the Behavioral Immune System Predicts Juror
Decision-Making. \emph{Evolutionary Psychological Science}, \emph{3}(4),
325--334.
doi:\href{https://doi.org/10.1007/s40806-017-0102-z}{10.1007/s40806-017-0102-z}

\hypertarget{ref-Burriss2015}{}
Burriss, R. P., Troscianko, J., Lovell, P. G., Fulford, A. J. C.,
Stevens, M., Quigley, R., \ldots{} Rowland, H. M. (2015). Changes in
Women's Facial Skin Color over the Ovulatory Cycle are Not Detectable by
the Human Visual System. \emph{PLOS ONE}, \emph{10}(7), e0130093.
doi:\href{https://doi.org/10.1371/journal.pone.0130093}{10.1371/journal.pone.0130093}

\hypertarget{ref-Button2015}{}
Button, K. S., Kounali, D., Thomas, L., Wiles, N. J., Peters, T. J.,
Welton, N. J., \ldots{} Lewis, G. (2015). Minimal clinically important
difference on the Beck Depression Inventory - II according to the
patient's perspective. \emph{Psychological Medicine}, \emph{45}(15),
3269--3279.
doi:\href{https://doi.org/10.1017/S0033291715001270}{10.1017/S0033291715001270}

\hypertarget{ref-R-pwr}{}
Champely, S. (2017). \emph{pwr: Basic functions for power analysis}.
Retrieved from \url{https://CRAN.R-project.org/package=pwr}

\hypertarget{ref-Cohen1988}{}
Cohen, J. (1988). \emph{Statistical power analysis for the behavioral
sciences} (2nd ed.). Hillsdale, N.J: L. Erlbaum Associates.

\hypertarget{ref-Delacre2017}{}
Delacre, M., Lakens, D., \& Leys, C. (2017). Why Psychologists Should by
Default Use Welch's \emph{t}-test Instead of Student's \emph{t}-test.
\emph{International Review of Social Psychology}, \emph{30}(1), 92.
doi:\href{https://doi.org/10.5334/irsp.82}{10.5334/irsp.82}

\hypertarget{ref-Goertzen2010}{}
Goertzen, J. R., \& Cribbie, R. A. (2010). Detecting a lack of
association: An equivalence testing approach. \emph{British Journal of
Mathematical and Statistical Psychology}, \emph{63}(3), 527--537.
doi:\href{https://doi.org/10.1348/000711009X475853}{10.1348/000711009X475853}

\hypertarget{ref-Hemphill2003}{}
Hemphill, J. F. (2003). Interpreting the magnitudes of correlation
coefficients. \emph{American Psychologist}, \emph{58}(1), 78--80.

\hypertarget{ref-R-purrr}{}
Henry, L., \& Wickham, H. (2017). \emph{purrr: Functional programming
tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}

\hypertarget{ref-Hyde2008}{}
Hyde, J. S., Lindberg, S. M., Linn, M. C., Ellis, A. B., \& Williams, C.
C. (2008). Gender Similarities Characterize Math Performance.
\emph{Science}, \emph{321}(5888), 494--495.
doi:\href{https://doi.org/10.1126/science.1160364}{10.1126/science.1160364}

\hypertarget{ref-jamovi2017}{}
jamovi project. (2017). Jamovi (Version 0.8) {[}Computer Software{]}.
Retrieved from https://www.jamovi.org.

\hypertarget{ref-Kahane2015}{}
Kahane, G., Everett, J. A., Earp, B. D., Farias, M., \& Savulescu, J.
(2015). ``Utilitarian'' judgments in sacrificial moral dilemmas do not
reflect impartial concern for the greater good. \emph{Cognition},
\emph{134}, 193--209.
doi:\href{https://doi.org/10.1016/j.cognition.2014.10.005}{10.1016/j.cognition.2014.10.005}

\hypertarget{ref-Kordsmeyer2017}{}
Kordsmeyer, T. L., \& Penke, L. (2017). The association of three
indicators of developmental instability with mating success in humans.
\emph{Evolution and Human Behavior}, \emph{38}(6), 704--713.
doi:\href{https://doi.org/10.1016/j.evolhumbehav.2017.08.002}{10.1016/j.evolhumbehav.2017.08.002}

\hypertarget{ref-Lakens2017a}{}
Lakens, D. (2017a). Equivalence Tests: A Practical Primer for t Tests,
Correlations, and Meta-Analyses. \emph{Social Psychological and
Personality Science}, \emph{8}(4), 355--362.
doi:\href{https://doi.org/10.1177/1948550617697177}{10.1177/1948550617697177}

\hypertarget{ref-R-TOSTER}{}
Lakens, D. (2017b). \emph{TOSTER: Two one-sided tests (TOST) equivalence
testing}. Retrieved from \url{https://CRAN.R-project.org/package=TOSTER}

\hypertarget{ref-Lakens2017b}{}
Lakens, D., Adolfi, F. G., Albers, C., Anvari, F., Apps, M. A. J.,
Argamon, S. E., \ldots{} Zwaan, R. (2017). Justify Your Alpha: A
Response to ``Redefine Statistical Significance''. \emph{PsyArXiv}.
doi:\href{https://doi.org/10.17605/OSF.IO/9S3Y6}{10.17605/OSF.IO/9S3Y6}

\hypertarget{ref-Lenth2007}{}
Lenth, R. V. (2007). Post hoc power: Tables and commentary. \emph{Iowa
City: Department of Statistics and Actuarial Science, University of
Iowa}.

\hypertarget{ref-Lynott2014}{}
Lynott, D., Corker, K. S., Wortman, J., Connell, L., Donnellan, M. B.,
Lucas, R. E., \& O'Brien, K. (2014). Replication of ``Experiencing
Physical Warmth Promotes Interpersonal Warmth'' by Williams and Bargh
(2008). \emph{Social Psychology}, \emph{45}(3), 216--222.
doi:\href{https://doi.org/10.1027/1864-9335/a000187}{10.1027/1864-9335/a000187}

\hypertarget{ref-Maxwell2015}{}
Maxwell, S. E., Lau, M. Y., \& Howard, G. S. (2015). Is psychology
suffering from a replication crisis? What does ``failure to replicate''
really mean? \emph{American Psychologist}, \emph{70}(6), 487--498.
doi:\href{https://doi.org/10.1037/a0039400}{10.1037/a0039400}

\hypertarget{ref-Meehl1978}{}
Meehl, P. E. (1978). Theoretical risks and tabular asterisks: Sir Karl,
Sir Ronald, and the slow progress of soft psychology. \emph{Journal of
Consulting and Clinical Psychology}, \emph{46}(4), 806--834.
doi:\href{https://doi.org/10.1037/0022-006X.46.4.806}{10.1037/0022-006X.46.4.806}

\hypertarget{ref-Meyners2012}{}
Meyners, M. (2012). Equivalence tests review. \emph{Food Quality and
Preference}, \emph{26}(2), 231--245.
doi:\href{https://doi.org/10.1016/j.foodqual.2012.05.003}{10.1016/j.foodqual.2012.05.003}

\hypertarget{ref-Moon2014}{}
Moon, A., \& Roeder, S. S. (2014). A Secondary Replication Attempt of
Stereotype Susceptibility (Shih, Pittinsky, \& Ambady, 1999).
\emph{Social Psychology}, \emph{45}(3), 199--201.
doi:\href{https://doi.org/10.1027/1864-9335/a000193}{10.1027/1864-9335/a000193}

\hypertarget{ref-Morey2016}{}
Morey, R., \& Lakens, D. (2016). \emph{Why Most Of Psychology Is
Statistically Unfalsifiable}. Zenodo.
doi:\href{https://doi.org/10.5281/zenodo.838685}{10.5281/zenodo.838685}

\hypertarget{ref-Murphy2014}{}
Murphy, K. R., Myors, B., \& Wolach, A. (2014). \emph{Statistical Power
Analysis: A Simple and General Model for Traditional and Modern
Hypothesis Tests, Fourth Edition}. Routledge.

\hypertarget{ref-Neyman1933}{}
Neyman, J., \& Pearson, E. S. (1933). On the Problem of the Most
Efficient Tests of Statistical Hypotheses. \emph{Philosophical
Transactions of the Royal Society of London A: Mathematical, Physical
and Engineering Sciences}, \emph{231}(694-706), 289--337.
doi:\href{https://doi.org/10.1098/rsta.1933.0009}{10.1098/rsta.1933.0009}

\hypertarget{ref-Norman2003}{}
Norman, G. R., Sloan, J. A., \& Wyrwich, K. W. (2003). Interpretation of
Changes in Health-related Quality of Life: The Remarkable Universality
of Half a Standard Deviation. \emph{Medical Care}, \emph{41}(5),
582--592.

\hypertarget{ref-Perugini2014}{}
Perugini, M., Gallucci, M., \& Costantini, G. (2014). Safeguard Power as
a Protection Against Imprecise Power Estimates. \emph{Perspectives on
Psychological Science}, \emph{9}(3), 319--332.
doi:\href{https://doi.org/10.1177/1745691614528519}{10.1177/1745691614528519}

\hypertarget{ref-Piaggio2006}{}
Piaggio, G., Elbourne, D. R., Altman, D. G., Pocock, S. J., Evans, S. J.
W., \& CONSORT Group, for the. (2006). Reporting of Noninferiority and
Equivalence Randomized Trials: An Extension of the CONSORT Statement.
\emph{JAMA}, \emph{295}(10), 1152--1160.
doi:\href{https://doi.org/10.1001/jama.295.10.1152}{10.1001/jama.295.10.1152}

\hypertarget{ref-Platt1964}{}
Platt, J. R. (1964). Strong inference. \emph{Science}, \emph{146}(3642),
347--353.

\hypertarget{ref-Quertemont2011}{}
Quertemont, E. (2011). How to Statistically Show the Absence of an
Effect. \emph{Psychologica Belgica}, \emph{51}(2), 109.
doi:\href{https://doi.org/10.5334/pb-51-2-109}{10.5334/pb-51-2-109}

\hypertarget{ref-R-base}{}
R Core Team. (2017). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\hypertarget{ref-Rogers1993}{}
Rogers, J. L., Howard, K. I., \& Vessey, J. T. (1993). Using
significance tests to evaluate equivalence between two experimental
groups. \emph{Psychological Bulletin}, \emph{113}(3), 553.

\hypertarget{ref-RStudioTeam2016}{}
RStudio Team. (2016). \emph{RStudio: Integrated Development Environment
for R}. Boston, MA: RStudio, Inc.

\hypertarget{ref-R-lattice}{}
Sarkar, D. (2008). \emph{Lattice: Multivariate data visualization with
R}. New York: Springer. Retrieved from
\url{http://lmdvr.r-forge.r-project.org}

\hypertarget{ref-Schumann2017}{}
Schumann, S., Klein, O., Douglas, K., \& Hewstone, M. (2017). When is
computer-mediated intergroup contact most promising? Examining the
effect of out-group members' anonymity on prejudice. \emph{Computers in
Human Behavior}, \emph{77}, 198--210.
doi:\href{https://doi.org/10.1016/j.chb.2017.08.006}{10.1016/j.chb.2017.08.006}

\hypertarget{ref-Seaman1998}{}
Seaman, M. A., \& Serlin, R. C. (1998). Equivalence confidence intervals
for two-group comparisons of means. \emph{Psychological Methods},
\emph{3}(4), 403--411.
doi:\href{https://doi.org/http://dx.doi.org.dianus.libr.tue.nl/10.1037/1082-989X.3.4.403}{http://dx.doi.org.dianus.libr.tue.nl/10.1037/1082-989X.3.4.403}

\hypertarget{ref-Senn2007}{}
Senn, S. (2007). \emph{Statistical issues in drug development} (2nd
ed.). Chichester, England ; Hoboken, NJ: John Wiley \& Sons.

\hypertarget{ref-Shih1999}{}
Shih, M., Pittinsky, T. L., \& Ambady, N. (1999). Stereotype
Susceptibility: Identity Salience and Shifts in Quantitative
Performance. \emph{Psychological Science}, \emph{10}(1), 80--83.
doi:\href{https://doi.org/10.1111/1467-9280.00111}{10.1111/1467-9280.00111}

\hypertarget{ref-Simonsohn2015}{}
Simonsohn, U. (2015). Small Telescopes Detectability and the Evaluation
of Replication Results. \emph{Psychological Science}, \emph{26}(5),
559--569.
doi:\href{https://doi.org/10.1177/0956797614567341}{10.1177/0956797614567341}

\hypertarget{ref-Tryon2001}{}
Tryon, W. W. (2001). Evaluating statistical difference, equivalence, and
indeterminacy using inferential confidence intervals: An integrated
alternative method of conducting null hypothesis statistical tests.
\emph{Psychological Methods}, \emph{6}(4), 371--386.

\hypertarget{ref-UnitedStates2016}{}
United States, \& Central Intelligence Agency. (2016). \emph{The CIA
World Factbook 2017}. Washington, DC.

\hypertarget{ref-Weber2012}{}
Weber, R., \& Popova, L. (2012). Testing Equivalence in Communication
Research: Theory and Application. \emph{Communication Methods and
Measures}, \emph{6}(3), 190--213.
doi:\href{https://doi.org/10.1080/19312458.2012.703834}{10.1080/19312458.2012.703834}

\hypertarget{ref-R-ggplot2}{}
Wickham, H. (2009). \emph{ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}

\hypertarget{ref-Williams2008}{}
Williams, L. E., \& Bargh, J. A. (2008). Experiencing Physical Warmth
Promotes Interpersonal Warmth. \emph{Science}, \emph{322}(5901),
606--607.
doi:\href{https://doi.org/10.1126/science.1162548}{10.1126/science.1162548}

\hypertarget{ref-R-knitr}{}
Xie, Y. (2015). \emph{Dynamic documents with R and knitr} (2nd ed.).
Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
\url{https://yihui.name/knitr/}

\hypertarget{ref-R-bookdown}{}
Xie, Y. (2016). \emph{bookdown: Authoring books and technical documents
with R markdown}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
\url{https://github.com/rstudio/bookdown}

\end{document}
